### 2018-11-11 R/ICES advice and assessment summaries.Rmd
Merging of iAssess to iAdvice

---
output: 
  word_document:
    reference_docx: ../PFA_report_template_v1.5.dotx
---
```{r, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
```

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=6) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)
library(docxtractr) # read word documents
library(icesSD)    # ICES Stock database
library(icesSAG)   # ICES Stock Assessment Graphs

# Load utils code
source("../../mptools/r/my_utils.r")

# set dropboxdirs
assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")

# =====================================================================================
# Load Stockkey and Rename data
# =====================================================================================

load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))

# read and process the iAdvice data
iAdvice <- 
  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20180704.xlsx",sep=""), 
             sheet     = "DATA",
             col_names = TRUE,
             col_types = "text",
             skip      = 0)  %>% 
  lowcase() %>% 
  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
             funs(as.integer)) %>% 
  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
  mutate( blim     = 1000 * blim, 
          msybtrig = 1000 * msybtrig) %>%

  mutate_at( c("stockices","speciestype", "assessmodel", 
               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
               "aggregated"), funs(tolower)) %>% 
  rename(stockkeylabelold = stockices, 
         assessmentyear = assessyear) %>% 
  arrange(stockkeylabelold, tacarea, year) %>%  
  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
  left_join(iStockkey, by="stockkeylabelold") %>%

  filter(adviceonstock == "y")

# =====================================================================================
# Load Excel and QCS data
# =====================================================================================

excelqcs <-
  readxl::read_excel(
    path= paste(assessdir, "/data/QCS and EXCEL Assessment Database combined.xlsx", sep=""), 
    sheet = "data", col_names = TRUE, col_types = "text", trim_ws   = FALSE ) %>%
  lowcase() %>% 
  mutate_at(c("assessmentyear","year", 
              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
              "lowssb", "ssb", "highssb", 
              "lowf","f","highf", "flandings","fdiscards",
              "catches","discards", "landings"),   funs(as.numeric)) %>% 
  mutate_at(c("stockkey"),    funs(as.integer)) %>% 
  mutate_at(c("unitofrecruitment", "recruitmentdescription"),    funs(tolower)) %>% 
  mutate(recruitment       = ifelse(unitofrecruitment == "thousands", recruitment/1000, recruitment),
         unitofrecruitment = ifelse(unitofrecruitment == "thousands", "millions", unitofrecruitment),
         recruitment       = ifelse(unitofrecruitment == "billions", recruitment*1000, recruitment),
         unitofrecruitment = ifelse(unitofrecruitment == "billions", "millions", unitofrecruitment),
         source            = "excel") %>% 
  filter(purpose == "Advice")

excelqcs_unique <-
  excelqcs %>% 
  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  ungroup()

# unique(excelqcs$purpose)
# unique(icesSAG$purpose)
# filter(excelqcs, stockkeylabelold == "bss-47") %>% View()

# =====================================================================================
# Load SAG database; also unpublished ones
# =====================================================================================

icesSAG <- 
  get(load(file=file.path(assessdir, "rdata/icesSAG 20180703.RData"))) %>% 
  lowcase() %>%
  rename(stockkeylabel = fishstock) %>%
  left_join(iRename, by="stockkeylabel") %>%
  left_join(iStockkey, by="stockkey") %>% 
  mutate_at(c("assessmentyear","year", 
              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
              "lowssb", "ssb", "highssb", 
              "lowf","f","highf", 
              "catches","discards", "landings"),   funs(as.numeric)) %>% 
  mutate(source = "sag") %>% 
  filter(purpose == "Advice")

icesSAG_unique <-
  icesSAG %>% 
  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
  ungroup()

# =====================================================================================
# Merge SAG and excel database 
# =====================================================================================

iAssess <-

  # create excelqcs dataset to add
  dplyr::anti_join(select(excelqcs_unique, stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear),
                   select(icesSAG_unique,  stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear)) %>% 
  left_join(excelqcs, by=c("stockkey", "stockkeylabelold","stockkeylabelnew", "assessmentyear") ) %>% 
  data.frame() %>%

  # add icesSAG
  bind_rows(., icesSAG)

# glimpse(excelqcs_unique)
# glimpse(icesSAG_unique)
# glimpse(iAssess)

```

```{r functions,  message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# stock="her-47d3"
# firstassessyear=2014
# lastassessyear=2018
# firstyear=2000
# lastyear=2020

# Function to plot historical assessments
plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear) {

  d <-
    iAssess %>% 
    ungroup() %>% 
    filter(stockkeylabelold == stock,
           assessmentyear   %in% c(firstassessyear:lastassessyear), 
           year             %in% firstyear:lastyear) %>%

    mutate(tyear     = substr(as.character(assessmentyear), 3,4),
           type      = ifelse(assessmentyear == lastassessyear, "last", "other")) %>%
    select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
           type, source,
           lowrecruitment, recruitment, highrecruitment, 
           lowf, f, highf, 
           lowssb, ssb, highssb) %>%
    data.frame()

  # plot ssb
  p1 <-
    d %>% 
    filter(!is.na(ssb)) %>%

    ggplot(aes(year,ssb, group=tyear)) +

    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowssb, ymax=highssb, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
               colour="red", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes = FALSE) +

    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=msybtrig), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, msybtrig, label = "msy btrigger", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "SSB")

  # F
  p2 <-
    d %>% 
    filter(!is.na(f)) %>%

    ggplot(aes(year,f, group=tyear)) +

    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowf, ymax=highf, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=fmsy), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
               aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "F")

  # Recruitment
  p3 <-
    d %>% 
    filter(!is.na(recruitment)) %>%

    ggplot(aes(year,recruitment, group=tyear)) +

    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type=="last"), 
                aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    scale_colour_manual  (values=c(last = "red",other="black")) +
    scale_fill_manual    (values=c(last = "red",other="black")) +
    scale_linetype_manual(values=c(last="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "Recruitment")

  # Combined plot
  # plot_grid(p1 + theme(legend.position  = "none", 
  #                    axis.title       = element_blank(),
  #                    strip.background = element_rect(colour=NA, fill=NA),
  #                    strip.text       = element_text(colour=NA)), 
  #         p2 + theme(axis.title       = element_blank()),
  #         p3 + theme(axis.title       = element_blank()),
  #         ncol=3, align = 'h')

    plot_grid(p1, p2,  ncol=2, align = 'hv')

  } # End of plot_history function

# firstadviceyear <- 2016
# lastadviceyear  <- 2019
# stock           <- "her-47d3"

# Function to plot advice overviews
table_advice <- function(stock,firstadviceyear,lastadviceyear) {

  d <-
    iAdvice %>% 
    filter(stockkeylabelold == stock, 
           year %in% firstadviceyear:lastadviceyear) %>%
    mutate(advicebasis= substr(advicebasis,1,30),

           advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),

           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"),

           catch      = ifelse(!is.na(icescatch), icescatch, NA),
           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),

           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),

           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),

           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           catch_c    = 100 * (catch / lag(catch, n=1) -1)
           ) %>%

    select(year, advicebasis, advice, fadvmax, advicemax, tac, catch, 
           advice_c, tac_c, catch_c,
           advicetype, tactype, catchtype)

  d %>% 
  select(1, 2, 4, 5, 6, 7) %>%
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               justify      = "right",
               missing      =" ",
               big.mark     = '', 
               round        = c(0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

table_advice2 <- function(stock,firstadviceyear,lastadviceyear) {

  iAdvice %>% 
    filter(stockkeylabelold == stock, 
           year %in% firstadviceyear:lastadviceyear) %>%
    mutate(advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),

           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"),

           catch      = ifelse(!is.na(icescatch), icescatch, NA),
           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),

           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),

           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),

           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           catch_c    = 100 * (catch / lag(catch, n=1) -1)
           ) %>%

    select(year, advice_c, tac_c, catch_c, advicetype, tactype, catchtype) %>% 
    pandoc.table(., 
                 style        = "simple",
                 split.tables = 100, 
                 split.cells  = c(rep(7,10)),
                 justify      = "right",
                 missing      =" ",
                 big.mark     = '', 
                 round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

# table_advice(stock="her-47d3",firstadviceyear=2015,lastadviceyear=2019)

```

**Summary of ICES advice and assessments**

*June 2018 advice round*

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

<!--Section 1: introduction ------------------------------------------------------------------- -->

# Introduction

In this document an overview is presented of the ICES assessment and advice for the most recent number of years for the stocks that are relevant to the PFA and Dutch fisheries in general.

<!--Section 2: material and methods ----------------------------------------------------------- -->

# Material and methods

The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

<!--Section 3: advice and assessment overviews ------------------------------------------------ -->

# Assessment and advice overviews

## North Sea herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}

s   <- "her-47d3"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)

```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Assessment benchmarked in 2018. New natural mortality estimates lead to higher SSB and lower F. New method to standardize natural mortality estimates (no more scaling revision in future?). Revision of reference points. Weak 2014 year class leads to substantially lower advice, which was already foreseen in 2017. 
[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.3a47d.pdf)

## Western Baltic Spring Spawning herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-3a22"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Assessment benchmarked in 2018. Revision of reference points. Redefinition of Blim from 90 kT to 120 kT which leads to stock having been below Blim for a number of years already. No possibility of rebuilding stock to Blim within one year (which is the unwritten new ICES rule), and therefore a zero catch advice. An additional analysis will be undertaken (by MP) to evaluate alternative options for the zero catch advice. 
[her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.20-24.pdf)

## Baltic Sea herring

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-2532-gor"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Substantially lower estimate of SSB and higher estimate of fishing mortality in the most recent assessment is probably due to the variability in the survey index. The 2014 year class is estimated to be strong.  
[her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.25-2932.pdf)

## Herring in 6a and 7bc

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-67bc"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The advice for a zero TAC is the same as in the previous years. Unfortunately ICES has not presented advice on the proposed rebuilding plan that has been developed by the PELAC and that has been submitted for review to ICES and that had been evaluated by the herring working group in March 2018. In previous years the monitoring TAC for herring in 6a and 7bc had been suggested by ICES as a means to maintain the knowledge base for this herring stock. The monitoring TAC is still mentioned under issues relevant for the advice. 
[her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.6a7bc.pdf)

## Herring in the Irish Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-nirs"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea. 
[her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.nirs.pdf)

## Herring in the Celtic Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "her-irls"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year. Evaluations conducted in 2018 showed that the agreed management strategy is no longer precautionary. Therefore the advice is according to the MSY approach. 
[her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.irls.pdf)

## Sprat in the North Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "spr-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The change in recruitment level between this year’s assessment and the previous years is caused by a recent update of natural mortality of sprat in the first six months of their life. The 2018 assessment also revised the 2017 SSB downwards. The assessment has a tendency to overestimate above-average recruitments.
[spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.4.pdf)

## Sprat in the Baltic Sea

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "spr-2232"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Advice according to the Baltic Multiannual Plan (MAP). 
[spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.22-32.pdf)

## Greater argentine in 5b and 6a

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "arg-5b6a"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Last assessment was in 2017. Biennial assessment based on survey trends")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019. 
[aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)

## Northern Hake

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "hke-nrtn"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size is well above MSY Btrigger and fishing mortality is below Fmsy. In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years. 
[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hke.27.3a46-8abd.pdf)

## North Sea cod

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "cod-347d"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years. According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey. The 2017 SSB is 32% lower compared to the estimate that was presented last year. This also leads to a substantially lower advice for 2019 compared to the advice for 2018. 
[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/cod.27.47d20.pdf)

## North Sea plaice

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "ple-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy. An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016). This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock. The advice is based on the MSY approach. 
[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/ple.27.420.pdf)

## North Sea sole

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "sol-nsea"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/sol.27.4.pdf)

## Atlantic Seabass

```{r, echo=FALSE, message=FALSE, warning=FALSE}
s   <- "bss-47"
fay <- 2014; lay <- 2018
fy  <- 2000; ly  <- 2018

cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
```

**History of stock assessments**

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
```

**Advice and management overview**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
```

**Comments**

ICES has reissued the advice for 2018 and issued an advice for 2019. ICES was requested by the EU to provide updated advice on fishing opportunities for 2018 because a benchmark of the assessment had been conducted in early 2018. The 2018 benchmark has not changed the perception of the current SSB (B below Blim). However, given the current low fishing mortality and the above-average recruitment in 2013 and 2014, SSB in 2019 and 2020 is predicted to be above Blim, allowing for some catches in 2018 and 2019. 
[bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/bss.27.4bc7ad-h.pdf)

# More information

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.

### 2018-11-11 R/old/ICES historical assessment analysis.Rmd
Merging of iAssess to iAdvice

---
output: 
  word_document:
    reference_docx: "C:/Users/Martin Pastoors/AppData/Roaming/Microsoft/Templates/Normal.dotm"
---

```{r setup, include=FALSE}
# -----------------------------------------------------------------------------------------------
# ICES historical assessment analysis.Rmd
#
# stocknames: lookup list to transfer old and new stock names
# stocklist : stock list with stock description and advice status by assessment year
# speciesdb : properties of species (names etc.)
# stockdb   : properties of stocks
# sagdb     : stock assessment database (by stock, assessmentyear and year)
#
# 30/03/2017 first coding during HAWG
# 07/07/2017 adapted for downloading as csv file
# 14/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
# 20/07/2017 thorough checking of databases and making sure all items are filled and OK
# 01/08/2017 added a number of diagnostic graphs; redone the excel database for nephrops
# 01/08/2017 added the stock assessment methods from stock advice database
# 10/08/2017 moved checking routines to separate code
# 11/08/2017 adapted for R3.4.1 and Tidyverse
# 14/08/2017 added assessmentmodel and assessmenttype categories
# 29/08/2017 converted to rmarkdown document
# -----------------------------------------------------------------------------------------------

rm(list=ls())

require("knitr")
knitr::opts_chunk$set(crop = TRUE, echo = FALSE, message=FALSE, warning=FALSE)
knit_hooks$set(crop = hook_pdfcrop)
opts_knit$set(root.dir="D:/Dropbox/ICES Assessment database")

# Libraries
library(rmarkdown)

library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours

library(icesSD)    # ICES Stock database
library(icesSAG)   # ICES Stock Assessment Graphs

# Load utils code
source("D:/XXX/PRF/r/my_utils.r")

```

# Generating and analyzing the ICES historical assessment database

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

## Introduction

Ambition: bringing together the ICES historical assessment database and the ICES webservices for the Standard Assessment Graphs (SAG).

The ICES historical assessment database was originally started by Martin Pastoors already in the early 2000s and later maintained by the ICES secretariat (Barbara Schoute) by collecting the annual assessment results from the standard graph database into an overall excel file. This was used to generate e.g. the historical spaghetti plots in the ACFM and later ACOM advice, before it was taken over by the SAG database.

In addition, there is the ICES Advice Database (IAD) which has been set up about five years ago to attempt to capture the ICES advice in combinations with metadata on the short term forecast, reference points and assessment methods. This will be used here to contribute some information on reference points and assessment methods. The advice part will not come into play (yet).

Ultimately the intention is to evaluate the long term historical performance of stock assessments and to consider variability, bias and changes in perception in relation to methods, benchmarks and other explanatory variables.

For now, the focus is on the method of combining the data from different sources.

## The first challenge

The first challenge is to agree on the names for the stocks. Recently ICES has renamed all the stocks to a new name that is more consistent with the area distribution, but also much more difficult to remember. Think about the move from **hom-west** to **hom.27.2a4a5b6a7a-ce-k8**. In order to maintain cohesion in the historical assessment database, a translation key is used to be able to convert old names to new names and vice versa.

```{r initialization, echo=TRUE, message=FALSE, warning=FALSE}
stocknames <-
  read.csv("D:/Dropbox/ICES Assessment database/ICES StocksOldCodesNewCodesEGs.csv", 
           stringsAsFactors = FALSE) %>% 
  select(1,2) %>% 
  setNames(., c("fishstockold","fishstock")) 
glimpse(stocknames)
```

## Getting information on stock assessment process and reference points from the ICES Advice Database

The ICES Advice Database (IAD) is a database set up by Martin Pastoors to capture the advice tables in the advisory reports. To this have been added: information on short term forecasts, information on reference points and information on the assessment process. Here we will only use the reference points and the assessment methods.

```{r iad, echo=TRUE, message=FALSE, warning=FALSE}
load(file="../ICES advice database/rdata/iad.RData")
iad <-
  iad %>% 
  select(fishstock, assessmentyear, assessmentmodel = assessmodel,
         flim, fpa, fmsy, blim, bpa, msybtrig,
         firstyearofdata, ncpueseries, nsurveyseries, 
         assessmenttype=assesstype, assessmentcomment=assesscomment) %>% 
  filter(!is.na(assessmentmodel), assessmentmodel != "")
glimpse(iad)
```

## Turning to the ICES webservices

So what is actually in the ICES webservices for Stock Assessment Graphs (or at least what I could find in it):

* icesSD::getSD (Get Stock List Data). 
* icesSAG::getListStocks (Again a list of stocks with properties; overlap with getSD?)
* icesSAG::getSAG(..., data="refpts", ...) (Reference points)
* icesSAG::getSAG(..., data="summary", ...) (Assessment results)

## Get Stock List Data (getSD)

The Stock List Data consists of information that is specific to species (e.g. scientific name, common name, fisheries guild etc.) that does not change over time and information on stocks that is year specific (advice type, assessment method, section etc.). In the analysis I have split them into two separate datasets. First I read in the overall stock list data

```{r getsd, echo=TRUE, message=FALSE, warning=FALSE}
s <-
  icesSD::getSD() %>%
  lowcase %>%
  rename(fishstock      = stockkeylabel,
         assessmentyear = activeyear,
         scientificname = speciesscientificname,
         commonname     = speciescommonname) %>%
  mutate   (fishstock   = gsub(" ", "", fishstock))
glimpse(s)
```

Then I have created a separate species database (not by year):

```{r speciesdb, echo=TRUE, message=FALSE, warning=FALSE}
speciesdb <-
  readxl::read_excel(path="downloads/species_list.xlsx", col_names=TRUE, col_types="text") %>%
  mutate_at(c("commonname","trophicguild","fisheriesguild","sizeguild"), funs(tolower)) %>%
  group_by(faocode, scientificname, commonname) %>%
  filter(row_number() == 1) %>%
  select(faocode, scientificname, commonname, trophicguild, fisheriesguild, sizeguild) %>%
  arrange(faocode) %>%
  filter(!is.na(faocode) & !is.na(scientificname) & !is.na(commonname))
glimpse(speciesdb)
```

And a dataset by stock

```{r stockdb, echo=TRUE, message=FALSE, warning=FALSE}
stockdb <-
  s %>%
  mutate(faocode = substr(fishstock,1,3)) %>%
  select(fishstock, assessmentyear, ecoregion, expertgroup, advicedraftinggroup, datacategory,
         assessmentfrequency, assessmenttype, advicecategory, advicetype, useofdiscardsinadvice,
         pabufferapplied) %>%

  # lowercase and remove na text strings
  mutate_at(vars("assessmenttype","advicecategory","advicetype",
                 "useofdiscardsinadvice", "pabufferapplied"), funs(tolower)) %>%
  mutate(assessmenttype  = ifelse(assessmenttype == "na",NA, assessmenttype),
         pabufferapplied = ifelse(pabufferapplied == "na", NA, pabufferapplied)) %>% 
  rename(assessmentmodel = assessmenttype)
glimpse(stockdb)
```

## icesSAG

However, icesSAG also has a stock list

```{r stocklist, echo=TRUE, message=FALSE, warning=FALSE}
# p <- icesSAG::getListStocks(year=myyear)

# shortcut
p <- read.csv(file="downloads/stocklist.csv", stringsAsFactors = FALSE)
glimpse(p)

stocklist <-
  p %>%
  lowcase %>%
  rename(fishstock = stockkeylabel) %>%
  select(fishstock, assessmentyear, stockdescription, status)
glimpse(stocklist)
```

Then we get to the reference points (straightforward)

```{r refpoints, echo=TRUE, message=FALSE, warning=FALSE}
# t <- getSAG(stock=NULL, year=0, data="refpts", combine=TRUE)

# shortcut
t <- read.csv(file="downloads/refpoints.csv", stringsAsFactors = FALSE)
refpoints <-
  t %>%
  lowcase %>%
  rename(fishstock = stockkeylabel) %>%
  dplyr::select(fishstock, assessmentyear,
                flim, blim, fmsy, msybtrigger, fpa, bpa,
                fmgt=fmanagement, bmgt=bmanagement,
                recruitmentage) %>%
  mutate_at(vars(flim:bmgt), funs(as.numeric))
glimpse(refpoints)
```

And then to the piece-de-resistance: the stock assessment database. After reading the database we factor in all the previous datasets to create one overall dataset of all available information by stocks, assessmentyear and year.

```{r sagdownload, echo=TRUE, message=FALSE, warning=FALSE}
# sagdownload <- icesSAG::getSAG(stock=NULL, year=0, data="summary", combine=TRUE)

# shortcut
sagdownload <- 
  read.csv(file="downloads/sagdownload.csv", stringsAsFactors =FALSE) %>% 
  lowcase %>% 
  dplyr:: select(fishstock, assessmentyear, year,
                 recruitment, highrecruitment, lowrecruitment,
                 ssb, highssb, lowssb,
                 f, highf, lowf, 
                 catches, landings, discards,
                 fage, 
                 units, 
                 stocksizedescription, stocksizeunits,
                 fishingpressuredescription, fishingpressureunits,
                 stockpublishnote) %>%

  # add fao code if missing (NEEDS A MORE SOFISTICATED APPROACH)
  mutate(faocode = substr(fishstock,1,3)) %>%

  # dealing with old and new stocknames
  left_join(stocknames, by = c("fishstock" = "fishstockold")) %>% 
  rename(fishstocknew = fishstock.y) %>% 
  left_join(stocknames, by = c("fishstock")) %>% 
  mutate(fishstocknew = ifelse(is.na(fishstocknew) & !is.na(fishstockold), fishstock, fishstocknew),
         fishstockold = ifelse(is.na(fishstockold) & !is.na(fishstocknew), fishstock, fishstockold),
         fishstockold = ifelse(is.na(fishstockold) & is.na(fishstocknew) , fishstock, fishstockold)) %>% 
  mutate_at(vars("stocksizedescription","stocksizeunits","fishingpressuredescription",
                 "fishingpressureunits"), funs(tolower)) %>%

  # join relevant datasets
  left_join(refpoints, by=c("fishstock","assessmentyear")) %>% 
  left_join(stocklist, by=c("fishstock","assessmentyear")) %>% 
  left_join(stockdb  , by=c("fishstock","assessmentyear")) %>% 
  left_join(speciesdb, by=c("faocode")) %>% 
  left_join(iad,       by=c("fishstock","assessmentyear")) %>%

  mutate(source   = "download", 
         flim     = ifelse(!is.na(flim.x), flim.x, flim.y),
         fpa      = ifelse(!is.na(fpa.x),  fpa.x, fpa.y),
         fmsy     = ifelse(!is.na(fmsy.x), fmsy.x, fmsy.y),
         blim     = ifelse(is.na(blim.x),  blim.x, blim.y),
         bpa      = ifelse(is.na(bpa.x),   bpa.x, bpa.y),
         msybtrig = ifelse(is.na(msybtrigger),  msybtrigger, msybtrig),
         assessmentmodel = ifelse(!is.na(assessmentmodel.y), assessmentmodel.y, assessmentmodel.x) ) %>%

  select(-flim.x, -flim.y, -fpa.x, -fpa.y, -fmsy.x, -fmsy.y, -blim.x, -blim.y, -bpa.x, -bpa.y, 
         -msybtrigger, -assessmentmodel.x, -assessmentmodel.y) %>% 
 data.frame() 
glimpse(sagdownload)
```

## ICES historical assessment database

The ICES historical assessment database consists of a large excel spreadsheet with the historial stock assessment of many stocks sometimes going back to the mid 1990s.

```{r sagexcel, echo=TRUE, message=FALSE, warning=FALSE}
sagexcel <-
  readxl::read_excel("ICES Assessment Summary database.xlsx",
             sheet = "DATA", col_names = TRUE, col_types = "text", skip = 0) %>%
  lowcase %>%
  rename(assessmentyear = assyear) %>% 
  mutate(fishstock = tolower(fishstock)) %>%
  mutate_at(vars("year","assessmentyear"), funs(as.integer)) %>% 
  mutate_at(vars("lowrecruitment", "recruitment","highrecruitment",
                 "lowssb","ssb","highssb",
                 "lowf", "f","highf",
                 "landings","catches","discards","ibc",
                 "flim","fpa","fmsy", "fmanagement",
                 "blim","bpa","msybtrigger","bmanagement",
                 "recruitmentage","recruitmentlength"), funs(as.numeric)) %>% 
  select(-contains("custom")) %>% 
  rename(scientificname = speciesname,
         commonname     = sgname,
         fmgt           = fmanagement,
         bmgt           = bmanagement,
         faocode        = species) %>%

  # dealing with old and new stocknames
  left_join(stocknames, by = c("fishstock" = "fishstockold")) %>% 
  rename(fishstocknew = fishstock.y) %>% 
  left_join(stocknames, by = c("fishstock")) %>% 
  mutate(fishstocknew = ifelse(is.na(fishstocknew) & !is.na(fishstockold), fishstock, fishstocknew),
         fishstockold = ifelse(is.na(fishstockold) & !is.na(fishstocknew), fishstock, fishstockold),
         fishstockold = ifelse(is.na(fishstockold) & is.na(fishstocknew) , fishstock, fishstockold)) %>%

  mutate_at(vars("stocksizedescription","stocksizeunits","fishingpressuredescription",
                 "fishingpressureunits","commonname", "scientificname"), funs(tolower)) %>% 
  select(-assessmentkey, -icesareas, -report, -(unitofrecruitment:tbiomassunit), 
         -catcheslandingsunits, -officiallandings, -(ibc:yieldssb), -(flandings:funallocated), 
         -(flength:area), -scientificname, -commonname) %>% 
  filter(year <= assessmentyear) %>%

  # add fao code if missing
  mutate(faocode = substr(fishstock,1,3)) %>%

  # add species information
  left_join(speciesdb, by=c("faocode")) %>%

  # add ICES Advice Database information
  select(-blim, -bpa, -msybtrigger,-flim, -fpa, -fmsy) %>% 
  left_join(iad, by=c("fishstock","assessmentyear")) %>%

  mutate(source = "excel") 
glimpse(sagexcel)
```

## Combining the ICES historical assessment database with the SAG download database

First step: we create unique combinations of fishstock and assessment year for each of the two databases

```{r uniquesets, echo=TRUE, message=FALSE, warning=FALSE}
sagdownload_unique <-
  sagdownload %>% 
  group_by(fishstock, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(fishstock, assessmentyear)

sagexcel_unique <-
  sagexcel %>% 
  group_by(fishstock, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  select(fishstock, assessmentyear)
```

And then we calculate the which combinations of fishstock and assessment year are not already included in the SAG download but are available in the Excel version.

```{r exceltoadd, echo=TRUE, message=FALSE, warning=FALSE}
sagexcel_toadd <-
  setdiff(sagexcel_unique,sagdownload_unique) %>% 
  left_join(sagexcel, by=c("fishstock","assessmentyear")) %>% 
  data.frame()
```

Finally we add the 'excel' data to the downloaded data and there are a number of cleanup actions to deal with some of the inconsistencies in the database (notably naming of methods, stocksize description etc. )

```{r sagdb, echo=TRUE, message=FALSE, warning=FALSE}
sagdb <- 
  rbind.all.columns(sagdownload, sagexcel_toadd) %>%

  # corrections to stock size descriptions
  mutate(
    stocksizedescription = gsub("stock size: "                 ,""                   , stocksizedescription),
    stocksizedescription = gsub("indices"                      ,"index"              , stocksizedescription),
    stocksizedescription = gsub("indicator"                    ,"index"              , stocksizedescription),
    stocksizedescription = gsub("^biomass$"                    ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^biomass index$"              ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^evhoe biomass index$"        ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^stock size index: abundance$","abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^abundance$"                  ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^density$"                    ,"density index"      , stocksizedescription),
    stocksizedescription = gsub("^tsb$"                        ,"total biomass"      , stocksizedescription),
    stocksizedescription = gsub("^total abundance index$"      ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^index$"                      ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^stock abundance$"            ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^density index$"              ,"abundance index"    , stocksizedescription),

    stocksizedescription = gsub("stock size index: biomass \\(ages 1-8\\)"   ,"total biomass index"      , stocksizedescription),
    stocksizedescription = gsub("stock size index: german survey"            ,"total biomass index"      , stocksizedescription),
    stocksizedescription = gsub("stock size index: smoothed greenland index" ,"total biomass index"      , stocksizedescription),

    stocksizedescription = ifelse(grepl("tv",stocksizedescription), "abundance index", stocksizedescription),
    stocksizedescription = ifelse(grepl("ssb & b",stocksizedescription) , "ssb", stocksizedescription),
    stocksizedescription = ifelse(grepl("total biomass/bmsy",stocksizedescription) , "b/bmsy", stocksizedescription),
    stocksizedescription = ifelse(stocksizedescription == "stock size" & stocksizeunits == "tonnes", "ssb", stocksizedescription),
    stocksizedescription = ifelse(stocksizedescription == "stock size" & grepl("kg/",stocksizeunits), "total biomass index", stocksizedescription),
    stocksizedescription = ifelse(grepl("relative", stocksizeunits, fixed=TRUE) & is.na(stocksizedescription), "total biomass index", stocksizedescription)
  ) %>%

  # corrections to stock units
  mutate(
    stocksizeunits = gsub("stock size: "                 ,""                   , stocksizeunits),
    stocksizeunits = gsub(" ", "", stocksizeunits),
    stocksizeunits = gsub("density(burrows/m2)", "burrows/m2", stocksizeunits, fixed=TRUE),
    stocksizeunits = gsub("cpue(kg/1000hooks)", "kg/1000hooks", stocksizeunits, fixed=TRUE),
    stocksizeunits = gsub("^abundance$", "millions", stocksizeunits),
    stocksizeunits = gsub("na(ratio)", "relative", stocksizeunits, fixed=TRUE),
    stocksizeunits = ifelse(grepl("kg/h", stocksizeunits, fixed=TRUE), "kg/hour", stocksizeunits),
    stocksizeunits = ifelse(grepl("n/h", stocksizeunits, fixed=TRUE), "n/hour", stocksizeunits)
  ) %>%

  # corrections to fishing pressure descriptions
  mutate(
    fishingpressuredescription = gsub("fishing pressure: ",""  , fishingpressuredescription),
    fishingpressuredescription = gsub(" ",""  , fishingpressuredescription),
    fishingpressuredescription = gsub("f&hr","f"  , fishingpressuredescription, fixed=TRUE),
    fishingpressuredescription = gsub("fishingpressure","f"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("finwinterrings","f"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("weightedf","f"  , fishingpressuredescription),

    fishingpressuredescription = gsub("harvestrate","hr"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("relativehr","hr/index"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("hrindex","hr/index"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("relativeexploitationrate","hr/index"  , fishingpressuredescription),

    fishingpressuredescription = ifelse(grepl("ages",fishingpressuredescription), "f", fishingpressuredescription),
    fishingpressuredescription = ifelse(grepl("null",fishingpressuredescription), NA, fishingpressuredescription),

    fishingpressureunits       = ifelse(grepl("relative",fishingpressuredescription) & is.na(fishingpressureunits), "relative", fishingpressureunits ),
    fishingpressuredescription = ifelse(grepl("relative",fishingpressuredescription) , "fproxy", fishingpressuredescription )
  ) %>%

  # corrections to fishing pressure units
  mutate(
    fishingpressureunits = gsub(" ",""  , fishingpressureunits),
    fishingpressureunits = gsub("peryear","year-1"  , fishingpressureunits),
    fishingpressureunits = gsub("%","percent"  , fishingpressureunits),
    fishingpressureunits = gsub("^f$","year-1"  , fishingpressureunits),
    fishingpressureunits = gsub("^catch/biomass$","relative"  , fishingpressureunits),

    fishingpressureunits = ifelse(grepl("cm",fishingpressureunits), "year-1",fishingpressureunits) ,
    fishingpressureunits = ifelse(grepl("null",fishingpressureunits), NA,fishingpressureunits) ,
    fishingpressureunits = ifelse(grepl("ratio",fishingpressureunits), "relative",fishingpressureunits) 
  ) %>%

  # Add assessment type2 category (assess, bench, old, alt, explore)
  # CHECK: Should I remove the labels from the fishstock variable  ??
  mutate(
    assessmenttype2 = ifelse(grepl("-bench$" , fishstock), "bench", "assess"),
    assessmenttype2 = ifelse(grepl("-old$"   , fishstock), "old"  , assessmenttype2),
    assessmenttype2 = ifelse(grepl("-alt$"   , fishstock), "alt"  , assessmenttype2)
  ) %>%

  # correction due to missing units
  mutate(
    stocksizeunits       = ifelse(stocksizedescription=="b/bmsy" & is.na(stocksizeunits),"relative",stocksizeunits),  
    fishingpressureunits = ifelse(fishingpressuredescription=="f/fmsy" & is.na(fishingpressureunits),"relative",fishingpressureunits)
  ) %>%

  # corrections to the assignments of specific stocks and years
  mutate(
    stocksizedescription       = ifelse(fishstock=="anb-8c9a" & assessmentyear==2013,"b/bmsy"  ,stocksizedescription),
    stocksizeunits             = ifelse(fishstock=="anb-8c9a" & assessmentyear==2013,"relative",stocksizeunits),
    fishingpressuredescription = ifelse(fishstock=="anb-8c9a" & assessmentyear==2013,"f/fmsy"  ,fishingpressuredescription),
    fishingpressureunits       = ifelse(fishstock=="anb-8c9a" & assessmentyear==2013,"relative",fishingpressureunits)
  ) %>%

  # remove double series (e.g. mac-nea 2013 is twice in the sag download)
  group_by(fishstock, assessmentyear, year) %>% 
  filter(row_number() == 1) %>%

  # convert to lowercase
  mutate_at(vars(assessmenttype), funs(tolower)) %>%

  # remove empty variables
  select(-recruitmentlength)
```

First overview of what we have (very ugly !!)

```{r firstoverview, echo=TRUE, message=FALSE, warning=FALSE}
sagdb %>% 
  group_by(fishstock, assessmentyear) %>% 
  filter(row_number()==1) %>% 
  filter(assessmentyear >= 2001) %>% 
  mutate(species = paste0(commonname," (", faocode,")"), sep="") %>% 
  group_by(species, assessmentyear) %>% 
  summarise(nassess = n()) %>%

  dcast(species ~ assessmentyear, sum, value.var="nassess", margins = TRUE) %>% 
  no.emphasis.table() %>% 
  select(-1) %>% 
  pandoc.table(.,
               style = "grid",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
```

Better to use a plot

```{r firstplot, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=15}
x <-
  sagdb %>% 
  filter(assessmentyear >= 1990) %>% 
  filter(fisheriesguild != "", !is.na(fisheriesguild)) %>% 
  filter(fisheriesguild %in% c("crustacean","elasmobranch","pelagic","demersal","benthic")) %>%
  mutate(fishstockold = ifelse(is.na(fishstockold)|fishstockold=="", fishstocknew, fishstockold)) %>% 
  group_by(fishstockold, fishstocknew, assessmentyear, source) %>% 
  data.frame() %>% 
  mutate(fishstockold = factor(fishstockold), 
         fishstockold = factor(fishstockold, levels = rev(levels(fishstockold)))) %>%

  # here is the bit to add the colour variable
  mutate(source = factor(source))

# define colour scale
myColors        <- brewer.pal(length(levels(x$source)),"Set1")
names(myColors) <- levels(x$source)

p1 <-
  filter(x, fisheriesguild %in% c("crustacean","elasmobranch")) %>% 
  ggplot(aes(x=assessmentyear, y=fishstockold)) +
  theme_publication() +
  theme(panel.spacing = unit(1, "lines"),
        text          = element_text(size=8),
        legend.title  = element_blank()) +
  geom_point(aes(colour = source)) +
  scale_colour_manual(name = "source", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  labs(x = " ", y = NULL ) +
  facet_wrap(~fisheriesguild, scales="free_y", shrink=TRUE, ncol=1)

p2a <-
  filter(x, fisheriesguild %in% c("pelagic")) %>% 
  ggplot(aes(x=assessmentyear, y=fishstockold)) +
  theme_publication() +
  theme(panel.spacing = unit(1, "lines"),
        text          = element_text(size=8),
        legend.title  = element_blank()) +
  geom_point(aes(colour = source)) +
  scale_colour_manual(name = "source", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  labs(x = NULL, y = NULL ) +
  facet_wrap(~fisheriesguild, scales="free_y", shrink=TRUE, ncol=1)

p2b <-
  filter(x, fisheriesguild %in% c("benthic")) %>% 
  ggplot(aes(x=assessmentyear, y=fishstockold)) +
  theme_publication() +
  theme(panel.spacing = unit(1, "lines"),
        text          = element_text(size=8),
        legend.title  = element_blank()) +
  geom_point(aes(colour = source)) +
  scale_colour_manual(name = "source", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  labs(x = "assessmentyear", y = NULL ) +
  facet_wrap(~fisheriesguild, scales="free_y", shrink=TRUE, ncol=1)

p2 <- plot_grid(p2a + 
                  theme(legend.position="none", 
                        axis.title.y=element_blank(), 
                        axis.text.x=element_blank(), 
                        plot.margin = unit(c(10, 5, 1, 5), "mm")), 
                p2b + 
                  theme(legend.position="none", 
                        axis.title.y=element_blank(), 
                        plot.margin = unit(c(1, 5, 5, 5), "mm")),
                ncol=1,align="v",
                rel_heights = c(4.7, 2))

p3 <-
  filter(x, fisheriesguild %in% c("demersal")) %>% 
  ggplot(aes(x=assessmentyear, y=fishstockold)) +
  theme_publication() +
  theme(panel.spacing    = unit(0.1, "lines"),
        text             = element_text(size=8),
        legend.position  = "right",
        legend.direction = "vertical",
        legend.title     = element_blank()) +
  geom_point(aes(colour = source)) +
  scale_colour_manual(name = "source", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  labs(y = "fishstock", x=" " ) +
  facet_wrap(~fisheriesguild, scales="free_y", shrink=TRUE)

plot_grid(p1 + theme(legend.position="none") + theme(axis.title.y=element_blank()), 
          p2 ,
          p3 ,
          ncol=3,
          align = 'v', rel_widths = c(1,1,1.2))

```

## To do

* Move the data and r code to github

```{r Filter trip data, echo=FALSE, message=FALSE, warning=FALSE}

```

### 2018-11-11 R/old/iAssess databases.Rmd
Merging of iAssess to iAdvice

---
output: 
  word_document:
    reference_docx: "my_template.dotx"
---

```{r, include=FALSE}

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knit_hooks$set(crop = hook_pdfcrop)
rm(list=ls())

```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# -----------------------------------------------------------------------------------------------
# iAssess databases.Rmd
#
# 27/11/2017 converted from iAssess analysis markdown
# 02/02/2017 generated first version for discussion with ICES
# 07/12/2017 now QCS data is also stored on dropbox; code adapted for that
# 02/01/2018 Working on assessment method coding; renamed some of the variables in the input files
# 04/01/2018 coding the cod assessments in iStock.xlsx
# 05/01/2018 now combined QCS and Excel data into one Excel file, to allow for manual alterations
# -----------------------------------------------------------------------------------------------

# Libraries
library(rmarkdown)

library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(docxtractr) # read word documents
library(icesSD)    # ICES Stock database
library(icesSAG)   # ICES Stock Assessment Graphs

# Load utils code
source("../../mptools/r/my_utils.r")

# set dropboxdir
dropboxdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
```

# Description of the databases leading up to the iAssess database

## Documenting the past of stock assessment in ICES

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

## 1. Introduction

The iAssess database system aims to bring together the past of stock assessments within the ICES advisory system. To that end, the iAssess database combines three main sources of data: (1) the historical Quality Control Sheets (QCS) that were used in the 1990s, (2) the annual data of the 'standard graphs' that were originally started by Martin Pastoors in the early 2000s and later maintained by the ICES secretariat (Barbara Schoute), and (3) the ICES Standard Assessment Graph webservices that have been started around 2014.

The development of the iAssess database is a joint effort by several fisheries scientists working in different parts of the ICES system, notably:

* Martin Pastoors, fisheries scientist at the Pelagic Freezer-Association, Netherlands 
* Einar Hjorleifsson and Siggi Þór Jónsson, MFRI, Iceland
* David Miller, Colin Millar, Carlos Pinto, Mike Drew and Adriana Villamor @ ICES secretariat
* Esther Schuch, PhD student at Wageningen University, Netherlands
* Barbara Schoute @ Natuurmonumenten, Netherlands.

## 2. General outline

### 2.1 Data storage and coding

Data storage is currently on the following Dropbox folder: https://www.dropbox.com/home/ICES%20Assessment%20database. Here the basic Excel files are stored and the rdata files generated in the processing of data.

The coding is stored on the following github folder: https://github.com/martinpastoors/iAssess. Here all the r codes are kept and also the Quality Control Sheets (not really consistent to keep them on github!)

In addition we use some utily codes from https://github.com/martinpastoors/mptools, e.g.: get_dropbox, ...

### 2.2 Naming conventions

ICES is an organization that was established in 1902. From the beginning it has had a strong focus on fisheries science and stock assessment. With that history, comes a tendency to develop standard practices to do and to call certain things in certain ways. For the purpose of the iAssess stock assessment database, the key element for linking objects has been the shortname association with the different stocks.

Unfortunately, the year 2017, when the iAssess effort really took off, has also been the year in which this tradition of calling the stocks in the same way, was left aside. New names, supposedly more consistent names, were developed for the stocks. Think about the move from **hom-west** to **hom.27.2a4a5b6a7a-ce-k8**. This brought about an extra challenge for the iAssess project for now only were we to bring together the historical assessment, we also needed to make sure that new names for similar stocks would not jeopardize the historical comparisons. In the renaming of the old names to the new names, not only the names changed, but also the stockkeys. Therefore an approach was followed to apply the same stockkey to the new name and old name, if the stocks were clearly the same, but with a new name. When there was no old stockkey available (because the old stockkey was not used prior to the renaming), a manual stockkey would be chosen, starting from 100000 and beyond.

In addition, for each of the stocks in the database, it is now possibly to denote them by their old name (stockkeylabelold) or the new name (stockkeylabelnew). The **iStock.xlsx** contains a summary of all assessments with the appropriate stockkeys and stockkeylabels.

```{r, out.width = "100%"}
include_graphics("../pictures/istock.png")
```

_Figure 1: iStock.xlsx (on dropbox). Example focussing on mackerel._

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Generate iStock, iRename and iStockkey

# read iStock object from excel file
t <- 
  readxl::read_excel(path= paste(dropboxdir, "/data/iStock.xlsx", sep=""), 
                     col_names = TRUE, col_types = "text", trim_ws   = TRUE) %>%
  lowcase() %>% 
  mutate_at(c("assessmentmodel","assessmenttype","advicetype",
              "expertgroup","useofdiscardsinadvice","pabufferapplied", "assessmentstatus"),
            funs(tolower)) %>% 
  mutate_at(c("fmax","f01","fmed","f35spr","flim","fpa","fmsy","blim","bpa","msybtrigger"), 
            funs(as.numeric)) %>% 
  mutate_at(c("stockkey", "assessmentyear", "firstyearofdata", "ncpueseries","nsurveyseries"),
            funs(as.integer)) %>%

  mutate(assessmenttype   = ifelse(grepl("explo",assessmentmodel), "exploratory",assessmenttype),
         assessmenttype   = ifelse(grepl("trends",assessmentmodel), "trends",assessmenttype) ) %>%

  mutate(assessmentmodel  = ifelse(grepl("(xsa"   , assessmentmodel, fixed=TRUE), "xsa"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("sxsa"   , assessmentmodel, fixed=TRUE), "sxsa"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(flxsa" , assessmentmodel, fixed=TRUE), "xsa"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(ica"   , assessmentmodel, fixed=TRUE), "ica"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(flica" , assessmentmodel, fixed=TRUE), "ica"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(sam"   , assessmentmodel, fixed=TRUE), "sam"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(flsam" , assessmentmodel, fixed=TRUE), "sam"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(tsa"   , assessmentmodel, fixed=TRUE), "tsa"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(adapt" , assessmentmodel, fixed=TRUE), "adapt" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(ss3"   , assessmentmodel, fixed=TRUE), "ss3"   , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(stock synthesis 3", assessmentmodel, fixed=TRUE), 
                                   "ss3", assessmentmodel),
         assessmentmodel  = ifelse(grepl("(gadget" , assessmentmodel, fixed=TRUE), "gadget", assessmentmodel),
         assessmentmodel  = ifelse(grepl("(asap"  , assessmentmodel, fixed=TRUE), "asap"  , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(amish"  , assessmentmodel, fixed=TRUE), "amish" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(aspic"  , assessmentmodel, fixed=TRUE), "aspic" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(mycc"   , assessmentmodel, fixed=TRUE), "mycc"  , assessmentmodel),
         assessmentmodel  = ifelse(grepl("multi-year catch curve", 
                                         assessmentmodel, fixed=TRUE), "mycc"  , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(aspic"  , assessmentmodel, fixed=TRUE), "aspic" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("aarts"  , assessmentmodel, fixed=TRUE), 
                                   "aarts_poos" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(cbbm"  , assessmentmodel, fixed=TRUE), "cbbm" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(scaa"  , assessmentmodel, fixed=TRUE), "scaa" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(sms"  , assessmentmodel, fixed=TRUE), "sms" , assessmentmodel),
         assessmentmodel  = ifelse(grepl("(tasacs"  , assessmentmodel, fixed=TRUE), "tasacs" , assessmentmodel),
         # NEED TO FINALIZE THIS LIST !!

         assessmentdate   = as.Date(as.numeric(assessmentdate), origin="1899-12-30"),
         advicereleasedate= as.Date(as.numeric(advicereleasedate), origin="1899-12-30"),
         modifieddate     = as.Date(as.numeric(modifieddate), origin="1899-12-30")

         # assessmenttype2 = ifelse(assessmenttype %in% c("exploratory","trends","trends only"), 
         #                          "trends",assessmenttype)
  ) %>% 
  arrange(stockkey, assessmentyear, stockkeylabel)

# t %>% filter(fao =="cod") %>% distinct(assessmentmodel) %>% View()

iStock <-
  t %>% 
  select(stockkey, stockkeylabel, 
         assessmentyear, assessmenttype, assessmentmodel, assessmentdate, assessmentcomment,
         stockarea, expertgroup, advicedraftinggroup, 
         firstyearofdata, agerange, ncpueseries, nsurveyseries, 
         datacategory, yearoflastassessment, assessmentfrequency, yearofnextassessment, 
         advicereleasedate, advicetype, useofdiscardsinadvice, pabufferapplied, 
         assessmentstatus, assessmentscale, sectionnumber, assessmentkey, modifieddate,
         fmax, f01, fmed, f35spr, flim, fpa, fmsy, blim, bpa, msybtrigger, checkeddate, checkedby)

# iStockkey
iStockkey <-
  t %>% 
  group_by(stockkey) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(stockkey, stockkeylabelnew, stockkeylabelold)

# iRename
iRename <-
  t %>% 
  # group_by(stockkey, stockkeylabel, stockkeydescription, speciesfaocode=fao) %>% 
  group_by(stockkey, stockkeylabel) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(stockkeylabel, stockkey, stockkeydescription, speciesfaocode=fao)

# iSpecies
iSpecies <-
  readxl::read_excel(path= paste(dropboxdir, "/data/species_list.xlsx", sep=""), 
                     col_names = TRUE, col_types = "text", trim_ws   = TRUE) %>%
  mutate_at(c("speciescommonname","trophicguild","fisheriesguild","sizeguild"), 
            funs(tolower)) %>%
  group_by(speciesfaocode, speciesscientificname, speciescommonname) %>%
  arrange(speciesfaocode)

save(iStock     , file=paste(dropboxdir, "/rdata/iStock.RData", sep=""))
save(iStockkey  , file=paste(dropboxdir, "/rdata/iStockkey.RData", sep=""))
save(iRename    , file=paste(dropboxdir, "/rdata/iRename.RData", sep=""))
save(iSpecies   , file=paste(dropboxdir, "/rdata/iSpecies.RData",sep=""))

# iStock %>% filter(stockkeylabel == "cod-skag") %>% View()
# iStockkey %>% filter(stockkeylabelold == "cod-skag") %>% View()
# iStockkey %>% filter(stockkey == 100027) %>% View()

```

iStock object:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(filter(iStock,    stockkey == 169137, assessmentyear >= 2014), width=95)
```

iStockkey object:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(filter(iStockkey,    stockkey == 169137))
```

iRename object:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(filter(iRename,    stockkey == 169137))
```

iSpecies object:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(filter(iSpecies,    speciesfaocode == "mac"))
```

##### page break

### 2.3 Quality Control Sheets (QCS)

```{r, eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}

# make list of filenames
file.list <- c(list.files(path       = paste(dropboxdir, "/data/qcs-docs-docx/", sep=""), 
                          pattern    = "docx",
                          recursive  = TRUE, 
                          full.names = TRUE,   
                          ignore.case= TRUE) )

# read files from filename list
for (i in 1:length(file.list)) {

  # extract stock name
  stockkeylabel <- gsub(".qcs.docx","", file.list[i], fixed=TRUE)
  stockkeylabel <- gsub("((?:[^/]*/)*)(.*)","\\2", stockkeylabel)

  # print(paste0(i,stockkeylabel,file.list[i],sep=" - "))

  # set docx object
  docx <- read_docx(path=file.list[i])

  # read tables
  tmp <- docx_extract_all_tbls(docx, guess_header = FALSE, trim = TRUE)

  # convert tables to data frames
  for (j in 1:length(tmp)) {

    t       <- tmp[j] %>%  as.data.frame()
    variable <- tolower(t[1,1])
    yeartype <- tolower(t[2,2])
    nc      <- ncol(t)
    nr      <- nrow(t)
    head    <- t[3,2:nc]
    rows    <- t[4:nr,1]  %>% data.frame()
    names(rows) <- "assessmentyear"
    t <- 
      t[4:nr,2:nc] %>% 
      setNames(head) %>% 
      cbind(rows) %>% 
      gather(key=year, value=value, 1:(nc-1)) %>% 
      mutate(stockkeylabel          = tolower(stockkeylabel), 
             variable       = variable,
             yeartype       = yeartype, 
             value          = as.numeric(gsub("\\s+","",value)),
             assessmentyear = substr(as.character(assessmentyear), 1,4),
             assessmentyear = as.integer(assessmentyear)) %>% 
      filter(!is.na(value), value != "")

    if (j == 1) { data <- t
    } else      { data <- rbind(data,t) }

  } # end of j for loop

  if (i == 1) { qcsdata <- data
  } else      { qcsdata <- rbind(qcsdata,data) }
} #end of i for loop

# check and convert
t <-
  qcsdata %>% 
  mutate(
    year    = as.integer(year),
    stockkeylabel   = gsub(".qcs.docx","",stockkeylabel), 
    stockkeylabel   = gsub("anb-89","anb-8c9a", stockkeylabel),
    stockkeylabel   = gsub("anp-89","anp-8c9a", stockkeylabel),
    stockkeylabel   = gsub("cod-coast","cod-coas", stockkeylabel),
    stockkeylabel   = gsub("had-icel","had-iceg", stockkeylabel),
    stockkeylabel   = gsub("had-irisde3","had-iris", stockkeylabel),
    stockkeylabel   = gsub("her-2532excgor","her-2532-gor", stockkeylabel),
    stockkeylabel   = gsub("mac-wes","mac-nea", stockkeylabel),

    yeartype= gsub("\\s","",yeartype),

    variable = gsub("  "," ", variable), 
    var     = ifelse(grepl("average f"  , variable), "f"  , NA),
    var     = ifelse(grepl("spawning"   , variable), "ssb", var),
    var     = ifelse(grepl("recruitment", variable), "r"  , var),
    var     = ifelse(grepl("fishable"   , variable), "fb" , var),

    unit    = ifelse(var=="f"                                  , "year-1"                            , NA),

    # extract the bits between brackets
    unit    = ifelse(var=="ssb" & grepl("\\(.+\\)$"  , variable), gsub(".+\\((.+)\\)$","\\1", variable), unit),
    unit    = ifelse(var=="fb"  & grepl("\\(.+\\)$"  , variable), gsub(".+\\((.+)\\)$","\\1", variable), unit),
    unit    = ifelse(             grepl("unit: (.+)$", variable), gsub(".+unit: (.+)$","\\1", variable), unit),
    unit    = ifelse(var=="r" & stockkeylabel=="cod-2532"              , "thousands"                         , unit),  
    unit    = ifelse(var=="r" & stockkeylabel=="cod-iceg"              , "millions"                          , unit),  
    unit    = ifelse(var=="r" & stockkeylabel=="ple-celt"              , "thousands"                         , unit),  
    unit    = ifelse(var=="r" & stockkeylabel=="hke-soth"              , "check"                             , unit),

    unit    = gsub("[^[:alnum:]]", "", unit), 
    unit    = gsub("^000s$|^000$","thousands", unit),
    unit    = gsub("^000 000s$|000000s","millions" , unit),
    unit    = gsub("^000t$|^000stonnes$|^000tonnes$","thousand tonnes" , unit),
    unit    = gsub("000million$","billions" , unit),
    unit    = gsub("^t$", "tonnes",unit),
    unit    = ifelse(unit == "year1","year-1",unit),

    age     = ifelse(var=="f" & grepl("\\(.+\\)$", variable), gsub(".+\\((.+)\\)$","\\1", variable), NA),
    age     = gsub(" +", "", age), 
    age     = gsub(",,u", "", age), 
    age     = gsub(",u", "", age), 
    age     = gsub(",w", "", age), 
    age     = ifelse(var=="f" & stockkeylabel=="anb-78ab" & assessmentyear <= 2000, "4-8" , age),
    age     = ifelse(var=="f" & stockkeylabel=="anb-78ab" & assessmentyear  > 2000, "6-10", age),
    age     = ifelse(var=="f" & stockkeylabel=="anp-78ab" & assessmentyear <= 2000, "3-7", age),
    age     = ifelse(var=="f" & stockkeylabel=="anp-78ab" & assessmentyear  > 2000, "3-8", age),
    age     = ifelse(var=="f" & stockkeylabel=="cod-347d" & assessmentyear <= 2002, "2-8", age),
    age     = ifelse(var=="f" & stockkeylabel=="cod-347d" & assessmentyear  > 2002, "2-4", age),
    age     = ifelse(var=="f" & stockkeylabel=="her-irls"                          ,"2-7" , age ),

    # extract recruitment age information between brackets
    age     = ifelse(var=="r" & grepl(".+\\(age (.)\\).+", variable), gsub(".+\\(age (.)\\).+","\\1", variable), age),
    age     = ifelse(var=="r" & grepl(".+\\(age (.) \\).+", variable), gsub(".+\\(age (.) \\).+","\\1", variable), age),
    age     = ifelse(var=="r" & stockkeylabel=="anb-78ab" & assessmentyear <= 2000, "1", age),
    age     = ifelse(var=="r" & stockkeylabel=="anb-78ab" & assessmentyear  > 2000, "2", age),
    age     = ifelse(var=="r" & stockkeylabel=="anp-78ab" & assessmentyear <= 2000, "0", age),
    age     = ifelse(var=="r" & stockkeylabel=="anp-78ab" & assessmentyear  > 2000, "1", age),
    age     = ifelse(var=="r" & stockkeylabel=="cod-2532", "2"    , age),  # looked up in ACFM report
    age     = ifelse(var=="r" & stockkeylabel=="her-irls", "1"    , age),  # looked up in ACFM report
    age     = ifelse(var=="r" & stockkeylabel=="mgw-78"  , "1"    , age),  # looked up in ACFM report
    age     = ifelse(var=="r" & stockkeylabel=="ple-celt", "1"    , age),  # looked up in ACFM report
    age     = ifelse(var=="r" & stockkeylabel=="whg-47d" , "check", age)
  )

setf   <- t %>% filter(var=="f")   %>% 
  select(stockkeylabel, assessmentyear, year, fishingpressuredescription=var, f=value, fishingpressureunits=unit, fage=age)

setr   <- t %>% filter(var=="r")   %>% 
  select(stockkeylabel, assessmentyear, year, recruitmentdescription=var, r=value, recruitmentunits=unit, 
         recruitmentage=age, yeartype) %>% 
  # correct for yearclass reporting of recruitment
  mutate(year    = ifelse(yeartype=="yearclass", year + as.integer(recruitmentage), year)) %>% 
  arrange(stockkeylabel, assessmentyear, year)

setssb <- t %>% filter(var=="ssb") %>% 
  select(stockkeylabel, assessmentyear, year, stocksizedescription=var, ssb=value, stocksizeunits=unit) 
# setfb  <- t %>% filter(var=="fb")  %>% 
#   select(stockkeylabel, assessmentyear, year, stocksizedescription=var, fb=value, stocksizeunits=unit, stocksizeage=age)

qcsdata <-

  select(setr, 
         stockkeylabel, assessmentyear, year, recruitment=r, 
         recruitmentunits, recruitmentage, recruitmentdescription) %>%

  full_join(select(setssb, 
                   stockkeylabel, assessmentyear, year, ssb, stocksizeunits, stocksizedescription), 
            by=c("stockkeylabel","assessmentyear","year")) %>%

  full_join(select(setf, 
                   stockkeylabel, assessmentyear, year, 
                   f, fishingpressureunits, fage, fishingpressuredescription) , 
            by=c("stockkeylabel","assessmentyear","year")) %>%

  arrange(stockkeylabel, assessmentyear, year) %>%

  mutate(ssb            = ifelse(stocksizeunits == "thousand tonnes", ssb * 1000, ssb),
         stocksizeunits = ifelse(stocksizeunits == "thousand tonnes", "tonnes", stocksizeunits)) %>%

  left_join(iRename, by=c("stockkeylabel")) %>%

  mutate(assessmenttype  = "assess",
         source          = "qcs")  %>%

  # add the assessment date
  left_join(select(iStock, 
                   stockkeylabel, assessmentyear, assessmentdate), 
            by = c("stockkeylabel","assessmentyear"))

# qcsdata %>%   filter(stockkeylabel == "cod-kat") %>% View()
# iRename %>%   filter(stockkeylabel == "cod-kat") %>% View()
# glimpse(iRename)
# glimpse(qcsdata)

# save dataset
save(qcsdata, file=paste(dropboxdir, "/rdata/qcsdata.RData",sep=""))

```

The first type of datasets are the old Quality Control Sheets as they have been in use up to around 2004. The Quality Control Sheets are contained in Word documents with three separate tables: one for fishing mortality, one for recruitment and one for SSB. Each of the tables is a matrix with year on the columns and assessmentyear on the rows. These type of tables, typically have a number of footnotes associated with them. For the purpose of reading in the data, all footnotes identifiers (i.e. the numbers in the tables) have been removed automatically in order to avoid having extra numbers being added to the values in the table. If not, a cell containing for example the value 100 and the footnote "5" would have been read as 1005. The original tables with the footnotes have been kept as a backup, so that more detailed transcription of the data is still possible.

```{r, out.width = "100%"}
include_graphics("../pictures/qcs.png")
```
_Figure 2: Example of Quality Control Sheet of Anglerfish L. budegassa (fishing mortality) with year in the columns and assessmentyear on the rows. Note the descriptions of the changes in ages over which average F is calculated (taken up in the code) and the comments on the bottom of the table (not yet taken up in the code)._

The variables used in the qcsdata object are shown in the text table below. There are `r length(unique(qcsdata$stockkeylabel))` stocks contained in the data object. The first assessment year is `r min(unique(qcsdata$assessmentyear))` and the last assessmentyear is `r max(unique(qcsdata$assessmentyear))`.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
glimpse(qcsdata, width=95)
```

The stocks involved are:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print((unique(qcsdata$stockkeylabel)), quote=FALSE, row.names=FALSE, width=95)
```

##### page break

### 2.4 Excel database of 'Standard Graphs'

```{r, echo=FALSE, message=FALSE, warning=FALSE}

exceldata <-
  readxl::read_excel(paste(dropboxdir, "/data/ICES Assessment Summary database.xlsx",sep=""),
             sheet = "DATA", 
             col_names = TRUE, 
             col_types = "text", 
             skip = 0) %>%
  lowcase %>%
  rename(stockkeylabel = fishstock) %>% 
  mutate(stockkeylabel = tolower(stockkeylabel),
         assessmentdate = as.Date(as.numeric(assessmentdate), origin="1899-12-30")) %>%
  mutate_at(vars("year","assessmentyear"), funs(as.integer)) %>% 
  mutate_at(vars("lowrecruitment", "recruitment","highrecruitment",
                 "lowssb","ssb","highssb",
                 "lowf", "f","highf",
                 "landings","catches","discards","ibc"), 
            funs(as.numeric)) %>%
  mutate_at(vars("stocksizedescription","stocksizeunits",
                 "fishingpressuredescription", "fishingpressureunits"), 
            funs(tolower)) %>% 
  # mutate(assessmenttype2 = ifelse(assessmentyear >= 2011 & assessmenttype2 == "assess", 
  #                                "update",assessmenttype2)) %>% 
  ungroup() %>%

  # select only the relevant fields (could be more though !!)
  select(stockkeylabel, assessmentyear, assessmenttype, year, assessmentdate,
         lowrecruitment, recruitment, highrecruitment, 
         lowssb, ssb, highssb, 
         lowf, f, highf, 
         landings, catches, discards, ibc, 
         stocksizeunits, stocksizedescription, fishingpressuredescription, fishingpressureunits,
         fage) %>%

  # dealing with old and new stocknames
  left_join(iRename, by = c("stockkeylabel")) %>% 
  filter(year <= assessmentyear) %>%

  mutate(source = "excel",
         units = "") %>% 
  arrange(stockkeylabel, assessmentyear, year)

# save dataset
save(exceldata, file=paste(dropboxdir, "/rdata/exceldata.RData",sep=""))
```

The origin of the Excel database started with a private collection of stock assessment results by Martin Pastoors, driven by an interest in the historical performance of stock assessments (e.g. Van Beek, F. A. and M. A. Pastoors (1999), Reeves, S. A. and M. A. Pastoors (2007)). The assessment results were taken from the annual assessments that were entered into the standard graph database, and initially the only additional effort was to add the assessmentyear to the variables already in the database. This private collection was later taken up by the ICES secretariat (Barbara Schoute) where it was used to populate the historial retrospective plots that were included in the advice documents from 2006 onwards.

Because the standard graph database in ICES was expanding (e.g. including more variables), the excel database was also expanded so that the data could be easily transferred from the standard graphs to the excel database.

The latest development has been that not only the assessments were included in the excel database but also the withdrawn assessments (e.g. because of errors detected) and the benchmark assessments (only to a very limited degree so far). To make it possible to distinguish between a final assessment and other assessments, the variable assessmenttype2 has been introduced with the following possible values: assess, update, benchmark, withdrawn, trends only. The number of benchmarks in the database is still very limited.

The variables used in the exceldata object are shown in the text table below. There are `r length(unique(exceldata$stockkeylabel))` stocks contained in the data object. The first assessment year is `r min(unique(exceldata$assessmentyear))` and the last assessmentyear is `r max(unique(exceldata$assessmentyear))`.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
glimpse(exceldata, width=95)
```

The stocks in the excel database are:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print((unique(exceldata$stockkeylabel)), quote=FALSE, row.names=FALSE, width=95)
```

##### page break

### 2.5 ICES Stock Assessment Graphs (SAG) webservices

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# get SAG from ICES webservices

# getSAG        <- icesSAG::getSAG(stock=NULL, year=0, data="summary", combine=TRUE)
# save(getSAG       , file=paste(dropboxdir, "/rdata/getSAG.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/getSAG.RData",sep=""))

# process standard graph data SAG and prepare for combining

sagdata <- 
  getSAG %>% 
  lowcase() %>% 
  dplyr:: select(stockkeylabel=fishstock, assessmentyear, year,
                 recruitment, highrecruitment, lowrecruitment,
                 ssb, highssb, lowssb,
                 f, highf, lowf, 
                 catches, landings, discards,
                 fage, 
                 units, 
                 stocksizedescription, stocksizeunits,
                 fishingpressuredescription, fishingpressureunits) %>%

  mutate(assessmenttype  = "assess",
         assessmentdate = as.Date(NA)) %>%

  # dealing with old and new stocknames
  left_join(iRename, by = c("stockkeylabel")) %>%

  # mutate tolower
  mutate_at(vars("stocksizedescription","stocksizeunits","fishingpressuredescription","fishingpressureunits"), 
            funs(tolower)) %>%

  mutate(source = "sag",
         ibc    = NA )%>% 
  ungroup() %>% 
  data.frame() %>%

  # add the assessment date
  left_join(select(iStock, 
                   stockkeylabel, assessmentyear, assessmentdate), 
            by = c("stockkeylabel","assessmentyear"))

# sagdata %>%   filter(stockkeylabel == "cod-kat") %>% View()

# save dataset
save(sagdata, file=paste(dropboxdir, "/rdata/sagdata.RData",sep=""))

```

The third element of the iAssess database are the assessments that are being collected now as part of the ICES Stock Assessment Graph Database and available through the ICES webservices. Although there are multiple elements to these databases, at this stage only the function getSAG has been used to obtain the assessment results. The function getSD (get Stock Database) has been used but only to convert it into the iStock.xlsx so that certain variables could be manually manipulated (e.g. the stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew and stockarea).

The variables used in the SAGdata object are shown in the text table below. There are `r length(unique(sagdata$stockkey))` stocks contained in the data object. The first assessment year is `r min(unique(sagdata$assessmentyear))` and the last assessmentyear is `r max(unique(sagdata$assessmentyear))`.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
glimpse(sagdata, width=95)
```

The stocks in the SAG database are:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
t <-
  sagdata %>% 
  distinct(stockkey) %>% 
  left_join(iStockkey, by="stockkey") %>% 
  arrange(stockkeylabelnew) %>% 
  mutate(stock = paste(stockkeylabelold, stockkeylabelnew,sep="/")) %>% 
  select(stock) %>% 
  filter(stock != "NA/NA") %>% 
  mutate(stock=gsub("/NA","", stock, fixed=TRUE))

print(sort(as.character(t(t))), quote=FALSE, col.names=FALSE, row.names=FALSE, width=95)
```

##### page break

## 3. Combining data

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# -----------------------------------------------------------------------------------------
# iAssess_part1: standard graph data SAG and prepare for combining
# -----------------------------------------------------------------------------------------
iAssess_part1 <- 
  sagdata %>%

  # temporary: removing her.27.1-24a514a 2017 because of error in the export of multiple versions. 
  filter(!(stockkeylabel == "her.27.1-24a514a" & assessmentyear == 2017))

# Extract list of fishstock and assessment years from SAG database
iAssess_part1_unique <-
  iAssess_part1 %>% 
  group_by(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  ungroup()

# -----------------------------------------------------------------------------------------
# iAssess_part2: Old excel assessment database and prepare for combining
# -----------------------------------------------------------------------------------------
iAssess_part2 <-
  exceldata

# Extract list of fishstock and assessment years from old database
iAssess_part2_unique <-
  iAssess_part2 %>% 
  group_by(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  ungroup()

# Select from iAssess_part2 the assessments to be added to iAssess_part1
iAssess_part2_toadd <-
  setdiff(select(iAssess_part2_unique, stockkey, assessmentyear, assessmenttype),
          select(iAssess_part1_unique, stockkey, assessmentyear, assessmenttype)) %>% 
  left_join(iAssess_part2, by=c("stockkey","assessmentyear", "assessmenttype")) %>% 
  data.frame()

# -----------------------------------------------------------------------------------------
# iAssess_part3: dataset from old Quality Control Sheets
# -----------------------------------------------------------------------------------------

iAssess_part3 <-
  qcsdata

# Extract list of fishstock and assessment years from old QCS database
iAssess_part3_unique <-
  iAssess_part3 %>% 
  group_by(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  filter(row_number()==1) %>% 
  select(stockkey, assessmentyear, stockkeylabel, assessmenttype) %>% 
  ungroup()

# Select from iAssess_part3 the assessments to be added to iAssess_part1
iAssess_part3_toadd <-
  setdiff(select(iAssess_part3_unique, stockkey, assessmentyear, assessmenttype),
          select(iAssess_part1_unique, stockkey, assessmentyear, assessmenttype)) %>% 
  left_join(iAssess_part3, by=c("stockkey","assessmentyear", "assessmenttype")) %>% 
  data.frame()

# -----------------------------------------------------------------------------------------
# Generate iAssess and do cleaning up
# -----------------------------------------------------------------------------------------

iAssess <- 
  bind_rows(iAssess_part1, iAssess_part2_toadd) %>%  
  bind_rows(., iAssess_part3_toadd) %>%

  # corrections to stock size descriptions
  mutate(
    stocksizedescription = gsub("stock size: "                 ,""                   , stocksizedescription),
    stocksizedescription = gsub("indices"                      ,"index"              , stocksizedescription),
    stocksizedescription = gsub("indicator"                    ,"index"              , stocksizedescription),
    stocksizedescription = gsub("^biomass$"                    ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^biomass index$"              ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^evhoe biomass index$"        ,"total biomass index", stocksizedescription),
    stocksizedescription = gsub("^stock size index: abundance$","abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^abundance$"                  ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^density$"                    ,"density index"      , stocksizedescription),
    stocksizedescription = gsub("^tsb$"                        ,"total biomass"      , stocksizedescription),
    stocksizedescription = gsub("^total abundance index$"      ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^index$"                      ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^stock abundance$"            ,"abundance index"    , stocksizedescription),
    stocksizedescription = gsub("^density index$"              ,"abundance index"    , stocksizedescription),

    stocksizedescription = gsub("stock size index: biomass \\(ages 1-8\\)"   ,"total biomass index"      , stocksizedescription),
    stocksizedescription = gsub("stock size index: german survey"            ,"total biomass index"      , stocksizedescription),
    stocksizedescription = gsub("stock size index: smoothed greenland index" ,"total biomass index"      , stocksizedescription),

    stocksizedescription = ifelse(grepl("tv",stocksizedescription), "abundance index", stocksizedescription),
    stocksizedescription = ifelse(grepl("ssb & b",stocksizedescription) , "ssb", stocksizedescription),
    stocksizedescription = ifelse(grepl("total biomass/bmsy",stocksizedescription) , "b/bmsy", stocksizedescription),
    stocksizedescription = ifelse(stocksizedescription == "stock size" & stocksizeunits == "tonnes", "ssb", stocksizedescription),
    stocksizedescription = ifelse(stocksizedescription == "stock size" & grepl("kg/",stocksizeunits), "total biomass index", stocksizedescription),
    stocksizedescription = ifelse(grepl("relative", stocksizeunits, fixed=TRUE) & is.na(stocksizedescription), "total biomass index", stocksizedescription)
  ) %>%

  # corrections to stock units
  mutate(
    stocksizeunits = gsub("stock size: "                 ,""                   , stocksizeunits),
    stocksizeunits = gsub(" ", "", stocksizeunits),
    stocksizeunits = gsub("density(burrows/m2)", "burrows/m2", stocksizeunits, fixed=TRUE),
    stocksizeunits = gsub("cpue(kg/1000hooks)", "kg/1000hooks", stocksizeunits, fixed=TRUE),
    stocksizeunits = gsub("^abundance$", "millions", stocksizeunits),
    stocksizeunits = gsub("na(ratio)", "relative", stocksizeunits, fixed=TRUE),
    stocksizeunits = ifelse(grepl("kg/h", stocksizeunits, fixed=TRUE), "kg/hour", stocksizeunits),
    stocksizeunits = ifelse(grepl("n/h", stocksizeunits, fixed=TRUE), "n/hour", stocksizeunits)
  ) %>%

  # corrections to fishing pressure descriptions
  mutate(
    fishingpressuredescription = gsub("fishing pressure: ",""  , fishingpressuredescription),
    fishingpressuredescription = gsub(" ",""  , fishingpressuredescription),
    fishingpressuredescription = gsub("f&hr","f"  , fishingpressuredescription, fixed=TRUE),
    fishingpressuredescription = gsub("fishingpressure","f"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("finwinterrings","f"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("weightedf","f"  , fishingpressuredescription),

    fishingpressuredescription = gsub("harvestrate","hr"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("relativehr","hr/index"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("hrindex","hr/index"  , fishingpressuredescription), 
    fishingpressuredescription = gsub("relativeexploitationrate","hr/index"  , fishingpressuredescription),

    fishingpressuredescription = ifelse(grepl("ages",fishingpressuredescription), "f", fishingpressuredescription),
    fishingpressuredescription = ifelse(grepl("null",fishingpressuredescription), NA, fishingpressuredescription),

    fishingpressureunits       = ifelse(grepl("relative",fishingpressuredescription) & is.na(fishingpressureunits), "relative", fishingpressureunits ),
    fishingpressuredescription = ifelse(grepl("relative",fishingpressuredescription) , "fproxy", fishingpressuredescription )
  ) %>%

  # corrections to fishing pressure units
  mutate(
    fishingpressureunits = gsub(" ",""  , fishingpressureunits),
    fishingpressureunits = gsub("peryear","year-1"  , fishingpressureunits),
    fishingpressureunits = gsub("%","percent"  , fishingpressureunits),
    fishingpressureunits = gsub("^f$","year-1"  , fishingpressureunits),
    fishingpressureunits = gsub("^catch/biomass$","relative"  , fishingpressureunits),

    fishingpressureunits = ifelse(grepl("cm",fishingpressureunits), "year-1",fishingpressureunits) ,
    fishingpressureunits = ifelse(grepl("null",fishingpressureunits), NA,fishingpressureunits) ,
    fishingpressureunits = ifelse(grepl("ratio",fishingpressureunits), "relative",fishingpressureunits) 
  ) %>%

  # correction due to missing units
  mutate(
    stocksizeunits       = ifelse(stocksizedescription=="b/bmsy" & is.na(stocksizeunits),"relative",stocksizeunits),  
    fishingpressureunits = ifelse(fishingpressuredescription=="f/fmsy" & is.na(fishingpressureunits),"relative",fishingpressureunits)
  ) %>%

  # corrections to the assignments of specific stocks and years
  mutate(
    stocksizedescription       = ifelse(stockkeylabel=="anb-8c9a" & assessmentyear==2013,"b/bmsy"  ,stocksizedescription),
    stocksizeunits             = ifelse(stockkeylabel=="anb-8c9a" & assessmentyear==2013,"relative",stocksizeunits),
    fishingpressuredescription = ifelse(stockkeylabel=="anb-8c9a" & assessmentyear==2013,"f/fmsy"  ,fishingpressuredescription),
    fishingpressureunits       = ifelse(stockkeylabel=="anb-8c9a" & assessmentyear==2013,"relative",fishingpressureunits)
  ) %>%

  # remove double series (e.g. mac-nea 2013 is twice in the sag download)
  group_by(stockkeylabel, assessmentyear, year, assessmenttype) %>% 
  filter(row_number() == 1) %>%

  # add old and new names
  left_join(iStockkey, by="stockkey") %>%

  # add species info
  left_join(iSpecies, by="speciesfaocode") %>%

  # add stock info
  # left_join(iStock, by=c("stockkey","stockkeylabel","assessmentyear","assessmenttype","assessmentdate")) %>%

  # convert to lowercase
  ungroup() %>% 
  as.data.frame()

save(iAssess, file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
```

We merge the three data sources by comparing all the combinations of **stockkey**, **assessmentyear**, **stockkeylabel** and **assessmenttype** in the different datasources. First all unique combinations that are in the _exceldata_ object but not in the _sagdata_ object, are added to the _sagdata_ object. Then all unique combinations that are in the _qcsdata_ object but not in the _exceldata_ object, are added to the _sagdata_ object. This combined object is called _iAssess_. Several manipulations are carried out to clean up the merged data, e.g. make consistent use of terminology for stocksizedescription, fishingpressuredescriptions etc.

The variables used in the _iAssess_ object are shown in the text table below. There are `r length(unique(iAssess$stockkey))` stocks contained in the data object. The first assessment year is `r min(unique(iAssess$assessmentyear))` and the last assessmentyear is `r max(unique(iAssess$assessmentyear))`.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
glimpse(iAssess, width=95)
```

##### page break

## 4. Quality control

##### page break

## 5. Data overviews

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.asp=1.25}

# data handling and plot by data source
x <-
  iAssess %>% 
  filter(assessmentyear >= 1987) %>% 
  filter(!is.na(stockkey)) %>% 
  group_by(stockkeylabelold, stockkeylabelnew, assessmentyear, assessmenttype, source) %>% 
  filter(row_number() ==1) %>% 
  select(-year) %>% 
  ungroup() %>% 
  arrange(stockkeylabelold) %>% 
  mutate(id = group_indices(., stockkeylabelold)) %>% 
  data.frame() %>% 
  mutate(stockkeylabelold = ifelse(is.na(stockkeylabelold), stockkeylabel, stockkeylabelold)) %>% 
  mutate(stockkeylabelold = factor(stockkeylabelold), 
         stockkeylabelold = factor(stockkeylabelold, levels = rev(levels(stockkeylabelold))),
         source           = factor(source, levels=c("qcs","excel","sag")),
         assessmenttype  = factor(assessmenttype),
         assessmentdate    = ifelse(is.na(assessmentdate), 
                                   make_date(year = assessmentyear, month = 6L, day = 30L),
                                   assessmentdate),
         assessmentdate    = as.Date(assessmentdate, origin="1970-01-01"),
         col              = ifelse(id <=  90            , 1, NA),
         col              = ifelse(id  >  90 & id <= 180, 2, col),
         col              = ifelse(id  > 180            , 3, col))

# define colour scale
myColors        <- brewer.pal(length(levels(x$source)),"Set1")
names(myColors) <- levels(x$source)

# define headers for columns
y <-
  x %>% 
  group_by(col) %>% 
  filter(row_number()==1| row_number() == n()) %>% 
  ungroup() %>% 
  mutate(id = group_indices(., col)) %>% 
  select(col, id, stockkeylabelold) %>% 
  group_by(col) %>% 
  summarise(code = paste(stockkeylabelold, collapse=" : "))

# plot by stock and data source
x %>% 
  left_join(y, by="col") %>% 
  ggplot(aes(x=assessmentdate, y=stockkeylabelold)) +
  theme_publication() +
  theme(panel.spacing = unit(1, "lines"),
        text          = element_text(size=8),
        legend.title  = element_blank(),
        axis.text.x   = element_text(angle = 90, vjust = 1)) +
  geom_point(aes(colour = source)) +
  scale_colour_manual(name = "source", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  scale_x_date(date_breaks = "1 year", date_labels = "%y") +
  labs(x = "assessmentyear", y = NULL ) +
  facet_wrap(~code, scales="free_y", shrink=TRUE, ncol=3)

```

_Figure 3. Overview of data sources for the different stocks (stockkeylabelold)._

##### page break

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.asp=1.25}

# plot by assessmenttype

myColors        <- brewer.pal(length(levels(x$assessmenttype)),"Set1")
names(myColors) <- levels(x$assessmenttype)

x %>% 
  left_join(y, by="col") %>% 
  ggplot(aes(x=assessmentdate, y=stockkeylabelold)) +
  theme_publication() +
  theme(panel.spacing = unit(1, "lines"),
        text          = element_text(size=8),
        legend.title  = element_blank(),
        axis.text.x   = element_text(angle = 90, vjust = 1)) +
  geom_point(aes(colour = assessmenttype)) +
  scale_colour_manual(name = "assessmenttype", values = myColors, na.value="lightgray") +
  scale_y_discrete(position="right") +
  scale_x_date(date_breaks = "1 year", date_labels = "%y") +
  labs(x = "assessmentyear", y = NULL ) +
  facet_wrap(~code, scales="free_y", shrink=TRUE, ncol=3)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.asp=1.25}

# table by assessmentyear and stock

iStock %>% 
  left_join(iStockkey, by=c("stockkey")) %>% 
  select(stockkeylabelold, assessmentyear, assessmenttype) %>% 
  mutate(tmp = "x") %>% 
  data.frame() %>% 
  # filter(stockkeylabelold =="cod-skag")

  dcast(stockkeylabelold ~ assessmentyear, margins=NA) %>% 
  write.csv(., file="D:/temp/istock.csv")

```

_Figure 4. Overview of assessment types for the different stocks (stockkeylabelold)._

## 6. Software development

## 7. Discussion

## 8. Future developments

## To do

* Move the data and r code to github

```{r, echo=FALSE, message=FALSE, warning=FALSE}

```

### 2018-11-11 R/DownloadDataFromSAG.r

@@ -1,0 +1,59 @@
+# ============================================================================
+# Download Assessment Data from SAG
+#
+# 03/07/2018 First coding of downloading data from SAG; adapted from addHistoricaldata...
+# 12/08/2018 Included raw downloads (all variables)
+# 17/08/2018 Update icesSAG@development because of errors in previous download
+# 23/10/2018 Now only download from SAG; do manipulations in iDatabases generator
+# ============================================================================
+
+# rm(list=ls())
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSAG")
+# devtools::install_github("ices-tools-prod/icesSD")
+# install_github("ices-tools-prod/icesSAG@development")
+
+library(icesSAG)
+library(tidyverse)
+library(stringr)       # for string manipulation
+
+# use token
+options(icesSAG.use_token = TRUE)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# =====================================================================================
+# Download all assessments in SAG in full
+# =====================================================================================
+
+assessmentkeys <- sort(findAssessmentKey())
+# getStockDownloadData(7306)
+
+# Download all the stock data - takes a long time
+t1  <- getStockDownloadData(assessmentkeys)
+
+t2 <- data.frame()
+for (i in 1:length(t1)) {
+  
+  print(i)
+  if (!is.null(t1[[i]])) {
+    tmp <-
+      t1[[i]] %>% 
+      mutate_all(funs("as.character"))
+    
+    t2  <- bind_rows(t2, tmp)
+  }
+}
+
+iSAGdownload <- 
+  t2 %>% 
+  lowcase()
+
+today <- format(Sys.time(), '%Y%m%d')
+save(iSAGdownload, file=paste0(assessdir, "/rdata/iSAGdownload ", today, ".RData"))
+

### 2018-11-11 R/HAWG data in SAG.r

@@ -1,0 +1,106 @@
+### ============================================================================
+### Add Historical Assessment Data to SAG
+###
+### 21/03/2018 First coding of adding historical data to standardgraphs
+### ============================================================================
+
+rm(list=ls())
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSAG")
+
+library(icesSAG)
+library(icesSD)
+library(tidyverse)
+library(directlabels)  # for printing labels at end of geom lines
+
+# use token
+options(icesSAG.use_token = TRUE)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# =====================================================================================
+# Download ICES Stock database
+# =====================================================================================
+
+# icesSD         <- icesSD::getSD()
+# save(icesSD, file=paste(dropboxdir, "/rdata/icesSD.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/icesSD.RData",sep=""))
+
+# get all the sandeel stocks
+filter(icesSD, substr(StockKeyLabel,1,3) == "san") %>% 
+  distinct(StockKeyLabel)
+
+# =====================================================================================
+# Download all the assessments in SAG; also unpublished ones
+# =====================================================================================
+
+# set all the hawgstocks
+hawgstocks <- c("her.27.3a47d", "her.27.nirs","her.27.irls","her.27.20-24", "her.27.6a7bc",
+                "spr.27.4","spr.27.3a","spr.27.7de",
+                "san.sa.1r","san.sa.2r","san.sa.3r","san.sa.4", "san.sa.5r", "san.sa.6", "san.sa.7r")
+
+# download all the stocks in HAWG 2018
+hawgdata <- 
+  getSAG(hawgstocks, year = 2018) %>% 
+  filter(Purpose == "Advice")
+
+save(hawgdata, file="hawgdata.RData")
+
+# Catch or landings
+hawgdata %>% 
[…]

### 2018-11-11 R/addHistoricalDataToSAG.r

@@ -1,0 +1,244 @@
+### ============================================================================
+### Add Historical Assessment Data to SAG
+###
+### 21/03/2018 First coding of adding historical data to standardgraphs
+### ============================================================================
+
+rm(list=ls())
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSAG")
+
+library(icesSAG)
+library(icesSD)
+library(tidyverse)
+library(directlabels)  # for printing labels at end of geom lines
+
+# use token
+options(icesSAG.use_token = TRUE)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# =====================================================================================
+# Download ICES Stock database
+# =====================================================================================
+
+icesSD         <- icesSD::getSD()
+save(icesSD, file=paste(dropboxdir, "/rdata/icesSD.RData",sep=""))
+
+# =====================================================================================
+# Download all the assessments in SAG; also unpublished ones
+# =====================================================================================
+
+
+icesSAG  <-
+  getSAG(stock=NULL, year = 0, data="summary", combine=TRUE) %>% 
+  lowcase()
+  # filter(Purpose == "Advice")
+save(icesSAG, file=file.path(dropboxdir, "rdata/icesSAG 20180427.RData"))
+
+sort(unique(icesSAG$Purpose))
+
+# create overview of assessments
+icesSAG %>% 
+  group_by(fishstock, assessmentyear, purpose) %>% 
+  filter(row_number() == 1) %>% 
+  group_by(assessmentyear) %>% 
+  summarize(n = n()) %>% 
+  View()
+
+# check the historical data
+icesSAG %>% 
+  filter(assessmentyear == 1989) %>% 
+  View()
+
+# plot specific assessment
+icesSAG %>% 
+  filter(year >= 1980) %>%
+  filter(fishstock %in% c("cod-347d")) %>%
+  mutate(decade            = 10*floor(assessmentyear/10),
+         tyear             = substr(as.character(assessmentyear),3,4)) %>% 
+  
[…]

### 2018-11-11 R/changeHistoricalDataInSAG.r

@@ -1,0 +1,81 @@
+# ============================================================================
+# Change Historical Assessment Data in SAG
+#
+# 21/03/2018 First coding of adding historical data to standardgraphs
+# 27/04/2018 Change to Historical in the Purpose Field
+# ============================================================================
+
+rm(list=ls())
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSAG")
+
+library(icesSAG)
+library(icesSD)
+library(tidyverse)
+library(directlabels)  # for printing labels at end of geom lines
+
+# use token
+options(icesSAG.use_token = TRUE)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# Download and mutate the data for stock and year
+rby <- 
+  getSAG("her-47d3",year = 2002) %>% 
+  mutate(landings = ifelse(landings == 0, NA, landings))
+  
+
+# Create the input data for uploading  
+info     <- stockInfo(StockCode      = unique(rby$fishstock), 
+                      AssessmentYear = unique(rby$AssessmentYear), 
+                      ContactPerson  = "martin.pastoors@gmail.com")
+
+info$StockCategory             <- NA
+info$MSYBtrigger               <- NA
+info$Blim                      <- 800000
+info$Bpa                       <- 1300000
+info$Flim                      <- NA
+info$Fpa                       <- 0.25 
+info$FMSY                      <- NA
+info$Fage                      <- "2-6"
+info$RecruitmentAge            <- 0
+info$CatchesCatchesUnits       <- unique(rby$units[!is.na(rby$units)])
+info$RecruitmentDescription    <- "wr"
+info$RecruitmentUnits          <- "thousands"
+info$FishingPressureDescription<- "F"
+info$FishingPressureUnits      <- "Year-1"
+info$StockSizeDescription      <- "SSB"
+info$StockSizeUnits            <- unique(rby$stockSizeUnits[!is.na(rby$stockSizeUnits)])
+info$Purpose                   <- "Advice"
+# info$CustomLimitName1         <- "MBAL"
+# info$CustomLimitValue1        <- 800000
+
+fishdata <- stockFishdata(min(rby$Year):max(rby$Year))
+
+fishdata$Landings              <- rby$landings
+fishdata$Catches               <- rby$catches
[…]

### 2018-11-11 R/compare databases istock advice assess.r

@@ -1,0 +1,282 @@
+# -----------------------------------------------------------------------------------------------
+# compare databases istock advice assessment
+#
+# 12/08/2018 first coding
+# 07/11/2018 updated coding after revisions to iAdvice and QCSExcel files
+# -----------------------------------------------------------------------------------------------
+
+rm(list=ls())
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(stringr)   # string manipulation
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+advicedir <- paste(get_dropbox(), "/ICES Advice database", sep="")
+
+# load datasets: sag, sagful, qcsexcel, istock, iadvice
+sag <- 
+  get(load(file=file.path(assessdir, "rdata/icesSAG 20180812.RData")))
+
+sagfull <- 
+  get(load(file=file.path(assessdir, "rdata/icesSAGfull 20181023.RData")))
+
+sd <-
+  get(load(file=file.path(assessdir, "rdata/icesSD 20181023.RData")))
+
+qcsexcel <-
+  readxl::read_excel(
+    path= paste(assessdir, "/excel/QCS and EXCEL Assessment Database combined.xlsx", sep=""), 
+    sheet = "data",
+    col_names = TRUE, 
+    col_types = "text", 
+    trim_ws   = FALSE) %>%
+  lowcase() 
+
+
+istock <-
+  readxl::read_excel(
+    path= paste(assessdir, "/excel/iStock.xlsx", sep=""), 
+    sheet     = "istock",
+    col_names = TRUE, 
+    col_types = "text", 
+    trim_ws   = FALSE) %>%
+  lowcase() 
+
+advice <-
+  readxl::read_excel(
+    path= paste(advicedir, "/excel/ICES Scientific Advice database 20181101.xlsx", sep=""), 
+    sheet     = "DATA",
+    col_names = TRUE, 
+    col_types = "text", 
+    trim_ws   = FALSE) %>%
+  lowcase() 
+
+# --------------------------------------------------------------------------------
+# summarize datasets (variables)
[…]

### 2018-11-11 R/dev/Convert Sframe to management areas.r

@@ -1,0 +1,106 @@
+# --------------------------------------------------------------------------------------------
+# convert STECF Sframe to management areas.r
+#
+# 28/04/2018 cleaned up the code
+# --------------------------------------------------------------------------------------------
+
+# devtools::install_github("fishvice/tidyices", dependencies = FALSE)
+
+# Note: geom_sf is only on the development version of ggplot2. Make sure to start rstudio in administrator mode before starting the command. 
+# library(devtools)
+# devtools::install_github("tidyverse/ggplot2")
+
+# devtools::install_github("einarhjorleifsson/iceshape", dependencies = FALSE)
+library(iceshape)   # ls("package:iceshape")
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(viridis)
+
+source("../mptools/r/my_utils.r")
+# load("../prf/rdata/world.df.RData")
+
+# get dropbox directory
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# open Sframe (STECF) data
+load(file = file.path(dropboxdir,"rdata", "STECF sframe.RData"))
+
+head(stockdef)
+head(fmz)
+
+
+# ----------------------------------------------------------------------------------------------------------
+# set world in sf format
+# ----------------------------------------------------------------------------------------------------------
+
+world1 <- sf::st_as_sf(
+  map('world', plot = FALSE, fill = TRUE, 
+               regions = c("netherlands","belgium","France(?!:Corsica)",
+                           "ireland","united kingdom", "UK", "Denmark","Germany",
+                           "norway","iceland", "greenland", "faroe islands", "spain","portugal",
+                           "sweden", "finland","poland"))
+)
+
+# ggplot(data=world1) + geom_sf()
+
+
+# ----------------------------------------------------------------------------------------------------------
+# plot the required stocks, combine with fao areas (in package iceshape) and generate plot
+# ----------------------------------------------------------------------------------------------------------
+
+fmz %>% 
+  
+  filter(grepl("COD/1N2AB.", TAC_id)) %>% 
+  # filter(grepl("COD/[A-Z0-9]", TAC_id)) %>% 
+  mutate(unit = tolower(area)) %>% 
+  arrange(TAC_id) %>% 
+
[…]

### 2018-11-11 R/dev/SAG DLS categories.r

@@ -1,0 +1,174 @@
+# -----------------------------------------------------------------------------------------------
+# SAG DLS categories
+#
+# 02/09/2018 first coding
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load datasets: sag, sagful, qcsexcel, istock, iadvice
+sagfull <- 
+  get(load(file=file.path(assessdir, "rdata/icesSAGfull 20180812.RData")))
+
+sagfull %>% 
+  group_by(assessmentyear, )
+  filter(row_number() == 1) %>% 
+  t() %>% 
+  pandoc.table(., 
+               style        = "simple",
+               split.tables = 80, 
+               split.cells  = c(rep(7,10)),
+               justify      = "left",
+               missing      =".",
+               big.mark     = ',', 
+               round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) )
+
+sagfull %>% 
+  filter(row_number() == 1) %>% 
+  t() %>% 
+  data.frame() %>% 
+  rownames_to_column() %>%  
+  setNames(c("var","tmp")) %>%
+  mutate(tmp = substr(tmp, 1, 50)) %>% 
+  arrange(var) %>% 
+  pandoc.table(., 
+               style        = "simple",
+               split.tables = 100, 
+               split.cells  = c(rep(7,10)),
+               justify      = "left",
+               missing      =".",
+               big.mark     = ',', 
+               round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) )
+
+qcsexcel %>% 
+  filter(row_number() == 1) %>% 
+  t() %>% 
+  pandoc.table(., 
+               style        = "simple",
+               split.tables = 80, 
+               split.cells  = c(rep(7,10)),
+               justify      = "left",
+               missing      =".",
+               big.mark     = ',', 
[…]

### 2018-11-11 R/dev/XML reader.r

@@ -1,0 +1,9 @@
+library(xml2)
+library(tidyverse)
+
+data <- read_xml("D:/Dropbox/ICES Assessment database/data/getStockDownloadData.xml")
+summary(data)
+point <- 
+  data %>% 
+  xml_find_all("//point")
+

### 2018-11-11 R/dev/plot iStock area.r

@@ -1,0 +1,111 @@
+# --------------------------------------------------------------------------------------------
+# plot iStock area.r
+#
+# plotting stock areas (over time)
+# 
+# 27/04/2018 cleaned up the code
+# --------------------------------------------------------------------------------------------
+
+# devtools::install_github("fishvice/tidyices", dependencies = FALSE)
+
+# Note: geom_sf is only on the development version of ggplot2. Make sure to start rstudio in administrator mode before starting the command. 
+# library(devtools)
+# devtools::install_github("tidyverse/ggplot2")
+
+# devtools::install_github("einarhjorleifsson/iceshape", dependencies = FALSE)
+library(iceshape)   # ls("package:iceshape")
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(viridis)
+search()
+
+source("../mptools/r/my_utils.r")
+# load("../prf/rdata/world.df.RData")
+
+# get dropbox directory
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# ----------------------------------------------------------------------------------------------------------
+# set world in sf format
+# ----------------------------------------------------------------------------------------------------------
+
+world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE, 
+                           regions = c("netherlands","belgium","France(?!:Corsica)",
+                                       "ireland","united kingdom", "UK", "Denmark","Germany",
+                                       "norway","iceland", "greenland", "faroe islands", "spain","portugal",
+                                       "sweden", "finland","poland")))
+# ggplot(data=world1) + geom_sf()
+
+
+# ----------------------------------------------------------------------------------------------------------
+# read the iStock data
+# ----------------------------------------------------------------------------------------------------------
+istock <-
+  read_excel(path=paste(dropboxdir, "/data/istock.xlsx",sep=""),
+             sheet="istock", col_names=TRUE, col_types="text") %>%
+  lowcase() %>%
+  mutate(stockarea = ifelse(substr(stockarea,1,2)=="21",toupper(stockarea), stockarea))
+
+
+# ----------------------------------------------------------------------------------------------------------
[…]

### 2018-11-11 R/iAdvice advice tac and catch.r

@@ -1,0 +1,88 @@
+# ===========================================================================
+# iAdvice refpoints.r
+#
+# 10/10/2018 initial coding
+# ===========================================================================
+
+# rm(list=ls())
+
+library(readxl)
+library(tidyverse)
+library(stringr)
+library(pander)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Source the utils file
+source("../mptools/R/my_utils.r")
+
+# set dropboxdirs
+assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
+advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")
+
+
+load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
+load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))
+
+iAdvice <- 
+  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20181001.xlsx",sep=""), 
+             sheet     = "DATA",
+             col_names = TRUE,
+             col_types = "text",
+             skip      = 0)  %>% 
+  lowcase() %>% 
+  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
+             funs(as.integer)) %>% 
+  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
+               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
+               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
+               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
+  mutate( blim     = 1000 * blim, 
+          msybtrig = 1000 * msybtrig) %>% 
+  
+  mutate_at( c("stockices","speciestype", "assessmodel", 
+               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
+               "aggregated"), funs(tolower)) %>% 
+  rename(stockkeylabelold = stockices, 
+         assessmentyear = assessyear) %>% 
+  arrange(stockkeylabelold, tacarea, year) %>%  
+  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
+  left_join(iStockkey, by="stockkeylabelold") 
+
+
+
+
+
+# ---------------------------------------------------------------------------
[…]

### 2018-11-11 R/iAdvice database revision.r

@@ -1,0 +1,120 @@
+# -----------------------------------------------------------------------------------------------
+# iDatabases checks
+#
+# 10/08/2017 created code from ICES SAG download.r
+# 25/09/2017 cleaned up; only checking code left
+# 23/10/2018 comparing iAdvice and iStock
+# 25/10/2018 revising the 20181001 iAdvice database in Excel
+# -----------------------------------------------------------------------------------------------
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+advicedir  <- paste(get_dropbox(), "/ICES Advice database", sep="")
+
+load(file=paste(assessdir, "/rdata/iRename.RData",sep=""))
+load(file=paste(assessdir, "/rdata/iStockkey.RData",sep=""))
+
+# Read and convert iAdvice spreadsheet
+iAdvice <-
+  
+  # read from excel
+  readxl::read_excel(
+    path= paste(advicedir, "/Excel/ICES Scientific Advice database 20181001.xlsx", sep=""), 
+    sheet     = "DATA",
+    col_names = TRUE, 
+    col_types = "text", 
+    trim_ws   = FALSE) %>%
+  
+  # make variables lowercase
+  lowcase() %>% 
+  
+  # rename variables
+  rename(tacyear = year, 
+         landings=iceslandings, ibc=icesindustrialbycatch, discards=icesdiscards,
+         catches = icescatch, msybtrigger=msybtrig, assessmentyear=assessyear,
+         assessmenttype = assesstype, assessmentcomment=assesscomment, assessmentmodel = assessmodel, 
+         catchlandingsunits = catchunits, stockkeylabel = stockices) %>% 
+  
+  # convert to numeric
+  mutate_at(c("advisedlandingsmax", "advisedcatchmax", "tal", "tac", 
+              "officiallandings", "landings", "ibc", "discards", "catches",
+              "fsqymin1", "ssbymin1", "fadvmax", 
+              "fmax", "f01", "fmed", "f35spr", "flim", "fpa", "fmsy",
[…]

### 2018-11-11 R/iAdvice refpoints.r

@@ -1,0 +1,89 @@
+# ===========================================================================
+# iAdvice refpoints.r
+#
+# 10/10/2018 initial coding
+# ===========================================================================
+
+# rm(list=ls())
+
+library(readxl)
+library(tidyverse)
+library(stringr)
+library(pander)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Source the utils file
+source("../mptools/R/my_utils.r")
+
+# set dropboxdirs
+assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
+advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")
+
+
+load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
+load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))
+
+iAdvice <- 
+  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20181001.xlsx",sep=""), 
+             sheet     = "DATA",
+             col_names = TRUE,
+             col_types = "text",
+             skip      = 0)  %>% 
+  lowcase() %>% 
+  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
+             funs(as.integer)) %>% 
+  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
+               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
+               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
+               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
+  mutate( blim     = 1000 * blim, 
+          msybtrig = 1000 * msybtrig) %>% 
+  
+  mutate_at( c("stockices","speciestype", "assessmodel", 
+               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
+               "aggregated"), funs(tolower)) %>% 
+  rename(stockkeylabelold = stockices, 
+         assessmentyear = assessyear) %>% 
+  arrange(stockkeylabelold, tacarea, year) %>%  
+  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
+  left_join(iStockkey, by="stockkeylabelold") 
+
+
+
+
+
+# ---------------------------------------------------------------------------
[…]

### 2018-11-11 R/iAssess figure Mohns rho.r

@@ -1,0 +1,335 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph plotting
+#
+# 30/03/2017 first coding during HAWG
+# 14/07/2017 adapted during HERAS
+# 11/08/2017 adapter for R 3.4.1 and tidyverse
+# 14/08/2017 added plot for assessment methods
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+library(lubridate)
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# Load dataset
+load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
+
+
+# ---------------------------------------------------------------------------------------------
+# Create subset of series (stocks) with at least 8 years of data and SSB in tonnes, F in year-1
+# ---------------------------------------------------------------------------------------------
+selected <-
+  iAssess %>% 
+  
+  group_by(stockkey, stockkeylabelold, assessmentyear) %>%
+  filter(row_number() == 1) %>%
+  filter(assessmentyear >= 1990, assessmentyear <= 2017) %>% 
+  
+  group_by(stockkey, stockkeylabelold) %>%
+  summarise(nyears = n()) %>% 
+  filter(nyears >= 15) %>% 
+  select(stockkey, stockkeylabelold)
+
+
+# Check certain assessments
+# filter(sagdb, stockkeylabelold == "cod-2532", assessmentyear==2017, stocksizeunits=="biomass") %>% View()
+
+# ---------------------------------------------------------------------------------------------
+# Calculate Mohn's rho; first in code then in function
+# ---------------------------------------------------------------------------------------------
+d <-
+  selected %>% 
[…]

### 2018-11-11 R/iAssess figure assessments units per stock overview.R

@@ -1,0 +1,136 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph data plotting
+#
+# 01/08/2017 extracted code from SAG download code
+# 11/08/2017 adapted for R 3.4.1
+# 28/08/2017 small modification to variable names
+# -----------------------------------------------------------------------------------------------
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+# Load dataset
+load(file="rdata/sagdb.RData")
+
+# ---------------------------------------------------------------------------------------------
+# plots of ssb or f descriptions by assessment year and stock and fisheries guild
+# ---------------------------------------------------------------------------------------------
+
+myvar  <- "stocksizedescription"
+myvar2 <- "ssb"
+
+myvar  <- "fishingpressuredescription"
+myvar2 <- "f"
+
+x <-
+  sagdb %>% 
+  filter(assessmentyear >= 1990) %>% 
+  filter(fisheriesguild != "", !is.na(fisheriesguild)) %>% 
+  filter(fisheriesguild %in% c("crustacean","elasmobranch","pelagic","demersal","benthic")) %>%
+  mutate(fishstockold = ifelse(is.na(fishstockold)|fishstockold=="", fishstocknew, fishstockold)) %>% 
+  group_by(fishstockold, fishstocknew, assessmentyear, source) %>% 
+  data.frame() %>% 
+  mutate(fishstockold = factor(fishstockold), 
+         fishstockold = factor(fishstockold, levels = rev(levels(fishstockold)))) %>% 
+
+  # here is the variable bit with myvar and myvar2
+  mutate_(myvar = ifelse(is.na(myvar2), NA, myvar))
+
+# x <-
+#   sagdb %>% 
+#   filter(assessmentyear >= 1990) %>% 
+#   filter(fisheriesguild != "", !is.na(fisheriesguild)) %>% 
[…]

### 2018-11-11 R/iAssess figure compare assessment to forecasts.r

@@ -1,0 +1,113 @@
+# -----------------------------------------------------------------------------------------------
+# iAssess figure compare assessment with forecasts
+#
+# 05/10/2017 first coding
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load the data
+load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iForecast.RData", sep=""))
+
+# convert iForecast to the appropriate years
+t <-
+  iForecast %>% 
+  mutate(year = lag(managementyear), 
+         ssbpreviousyear = ssbpreviousyear * 1000)
+
+# ---------------------------------------------------------------------------------------------
+# plot recent assessment with historic forecast values 
+# ---------------------------------------------------------------------------------------------
+
+d <-
+  iAssess %>% 
+  filter(grepl("mac-nea", stockkeylabel),
+         assessmentyear == 2017,
+         assessmenttype2 == "assess") %>% 
+  ungroup() %>% 
+  filter(year             >  1980,
+         year             <= assessmentyear) %>% 
+  select(assessmentyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, 
+         recruitment:lowrecruitment, f:lowf, ssb:lowssb,
+         assessmenttype2) %>%    
+  left_join(t, by=c("stockkey","stockkeylabel","year")) %>% 
+  data.frame()
+
+# plot ssb
+p1 <-
+  d %>% 
+  filter(!is.na(ssb)) %>%  
+  
+  ggplot(aes(year,ssb)) +
+  
+  theme_publication() +
+  theme(legend.title=element_blank(),
+        axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
+        axis.text.y = element_text(size=9),
+        # strip.background = element_blank(),
+        legend.position = "null") +
+  
[…]

### 2018-11-11 R/iAssess figure historic retro by stock.r

@@ -1,0 +1,223 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph plotting
+#
+# 30/03/2017 first coding during HAWG
+# 14/07/2017 adapted during HERAS
+# 11/08/2017 adapter for R 3.4.1 and tidyverse
+# 14/08/2017 added plot for assessment methods
+# 04/10/2017 only plot for retro. Updated for iAssess
+# -----------------------------------------------------------------------------------------------
+
+rm(list=ls())
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load the data
+load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+
+# filter(exceldata, stockkeylabel=="her-noss", assessmentyear==2017)
+# filter(sagdata, stockkeylabel %in% c("her-noss","her.27.1-24a514a"), assessmentyear==2017) %>% 
+#   View()
+
+# ---------------------------------------------------------------------------------------------
+# Historic retros: plot stock data over different assessment years 
+# ---------------------------------------------------------------------------------------------
+
+d <-
+  iAssess %>% 
+
+  # filter(grepl("cod-iceg", stockkeylabelold) ) %>% 
+  # filter(grepl("mac-nea", stockkeylabelold),assessmenttype %in% c("update","bench"),assessmentyear>=2016) %>% 
+  # filter(grepl("hom-west", stockkeylabelold),assessmenttype %in% c("update","bench"),assessmentyear>=2016) %>% 
+  # filter(grepl("ple-nsea", stockkeylabelold) ) %>% 
+  # filter(grepl("cod-347d", stockkeylabelold) ) %>% 
+  filter(grepl("hom-west", stockkeylabelold) ) %>% 
+  
+  # filter(grepl("mac-nea|hom-west|whb-comb|her-noss", stockkeylabelold) ) %>% 
+  # filter(grepl("whb", stockkeylabelold) ) %>% 
+  # filter(grepl("noss", stockkeylabelold) ) %>% 
+  
+  ungroup() %>% 
+  filter(year             >   1985, 
[…]

### 2018-11-11 R/iDatabases checks.r

@@ -1,0 +1,234 @@
+# -----------------------------------------------------------------------------------------------
+# iDatabases checks
+#
+# 10/08/2017 created code from ICES SAG download.r
+# 25/09/2017 cleaned up; only checking code left
+# 23/10/2018 comparing iAdvice and iStock
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(directlabels)  # for printing labels at end of geom lines
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+advicedir  <- paste(get_dropbox(), "/ICES Advice database", sep="")
+
+# load(file=paste(assessdir, "/rdata/iStock.RData",sep=""))
+# load(file=paste(assessdir, "/rdata/qcsexcel.RData",sep=""))
+
+# load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+
+# -----------------------------------------------------------------------------------------
+# Compare iStock and iAdvice 
+# -----------------------------------------------------------------------------------------
+
+t <- 
+  iAdvice %>% 
+  dplyr::select(assessmentyear, tacyear, stockkey, stockarea, stocksubarea) %>% 
+  left_join(dplyr::select(iStock, 
+                          assessmentyear, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, stockarea), 
+            by = c("assessmentyear","stockkey"))
+
+# -----------------------------------------------------------------------------------------
+# Trial for anglerfish  
+# -----------------------------------------------------------------------------------------
+
+# t <- 
+#   iAlloc %>% 
+#   filter(!is.na(stockkey)) %>% 
+#   group_by(stockkeyscombined) %>% 
+#   filter(row_number() ==1) %>% 
[…]

### 2018-11-11 R/iMerge iStock to iAdvice.r

@@ -1,0 +1,186 @@
+# -----------------------------------------------------------------------------------------------
+# iMerge iStock to iAdviced.r
+#
+# Getting rid of the iStock database by merging unique records into iAdvice 
+# 
+# 07/11/2018 No longer need iStock
+# -----------------------------------------------------------------------------------------------
+
+rm(list=ls())
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+assessdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+advicedir  <- paste(get_dropbox(), "/ICES Advice database", sep="")
+
+# -----------------------------------------------------------------------------------------
+# load the Advice database
+# -----------------------------------------------------------------------------------------
+
+iAdvice <-
+  readxl::read_excel(
+    path= paste(advicedir, "/Excel/ICES Scientific Advice database 20181108.xlsx", sep=""), 
+    sheet     = "DATA",
+    col_names = TRUE, 
+    col_types = "text", 
+    trim_ws   = FALSE) %>%
+  lowcase() %>% 
+  
+  # split up advisedlandings
+  
+  # mutate(advisedlandings = gsub("\\[|\\]","",advisedlandings)) %>% 
+  # separate(advisedlandings, into=c("advisedlandingsmin","advisedlandingsmax"), sep="-") %>% 
+  # mutate(advisedlandingsmax = ifelse(is.na(advisedlandingsmax), advisedlandingsmin, advisedlandingsmax)) %>% 
+  # mutate(advisedlandingsmin = ifelse(advisedlandingsmin == advisedlandingsmax, NA, advisedlandingsmin)) %>% 
+  
+  # split up advisedcatch
+  
+  # mutate(advisedcatch = gsub("\\[|\\]","",advisedcatch)) %>% 
+  # separate(advisedcatch, into=c("advisedcatchmin","advisedcatchmax"), sep="-") %>% 
[…]

### 2018-11-11 R/old/ICES Assessment Summary database code1.r

@@ -1,0 +1,203 @@
+# =================================================================
+# R-script for analyzing and plotting the ICES quality database and
+# the ICES advice database
+#
+# Martin Pastoors
+#
+# Original: 07/05/2014
+# Updated : 01/06/2015
+# =================================================================
+
+# Reset lists
+rm(list=ls())
+
+# Add relevant (graphic) libraries
+library(graphics)
+library(lattice)
+library(reshape2)
+library(dplyr)
+
+source ("multiplot.r")
+
+# Set working directory
+setwd("C://data//data//ICES")
+
+# Read ICES quality database in csv format ------------------------------------
+so             <- read.csv("ICES quality database 20150429.csv", stringsAsFactors=F)
+
+so$F           <- as.numeric(so$F)
+so$Recruitment <- as.numeric(so$Recruitment)
+so$SSB         <- as.numeric(so$SSB)
+so$Landings    <- as.numeric(so$Landings)
+names(so)
+
+# Read ICES advice database in csv format
+to             <- read.csv("20150428 History of scientific advice and management.csv", stringsAsFactors=F)
+ta             <- select(to, Year, Stock, Stocktype, Area, Management.authority, 
+                         Assess.year)
+names(ta)      <- c("y","FishStock","stkt", "area", "ma", "AssYear")
+
+# merge quality and advice datasets
+ua <- merge (sa,ta,by=c("AssYear", "FishStock"),all.x= TRUE)
+# names(ua)
+
+# select only relevant columns
+# using select and names(); to be done
+
+# =========================================================================
+
+# Calculate Mohn's rho on the dataset
+
+# =========================================================================
+
+# Plot for a certain species
+stk   <- "her-47d3"
+minay <- 2000
+maxay <- 2015
+miny  <- 1990
+maxy  <- 2015
+g   <- ua[ua$FishStock == stk & 
+          ua$Year    >= miny  & ua$Year <= maxy & 
+          ua$AssYear >= minay & ua$AssYear <= maxay ,]
+lng <- length(unique(g$AssYear))
+
+# generate plot of retro's
[…]

### 2018-11-11 R/old/ICES Assessment Summary database code2.r

@@ -1,0 +1,44 @@
+# R-script for ICES fishdata base 2013
+# Martin Pastoors
+# 07/05/2014
+
+# Reset lists
+rm(list=ls())
+
+# Add graphics library
+library(graphics)
+library(lattice)
+
+# Set working directory
+# setwd("C://Users//Martin Pastoors//Documents//ZZZ//ICES fishdata")
+setwd("C://data//data//ICES")
+
+fishdata <- read.csv("ICES quality database 20150429.csv", stringsAsFactors=F)
+str(fishdata)
+# summary(fishdata)
+
+# subset data to north sea
+fishdataNS2013 <- fishdata[
+                      fishdata$AssYear==2013 & 
+                      fishdata$Ecoregion== "north sea" & 
+                      fishdata$Species %in% c("her","ple","cod","sol", "had", "tur", "whg"),]
+# add the decades
+decade <- floor(fishdataNS2013$Year / 10) * 10
+fishdataNS2013 <- cbind(fishdataNS2013,decade)
+
+# str(fishdataNS2013)
+# summary(fishdataNS2013)
+
+# Plot the trends in F (with smoother)
+xyplot( F~Year|FishStock ,data=fishdataNS2013, xlim=c(1950,2020), ylim=c(0,1.2), scales=list(0.3, 0.6, 0.9), pch="+")
+
+# Calculate averages
+mean.fishdata <- aggregate(fishdataNS2013$MeanF, list(fishdataNS2013$decade,fishdataNS2013$FishStock), mean, na.rm=TRUE)
+# mean.fishdata.tapply <- tapply(fishdataNS2013$MeanF, list(fishdataNS2013$decade,fishdataNS2013$FishStock), mean, na.rm=TRUE)
+# head(mean.fishdata)
+
+# Plot decade averages
+xyplot( x~Group.1|Group.2 ,data=mean.fishdata, xlim=c(1950,2020), ylim=c(0,1.2), scales=list(0.3, 0.6, 0.9))
+
+
+

### 2018-11-11 R/old/exceldata.r

@@ -1,0 +1,93 @@
+# -----------------------------------------------------------------------------------------------
+# exceldata.r
+#
+# 30/03/2017 first coding during HAWG
+# 07/07/2017 adapted for downloading as csv file
+# 14/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
+# 20/07/2017 thorough checking of databases and making sure all items are filled and OK
+# 01/08/2017 added a number of diagnostic graphs; redone the excel database for nephrops
+# 01/08/2017 added the stock assessment methods from stock advice database
+# 10/08/2017 moved checking routines to separate code
+# 11/08/2017 adapted for R3.4.1 and Tidyverse
+# 14/08/2017 added assessmentmodel and assessmenttype categories
+# 04/09/2017 updated for western horse mackerel 2017
+# 06/09/2017 added automatic link to dropbox folder
+# 10/09/2017 split all data into new databases: iAssess, iAdvice, iManage, iForecast, iSpecies, iStock
+# 04/10/2017 updated with new ICES data on WGWIDE assessments
+# 06/10/2017 added assessmenttype2 column in excel spreadsheet
+# 09/10/2017 added qcsdata integration and added istock read from spreadsheet
+# 16/10/2017 final check before meeting in Reykjavik
+# 17/10/2017 updates during workshop in Reykjavik
+# 23/10/2017 moved iStock.r to separate code file
+# 03/11/2017 split excel datafile into separate code file
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load iRename
+load(file=paste(dropboxdir, "/rdata/iRename.RData", sep=""))
+
+# -----------------------------------------------------------------------------------------
+# exceldata: Old excel assessment database and prepare for combining
+# -----------------------------------------------------------------------------------------
+
+exceldata <-
[…]

### 2018-11-11 R/old/grid_arrange_shared_legend.r

@@ -1,0 +1,50 @@
+# ----------------------------------------------------------------------------------------------
+# grid_arrange_shared_legend.r 
+# 
+# https://github.com/tidyverse/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
+# 24/3/2017
+# ----------------------------------------------------------------------------------------------
+
+grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right")) {
+
+  require(ggplot2)
+  require(gridExtra)
+  require(grid)
+  
+  plots    <- list(...)
+  position <- match.arg(position)
+  g        <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
+  legend   <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
+  lheight  <- sum(legend$height)
+  lwidth   <- sum(legend$width)
+  gl       <- lapply(plots, function(x) x + theme(legend.position="none"))
+  gl       <- c(gl, ncol = ncol, nrow = nrow)
+  
+  combined <- switch(position,
+                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
+                                            legend,
+                                            ncol = 1,
+                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
+                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
+                                           legend,
+                                           ncol = 2,
+                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))
+  
+  grid.newpage()
+  grid.draw(combined)
+  
+  # return gtable invisibly
+  invisible(combined)
+  
+}
+
+
+# dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
+# 
+# p1 <- qplot(carat, price, data = dsamp, colour = clarity)
+# p2 <- qplot(cut, price, data = dsamp, colour = clarity)
+# p3 <- qplot(color, price, data = dsamp, colour = clarity)
+# p4 <- qplot(depth, price, data = dsamp, colour = clarity)
+# 
+# grid_arrange_shared_legend(p1, p2, p3, p4, ncol = 4, nrow = 1)
+# grid_arrange_shared_legend(p1, p2, p3, p4, ncol = 2, nrow = 2)

### 2018-11-11 R/old/historic retro dev.r

@@ -1,0 +1,574 @@
+# ===========================================================================
+# R-script for ICES assessment summary 1982-2016
+# Martin Pastoors
+
+# Original:09/07/2014
+# updated: 08/09/2014
+# updated: 10/03/2015
+# updated: 05/04/2016
+# 10/04/2016 removed forecast years to separate file
+# 04/09/2016 updated during WGWIDE
+# 24/02/2014 adapted for all lowercase; added directlabels (optional)
+# 27/06/2017 more plots for western horse mackerel
+# ===========================================================================
+
+# Reset lists
+rm(list=ls())
+
+library(directlabels)  # for printing labels at end of geom lines
+library(readxl)
+require(ggplot2)
+library(dplyr)
+library(tidyr)
+
+# Set working directory
+setwd("D:/Dropbox/ICES Assessment database")
+
+# code for combining plots
+source("r/multiplot.r")
+
+# code for combining plots with single legend
+source("r/grid_arrange_shared_legend.r")
+
+# publication theme
+source("r/theme_publication.r")
+
+
+upcase <- function(df) {
+  names(df) <- toupper(names(df)) %>% gsub("\\s+","_",.) 
+  df
+}
+
+lowcase <- function(df) {
+  names(df) <- tolower(names(df)) %>% gsub("\\s+","_",.) 
+  df
+}
+
+# ===========================================================================
+# Read and convert data; or load the dataframes
+# ===========================================================================
+
+load("../ICES advice database/rdata/IAD.RData")
+
+iad <-
+  iad %>% 
+  lowcase 
+  
+
+# rby <-
+#   read_excel("ICES Assessment Summary database.xlsx",
+#              sheet = "DATA",
+#              col_names = TRUE,
+#              skip = 0) %>%
+#   lowcase %>%
+#   mutate_each(funs(as.numeric), c(low_f:high_f,
+#                                   low_recruitment:high_recruitment,
+#                                   low_ssb:high_ssb,
+#                                   landings, catches,
+#                                   year, assyear)) %>%
+#   select(-matches("custom"))
+ 
+# save(rby, file="ICES Assessment Summary database.RData")
[…]

### 2018-11-11 R/old/historic retro mac-nea.r

@@ -1,0 +1,199 @@
+# ===========================================================================
+# R-script for ICES assessment summary 1982-2016
+# Martin Pastoors
+
+# Original:09/07/2014
+# updated: 08/09/2014
+# updated: 10/03/2015
+# updated: 05/04/2016
+# 10/04/2016 removed forecast years to separate file
+# 04/09/2016 updated during WGWIDE
+# 24/02/2014 adapted for all lowercase
+# ===========================================================================
+
+# Reset lists
+rm(list=ls())
+
+library(readxl)
+require(ggplot2)
+library(dplyr)
+library(tidyr)
+library(directlabels)  # for printing labels at end of geom lines
+
+# Set working directory
+setwd("C:/Users/Martin Pastoors/Documents/Dropbox/ICES Assessment database")
+
+# code for combining plots
+source("r/multiplot.r")
+source("r/theme_publication.r")
+
+# ===========================================================================
+# Read and convert data; or load the dataframe
+# ===========================================================================
+
+load("ICES Assessment Summary database.RData")
+
+# rby <-
+#   read_excel("ICES Assessment Summary database.xlsx",
+#              sheet = "DATA",
+#              col_names = TRUE,
+#              skip = 0) %>%
+#   mutate_each(funs(as.numeric), c(F, Recruitment, SSB,
+#                                   Landings, Catches,
+#                                   Year, AssYear))
+# 
+# names(rby) <- tolower(names(rby))
+
+# save(rby, file="ICES Assessment Summary database.RData")
+
+# Store the forecast years
+rbyF <-
+  rby %>% 
+  tbl_df() %>% 
+  filter(year >= assyear) %>% 
+  mutate(forecastyear = year - assyear)
+# write.csv(rbyF,"rbyF.csv")
+
+# Store the assessments
+rbyA <-
+  rby %>% 
+  tbl_df() %>% 
+  filter(year <= assyear)
+# write.csv(rbyA,"rbyA.csv")
+
+# ========================================================================================
+# Historic retros: plot stock data over different assessment years and assessment versions
+# ========================================================================================
+
+d <-
+  rbyA %>% 
[…]

### 2018-11-11 R/old/historic retro.r

@@ -1,0 +1,283 @@
+# ===========================================================================
+# R-script for ICES assessment summary 1982-2016
+# Martin Pastoors
+
+# Original:09/07/2014
+# updated: 08/09/2014
+# updated: 10/03/2015
+# updated: 05/04/2016
+# 10/04/2016 removed forecast years to separate file
+# 04/09/2016 updated during WGWIDE
+# 24/02/2014 adapted for all lowercase; added directlabels (optional)
+# ===========================================================================
+
+# Reset lists
+rm(list=ls())
+
+library(readxl)
+require(ggplot2)
+library(dplyr)
+library(tidyr)
+library(directlabels)  # for printing labels at end of geom lines
+
+# Set working directory
+setwd("C:/Users/Martin Pastoors/Documents/Dropbox/ICES Assessment database")
+
+# code for combining plots
+source("r/multiplot.r")
+
+# ===========================================================================
+# Read and convert data; or load the dataframe
+# ===========================================================================
+
+load("ICES Assessment Summary database.RData")
+
+# rby <-
+#   read_excel("ICES Assessment Summary database.xlsx",
+#              sheet = "DATA",
+#              col_names = TRUE,
+#              skip = 0) %>%
+#   mutate_each(funs(as.numeric), c(F, Recruitment, SSB,
+#                                   Landings, Catches,
+#                                   Year, AssYear))
+# 
+# names(rby) <- tolower(names(rby))
+# 
+# save(rby, file="ICES Assessment Summary database.RData")
+
+# Store the forecast years
+rbyF <-
+  rby %>% 
+  tbl_df() %>% 
+  filter(year >= assyear) %>% 
+  mutate(forecastyear = year - assyear)
+# write.csv(rbyF,"rbyF.csv")
+
+# Store the assessments
+rbyA <-
+  rby %>% 
+  tbl_df() %>% 
+  filter(year <= assyear)
+# write.csv(rbyA,"rbyA.csv")
+
+# ===========================================================================
+# Historic retros: plot stock data over different assessment years 
+# ===========================================================================
+
+
+d <-
+  rbyA %>% 
+  filter(fishstock %in% c("her-noss", "whb-comb","mac-nea","hom-west"),
[…]

### 2018-11-11 R/old/iAssess figure assessments per stock overview.R

@@ -1,0 +1,194 @@
+# -----------------------------------------------------------------------------------------------
+# iAssess figure all assessments per stock
+#
+# 25/09/2017 taken from ices SAG data plot
+# 16/10/2017 now includes datapublished
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+library(lubridate)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# Load dataset
+load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
+
+# ---------------------------------------------------------------------------------------------
+# plots of assessments by assessment year and stock and fisheries guild (for a single variable)
+# ---------------------------------------------------------------------------------------------
+
+# To do: convert this into a function that can plot the different variables. I have struggled with this a lot but 
+# without success. 1/8/2017
+
+x <-
+  iAssess %>% 
+  # filter(assessmentyear >= 1986) %>% 
+  # filter(fisheriesguild != "", !is.na(fisheriesguild)) %>% 
+  # filter(fisheriesguild %in% c("crustacean","elasmobranch","pelagic","demersal","benthic")) %>%
+  group_by(stockkeylabelold, stockkeylabelnew, assessmentyear, assessmenttype2, source) %>% 
+  filter(row_number() ==1) %>% 
+  select(-year) %>% 
+  ungroup() %>% 
+  arrange(stockkeylabelold) %>% 
+  mutate(id = group_indices(., stockkeylabelold)) %>% 
+  data.frame() %>% 
+  mutate(stockkeylabelold = ifelse(is.na(stockkeylabelold), stockkeylabel, stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = factor(stockkeylabelold), 
+         stockkeylabelold = factor(stockkeylabelold, levels = rev(levels(stockkeylabelold))),
[…]

### 2018-11-11 R/old/iAssess test.r

@@ -1,0 +1,145 @@
+# -----------------------------------------------------------------------------------------------
+# IAssess test.r
+#
+# 22/10/2017 test consistency of iAssess components
+# -----------------------------------------------------------------------------------------------
+
+# Note requires running iDatabases generator.r beforehand
+
+library(RColorBrewer)
+library(tidyverse)
+
+filter(iAssess_test, stockkeylabel == "her-47d3" & source=="qcs") %>% View
+
+iAssess_test <- 
+  bind_rows(iAssess_part1, iAssess_part2) %>%  
+  bind_rows(., iAssess_part3) %>% 
+  
+  # corrections to stock size descriptions
+  mutate(
+    stocksizedescription = gsub("stock size: "                 ,""                   , stocksizedescription),
+    stocksizedescription = gsub("indices"                      ,"index"              , stocksizedescription),
+    stocksizedescription = gsub("indicator"                    ,"index"              , stocksizedescription),
+    stocksizedescription = gsub("^biomass$"                    ,"total biomass index", stocksizedescription),
+    stocksizedescription = gsub("^biomass index$"              ,"total biomass index", stocksizedescription),
+    stocksizedescription = gsub("^evhoe biomass index$"        ,"total biomass index", stocksizedescription),
+    stocksizedescription = gsub("^stock size index: abundance$","abundance index"    , stocksizedescription),
+    stocksizedescription = gsub("^abundance$"                  ,"abundance index"    , stocksizedescription),
+    stocksizedescription = gsub("^density$"                    ,"density index"      , stocksizedescription),
+    stocksizedescription = gsub("^tsb$"                        ,"total biomass"      , stocksizedescription),
+    stocksizedescription = gsub("^total abundance index$"      ,"abundance index"    , stocksizedescription),
+    stocksizedescription = gsub("^index$"                      ,"abundance index"    , stocksizedescription),
+    stocksizedescription = gsub("^stock abundance$"            ,"abundance index"    , stocksizedescription),
+    stocksizedescription = gsub("^density index$"              ,"abundance index"    , stocksizedescription),
+    
[…]

### 2018-11-11 R/old/iDatabases generator old.r

@@ -1,0 +1,599 @@
+# -----------------------------------------------------------------------------------------------
+# IDatabases generator.r
+#
+# iSpecies  : properties of species (names etc.)
+# iStock    : properties of stocks and assessment
+# iAssess   : stock assessment database (by stock, assessmentyear and year and date)
+# iAdvice   : ICES advice and advice basis
+# iForecast : short term forecasts
+# iManage   : management actions and implementation
+# iKeys     : transferring names of stocks
+# 
+# stocklist : stock list with stock description and advice status by assessment year - integrate into iStock
+#
+# 30/03/2017 first coding during HAWG
+# 07/07/2017 adapted for downloading as csv file
+# 14/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
+# 20/07/2017 thorough checking of databases and making sure all items are filled and OK
+# 01/08/2017 added a number of diagnostic graphs; redone the excel database for nephrops
+# 01/08/2017 added the stock assessment methods from stock advice database
+# 10/08/2017 moved checking routines to separate code
+# 11/08/2017 adapted for R3.4.1 and Tidyverse
+# 14/08/2017 added assessmentmodel and assessmenttype categories
+# 04/09/2017 updated for western horse mackerel 2017
+# 06/09/2017 added automatic link to dropbox folder
+# 10/09/2017 split all data into new databases: iAssess, iAdvice, iManage, iForecast, iSpecies, iStock
+# 04/10/2017 updated with new ICES data on WGWIDE assessments
+# 06/10/2017 added assessmenttype2 column in excel spreadsheet
+# 09/10/2017 added qcsdata integration and added istock read from spreadsheet
+# 16/10/2017 final check before meeting in Reykjavik
+# 17/10/2017 updates during workshop in Reykjavik
+# 23/10/2017 moved iStock.r to separate code file
+# 03/11/2017 now merges from separate datasets: qcsdata, exceldata, sagdata
+# -----------------------------------------------------------------------------------------------
+
+# ICES Stock database
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSD")
[…]

### 2018-11-11 R/old/iStock.r

@@ -1,0 +1,131 @@
+# -----------------------------------------------------------------------------------------------
+# iStock.r
+#
+# Generate iStock, iRename and iStockkey
+#
+# 23/10/2017 copied code out of iDatabases generator.r
+# -----------------------------------------------------------------------------------------------
+
+# ICES Stock database libraries
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSD")
+# devtools::install_github("ices-tools-prod/icesSAG")
+# library(icesSD)  # ICES Stock database
+# library(icesSAG)  # ICES Stock Assessment Graphs
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# -----------------------------------------------------------------------------------------
+# Download or load the ICES webservices data
+# -----------------------------------------------------------------------------------------
+
+# getSD         <- icesSD::getSD()
+# getListStocks <- icesSAG::getListStocks(year=0)
+
+# save(getSD        , file=paste(dropboxdir, "/rdata/getSD.RData",sep=""))
+# save(getListStocks, file=paste(dropboxdir, "/rdata/getListStocks.RData",sep=""))
+
+# load(file=paste(dropboxdir, "/rdata/getSD.RData",sep=""))
+# load(file=paste(dropboxdir, "/rdata/getListStocks.RData",sep=""))
+
+# -------------------------------------------------------------------------------------------------
+# iStock, iStockkey, iRename
+# Generate iStock (by stock, assessment year and date) - information on stock assessment properties
+# Note: iStock read from Excel file because substantial amount of information added, e.g. areas
+# -------------------------------------------------------------------------------------------------
+
[…]

### 2018-11-11 R/old/ices SAG analysis.r

@@ -1,0 +1,539 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph plotting
+#
+# 30/03/2017 first coding during HAWG
+# 14/07/2017 adapted during HERAS
+# 11/08/2017 adapter for R 3.4.1 and tidyverse
+# 14/08/2017 added plot for assessment methods
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("D:/GIT/mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- get_dropbox()
+setwd(paste(dropboxdir, "/ICES Assessment database", sep=""))
+
+# load the data
+load(file="rdata/sagdb.RData")
+
+# ---------------------------------------------------------------------------------------------
+# plots for single year over multiple species
+# ---------------------------------------------------------------------------------------------
+
+# ssb over MSY Btrigger
+p1 <-
+  sagdb %>% 
+  filter(fishstocknew %in% c("her.27.20-24", 
+                             "her.27.3a47d", 
+                             "her.27.6a7bc",
+                             "her.27.1-24a514a",
+                             "mac.27.nea", 
+                             "whb.27.1-91214", 
+                             "hom.27.2a4a5b6a7a-ce-k8", 
+                             "spr.27.4") ) %>% 
+  filter(assessmentyear == 2016) %>% 
+  ggplot(aes(year,ssb, group=assessmentyear)) +
+  theme_publication() +
+  geom_line(aes(col = ifelse(ssb > msybtrig,'red','green')), size=1.5) +
+  # geom_line(col = "black") +
+  geom_hline(aes(yintercept=msybtrig), colour="gray80", linetype="dashed", size=0.5) +
+  expand_limits(y = 0) +
+  theme(legend.position  = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank(),
+        axis.ticks       = element_blank(),
+        legend.title     = element_blank(),
+        axis.text.x      = element_text(angle = 0, vjust = 0.5, size=9),
+        axis.text.y      = element_blank(),
[…]

### 2018-11-11 R/old/ices SAG checker.r

@@ -1,0 +1,298 @@
+# -----------------------------------------------------------------------------------------------
+# ICES SAG checker
+#
+# 10/08/2017 created code from ICES SAG download.r
+# -----------------------------------------------------------------------------------------------
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+load(file="rdata/sagdb.RData")
+
+count_not_na(sagdb)
+
+sort(unique(sagdb$assessmentcat))
+sort(unique(sagdb$stocksizedescription))
+sort(unique(sagdb$stocksizeunits))
+sort(unique(sagdb$fishingpressuredescription))
+sort(unique(sagdb$fishingpressureunits))
+sort(unique(sagdb$assesstype))
+sort(unique(sagdb$assessmenttype))
+sort(unique(sagdb$assessmentmodel))
+
+sagdb %>% 
+  filter(assessmentmodel %in% c("xsa","ica","vpa","ls","sam","xsm", "tsa", "trends only", "surba","sms")) %>% 
+  View()
+
+
+# check assessmentypes used
+sagdb %>% 
+  group_by(fishstock, assessmentyear) %>% 
+  filter(row_number() == 1) %>% 
+  filter(!is.na(assesstype)) %>% 
+  select(fishstock, assessmentyear, stockpublishnote, status, assesstype) %>% 
+  View()
+
+# check stocksizedescriptions used
+sagdb %>% 
+  group_by(fishstock, assessmentyear) %>% 
+  filter(row_number() == 1) %>% 
+  group_by(stocksizedescription) %>% 
+  summarise(n = n()) %>% 
+  arrange(stocksizedescription) %>% 
+  View()
+
+filter(sagdb, is.na(commonname)) %>% View()
+
+sort(unique(sagdb$stocksizeunits))
+filter(sagdb, is.na(fishstocknew)) %>% View()
+filter(sagdb, is.na(commonname)) %>% View()
+filter(sagdb, is.na(stocksizeunits)) %>% View()
+filter(sagdb, stocksizedescription == "stock size" & stocksizeunits == "tonnes") %>% View()
+filter(sagdb, grepl("ratio", stocksizeunits)) %>% View()
[…]

### 2018-11-11 R/old/ices SAG data checker.r

@@ -1,0 +1,2004 @@
+
+
+
+
+
+
+<!DOCTYPE html>
+<html lang="en">
+  <head>
+    <meta charset="utf-8">
+  <link rel="dns-prefetch" href="https://assets-cdn.github.com">
+  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
+  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">
+
+
+
+  <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/frameworks-bedfc518345498ab3204d330c1727cde7e733526a09cd7df6867f6a231565091.css" integrity="sha256-vt/FGDRUmKsyBNMwwXJ83n5zNSagnNffaGf2ojFWUJE=" media="all" rel="stylesheet" />
+  <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/github-a1f1041276ec59b7ad51bdbd35d2d73f15f99aebe2686a60e9cd9f705b57d220.css" integrity="sha256-ofEEEnbsWbetUb29NdLXPxX5muviaGpg6c2fcFtX0iA=" media="all" rel="stylesheet" />
+  
+  
+  
+  
+
+  <meta name="viewport" content="width=device-width">
+  
+  <title>iAssess/ices SAG data checker.r at master · martinpastoors/iAssess</title>
+  <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
+  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
+  <meta property="fb:app_id" content="1401488693436528">
+
+    
+    <meta content="https://avatars2.githubusercontent.com/u/8004587?v=4&amp;s=400" property="og:image" /><meta content="GitHub" property="og:site_name" /><meta content="object" property="og:type" /><meta content="martinpastoors/iAssess" property="og:title" /><meta content="https://github.com/martinpastoors/iAssess" property="og:url" /><meta content="iAssess - ICES Historical Assessment Database" property="og:description" />
+
+  <link rel="assets" href="https://assets-cdn.github.com/">
[…]

### 2018-11-11 R/old/ices SAG data plot.r

@@ -1,0 +1,344 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph data plotting
+#
+# 01/08/2017 extracted code from SAG download code
+# 11/08/2017 adapted for R 3.4.1
+# 28/08/2017 small modification to variable names
+# -----------------------------------------------------------------------------------------------
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+# Load dataset
+load(file="rdata/sagdb.RData")
+
+# ---------------------------------------------------------------------------------------------
+# summarize dataset
+# ---------------------------------------------------------------------------------------------
+
+sagdb %>% 
+  group_by(fishstock, assessmentyear) %>% 
+  filter(row_number()==1) %>% 
+  filter(assessmentyear >= 2001) %>% 
+  mutate(species = paste0(commonname," (", faocode,")"), sep="") %>% 
+  group_by(species, assessmentyear) %>% 
+  summarise(nassess = n()) %>% 
+  
+  dcast(species ~ assessmentyear, sum, value.var="nassess", margins = TRUE) %>% 
+  no.emphasis.table() %>% 
+  select(-1) %>% 
+  pandoc.table(.,
+               style = "simple",
+               split.tables=400, 
+               justify = "right",
+               missing=".",
+               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
+               caption="number of assessments per species")
+
+
+
+sagdb %>% 
+  filter(assessmentyear >= 2001) %>% 
+  filter(assessmentcat == "assess") %>% 
+  group_by(fishstockold, assessmentyear) %>%
+  summarise(nyears = n()) %>% 
+  dcast(fishstockold ~ assessmentyear, value.var="nyears", fun.aggregate=sum, margins=TRUE) %>% 
+  View()
+
+
[…]

### 2018-11-11 R/old/ices SAG download.r

@@ -1,0 +1,494 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph download
+#
+# stocknames: lookup list to transfer old and new stock names
+# stocklist : stock list with stock description and advice status by assessment year
+# speciesdb : properties of species (names etc.)
+# stockdb   : properties of stocks
+# sagdb     : stock assessment database (by stock, assessmentyear and year)
+#
+# 30/03/2017 first coding during HAWG
+# 07/07/2017 adapted for downloading as csv file
+# 14/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
+# 20/07/2017 thorough checking of databases and making sure all items are filled and OK
+# 01/08/2017 added a number of diagnostic graphs; redone the excel database for nephrops
+# 01/08/2017 added the stock assessment methods from stock advice database
+# 10/08/2017 moved checking routines to separate code
+# 11/08/2017 adapted for R3.4.1 and Tidyverse
+# 14/08/2017 added assessmentmodel and assessmenttype categories
+# 04/09/2017 updated for western horse mackerel 2017
+# 06/09/2017 added automatic link to dropbox folder
+# -----------------------------------------------------------------------------------------------
+
+# ICES Stock database
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSD")
+# devtools::install_github("ices-tools-prod/icesSAG")
+# library(icesSD)  # ICES Stock database
+# library(icesSAG)  # ICES Stock Assessment Graphs
+
+# which functions in a package?
+# lsf.str("package:icesSAG")
+# lsf.str("package:icesSD")
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+
+# Load utils code
+source("D:/GIT/mptools/r/my_utils.r")
+
[…]

### 2018-11-11 R/old/ices SAG merge.r

@@ -1,0 +1,112 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph merge with 'old' dataset
+#
+# 15/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
+# -----------------------------------------------------------------------------------------------
+
+library(dplyr)     # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+# load csv with old and new stocknames
+stocknames <-
+  read.csv("D:/Dropbox/ICES Assessment database/ICES StocksOldCodesNewCodesEGs.csv", stringsAsFactors = FALSE) %>% 
+  select(1,2) %>% 
+  setNames(., c("fishstockold","fishstock")) 
+
+# load SAG database and combined database and then stop from here
+load(file="rdata/sagdb.RData")
+load(file="rdata/stockdb.RData")
+load(file="rdata/combdb.RData")
+
+# ----------------------------------------------------------------------
+
+# keep only fishstock and assessment year of the SAG database
+x <-
+  sagdb %>% 
+  group_by(fishstock, assessmentyear) %>% 
+  filter(row_number()==1) %>% 
+  select(fishstock, assessmentyear)
+
+# read old database
+rbya <-
+  read_excel("ICES Assessment Summary database.xlsx",
+             sheet = "DATA",
+             col_names = TRUE,
+             col_types = "text",
+             skip = 0) %>%
+  lowcase %>%
+  rename(assessmentyear = assyear) %>% 
+  mutate(fishstock = tolower(fishstock)) %>%
+  mutate_at(vars("year","assessmentyear"), funs(as.integer)) %>% 
+  mutate_at(vars("lowrecruitment", "recruitment","highrecruitment",
+                 "lowssb","ssb","highssb",
+                 "lowf", "f","highf",
+                 "landings","catches","discards","ibc",
+                 "flim","fpa","fmsy", "fmanagement",
+                 "blim","bpa","msybtrigger","bmanagement",
+                 "recruitmentage","recruitmentlength"), funs(as.numeric)) %>% 
+  select(-contains("custom")) %>% 
[…]

### 2018-11-11 R/old/ices SAG plot.r

@@ -1,0 +1,537 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph plotting
+#
+# 30/03/2017 first coding during HAWG
+# 14/07/2017 adapted during HERAS
+# 11/08/2017 adapter for R 3.4.1 and tidyverse
+# 14/08/2017 added plot for assessment methods
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+setwd("D:/Dropbox/ICES Assessment database")
+
+# Load utils code
+source("D:/XXX/PRF/r/my_utils.r")
+
+# load the data
+load(file="rdata/sagdb.RData")
+
+# ---------------------------------------------------------------------------------------------
+# plots for single year over multiple species
+# ---------------------------------------------------------------------------------------------
+
+# ssb over MSY Btrigger
+p1 <-
+  sagdb %>% 
+  filter(fishstocknew %in% c("her.27.20-24", 
+                             "her.27.3a47d", 
+                             "her.27.6a7bc",
+                             "her.27.1-24a514a",
+                             "mac.27.nea", 
+                             "whb.27.1-91214", 
+                             "hom.27.2a4a5b6a7a-ce-k8", 
+                             "spr.27.4") ) %>% 
+  filter(assessmentyear == 2016) %>% 
+  ggplot(aes(year,ssb, group=assessmentyear)) +
+  theme_publication() +
+  geom_line(aes(col = ifelse(ssb > msybtrig,'red','green')), size=1.5) +
+  # geom_line(col = "black") +
+  geom_hline(aes(yintercept=msybtrig), colour="gray80", linetype="dashed", size=0.5) +
+  expand_limits(y = 0) +
+  theme(legend.position  = "none",
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank(),
+        axis.ticks       = element_blank(),
+        legend.title     = element_blank(),
+        axis.text.x      = element_text(angle = 0, vjust = 0.5, size=9),
+        axis.text.y      = element_blank(),
+        strip.text.x     = element_text(colour="black", size=8, hjust=0),
[…]

### 2018-11-11 R/old/merge excel and qcs data.r

@@ -1,0 +1,564 @@
+# -----------------------------------------------------------------------------------------
+# temporary coding: merging of excel and qcs files
+# 05/01/2018
+# -----------------------------------------------------------------------------------------
+
+rm(list=ls())
+
+# Libraries
+library(rmarkdown)
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+library(lubridate)
+library(docxtractr) # read word documents
+library(icesSD)    # ICES Stock database
+library(icesSAG)   # ICES Stock Assessment Graphs
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# set dropboxdir
+dropboxdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
+
+# Generate iStock, iRename and iStockkey
+t <- 
+  readxl::read_excel(path= paste(dropboxdir, "/data/iStock.xlsx", sep=""), sheet = "istock",
+                     col_names = TRUE, col_types = "text", trim_ws   = FALSE) %>%
+  lowcase() %>% 
+  mutate_at(c("assessmentmodel","assessmenttype","advicetype",
+              "expertgroup","useofdiscardsinadvice","pabufferapplied", "assessmentstatus"),
+            funs(tolower)) %>% 
+  mutate_at(c("fmax","f01","fmed","f35spr","flim","fpa","fmsy","blim","bpa","msybtrigger"), 
+            funs(as.numeric)) %>% 
+  mutate_at(c("stockkey", "assessmentyear", "firstyearofdata", "ncpueseries","nsurveyseries"),
+            funs(as.integer)) %>% 
+  
+  mutate(assessmenttype   = ifelse(grepl("explo",assessmentmodel), "exploratory",assessmenttype),
+         assessmenttype   = ifelse(grepl("trends",assessmentmodel), "trends",assessmenttype) ) %>% 
+  
+  mutate(assessmentmodel  = ifelse(grepl("(xsa"   , assessmentmodel, fixed=TRUE), "xsa"   , assessmentmodel),
[…]

### 2018-11-11 R/old/plot exceldata and qcsdata.r

@@ -1,0 +1,365 @@
+# -----------------------------------------------------------------------------------------------
+# Plot Exceldata and QCS data
+#
+# 02/02/2018 first coding
+# 08/02/2018 fixing lot's of small things in the excel spreadsheet
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# ---------------------------------------------------------------------------------------------
+# Historic retros: plot stock data over different assessment years 
+# ---------------------------------------------------------------------------------------------
+
+d <-
+  readxl::read_excel(path= paste(dropboxdir, "/data/QCS and EXCEL Assessment Database combined.xlsx", sep=""), 
+                     sheet = "data",
+                     col_names = TRUE, 
+                     col_types = "text", 
+                     trim_ws   = FALSE) %>%
+  lowcase() %>% 
+  mutate_at(c("assessmentyear","year", "recruitment","ssb","f", "catches"),   funs(as.numeric)) %>% 
+  mutate_at(c("stockkey"),    funs(as.integer)) %>% 
+  mutate_at(c("unitofrecruitment", "recruitmentdescription"),    funs(tolower)) %>% 
+  mutate(recruitment       = ifelse(unitofrecruitment == "thousands", recruitment/1000, recruitment),
+         unitofrecruitment = ifelse(unitofrecruitment == "thousands", "millions", unitofrecruitment),
+         recruitment       = ifelse(unitofrecruitment == "billions", recruitment*1000, recruitment),
+         unitofrecruitment = ifelse(unitofrecruitment == "billions", "millions", unitofrecruitment),
+         tyear             = substr(as.character(assessmentyear),3,4),
[…]

### 2018-11-11 R/old/qcsdata.r

@@ -1,0 +1,293 @@
+# --------------------------------------------------------------------------------------------
+# qcsdata.r
+# 
+# Read Quality Control Sheets word tables
+#
+# 05/10/2017 coded up the reading in of all data files; datafiles converted to docx format; all superscripts removed prior to reading
+# 05/10/2017 converted to iAssess github
+# 03/11/2017 renamed file
+# --------------------------------------------------------------------------------------------
+
+library(docxtractr)
+library(tidyverse)
+library(stringr)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load rename datafiles
+load(file=paste(dropboxdir, "/rdata/iStockkey.RData", sep=""))
+load(file=paste(dropboxdir, "/rdata/iRename.RData", sep=""))
+
+# make list of filenames
+file.list <- c(list.files(path       = "data/qcs-docs-docx/", 
+                          pattern    = "docx",
+                          recursive  = TRUE, 
+                          full.names = TRUE,   
+                          ignore.case= TRUE) )
+
+# read files from filename list
+# i <- 1
+# j <- 1
+for (i in 1:length(file.list)) {
+  
+  # extract stock name
+  stockkeylabel <- gsub(".qcs.docx","", file.list[i], fixed=TRUE)
+  stockkeylabel <- gsub("((?:[^/]*/)*)(.*)","\\2", stockkeylabel)
+  
+  print(paste0(i,stockkeylabel,file.list[i],sep=" - "))
+  
+  # set docx object
+  docx <- read_docx(path=file.list[i])
+  # docx_tbl_count(docx)
+  # docx_describe_tbls(docx)
+  
+  # read tables
+  tmp <- docx_extract_all_tbls(docx, guess_header = FALSE, trim = TRUE)  
+  
+  # convert tables to data frames
+  for (j in 1:length(tmp)) {
+    
+    # print(j)
+    
+    t       <- tmp[j] %>%  as.data.frame()
+    variable <- tolower(t[1,1])
+    yeartype <- tolower(t[2,2])
+    nc      <- ncol(t)
+    nr      <- nrow(t)
+    head    <- t[3,2:nc]
[…]

### 2018-11-11 R/old/sagdata.r

@@ -1,0 +1,122 @@
+# -----------------------------------------------------------------------------------------------
+# sagdata.r
+#
+# download SAG data
+# 
+# 30/03/2017 first coding during HAWG
+# 07/07/2017 adapted for downloading as csv file
+# 14/07/2017 integrated the new and old stocknames; only data download and storage as rdata files
+# 20/07/2017 thorough checking of databases and making sure all items are filled and OK
+# 01/08/2017 added a number of diagnostic graphs; redone the excel database for nephrops
+# 01/08/2017 added the stock assessment methods from stock advice database
+# 10/08/2017 moved checking routines to separate code
+# 11/08/2017 adapted for R3.4.1 and Tidyverse
+# 14/08/2017 added assessmentmodel and assessmenttype categories
+# 04/09/2017 updated for western horse mackerel 2017
+# 06/09/2017 added automatic link to dropbox folder
+# 10/09/2017 split all data into new databases: iAssess, iAdvice, iManage, iForecast, iSpecies, iStock
+# 04/10/2017 updated with new ICES data on WGWIDE assessments
+# 06/10/2017 added assessmenttype2 column in excel spreadsheet
+# 09/10/2017 added qcsdata integration and added istock read from spreadsheet
+# 16/10/2017 final check before meeting in Reykjavik
+# 17/10/2017 updates during workshop in Reykjavik
+# 23/10/2017 moved iStock.r to separate code file
+# 03/11/2017 split off into separate SAG download code file
+# -----------------------------------------------------------------------------------------------
+
+# ICES Stock database
+
+# library(devtools)
+# devtools::install_github("ices-tools-prod/icesSD")
+# devtools::install_github("ices-tools-prod/icesSAG")
+library(icesSD)  # ICES Stock database
+library(icesSAG)  # ICES Stock Assessment Graphs
+
+library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# -----------------------------------------------------------------------------------------
[…]

### 2018-11-11 R/plot TAC areas test.r

@@ -1,0 +1,166 @@
+library(tidyverse)
+library(sf)
+library(maps)
+
+source("https://raw.githubusercontent.com/einarhjorleifsson/gisland/master/R/read_sf_ftp.R")
+
+
+# read ICES/FAO areas
+ices <- 
+  read_sf_ftp("fao-areas_nocoastline") %>% 
+  mutate(x = str_sub(name, 1, 2)) %>%
+  filter(x == "27") %>%
+  select(-x)
+# plot(ices)
+
+# clip map to include only ICES bound
+bb = st_bbox(ices)
+
+# actually, lets just focus on the central area, for now
+bb[[1]] <- -30
+bb[[2]] <- 45
+bb[[3]] <- 20
+bb[[4]] <- 70
+box = st_as_sfc(bb)
+
+# update ICES areas to box
+ices <-
+  st_intersection(ices, box) %>% 
+  left_join(iceshape::faolink)
+
+# generate country map
+world1 <- 
+  sf::st_as_sf(map('world', plot = FALSE, fill = TRUE, 
+                   regions = c("netherlands","belgium","France(?!:Corsica)",
+                               "ireland","united kingdom", "UK", "Denmark","Germany",
+                               "norway","iceland", "greenland", "faroe islands", "spain","portugal",
+                               "sweden", "finland","poland"))) %>% 
+  st_intersection(., box)
+
+# read the eez_iho interaction datasets
+eez_iho <-
+  read_sf("D:/Dropbox/ICES advice database/GIS/Intersect_EEZ_IHO_v3_2018/EEZ_IHO_v3.shp",
+          stringsAsFactors = FALSE) %>%
+  rename_all(., tolower) %>% 
+  st_intersection(., box) %>% 
+  
+  # generate INT code for international waters
+  mutate(iso_ter1 = ifelse(is.na(iso_ter1), "INT", iso_ter1)) %>% 
+  
+  # Convert EU countries to EU and make one area out of it
+  mutate(iso_ter1 = ifelse(iso_ter1 %in% c("BEL", "DEU","DNK", "ESP", "FIN","FRA", "GGY", "GBR", "HRV", "IRL",
+                                           "ITA", "JEY", "LTU","LVA", "NLD","POL","SVN", "SWE"), "EU", iso_ter1)) %>% 
[…]

### 2018-11-11 R/plot sag.r

@@ -1,0 +1,252 @@
+# -----------------------------------------------------------------------------------------------
+# Plot SAG
+# 02/02/2018 first coding
+# 08/02/2018 fixing lot's of small things in the excel spreadsheet
+# 03/09/2018 based on SAGfull
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# Load data
+# load(file=file.path(dropboxdir, "rdata/icesSAGfull 20180902.RData"))
+icesSAGfull <-
+  get(load(file=file.path(dropboxdir, "rdata/icesSAGfull 20181016.RData"))) %>% 
+  mutate(speciesfaocode = substr(stockkeylabel, 1, 3)) %>% 
+  left_join(iSpecies, by="speciesfaocode") %>% 
+  
+  # corrections
+  mutate(
+    catches        = ifelse(stockkeylabelold == "ghl-grn" & assessmentyear == 2018,catches/1000,catches ),
+    fisheriesguild = ifelse(stockkeylabelold == "whb-comb", "pelagic", fisheriesguild),
+    fisheriesguild = ifelse(speciesfaocode == "san", "demersal", fisheriesguild)
+  ) %>% 
+  
+  # remove duplicated rows
+  group_by(stockkey, assessmentyear, year, catches, landings, stocksize, fishingpressure, recruitment) %>% 
+  filter(row_number() == 1)
+
+# glimpse(icesSAGfull)
+
+
+
+# ----------------------------------------------------------------------------------
+# function for subplots
+# ----------------------------------------------------------------------------------
+
+history_plot <- function(mystock = "her-noss", myvar="stocksize", 
+                         firstyear = 2000, firstassessmentyear = 2010,
+                         lastassessmentyear = 2050, 
+                         firstrecruitmentyear = 2000) {
+  
+  d <-
+    icesSAGfull %>% 
+    group_by(stockkey, assessmentyear) %>% 
+    filter(assessmentkey == max(assessmentkey, na.rm=TRUE)) %>% 
[…]

### 2018-11-11 R/plots for pelagic forum.r

@@ -1,0 +1,673 @@
+# -----------------------------------------------------------------------------------------------
+# plots for pelagic forum
+#
+# 03/11/2017 first coding
+# 15/10/2018 updated for 2018 version of the talk
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+library(stringr)
+library(readxl)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load the data
+
+#load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iForecast.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))
+iSpecies <-
+  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
+  ungroup() %>% 
+  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild) 
+
+icesSAGfull <-
+  get(load(file=file.path(dropboxdir, "rdata/icesSAGfull 20181016.RData"))) %>% 
+  mutate(speciesfaocode = substr(stockkeylabel, 1, 3)) %>% 
+  left_join(iSpecies, by="speciesfaocode") %>% 
+  
+  # corrections
+  mutate(
+    catches        = ifelse(stockkeylabelold == "ghl-grn" & assessmentyear == 2018,catches/1000,catches ),
+    fisheriesguild = ifelse(stockkeylabelold == "whb-comb", "pelagic", fisheriesguild),
+    fisheriesguild = ifelse(speciesfaocode == "san", "demersal", fisheriesguild)
+  ) %>% 
+  
+  # remove duplicated rows
+  group_by(stockkey, assessmentyear, year, catches, landings, stocksize, fishingpressure, recruitment) %>% 
+  filter(row_number() == 1)
+
+  # %>%  dplyr::select(-c(grepl("custom", colnames(.))))
+
+
+# show catch in 2017
+icesSAGfull %>% 
[…]

### 2018-11-11 R/plots for relative stock trends.r

@@ -1,0 +1,161 @@
+# -----------------------------------------------------------------------------------------------
+# plots for pelagic forum
+#
+# 03/11/2017 first coding
+# 15/10/2018 updated for 2018 version of the talk
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+library(stringr)
+library(readxl)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/ICES Assessment database", sep="")
+
+# load the data
+
+#load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iForecast.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))
+# write.csv(iStockkey, file="iStockkey,csv", row.names = FALSE)
+
+iSpecies <-
+  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
+  ungroup() %>% 
+  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild) 
+
+icesSAGfull <-
+  get(load(file=file.path(dropboxdir, "rdata/icesSAGfull 20181016.RData"))) %>% 
+  mutate(speciesfaocode = substr(stockkeylabel, 1, 3)) %>% 
+  left_join(iSpecies, by="speciesfaocode") %>% 
+  
+  # corrections
+  mutate(
+    catches        = ifelse(stockkeylabelold == "ghl-grn" & assessmentyear == 2018,catches/1000,catches ),
+    fisheriesguild = ifelse(stockkeylabelold == "whb-comb", "pelagic", fisheriesguild),
+    fisheriesguild = ifelse(speciesfaocode == "san", "demersal", fisheriesguild)
+  ) %>% 
+  
+  # remove duplicated rows
+  group_by(stockkey, assessmentyear, year, catches, landings, stocksize, fishingpressure, recruitment) %>% 
+  filter(row_number() == 1)
+
[…]

### 2018-12-17 R/ICES advice and assessment summaries.Rmd
Updates 17/12/2018 (morning)

@@
   word_document:
-    reference_docx: ../PFA_report_template_v1.5.dotx
+    reference_docx: ../report_template_v1.5.dotx
 ---
-```{r, message=FALSE, warning=FALSE, include=FALSE}
-rm(list=ls())
-```
 
@@
 # 06/07/2018 first coding
+# 14/12/2018 updated for the new database structure
 # -----------------------------------------------------------------------------------------------
@@
 knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
-knitr::opts_chunk$set(fig.width=6) 
+knitr::opts_chunk$set(fig.width=10) 
 knit_hooks$set(crop = hook_pdfcrop)
@@
 library(lubridate)
-library(docxtractr) # read word documents
-library(icesSD)    # ICES Stock database
-library(icesSAG)   # ICES Stock Assessment Graphs
 
@@
 
-# set dropboxdirs
-assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
-advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# Load dataset
+load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
 
@@
 # =====================================================================================
-
-load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
-load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))
-
-# read and process the iAdvice data
-iAdvice <- 
-  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20180704.xlsx",sep=""), 
-             sheet     = "DATA",
-             col_names = TRUE,
-             col_types = "text",
-             skip      = 0)  %>% 
-  lowcase() %>% 
-  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
-             funs(as.integer)) %>% 
-  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
-               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
-               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
-               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
-  mutate( blim     = 1000 * blim, 
-          msybtrig = 1000 * msybtrig) %>%
-
-  mutate_at( c("stockices","speciestype", "assessmodel", 
-               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
-               "aggregated"), funs(tolower)) %>% 
-  rename(stockkeylabelold = stockices, 
-         assessmentyear = assessyear) %>% 
-  arrange(stockkeylabelold, tacarea, year) %>%  
-  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
-  left_join(iStockkey, by="stockkeylabelold") %>%
-
-  filter(adviceonstock == "y")
-
-# =====================================================================================
-# Load Excel and QCS data
-# =====================================================================================
-
-excelqcs <-
-  readxl::read_excel(
-    path= paste(assessdir, "/data/QCS and EXCEL Assessment Database combined.xlsx", sep=""), 
-    sheet = "data", col_names = TRUE, col_types = "text", trim_ws   = FALSE ) %>%
-  lowcase() %>% 
-  mutate_at(c("assessmentyear","year", 
-              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
-              "lowssb", "ssb", "highssb", 
-              "lowf","f","highf", "flandings","fdiscards",
-              "catches","discards", "landings"),   funs(as.numeric)) %>% 
-  mutate_at(c("stockkey"),    funs(as.integer)) %>% 
-  mutate_at(c("unitofrecruitment", "recruitmentdescription"),    funs(tolower)) %>% 
-  mutate(recruitment       = ifelse(unitofrecruitment == "thousands", recruitment/1000, recruitment),
-         unitofrecruitment = ifelse(unitofrecruitment == "thousands", "millions", unitofrecruitment),
-         recruitment       = ifelse(unitofrecruitment == "billions", recruitment*1000, recruitment),
-         unitofrecruitment = ifelse(unitofrecruitment == "billions", "millions", unitofrecruitment),
-         source            = "excel") %>% 
-  filter(purpose == "Advice")
-
-excelqcs_unique <-
-  excelqcs %>% 
-  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
-  filter(row_number()==1) %>% 
-  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
-  ungroup()
-
-# unique(excelqcs$purpose)
-# unique(icesSAG$purpose)
-# filter(excelqcs, stockkeylabelold == "bss-47") %>% View()
-
-# =====================================================================================
-# Load SAG database; also unpublished ones
-# =====================================================================================
-
-icesSAG <- 
-  get(load(file=file.path(assessdir, "rdata/icesSAG 20180703.RData"))) %>% 
-  lowcase() %>%
-  rename(stockkeylabel = fishstock) %>%
-  left_join(iRename, by="stockkeylabel") %>%
-  left_join(iStockkey, by="stockkey") %>% 
-  mutate_at(c("assessmentyear","year", 
-              "lowrecruitment", "recruitment", "highrecruitment", "recruitmentage",
-              "lowssb", "ssb", "highssb", 
-              "lowf","f","highf", 
-              "catches","discards", "landings"),   funs(as.numeric)) %>% 
-  mutate(source = "sag") %>% 
-  filter(purpose == "Advice")
-
-icesSAG_unique <-
-  icesSAG %>% 
-  group_by(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
-  filter(row_number()==1) %>% 
-  select(stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear) %>% 
-  ungroup()
-
-# =====================================================================================
-# Merge SAG and excel database 
-# =====================================================================================
-
-iAssess <-
-
-  # create excelqcs dataset to add
-  dplyr::anti_join(select(excelqcs_unique, stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear),
-                   select(icesSAG_unique,  stockkey, stockkeylabelold, stockkeylabelnew, assessmentyear)) %>% 
-  left_join(excelqcs, by=c("stockkey", "stockkeylabelold","stockkeylabelnew", "assessmentyear") ) %>% 
-  data.frame() %>%
-
-  # add icesSAG
-  bind_rows(., icesSAG)
-
-# glimpse(excelqcs_unique)
-# glimpse(icesSAG_unique)
-# glimpse(iAssess)
 
@@
     ungroup() %>% 
+    filter(purpose == "advice") %>% 
     filter(stockkeylabelold == stock,
@@
            lowrecruitment, recruitment, highrecruitment, 
-           lowf, f, highf, 
-           lowssb, ssb, highssb) %>%
+           lowfishingpressure, fishingpressure, highfishingpressure, 
+           lowstocksize, stocksize, highstocksize) %>%
     data.frame()
 
-  # plot ssb
+  # plot stocksize
   p1 <-
     d %>% 
-    filter(!is.na(ssb)) %>%
+    filter(!is.na(stocksize)) %>%
 
-    ggplot(aes(year,ssb, group=tyear)) +
+    ggplot(aes(year,stocksize, group=tyear)) +
 
     theme_publication() +
+
     theme(legend.title=element_blank(),
@@
           axis.text.y = element_text(size=9),
+          plot.title  = element_text(size=10), 
           # strip.background = element_blank(),
@@
     geom_ribbon(data=filter(d, type=="last"), 
-                aes(x=year, ymin=lowssb, ymax=highssb, fill=type), 
+                aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill=type), 
                 alpha=0.2) +
@@
     geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=msybtrig), 
+               aes(x=assessmentyear, yintercept=msybtrigger), 
                colour="red", linetype="dashed", inherit.aes=FALSE) +
     geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
-               aes( firstyear, msybtrig, label = "msy btrigger", vjust = -0.3, hjust = 0),
+               aes( firstyear, msybtrigger, label = "msy btrigger", vjust = -0.3, hjust = 0),
                colour="red", inherit.aes=FALSE) +
@@
     xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "SSB")
+    labs(x = NULL, y = NULL , title = "stocksize")
 
-  # F
+  # fishingpressure
   p2 <-
     d %>% 
-    filter(!is.na(f)) %>%
+    filter(!is.na(fishingpressure)) %>%
 
-    ggplot(aes(year,f, group=tyear)) +
+    ggplot(aes(year,fishingpressure, group=tyear)) +
 
@@
     geom_ribbon(data=filter(d, type=="last"), 
-                aes(x=year, ymin=lowf, ymax=highf, fill=type), 
+                aes(x=year, ymin=lowfishingpressure, ymax=highfishingpressure, fill=type), 
                 alpha=0.2) +
@@
     xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "F")
+    labs(x = NULL, y = NULL , title = "fishingpressure")
 
@@
     filter(stockkeylabelold == stock, 
-           year %in% firstadviceyear:lastadviceyear) %>%
-    mutate(advicebasis= substr(advicebasis,1,30),
+           tacyear %in% firstadviceyear:lastadviceyear,
+           adviceonstock == TRUE) %>%
+    mutate(advicebasis= substr(advicebasis,1,25),
 
-           advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
+           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
            advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),
+           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),
 
@@
 
-           catch      = ifelse(!is.na(icescatch), icescatch, NA),
-           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),
+           catch      = ifelse(!is.na(catches), catches, NA),
+           catchtype  = ifelse(!is.na(catches), "catches", ""),
 
-           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
-           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),
+           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
+           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),
 
@@
 
-           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
+           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
            tac_c      = 100 * (tac /lag(tac, n=1) - 1),
-           catch_c    = 100 * (catch / lag(catch, n=1) -1)
+           cat_c       = 100 * (catch / lag(catch, n=1) -1)
            ) %>%
 
-    select(year, advicebasis, advice, fadvmax, advicemax, tac, catch, 
-           advice_c, tac_c, catch_c,
+    select(tacyear, advicebasis, advice, fadvmax, advicemax, tac, catch, 
+           adv_c, tac_c, cat_c,
            advicetype, tactype, catchtype)
@@
   d %>% 
-  select(1, 2, 4, 5, 6, 7) %>%
+  select(1, 2, 4, 5, 6, 7, 8, 9, 10) %>%
   pandoc.table(., 
                style        = "simple",
-               split.tables = 100, 
+               split.tables = 200, 
                justify      = "right",
@@
 
-table_advice2 <- function(stock,firstadviceyear,lastadviceyear) {
-
-  iAdvice %>% 
-    filter(stockkeylabelold == stock, 
-           year %in% firstadviceyear:lastadviceyear) %>%
-    mutate(advice     = ifelse(!is.na(advisedcatch), advisedcatch, advisedlandings),
-           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicetype = ifelse(!is.na(advisedcatch), "catch", "landings"),
-
-           tac        = ifelse(!is.na(tac), tac, tal),
-           tactype    = ifelse(!is.na(tac), "catch", "landings"),
-
-           catch      = ifelse(!is.na(icescatch), icescatch, NA),
-           catchtype  = ifelse(!is.na(icescatch), "icescatch", ""),
-
-           catchtype  = ifelse(!is.na(iceslandings) & is.na(catch), "iceslandings", catchtype),
-           catch      = ifelse(!is.na(iceslandings) & is.na(catch), iceslandings, catch),
-
-           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
-           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),
-
-           advice_c   = 100 * (advicemax / lag(advicemax, n=1) - 1),
-           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
-           catch_c    = 100 * (catch / lag(catch, n=1) -1)
-           ) %>%
-
-    select(year, advice_c, tac_c, catch_c, advicetype, tactype, catchtype) %>% 
-    pandoc.table(., 
-                 style        = "simple",
-                 split.tables = 100, 
-                 split.cells  = c(rep(7,10)),
-                 justify      = "right",
-                 missing      =" ",
-                 big.mark     = '', 
-                 round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-}
-
-# table_advice(stock="her-47d3",firstadviceyear=2015,lastadviceyear=2019)
-
 ```
 
-**Summary of ICES advice and assessments**
-
-*June 2018 advice round*
+**Summary of ICES 2018 advice and assessments**
 
@@
 s   <- "her-47d3"
+fay <- 2010; lay <- 2018
+fy  <- 2000; ly  <- 2018
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+```
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+
+```
+
+Two weak year classes of North Sea herring are causing a decline in stock size and short-term prospects.
+The TAC and catches in 2018 have been substantially higher than the recommendations from ICES (600 000 instead of 518 000t).
+Due to the benchmark assessment of North Sea herring, the reference points for fishing mortality have been revised downwards.
+The combination of low recruitment, fishing harder than recommended and a lower Fmsy, lead to a substantially lower advice for 2019 (291 000 tonnes).
+
+The assessment of North Sea herring was benchmarked in February 2018.
+There have been some revisions in the model and in the data that is being used for North Sea herring.
+Most important it the way of handling natural mortality estimates.
+This is one of the few assessments where natural mortality is derived from the multispecies model of the North Sea.
+This multispecies model is rerun every three years leading  to new estimates of natural mortality for the whole time-series.
+Because of changes in the multispecies model and in the underlying data for predator species (e.g.
+hake, cod, saithe), the natural mortality estimates of herring have been revised upwards and downwards several times.
+This then also leads to rescaling of the estimates of stock size and fishing mortality.
+During the benchmark, a new method to standardize natural mortality estimates has been introduced which would hopefully circumvent re-scaling in the future?).
+However, this year there has been a revision of reference points again, leading to a lower Fmsy (0.26) compared to the previous estimate (0.33).
+Weak 2014 year class leads to substantially lower advice, which was already foreseen in 2017.
+
+[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.3a47d.pdf)
+
+## Atlanto-scandian herring
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+s   <- "her-noss"
 fay <- 2014; lay <- 2018
@@
 
-**History of stock assessments**
-
-```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
@@
 ```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
-```
-
-**Advice and management overview**
 
@@
 
+A new long-term management strategy for Atlanto-scandian herring has been evaluated and agreed.
+Fmsy and the fishing mortality target are now higher than the previous values being used.
+The advice for 2019 is based on following the new management plan, with a TAC of around 590 000 tonnes (compared to an advice of 384 000 tonnes that was advised for 2018).
+
+The stock assessment is consistent with last year's assessment.
+The stock appears to be slowly declining.
+Fishing mortality has risen sharply in the last year (due to the erroneous advice in 2016).
+There are signs of a reasonable sized incoming year class.
+The reason that the advice is substantially higher than last year is due to the new management plan that has a higher target fishing mortality.
+Last year, the advice was based on a target fishing mortality of 0.086, this year it is based on 0.14.
+
+[her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.1-24a514a.pdf)
+
+## Northeast Atlantic Mackerel
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+
+s   <- "mac-nea"
+fay <- 2014; lay <- 2018
+fy  <- 2000; ly  <- 2018
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
 ```
 
-**Comments**
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
-Assessment benchmarked in 2018.
-New natural mortality estimates lead to higher SSB and lower F.
-New method to standardize natural mortality estimates (no more scaling revision in future?).
-Revision of reference points.
-Weak 2014 year class leads to substantially lower advice, which was already foreseen in 2017.
-[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.3a47d.pdf)
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+
+```
+
+During the WGWIDE a press release came out on the Nordic swept-area survey that had shown that the biomass was 40% lower than the previous year.
+This survey result has an important effect on the stock assessment, as all the available indices now point to a lower stock size.
+This provides a much lower overall stock size compared to last year.
+Below is a comparison of the stock estimates of the current assessment (red line and uncertainty bands) in comparison the estimates of previous years.
+
+A major finding has been how sensitive the model results are to the inclusion of the RFID tags.
+Leaving out only the RFID tags, gives the perception of a much higher stock (see below).
+The RFID tags only  cover the most recent period.
+Compared to the old steel tags they exhibit a very strange phenomenon that the tag survival appears to be very low (~10%) compared to around 40% for the steel tags.
+Could it be that there is something wrong with the RFID tags or scanners in that they are missing fish?
+The problem is that the RFID tagging data are currently not assembled in a working group and are therefore not well evaluated.
+A proposal has been made to set up such a group.
+
+To deal with this problem of RFID tagging, an inter-benchmark has been set up which is leading up to a final meeting in the beginning of March 2019.
+
+[mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/mac.27.nea.pdf)
+
+## Western horse mackerel
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+s   <- "mac-nea"
+fay <- 2014; lay <- 2018
+fy  <- 2000; ly  <- 2018
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+```
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+
+```
+
+The stock assessment of Western horse mackerel is comparable to last year but there has been another revision upwards of the stock.
+This is especially relevant in conjunction with the biomass reference points that have been set after the benchmark in 2017 and where the biomass in 2015 (the last year in the time series) was selected as Blim.
+Because of the upward revisions in the assessments of last year and this year, the stock is now well above the absolute Blim and in fact close to MSY Btrigger even though the stock is at the lowest point in the time-series!
+Fishing mortality is estimated to be well below the Fmsy.
+The advice for 2019 is 145 000 tonnes, which is 24% higher than the advice for 2018.
+The increase in the advice is because the stock size has been revised upwards and because the recent catches have been taken with a fishing mortality which was lower than Fmsy.
+The advice is based on the application of Fmsy.
+
+The assessment of western horse mackerel is very uncertain and there has been frequent re-scaling of the assessment during every new assessment.
+This re-scaling is still happening.
+For western horse mackerel this results in the overall stock to be estimated higher than the year before.
+That is why the stock is now estimated to be close to MSY Btrigger while the stock is at the same time around the lowest point in the time series.
+The resulting advice is a direct application of the ICES advice rule to an assessment that could very well be too optimistic.
+
+[hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hom.27.2a4a5b6a7a-ce-k8.pdf)
+
+## Western horse mackerel
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+s   <- "whb-comb"
+fay <- 2014; lay <- 2018
+fy  <- 2000; ly  <- 2018
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+```
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+
+```
+
+The stock size of blue whiting was high in 2018 but is predicted to be substantially lower in 2019 due to low recruitments coming in.
+The stock size of blue whiting appears to have been overestimated.
+Catches have been well above 1 Million tonnes over the last few years and above the scientific advice.
+Fishing mortality is estimated to be above Fmsy.
+The advice for 2019 is 1 144 000 tonnes, which is 18% lower than the advice for 2018.
+
+The blue whiting survey went well in 2018 and showed the highest biomass of the time series.
+Despite the high catches, the stock is still at a very high level although a decline does appear to have started.
+The most recent year classes of blue whiting are very weak so that decline in catching opportunities in the further years is foreseen.
+
+[whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/whb.27.1-91214.pdf)
 
@@
 
-**History of stock assessments**
-
-```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
-```
-
-**Advice and management overview**
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
 table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
 ```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
 
@@
 An additional analysis will be undertaken (by MP) to evaluate alternative options for the zero catch advice.
+
 [her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.20-24.pdf)
@@
 
-**History of stock assessments**
-
-```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
 ```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-cat("Most recent assessment in red (with 95% conf. intervals). Earlier assessments is black.")
-```
-
-**Advice and management overview**
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 Substantially lower estimate of SSB and higher estimate of fishing mortality in the most recent assessment is probably due to the variability in the survey index.
 The 2014 year class is estimated to be strong.
+
 [her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.25-2932.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 The advice for a zero TAC is the same as in the previous years.
@@
 The monitoring TAC is still mentioned under issues relevant for the advice.
+
 [her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.6a7bc.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
+Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea.
 
-**Comments**
-
-Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea. 
 [her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.nirs.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year.
@@
 Therefore the advice is according to the MSY approach.
+
 [her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.irls.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 The change in recruitment level between this year’s assessment and the previous years is caused by a recent update of natural mortality of sprat in the first six months of their life.
@@
 The assessment has a tendency to overestimate above-average recruitments.
+
 [spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.4.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
+Advice according to the Baltic Multiannual Plan (MAP).
 
-**Comments**
-
-Advice according to the Baltic Multiannual Plan (MAP). 
 [spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.22-32.pdf)
@@
 
-**History of stock assessments**
-
-```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
 ```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-cat("Last assessment was in 2017. Biennial assessment based on survey trends")
-```
-
-**Advice and management overview**
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
+Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019.
 
-**Comments**
-
-Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019. 
 [aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
 In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
+
 [hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hke.27.3a46-8abd.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
@@
 This also leads to a substantially lower advice for 2019 compared to the advice for 2018.
+
 [cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/cod.27.47d20.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
@@
 The advice is based on the MSY approach.
+
 [ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/ple.27.420.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
+Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea.
 
-**Comments**
-
-Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
 [sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/sol.27.4.pdf)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice2(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-**Comments**
-
 ICES has reissued the advice for 2018 and issued an advice for 2019.
@@
 However, given the current low fishing mortality and the above-average recruitment in 2013 and 2014, SSB in 2019 and 2020 is predicted to be above Blim, allowing for some catches in 2018 and 2019.
+
 [bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/bss.27.4bc7ad-h.pdf)

### 2018-12-17 R/Analyse scale differences.r

@@ -27,7 +27,7 @@ dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
 load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
 
-d.stockdefiniions <-
+d.stockdefinitions <-
   
   iAdvice %>% 
   filter(grepl("ple-n|her-47|mac-67|mac-nea|mac-west|cod-34|cod-nsea|whb-c|whb-n|sol-ns|hom-c|hom-w|had-34|had-n|hke-n", 
@@ -284,14 +284,6 @@ d.merged2 <-
   mutate(n=n()) %>% 
   filter(n() == 2) 
 
-
-
-d.merged  %>%   filter(variable == myvar, grepl(myspecies, stockkeylabelold), 
-                       assessmentpair == "2017-2018") %>% View()
-
-d.merged2 %>%   filter(variable == myvar, grepl(myspecies, stockkeylabelold), 
-                       assessmentpair == "2017-2018") %>% View()
-
 # create dataset of means
 d.means <-
   d.merged2 %>% 
@@ -302,9 +294,6 @@ d.means <-
   summarise(value = mean(value, na.rm=TRUE)) %>% 
   gather(key=tmp, value=year, miny:maxy) %>% 
   arrange(stockkeylabelold, assessmentpair, variable, assessmentyear, label)
-
-d.means %>%   filter(variable == myvar, grepl(myspecies, stockkeylabelold), 
-                       assessmentpair == "2017-2018") %>% View()
 
 # calculate scaling parameters per stock and per year
 d.year <- 
@@ -360,14 +349,6 @@ d.scalebydecade %>%
 
 
 
-# ---------------------------------------------------------------------------------------------
-# plot between differences in stock definition
-# ---------------------------------------------------------------------------------------------
-
-d.stockdefinitionschange %>% 
-  left_join(d.year, by=c("stockkeylabelold", "assessmentyear")) %>% 
-  filter(!is.na(assessmentpair)) %>% 
-  filter(variable == "stocksize") 
   
 
 # ---------------------------------------------------------------------------------------------
@@ -408,7 +389,7 @@ d.merged %>%
              aes(colour=factor(label)), size=1) +
   
   scale_y_continuous(labels=scientific_format(digits=2)) +
-  scale_x_continuous(breaks = pretty_breaks()) +
+  scale_x_continuous(breaks = pretty_breaks(n=3)) +
[…]

### 2019-01-15 R/Analyse changes in pelagic stocks.Rmd
Updates 15/1/2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)
library(directlabels)  # for printing labels at end of geom lines

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# create subset for advice
d <-
  iAdvice %>%

  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss", 
               stockkeylabelold)) %>%
  filter(!grepl("benchmark", purpose)) %>% 
  filter(assessmentyear >= 2008) %>% 
  ungroup()

# create subset for assessments
e <-
  iAssess %>% 
  filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
  filter(assessmentyear == 2018) %>% 
  filter(year >= 2008) %>% 
  filter(year <= assessmentyear) %>% 
  filter(adviceonstock == TRUE)

```

**Analysis of advice changes in a number of key pelagic stocks**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**Introduction**

In this document an overview is presented of the ICES advice for a number of key pelagic stocks relevant to the Pelagic Advisory Council (her-347d, her-noss, her-3a22, hom-west, mac-nea, whb-comb). The analysis is based on ICES advice database (Pastoors et al, in prep.) and the ICES stock assessment graph database ( [ICES SAG](www.http://ices.dk/marine-data/tools/Pages/stock-assessment-graphs.aspx)). The document outlines changes in reference points, changes in advice and a comparison between stock size assumed as the basis for the forecast and the most recent estimates of stock size from the 2018 assessments.

**Changes in F reference points**

Fpa (blue) and Fmsy (green) reference points points as estimated in the different assessment years. Note that for a number of years it was no longer deemed necessary to provide Fpa estimates, and they were consequently removed from the advice documents. In the more recent years, Fpa values have been added again on the request of ICES clients.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>%

  ggplot(aes(x=assessmentyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_line(aes(y=value, group=var, colour=factor(var))) +
  geom_point(aes(y=value, group=var, colour=factor(var))) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="F") +
  scale_color_manual(values = c("green", "blue")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

```

**Changes in biomass points**

Blim (red) and MSY Btrigger (green) reference points points as estimated in the different assessment years.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
  gather(key=var, value=value, blim:msybtrigger) %>%

  ggplot(aes(x=assessmentyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_line(aes(y=value, group=var, colour=factor(var))) +
  geom_point(aes(y=value, group=var, colour=factor(var))) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="SSB") +
  scale_color_manual(values = c("red", "green")) +
  facet_wrap(~stockkeylabelold, scales="free_y")
```

**Changes in advice**

Recommended catches (in fact: max of advised landings) presented for each of the different TAC years. The red dots and line refer to the most recent published advice, the blue dots refer to replaced advices, due to errors or revisions in the stock assessments.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>% 
  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_point(aes(y       = advisedlandingsmax, 
                 colour  = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(aes(y        = advisedlandingsmax, 
                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'),
                linetype = ifelse(adviceonstock==TRUE,'advice',"replaced") )) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advice Landings Max (tonnes)") +
  scale_color_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c("solid", "blank")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

```

**Comparing perceived stock sizes with most recent estimates**

Comparison of the most recent estimates of stock size from the 2018 assessments (black line) with the SSB values assumed as part of the forecasts in each of the subsequent years. The red dots refer to the most recent published advice, the blue dots refer to replaced advices, due to errors or revisions in the stock assessments.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, ssbay) %>% 
  rename(year = assessmentyear) %>%

  ggplot(aes(x=year)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_point(aes(y=ssbay,
                 colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(data=e,
            aes(y = stocksize )) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="SSB") +
  scale_color_manual(values = c("red", "blue")) +
  facet_wrap(~stockkeylabelold, scales="free_y")
```

**References**

Pastoors, M.A. et al (*in prep*) Documenting the past of scientific advice: the ICES advice database.

**Advice tables**

The tables below are the summaries of advice for the stocks mentioned in this paper. Most of the columns are self-explanatory. However, the columns 'purpose' and 'advice' have a specific meaning: purpose refers to the purpose of the advice, either used for advice (purpose is 'advice' and advice is TRUE) or the advice has been replaced or withdrawn (purpose is 'replaced' and advice is FALSE). The column SSB_STF refers to the SSB as estimated in the short term forecasts in the diferent years. All quantities are in tonnes, except for fishing mortality which is in 1/year.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  pandoc.table(., 
             style        = "simple",
             split.tables = 180, 
             justify      = "right",
             missing      =".",
             big.mark     = '', 
             round        = c(0,0,0,2,2,0,0,0,0,2,0,0,0,0,0,0,0,0,0,0) )

```

```{r eval=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}

d %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  write.csv(., file="advice.csv")

```

### 2019-01-15 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Updates 15/1/2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
bibliography: paper_02_scalejumps.bib
csl: ices-journal-of-marine-science.csl
link-citations: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 13/01/2019 now only figures and tables
# 15/01/2019 added scale jump figure with explanatory variables
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)
library(directlabels)  # for printing labels at end of geom lines

# citations library
library(knitcitations) # use citations in rmarkdown
options("citation_format" = "pandoc")

# captioner library
library(captioner) # deal with captions
tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# Create subset of series (stocks) to be used for analysis
d <-
  iAssess %>%

  # Exclude prediction years
  filter(year <= assessmentyear) %>%

  # Define start year
  filter( year >= 1980) %>%

  # Only assessments with stock size in tonnes
  filter(stocksizeunits  == "tonnes") %>%  # Note: this is filtering on absolute vs relative; but not really

  # filter stocks
  filter(grepl("ple-n|her-47|mac-67|mac-nea|mac-west|cod-34|cod-nsea|whb-c|whb-n|sol-ns|hom-c|hom-w|had-3|had-n|hke-n", stockkeylabelold)) %>%

  mutate(
    stock = as.character(NA),
    stock = ifelse(grepl("ple-n",stockkeylabelold), "ple (north sea)",stock),
    stock = ifelse(grepl("her-4",stockkeylabelold), "her (north sea)",stock),
    stock = ifelse(grepl("mac-67|mac-nea|mac-west",stockkeylabelold), "mac (west)",stock),
    stock = ifelse(grepl("cod-34|cod-nsea",stockkeylabelold), "cod (north sea)",stock),
    stock = ifelse(grepl("whb-c|whb-n",stockkeylabelold), "whb (comb)",stock),
    stock = ifelse(grepl("sol-ns",stockkeylabelold), "sol (north sea)",stock),
    stock = ifelse(grepl("hom-c|hom-w",stockkeylabelold), "hom (west)",stock),
    stock = ifelse(grepl("had-3|had-n",stockkeylabelold), "had (north sea)",stock),
    stock = ifelse(grepl("hke-n",stockkeylabelold), "hke (north)",stock)
  ) %>%

  arrange(stock, purpose, desc(assessmentyear)) %>%

  # filter(purpose         == "advice") %>%

  select(assessmentyear, year, stock, stockkeylabelold, purpose, benchmark,
         stocksize, recruitment, fishingpressure) %>% 
  distinct() %>% 
  gather(key=variable, value=value, stocksize:fishingpressure) %>% 
  filter(!is.na(value), value != 0.0)  %>%

  # Keep this variable
  filter( variable %in% c("stocksize")) %>%

  arrange(stock, variable, desc(assessmentyear)) %>% 
  ungroup() %>%

  mutate(plot=ifelse(assessmentyear == max(assessmentyear), "last", "other"))

distinct(iAssess, stockkeylabelold, assessmentyear, purpose, benchmark) %>% filter(grepl("ple-n", stockkeylabelold)) %>% View()

# write.csv(distinct(d,stock), "temp.csv")

# Data handling ===========================================================================

# select data
d.selection <-
  d %>%
  filter(purpose == "advice")

# benchmark data
d.benchmarks <-
  d %>% 
  filter(benchmark == TRUE) %>% 
  distinct(stock, stockkeylabelold, assessmentyear, purpose, .keep_all = FALSE)

# assessment data
d.assessments <-
  d.selection %>% 
  distinct(stock, stockkeylabelold, assessmentyear, purpose, .keep_all = FALSE) %>%

  left_join(iAdvice, by=c("stockkeylabelold","assessmentyear", "purpose")) %>%
  mutate(assessmentmodel = tolower(assessmentmodel)) %>%

  distinct(stock, stockkeylabelold, assessmentyear, purpose, benchmark, assessmentmodel, 
           ncpueseries, nsurveyseries, .keep_all = FALSE) %>%

  # determine stock definition change and model definition change
  group_by(stock) %>% 
  arrange(stock, assessmentyear) %>% 
  mutate(assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel), 1, 0),
         assessmentmodelchange = ifelse(is.na(assessmentmodelchange), 0, assessmentmodelchange)) %>%

  group_by(stock) %>% 
  arrange(stock, assessmentyear) %>% 
  mutate(stockdefinitionchange = ifelse(stockkeylabelold != lag(stockkeylabelold), 1, 0),
         stockdefinitionchange = ifelse(is.na(stockdefinitionchange), 0, stockdefinitionchange))

# most recent assessments
d.recent <-
  d.selection %>% 
  filter(assessmentyear == 2018) %>% 
  ungroup() %>% 
  select(year, stock, purpose, variable, value_last=value)

# segments to endpoints
d.segment <-
  d.selection %>% 
  filter( (assessmentyear != 2018 & year == max(year)) ) %>%

  # add the most recent assessment data
  left_join(d.recent, by=c("year", "stock", "purpose", "variable")) %>% 
  arrange(stock, desc(year))

# endpoints
d.points <-
  d.selection %>% 
  filter( (assessmentyear != 2018 & year == max(year)) )

# CALCULATE IN ONE

# Mohn's and Jonsson's metrics all compared to the most recent assessment

d.mohnjonsson <-

  d.selection %>%

  group_by(stock, assessmentyear, variable) %>% 
  arrange(stock, variable, desc(assessmentyear)) %>%

  # remove most recent year 2018 and keep only endpoints
  filter( (assessmentyear != 2018 & year == max(year)) ) %>%

  filter( variable == "stocksize") %>% 
  filter( year >= 1980) %>%

  # add the most recent assessment data
  left_join(d.recent, by=c("year", "stock", "purpose", "variable")) %>%

  # calculate metrics
  mutate(jonsson_ab  = log(value/value_last),
         mohn_rho    = ((value - value_last) / value_last),
         mohn_rhoabs= abs(mohn_rho)) %>%

  group_by(stock, variable) %>% 
  mutate(jonsson_dev  = (log(value/value_last) - mean(jonsson_ab)),
         jonsson_dev2 = (jonsson_dev)^2,
         jonsson_asd  = abs(jonsson_dev)) %>%

  mutate(calctype = "in_one")

# Calculations by stock

d.mohnjonsson.bystock <-
  d.mohnjonsson %>%

  group_by(stock, variable, calctype) %>% 
  summarize(jonsson_ab   = mean(jonsson_ab),
            jonsson_asd  = sqrt(mean(jonsson_dev2)),
            mohn_rhoabs = mean(mohn_rhoabs),
            mohn_rho     = mean(mohn_rho), 
            firstyear    = min(year),
            lastyear     = max(year))

# Calculate Ralston's retrospective measures

d.ralston <-

  d.selection %>%

  arrange(stock, year, assessmentyear) %>% 
  group_by(stock, year) %>%

  filter(n() >= 2) %>%

  mutate(logB     = log(value),
         meanlogB = mean(log(value)),
         logBdev  = logB - meanlogB,
         logBdev2 = (logBdev)^2,
         n        = n(),
         n_1      = n - 1) %>%

  # now do summing over years
  mutate(sumlogBdev2 = sum(logBdev2) ) %>%

  # now summarize by year
  summarize(
    meanlogBdev = mean(logBdev),
    sumlogBdev2 = sum(logBdev2),
    n_1         = mean(n_1), 
    ralston_sigma = sqrt(sumlogBdev2/n_1)) %>%

  mutate(calctype = "in_one")

# Calculate by stock

d.ralston.bystock <-
  d.ralston %>%

  group_by(stock, calctype) %>%
  distinct(stock, year, n_1, sumlogBdev2, .keep_all = FALSE) %>% 
  summarize(
    sumlogBdev2all = sum(sumlogBdev2),
    n_1            = sum(n_1), 
    firstyear    = min(year),
    lastyear     = max(year)
  ) %>%

  # calculate overall metrics
  mutate(
    ralston_sigma = sqrt(sumlogBdev2all / n_1),
    cv            = percent(sqrt(exp(ralston_sigma^2)-1))
  )

# CALCULATE IN PEELS

# create meta data on peels
d.peels <-
  d.selection %>% 
  distinct(stock, assessmentyear, variable) %>% 
  group_by(stock, variable) %>% 
  arrange(stock, variable, assessmentyear) %>% 
  mutate(assessmentyear_0   = assessmentyear,
         assessmentyear_1   = lag(assessmentyear, n=1),
         assessmentyear_2   = lag(assessmentyear, n=2),
         assessmentyear_3   = lag(assessmentyear, n=3),
         assessmentyear_4   = lag(assessmentyear, n=4),
         assessmentyear_5   = lag(assessmentyear, n=5),
         assessmentyear_6   = lag(assessmentyear, n=6),
         assessmentyear_7   = lag(assessmentyear, n=7),
         assessmentyear_8   = lag(assessmentyear, n=8),
         assessmentyear_9   = lag(assessmentyear, n=9),
         assessmentyear_10   = lag(assessmentyear, n=10)
         ) %>% 
  rename(assessmentyearbase = assessmentyear) %>%
  gather(key=peel, value=assessmentyear, assessmentyear_0:assessmentyear_10) %>% 
  mutate(peel = as.integer(str_extract(peel, "[[:digit:]]+"))) %>% 
  arrange(stock, assessmentyearbase, variable, peel, assessmentyear) %>% 
  filter(!is.na(assessmentyear)) %>%

  group_by(stock, assessmentyearbase, variable) %>%
  mutate(maxpeel = max(peel)) %>% 
  ungroup()

# Create datasets with peels

d.inpeels <-
  d.peels %>% 
  left_join(d.selection, by=c("stock", "assessmentyear", "variable")) %>%
  arrange(stock, desc(assessmentyearbase), variable, desc(year), peel) %>% 
  group_by(stock, assessmentyearbase, variable, peel)  %>% 
  arrange(stock, variable, desc(assessmentyearbase), peel, desc(assessmentyear), desc(year))

# settings for number of years in peels

minnumberofyearsinpeel <- 5
maxnumberofyearsinpeel <- 5

# Calculate Mohn and Jonsson's retrospective metrics in peels

d.mohnjonsson_inpeels <-

  d.inpeels %>%

  filter( (peel != 0 & year == max(year)) ) %>% 
  filter(peel <= maxnumberofyearsinpeel, maxpeel >= minnumberofyearsinpeel) %>%

  # add the most recent estimates of the specific endyears
  left_join(select(d.selection, stock, assessmentyearbase=assessmentyear, year, purpose, 
                   variable, value_last=value),
            by=c("stock","assessmentyearbase","year","purpose","variable")) %>%

  group_by(stock, assessmentyearbase, variable, year) %>% 
  arrange(stock, desc(assessmentyearbase), variable, desc(year)) %>% 
  mutate(jonsson_ab  = log(value/value_last),
         mohn_rho    = ((value - value_last) / value_last),
         mohn_rhoabs= abs(mohn_rho)) %>%

  group_by(stock, assessmentyearbase, variable) %>% 
  mutate(jonsson_dev  = (log(value/value_last) - mean(jonsson_ab)),
         jonsson_dev2 = (jonsson_dev)^2 ,
         jonsson_asd  = abs(jonsson_dev)) %>%

  summarize(jonsson_ab   = mean(jonsson_ab),
            jonsson_asd  = sqrt(mean(jonsson_dev2)),
            mohn_rhoabs = mean(mohn_rhoabs),
            mohn_rho     = mean(mohn_rho) ) %>%

  mutate(calctype = "in_peels")

# Average by stock
d.mohnjonsson_inpeels.bystock <-
  d.mohnjonsson_inpeels %>% 
  group_by(stock, variable, calctype) %>% 
  summarize(jonsson_ab   = mean(jonsson_ab),
            jonsson_asd  = mean(jonsson_asd),
            mohn_rhoabs = mean(mohn_rhoabs),
            mohn_rho     = mean(mohn_rho), 
            firstyear    = min(assessmentyearbase),
            lastyear     = max(assessmentyearbase) )

# Calculate adapted Ralston's retrospective measures in peels

d.ralston_inpeels <-

  d.inpeels %>%

  # filter(year >= assessmentyearbase - 20) %>%
  filter(year >= assessmentyearbase - maxnumberofyearsinpeel, maxpeel >= minnumberofyearsinpeel) %>%

  group_by(stock, assessmentyearbase, variable, year) %>%
  arrange(stock, desc(assessmentyearbase), desc(year) ) %>%

  filter(n() >= 2) %>%

  mutate(logB     = log(value),
         meanlogB = mean(log(value)),
         logBdev  = logB - meanlogB,
         logBdev2 = (logBdev)^2,
         n        = n(),
         n_1      = n - 1) %>%

  # now do summing over years by assessmentyearbase
  mutate(sumlogBdev2 = sum(logBdev2) ) %>%

  # summarize over years
  summarize(
    meanlogBdev = mean(logBdev),
    sumlogBdev2 = sum(logBdev2),
    n_1         = mean(n_1), 
    avglogBdev2 = sumlogBdev2/n_1, 
    ralston_sigma = sqrt(sumlogBdev2/n_1)) %>%

  group_by(stock, assessmentyearbase, variable) %>%
  distinct(stock, assessmentyearbase, variable, year, n_1, sumlogBdev2, .keep_all = FALSE) %>% 
  summarize(
    sumlogBdev2all = sum(sumlogBdev2),
    n_1            = sum(n_1)) %>%

  # calculate overall metrics
  mutate(
    ralston_sigma = sqrt(sumlogBdev2all / n_1),
    cv            = percent(sqrt(exp(ralston_sigma^2)-1))) %>%

  mutate(calctype = "in_peels")

# Average over assessmentyearbase
d.ralston_inpeels.bystock <-
  d.ralston_inpeels %>% 
  group_by(stock, variable, calctype) %>% 
  summarize(ralston_sigma = mean(ralston_sigma), 
  firstyear               = min(assessmentyearbase),
  lastyear                = max(assessmentyearbase))

# combine the metric by year
d.metrics_annual <-
  bind_rows(d.mohnjonsson, d.ralston) %>% 
  bind_rows(select(d.mohnjonsson_inpeels, 
                   year = assessmentyearbase, stock, variable, calctype,
                   jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs)) %>% 
  bind_rows(select(d.ralston_inpeels,
                   year = assessmentyearbase, stock, variable, calctype,
                   ralston_sigma)) %>% 
  ungroup() %>% 
  select(calctype, stock, year, jonsson_ab, jonsson_asd, mohn_rho, ralston_sigma) %>% 
  gather(key=metric, value=value, jonsson_ab:ralston_sigma) %>%
  filter(!is.na(value)) %>% 
  separate(metric, into=c("author", "metric"), by="_") %>%
  group_by(calctype, stock, author, metric, year) %>% 
  mutate(n=n(),
         stock2 = substr(stock,1,3))

# combine the metrics by stock
d.metrics_bystock <-
  bind_rows(d.mohnjonsson.bystock, d.ralston.bystock) %>% 
  bind_rows(select(d.mohnjonsson_inpeels.bystock, 
                   stock, variable, calctype, firstyear, lastyear, 
                   jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs)) %>% 
  bind_rows(select(d.ralston_inpeels.bystock,
                   stock, variable, calctype, firstyear, lastyear, 
                   ralston_sigma)) %>% 
  ungroup() %>% 
  select(calctype, stock, firstyear, lastyear, jonsson_ab, jonsson_asd, mohn_rho, ralston_sigma) %>% 
  gather(key=metric, value=value, jonsson_ab:ralston_sigma) %>%
  filter(!is.na(value)) %>% 
  separate(metric, into=c("author", "metric"), by="_") %>%
  group_by(calctype, stock, author, metric) %>% 
  mutate(stock2 = substr(stock,1,3))

# Calculations of scale-differences and scale jumps ==========================================

removelastyears <- 0
numberofyears   <- 10
threshold       <- 0.1   # threshold for scale differences

# Create a set of assessment pairs: two subsequent assessments  
d.pairs <-
  d.selection %>% 
  distinct(stock, assessmentyear, variable) %>% 
  group_by(stock, variable) %>% 
  arrange(stock, variable, assessmentyear) %>% 
  mutate(assessmentyearprior = lag(assessmentyear),
         assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
  rename(assessmentyearlast  = assessmentyear) %>% 
  filter(!is.na(assessmentyearprior))

# data series with most recent (last) estimation
d.last <-
  d.pairs %>% 
  select(-assessmentyearprior) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="last") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation
d.prior <-
  d.pairs %>% 
  select(-assessmentyearlast) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="prior") %>% 
  rename(assessmentyear = assessmentyearprior)

d.jumps <-
  bind_rows(d.last, d.prior) %>% 
  group_by(stock, assessmentpair, variable) %>%

  # filter the appropriate period
  filter(year <= max(assessmentyear) - removelastyears,
         year >= max(assessmentyear) - removelastyears - numberofyears) %>%

  # sort
  arrange(stock, assessmentpair, assessmentyear, year) %>%

  # filter only years that are in common in last and prior
  group_by(stock, assessmentpair, variable, year) %>% 
  mutate(n = n()) %>% 
  filter(n() == 2) %>%

  # calculate means per period
  group_by(stock, assessmentpair, variable) %>% 
  mutate(   firstyear      = min(year),
            lastyear       = max(year)) %>% 
  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
  summarise(value = mean(value, na.rm=TRUE)) %>% 
  gather(key=tmp, value=year, firstyear:lastyear) %>% 
  arrange(stock, assessmentpair, variable, assessmentyear, label) %>%

  # calculate metrics
  filter(tmp=="firstyear") %>% 
  ungroup() %>% 
  select(stock, assessmentpair, variable, label, value) %>% 
  spread(key=label, value=value) %>% 
  mutate(phi            = (prior-last)/last,
         phi_abs        = abs(phi),
         assessmentyear = as.integer(substr(assessmentpair, 6,9)))

# calculate average jump per stock
d.jumps.bystock <-
  d.jumps %>% 
  group_by(stock, variable) %>% 
  summarize(
    pastoors_njumps    = sum(phi_abs > threshold),
    pastoors_percjumps = pastoors_njumps / n(),
    pastoors_phi       = mean(phi),
    pastoors_phiabs    = mean(phi_abs), 
    firstyear          = min(assessmentyear),
    lastyear           = max(assessmentyear),
    calctype           = "in_peels")

# jumps by decade
d.jumps.bydecade <-
  d.jumps %>% 
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  group_by(stock, variable, decade) %>% 
  summarise(n         = n(),
            n2        = sum(phi_abs > 0.1, na.rm=TRUE),
            phi     = mean(phi, na.rm=TRUE),
            phi_abs = mean(phi_abs, na.rm=TRUE)
  ) %>%
  mutate(phi_abs    = percent_format(accuracy=1)(phi_abs),
         ratio        = paste(n2,n,sep="/"),
         ratio2       = percent_format(accuracy=1) (n2/n),
         ratio3       = paste(ratio2," (", n, ")", sep="") )

# update d.metrics_bystock
d.metrics_bystock2 <-
  bind_rows(d.mohnjonsson.bystock, d.ralston.bystock) %>% 
  bind_rows(select(d.mohnjonsson_inpeels.bystock, 
                   stock, variable, calctype, firstyear, lastyear, 
                   jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs)) %>% 
  bind_rows(select(d.ralston_inpeels.bystock,
                   stock, variable, calctype, firstyear, lastyear, 
                   ralston_sigma)) %>% 
  bind_rows(select(d.jumps.bystock,
                   stock, variable, calctype, firstyear, lastyear, 
                   pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs)) %>% 
  ungroup() %>% 
  select(calctype, stock, firstyear, lastyear, jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs, ralston_sigma,
         pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs) %>% 
  gather(key=metric, value=value, jonsson_ab:pastoors_phiabs) %>%
  filter(!is.na(value)) %>% 
  separate(metric, into=c("author", "metric"), by="_") %>%
  group_by(calctype, stock, author, metric) %>% 
  mutate(stock2 = substr(stock,1,3))

# Compare scale jumps when stock definition has changed.

t <-
  d %>% 
  select(assessmentyear, stock, stockkeylabelold) %>% 
  distinct() %>% 
  mutate(
    stock = factor(stock),
    stock = factor(stock, levels=rev(levels(stock)))) %>% 
  group_by(stock, stockkeylabelold) %>% 
  summarize(
    firstyear = min(assessmentyear),
    lastyear  = max(assessmentyear)) %>% 
  arrange(stock, firstyear) %>% 
  group_by(stock) %>% 
  mutate(
    previous       = lag(lastyear),
    assessmentpair = paste(previous, firstyear, sep="-"),
    stockold       = lag(stockkeylabelold)) %>% 
  filter(!is.na(previous))  %>% 
  select(stock, assessmentpair)

t1 <-
  t %>% 
  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
  select(-assessmentyearprior) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
  mutate(label="after") %>% 
  rename(assessmentyear = assessmentyearlast)

# data series with previous (prior) estimation
t2 <-
  t %>% 
  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
  select(-assessmentyearlast) %>% 
  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
  mutate(label="before") %>% 
  rename(assessmentyear = assessmentyearprior)

# combine
t3 <-
  bind_rows(t2, t1) %>% 
  group_by(stock, assessmentpair, variable) %>%

  # filter the appropriate period
  filter(year <= max(assessmentyear) - removelastyears,
         year >= max(assessmentyear) - removelastyears - numberofyears) %>%

  # sort
  arrange(stock, assessmentpair, assessmentyear, year) %>%

  # filter only years that are in common in last and prior
  group_by(stock, assessmentpair, variable, year) %>% 
  mutate(
    n = n(),
    header = paste(stock, assessmentpair, sep=" ")
  ) %>% 
  filter(n() == 2)

# Captions used in this document ==========================================================

# section 1 -----------------------------------------------------

tab_nums(name    = "section1", level = 1, display = FALSE)

fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")

fig_nums(name    = "differentmethods", level = 2, display = FALSE, 
         caption = "Comparison of three different types of retrospective metrics. t refers to the year for a stock metric X, i refers to the assessment year. Diagram A is applicable to Mohn’s rho and Jonsson and Hjorleifsson’s retrospective metrics while Ap refers to the same metrics calculated in peels (in this example of 4 years), B refers to Ralston’s et al metrics and Bp to the same metrics in peels and C refers to the scale differences calculated in this example over 4 years (this paper)")

fig_nums(name    = "retroplots", level = 2, display = FALSE, 
         caption = "Historic retrospective analysis of the nine stocks analysed. For each stock, the most recent assessment is in red and the previous assessments in black. The end point of each assessment is indicated by a dot. Stock definitions may have changed over time, but this is not included in the graphs")

# section2 -----------------------------------------------------

tab_nums(name    = "section2", level = 1, display = FALSE)

fig_nums(name    = "retrometrics", level = 1, display = FALSE, 
         caption = " Retrospective metrics for 9 species derived from historical assessment records. Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011. In red: calculated values based on the whole time-series of historical data. In blue: based on retrospective peels of 5 years. Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")

fig_nums(name    = "retrometrics2", level = 2, display = FALSE, 
         caption = "Summary of retrospective metrics for 9 species derived from historical assessment records. Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011. In red: calculated values based on the whole time-series of historical data. In blue: based on retrospective peels of 5 years. Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")

# section 3 -----------------------------------------------------

tab_nums(name    = "scalejumps", level = 1, display = FALSE, 
         caption = "Percentage assessments with scale jumps of more than 10% expressed by decade. Between brackets: number of assessments within the decade.")

fig_nums(name    = "scalejumps", level = 1, display = FALSE, 
         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records. Results from pairwise comparisons over periods of 8 years. Large dots  refer to years when the scale difference is larger than 10%. Numbers at the bottom left of each panel refer to the number of assessments with a scale difference larger than 10% relative to the total number of assessments available for that stock. Colours indicate the decades. Black triangles and dashed lines indicate benchmark assessment years. Red triangles and dashed lines indicate years when assessment models were changed compared to the year before")

fig_nums(name    = "metricsinpeels", level = 2, display = FALSE, 
         caption = "Comparison of retrospective uncertainty metrics from historical assessment records, calculated in peels, normalized between 0 and 1. Asd from Jonsson & Hjorleifsson 2000, rhoabs as absolute value of rho from Mohn 1999, sigma from Ralston et al 2011, percjumps as the percentage of jumps (this paper) and phiabs as the absolute value of phi (this paper).")

# section 4 -----------------------------------------------------

fig_nums(name    = "stockdefchange", level = 1, display = FALSE, 
         caption = "Comparisons of assessments (SSB) where stock definitions changed. Green is old stock definition, red is new stock definition")

```

**Scale-jumps: a new metric to describe historical retrospective patterns in ICES stock assessments**

<!-- Tables ------------------------------------------------------ -->

**Tables**

*`r tab_nums("scalejumps")`*

```{r}

# print table
d.jumps.bydecade %>% 
  filter(variable == "stocksize") %>% 
  ungroup() %>% 
  select(stock, decade, ratio3) %>% 
  spread(key=decade, value=ratio3) %>% 
  pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "left",
               missing=".",
               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))

```

<!-- Figures ------------------------------------------------------ -->

#####page break

**Figures**

<!-- ################################################################################## -->
<!-- Plot Stock definitions      ------------------------------------------------------ -->
<!-- ################################################################################## -->

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

d %>% 
  select(assessmentyear, stock, stockkeylabelold) %>% 
  distinct() %>% 
  mutate(stock = factor(stock),
         stock = factor(stock, levels=rev(levels(stock)))) %>% 
  group_by(stock, stockkeylabelold) %>% 
  summarize(firstyear = min(assessmentyear),
            lastyear = max(assessmentyear)) %>%

  ggplot(aes(x=assessmentyear, y=stock)) +
  theme_publication() +
  ggalt::geom_dumbbell(aes(xend = lastyear, x=firstyear, group=stock)) +
  geom_text(aes(x=firstyear, label=stockkeylabelold), hjust=0, nudge_y=0.15, size=3)

```

*`r fig_nums("stockdefinitions")`*

#####page break

<!-- ################################################################################## -->
<!-- Plot different methods of creating historical retrospectives --------------------- -->
<!-- ################################################################################## -->

```{r echo=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE}

knitr::include_graphics("fig1.2_retromethods.jpg")

```

*`r fig_nums("differentmethods")`*

#####Page break

<!-- Plot Retroplots --------------------------- -->

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot
d.selection %>%   
  ggplot(aes(x=year, y=value)) +
  theme_publication() +
  theme(
    panel.spacing.x = unit(1, "mm"),
    panel.spacing.y = unit(1, "mm"),
    strip.background = element_blank(),
    strip.text       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    legend.position  = "none"
  ) +

  # geom_boxplot(aes(group=year), outlier.shape=NA, colour="gray50", fill=NA) +
  geom_segment(data=d.segment, 
               aes(x=year, xend=year,y=value,yend=value_last), colour="gray", linetype="dashed", size=0.5) +
  geom_line(aes(colour = plot, size=plot, group=assessmentyear)) +
  geom_point(data=d.points) +
  scale_colour_manual(values=c(last = "red", other="black")) +
  scale_size_manual (values=c(last=1, other=0.5)) +
  expand_limits(y=0) +
  scale_y_continuous(breaks = pretty_breaks(),  labels = scales::scientific_format(digits=2)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  labs(y="ssb (tonnes)") +
  facet_wrap(~stock, scales="free")

```

*`r fig_nums("retroplots")`*

#####Page break

<!-- ################################################################################## -->
<!-- Plot Retro metrics                                     --------------------------- -->
<!-- ################################################################################## -->

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

d.metrics_annual %>%

  ggplot() +
  theme_publication() +
  theme(
    panel.spacing.x = unit(1, "mm"),
    panel.spacing.y = unit(1, "mm"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(angle = 0),
    axis.text.x      = element_text(angle = 90, vjust = 0.5), 
    legend.position  = "none"
  ) +

  geom_hline(aes(yintercept=0), colour="gray", linetype="dashed") +
  # geom_line(aes(x=year, y = value, colour=calctype)) +

  geom_point(aes(x=year, y = value, colour=calctype), shape=1) +

  geom_segment(data=d.metrics_bystock, 
               aes(x=firstyear, xend=lastyear,y=value,yend=value, colour=calctype), size=1) +

  # geom_hline(data=d.metrics_bystock, aes(yintercept = value, colour=calctype), size=1) +

  scale_y_continuous(breaks=pretty_breaks()) +
  scale_x_continuous(breaks=pretty_breaks(n=4)) +
  labs(y="") +
  facet_grid(metric~stock2, scales="free_y")

```

*`r fig_nums("retrometrics")`*

#####Page break

<!-- ################################################################################## -->
<!-- Plot Summary of retro metrics                          --------------------------- -->
<!-- ################################################################################## -->

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

d.metrics_bystock %>%

  ggplot() +
  theme_publication() +
  theme(
    panel.spacing.x = unit(1, "mm"),
    panel.spacing.y = unit(1, "mm"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(angle = 0),
    axis.text.x      = element_text(angle = 90, vjust = 0.5), 
    legend.position  = "none"
  ) +

  geom_hline(aes(yintercept=0), colour="gray", linetype="dashed") +

  geom_bar(aes(x=stock2, y = value, fill=calctype), stat="identity", position="dodge") +

  scale_y_continuous(breaks=pretty_breaks()) +
  # scale_x_continuous(breaks=pretty_breaks(n=4)) +
  labs(y="", x="") +
  facet_grid(metric~calctype, scales="free_y")

```

*`r fig_nums("retrometrics2")`*

#####Page break

<!-- ################################################################################## -->
<!-- Plot scale jumps                                                                   -->
<!-- ################################################################################## -->

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot scale jumps between assessments
d.jumps %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

  ggplot(aes(x=assessmentyear, y=value)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +

  # geom_point(data=d.benchmarks, 
  #            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +

  # geom_point(data=d.assessmentmodelchange, 
  #            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +

  # geom_segment(data=d.benchmarks, 
  #              aes(x=assessmentyear+0.1, xend=assessmentyear+0.1, y=-1.0, yend=1.0), 
  #              colour="gray20", size=0.5, linetype="dashed") +

  # geom_segment(data=d.assessmentmodelchange, 
  #              aes(x=assessmentyear-0.1, xend=assessmentyear-0.1, y=-1.0, yend=1.0), 
  #              colour="red", size=0.5, linetype="dashed") +

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +

  # geom_text(data=filter(d.scale, variable == myvar), 
  #           aes(label=ratio),  
  #           x=-Inf, y=-Inf, hjust=-0.2, vjust=-0.8, inherit.aes=FALSE) +

  coord_cartesian(ylim=c(-1, 1)) +  # set the limits in cartesian space prevents outlier lines to be excluded
  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
  scale_x_continuous(breaks = pretty_breaks(), limits=c(1982, NA)) +
  labs(x=" ", y="") +
  facet_wrap(~stock)

```

*`r fig_nums("scalejumps")`*

<!-- ################################################################################## -->
<!-- COmpare all metrics in peels                                                       -->
<!-- ################################################################################## -->

```{r metricsinpeels, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

d.metrics_bystock2 %>% 
  filter(metric %in% c("asd", "rhoabs", "sigma", "percjumps", "phiabs")) %>% 
  filter(calctype == "in_peels") %>% 
  group_by(calctype, author, metric) %>% 
  mutate(value_std = (value - min(value)) / (max(value) - min(value))) %>% 
  ungroup %>%

  ggplot() +
  theme_publication() +
  theme(
    panel.spacing.x = unit(1, "mm"),
    panel.spacing.y = unit(1, "mm"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(angle = 0),
    axis.text.x      = element_text(angle = 90, vjust = 0.5)
  ) +

  geom_boxplot(aes(x=stock2, y=value_std), stat="boxplot", fill=NA, outlier.colour = NA, width=0.5) +
  geom_jitter(aes(x=stock2, y=value_std, fill=metric, colour=metric), size=3, alpha = 1.0, width=0.1) +
  scale_y_continuous(breaks=pretty_breaks()) +
  labs(y="normalized metrics", x="")

# d.metrics_bystock2 %>% 
#   filter(metric %in% c("asd", "rhoabs", "sigma", "percjumps", "phiabs")) %>% 
#   filter(calctype == "in_peels") %>% 
#   group_by(calctype, author, metric) %>% 
#   mutate(value_std = (value - min(value)) / (max(value) - min(value))) %>% 
#   ungroup %>% 
#   ggplot() +
#   theme_publication() +
#   theme(
#     panel.spacing.x = unit(1, "mm"),
#     panel.spacing.y = unit(1, "mm"),
#     strip.background = element_blank(),
#     strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
#     strip.text.y     = element_text(angle = 0),
#     axis.text.x      = element_text(angle = 90, vjust = 0.5), 
#     legend.position  = "none"
#   ) +  
#   geom_bar(aes(x=stock2, y = value, fill=calctype), stat="identity", position="dodge") +
#   scale_y_continuous(breaks=pretty_breaks()) +
#   labs(y="", x="") +
#   facet_wrap(~metric, scales="free_y")

```

*`r fig_nums("metricsinpeels")`*

<!-- ################################################################################## -->
<!-- COmpare scale jumps when stock definition changed                                  -->
<!-- ################################################################################## -->

```{r metricsatstockchange, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# plot scale jumps between assessments when stock definition changed
t3 %>%

  ggplot(aes(x=year, y=value, group=label)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
    strip.placement = "outside",
    legend.title    = element_blank(),
    legend.position  = "none"
  ) +

  geom_line(aes(colour=label), size=0.5) +

  geom_dl(aes(label  = stockkeylabelold, colour = label),
          method = list(dl.combine("first.bumpup"), cex = 0.8, hjust=0)) +

  scale_y_continuous(labels=scientific_format(digits=2)) +
  scale_x_continuous(breaks = pretty_breaks(n=3)) +
  expand_limits(y=0) +
  labs(x="", y="ssb") +
  facet_wrap( ~ header, scales="free")

```

*`r fig_nums("stockdefchange")`*

### 2019-01-15 R/pdf analysis of pages in ICES advice.Rmd
Updates 15/1/2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# PDF analysis of pages in ICES advice reports.Rmd
#
# 11/01/2019 first coding; 
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(pdftools)
library(tidyverse)
library(stringr)

# captioner library
library(captioner) # deal with captions
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"))

# Load utils code
source("../../mptools/r/my_utils.r")

# file path
my.path <- "E:/ICES/advice"

y <- 2018

# counter for loop
i <- 0

# loop over years
for (y in 2000:2018) {

  fl <- list.files(path=file.path(my.path, y), 
                   pattern = "pdf$", 
                   full.names = TRUE )
  print(y)

  for (fn in fl) {

    if (file.exists(fn) ) {
      # print(pdf_info(fn)[2])
      t <- 
        data.frame(fn=fn, year=y, npages = pdf_info(fn)[2], stringsAsFactors = FALSE) %>% 
        mutate(
          species = str_extract(basename(fn), "^[a-z]{3}"),
          species = ifelse(is.na(species), "ZZZ", species)       
        )
      if (i == 0) {acom_pages <- t} else {acom_pages <- bind_rows(acom_pages, t)}
      i <- i + 1

    } # end of file exists
  } # end of file loop

  gc()

} # end of y loop

# print(wg_pages)

# summarize by pentad of 5 years
acom_pages_bypentad <-
  acom_pages %>% 
  group_by(year) %>% 
  summarize(pages = sum(pages)) %>% 
  mutate(
    pentad = 10 / 2 * floor((year*2)/10)
  ) %>% 
  group_by(pentad) %>% 
  mutate(avg = mean(pages),
         end = pentad + 4)

# Figure numbers
fig_nums(name    = "pages_byspecies", level = 1, display = FALSE,
         caption = "Number of pages by species")

fig_nums(name    = "pages_byyear", level = 1, display = FALSE,
         caption = "Number of pages by year and averages by pentad")

```

**Number of pages in advice documents**

M.A. Pastoors

`r format(Sys.time(), '%d/%m/%Y')`

Number of pages in advice document. Species have been automatically derived from the first three letters of the file name (and may not have worked flawlessly). A selection of species is shown in  
`r fig_nums("pages_byspecies", display="cite")`. All pages by 'species' with annual averages by decade is shown in `r fig_nums("pages_byyear", display="cite")`

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>% 
  filter(species %in% c("ang", "cap", "cod", "eel", "had", "her", "hke", "hom", "mac", "meg", "nep",
                        "ple", "san", "sol", "whg", "whb")) %>% 
  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(fill=species), stat="identity") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  facet_wrap(~species)

```

*`r fig_nums("pages_byspecies")`*

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>%

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(aes(fill=species), stat="identity") +
  geom_segment(data=acom_pages_bypentad, aes(x=pentad, xend=end, y=avg,yend=avg), colour="black", size=2) +
  geom_text   (data=acom_pages_bypentad, aes(x=pentad, y=avg, label=as.integer(avg)), 
               vjust=-0.5, hjust=0) +
  expand_limits(y=0) +
  guides(fill = guide_legend(nrow = 8)) + 
  scale_x_continuous(breaks=scales::pretty_breaks())

```

*`r fig_nums("pages_byyear")`*

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

acom_pages %>% 
  group_by(year) %>% 
  summarize(nfiles = n()) %>%

  ggplot(aes(x=year,y=nfiles)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(stat="identity") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks())

```

### 2019-01-15 R/pdf analysis of pages in wg reports.Rmd
Updates 15/1/2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# PDF analysis of pages in WG reports.Rmd
#
# 11/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(pdftools)
library(tidyverse)

# captioner library
library(captioner) # deal with captions
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"))

# Load utils code
source("../../mptools/r/my_utils.r")

# file path
my.path <- "D:/ZZZ"

# counter for loop
i <- 0

# loop over years
for (wg in c("HAWG", "WGNSSK", "AFWG", "WGBFAS", "WGCSE", "WGSSDS", "WGNSDS", "WGWIDE", "WGMHSA","WGDEMPELBAL", 
             "WGHMM", "WGBIE", "WGDEEP", "WGANSA", "WGHANSA", "WGEF", "NWWG", "WGNEW")) {

  for (y in 2000:2018) {

    fn <- file.path(my.path, paste0(wg," ",y,".pdf")) 
    # print(fn)
    if (file.exists(fn)) {
      # print(pdf_info(fn)[2])
      t <- data.frame(wg=wg, year=y, npages = pdf_info(fn)[2], stringsAsFactors = FALSE)
      if (i == 0) {wg_pages <- t} else {wg_pages <- bind_rows(wg_pages, t)}
      i <- i + 1
    }
  } # end of y loop
}  # end of wg loop

# print(wg_pages)

# summarize by pentad of 5 years
wg_pages_bypentad <-
  wg_pages %>% 
  group_by(year) %>% 
  summarize(pages = sum(pages)) %>% 
  mutate(
    pentad = 10 / 2 * floor((year*2)/10)
  ) %>% 
  group_by(pentad) %>% 
  mutate(avg = mean(pages),
         end = pentad + 4)

# Figure numbers
fig_nums(name    = "pages_bywg", level = 1, display = FALSE,
         caption = "Number of pages by WG")

fig_nums(name    = "pages_byyear", level = 1, display = FALSE,
         caption = "Number of pages by year and averages by pentad")

```

**Number of pages in WG reports**

M.A. Pastoors

`r format(Sys.time(), '%d/%m/%Y')`

Number of pages in WG reports of the following ICES working groups: `r unique(wg_pages$wg)`.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

wg_pages %>%

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(fill=wg), stat="identity", colour="white", size=0.2) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  facet_wrap(~wg)

```

*`r fig_nums("pages_bywg")`*

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

wg_pages %>%

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(aes(fill=wg), stat="identity", colour="white", size=0.2) +
  geom_segment(data=wg_pages_bypentad, aes(x=pentad, xend=end, y=avg,yend=avg), colour="black", size=2) +
  geom_text   (data=wg_pages_bypentad, aes(x=pentad, y=avg, label=as.integer(avg)), 
               vjust=-0.5, hjust=0) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks())

```

*`r fig_nums("pages_byyear")`*

### 2019-01-15 R/Analyse changes in pelagic stocks.r

@@ -1,0 +1,122 @@
+# -----------------------------------------------------------------------------------------------
+# Analyse changes in pelagic stocks
+#
+# 07/01/2018 first coding
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(RColorBrewer) # colours
+library(lubridate)
+library(scales)
+library(directlabels)  # for printing labels at end of geom lines
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# Load dataset
+load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+
+# create subset
+d <-
+  iAdvice %>% 
+  
+  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss", 
+               stockkeylabelold)) %>%
+  filter(!grepl("benchmark", purpose)) %>% 
+  filter(assessmentyear >= 2008) %>% 
+  ungroup()
+
+# plot F reference points
+d %>% 
+  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
+  gather(key=var, value=value, fpa:fmsy) %>% 
+  
+  ggplot(aes(x=assessmentyear)) +
+  theme_publication() +
+  theme(legend.title = element_blank()) +
+  
+  geom_line(aes(y=value, group=var, colour=factor(var))) +
+  geom_point(aes(y=value, group=var, colour=factor(var))) +
+  
+  scale_x_continuous(breaks=pretty_breaks()) +
+  expand_limits(y=0) + 
+  labs(y="F") +
+  scale_color_manual(values = c("blue", "green")) +
+  facet_wrap(~stockkeylabelold, scales="free_y")
+
+# plot B reference points
+d %>% 
+  dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
[…]

### 2019-01-15 R/dev/Generate lat long map.r

@@ -1,0 +1,80 @@
+# --------------------------------------------------------------------------------------------
+# generate lat long map.r
+#
+# 27/12/2018 first coding
+# --------------------------------------------------------------------------------------------
+
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(rgeos)
+
+source("../mptools/r/my_utils.r")
+
+# get dropbox directory
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# ----------------------------------------------------------------------------------------------------------
+# set world in sf format
+# ----------------------------------------------------------------------------------------------------------
+world1 <- 
+  map('world', plot=FALSE) %>% 
+  summary()
+filter(grepl("spain",tolower(names)))
+View(world1)
+
+nams <- 
+  map("world", namesonly=TRUE, plot=FALSE) 
+nams[grep("island", tolower(nams))]
+
+world1 <- 
+  map('world',
+      plot = FALSE, 
+      fill = TRUE, 
+      regions = c("Netherlands","belgium","France(?!:Corsica)","ireland","united kingdom", "UK", 
+                  "Denmark","Germany","norway","iceland", "greenland", "faroe islands", "spain",
+                  "portugal","sweden", "finland","poland", 
+                  "canary islands:gran canaria", "canary islands:tenerife", "canary islands:la palma",
+                  "morocco", "mauritania", "western sahara", "chile", "peru", "chile:easter island") ) %>% 
+  map_data()  
+  # sample_n(5000, replace=FALSE) %>% 
+  # arrange(group, order)
+
+
+ggplot(data=world1, aes(x=long, y=lat, group=group)) +
+  theme_publication() +
+  geom_path()
+
+w1 <-
+  world1 %>% 
+  group_by(group, region, subregion) %>%
+  filter(n() > 10) %>% 
+  distinct(group, region, subregion)
+
+w1a <-
+  w1 %>% 
+  left_join(world1, by=c("group","region","subregion")) %>% 
+  ungroup() %>% 
+  sample_n(7000, replace=FALSE) 
+
+w1b <-
+  world1 %>% 
+  filter(!(group %in% w1$group & region %in% w1$region & subregion %in% w1$subregion))
+
+w2 <-
+  bind_rows(w1a, w1b) %>% 
+  arrange(group, order) %>% 
[…]

### 2019-01-15 R/iDatabases generator.r

@@ -294,6 +294,12 @@ sag <-
   # filter(row_number() == 1) %>% 
   ungroup() %>% 
   
+  # Deal with plaice in the North Sea (assessments prior to 2015 had different area allocation)
+  mutate(
+    stockkey         = ifelse(stockkeylabel == "ple-nsea" & assessmentyear %in% 2013:2014, 100103, stockkey),
+    stockkeylabel    = ifelse(stockkeylabel == "ple-nsea" & assessmentyear %in% 2013:2014, "ple-nsea2", stockkeylabel)
+  ) %>% 
+  
   # Only keep distinct rows
   distinct() %>% 
   
@@ -309,6 +315,7 @@ sag <-
   group_by(stockkey, stockkeylabel, assessmentyear, purpose, year) %>% 
   filter(row_number() == 1) %>% 
   ungroup()
+glimpse(sag)
 
 save(sag, file=paste(advicedir, "/rdata/iSAG.RData",sep=""))
 
@@ -656,6 +663,7 @@ save(iAssess, file=paste(advicedir, "/rdata/iAssess.RData",sep=""))
 #   scale_y_continuous(labels = scales::scientific_format(digits=2)) +
 #   facet_wrap(~stockkeylabel, scales="free_y")
 
+# distinct(iAssess, stockkeylabel, assessmentyear) %>% filter(grepl("ple-nsea", stockkeylabel)) %>% View()
 
 # -----------------------------------------------------------------------------------------
 # load Esther's data

### 2019-01-15 R/plot TAC area.r

@@ -1,0 +1,112 @@
+# --------------------------------------------------------------------------------------------
+# plot TAC area.r
+#
+# plotting stock areas (over time)
+# 
+# 27/04/2018 cleaned up the code
+# 28/11/2018 updated the code to work with new iAdvice database
+# 19/12/2018 using new FAO_IHO dataset for international and EEZ waters
+# --------------------------------------------------------------------------------------------
+
+# devtools::install_github("fishvice/tidyices", dependencies = FALSE)
+
+# devtools::install_github("einarhjorleifsson/iceshape", dependencies = FALSE)
+# devtools::install_github("einarhjorleifsson/ggmisc", dependencies = FALSE)
+library(iceshape)   # ls("package:iceshape")
+
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(viridis)
+
+source("../mptools/r/my_utils.r")
+# load("../prf/rdata/world.df.RData")
+
+# get dropbox directory
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# ----------------------------------------------------------------------------------------------------------
+# set world in sf format
+# ----------------------------------------------------------------------------------------------------------
+
+world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE, 
+                           regions = c("netherlands","belgium","France(?!:Corsica)",
+                                       "ireland","united kingdom", "UK", "Denmark","Germany",
+                                       "norway","iceland", "greenland", "faroe islands", "spain","portugal",
+                                       "sweden", "finland","poland")))
+# ggplot(data=world1) + geom_sf()
+
+
+# ----------------------------------------------------------------------------------------------------------
+# load the iAdvice data
+# ----------------------------------------------------------------------------------------------------------
+
+load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+load(file.path(advicedir,"rdata/fao_iho.RData"))
+
+fao_iho %>% filter(grepl("27.2.a", name)) %>% View()
+
+
+# ----------------------------------------------------------------------------------------------------------
+# generate the stk object: stock, assessmentyear and area(s) 
[…]

### 2019-01-21 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Updates until 20/1/2019

@@
   iAssess %>%
-
-  # Exclude prediction years
-  filter(year <= assessmentyear) %>%
-
-  # Define start year
-  filter( year >= 1980) %>%
-
-  # Only assessments with stock size in tonnes
-  filter(stocksizeunits  == "tonnes") %>%  # Note: this is filtering on absolute vs relative; but not really
 
@@
 
+  # Exclude prediction years (but keep the benchmark information that has no year attached to it)
+  filter(is.na(year) | (year <= assessmentyear & year >= 1980) ) %>% 
+  filter(assessmentyear >= 1982) %>% 
+  filter(is.na(year) | !(is.na(stocksize) & is.na(fishingpressure))) %>%
+
+  # Only assessments with stock size in tonnes
+  filter(is.na(year) |stocksizeunits  == "tonnes") %>%
+
+  # Sorting
   arrange(stock, purpose, desc(assessmentyear)) %>%
-
-  # filter(purpose         == "advice") %>%
 
@@
   gather(key=variable, value=value, stocksize:fishingpressure) %>% 
-  filter(!is.na(value), value != 0.0)  %>%
+  filter(is.na(year) | (!is.na(value) & value != 0.0) )  %>%
 
@@
 
-  arrange(stock, variable, desc(assessmentyear)) %>% 
+  arrange(stock, variable, (assessmentyear)) %>% 
   ungroup() %>%
@@
   mutate(plot=ifelse(assessmentyear == max(assessmentyear), "last", "other"))
-
-distinct(iAssess, stockkeylabelold, assessmentyear, purpose, benchmark) %>% filter(grepl("ple-n", stockkeylabelold)) %>% View()
 
@@
   d %>%
-  filter(purpose == "advice")
+  filter(purpose == "advice") %>% 
+  filter(!is.na(year))
 
@@
   filter(benchmark == TRUE) %>% 
-  distinct(stock, stockkeylabelold, assessmentyear, purpose, .keep_all = FALSE)
+  distinct(stock, stockkeylabelold, assessmentyear, .keep_all = FALSE) %>% 
+  mutate(benchmark = 1)
 
@@
   left_join(iAdvice, by=c("stockkeylabelold","assessmentyear", "purpose")) %>%
+  select(-benchmark) %>% 
   mutate(assessmentmodel = tolower(assessmentmodel)) %>%
@@
   # determine stock definition change and model definition change
+  arrange(stock, assessmentyear) %>% 
   group_by(stock) %>% 
-  arrange(stock, assessmentyear) %>% 
-  mutate(assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel), 1, 0),
-         assessmentmodelchange = ifelse(is.na(assessmentmodelchange), 0, assessmentmodelchange)) %>%
+  mutate(
+    assessmentmodelchange = ifelse(assessmentmodel != lead(assessmentmodel), 2, 0),
+    assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel) , 1, assessmentmodelchange),
+    assessmentmodelchange = ifelse(assessmentmodel != lag(assessmentmodel) & 
+                                   assessmentmodel != lead(assessmentmodel), 3, assessmentmodelchange),
+
+    # add first and last year to stock definitions
+    assessmentmodelchange = 
+      case_when(
+        assessmentyear == min(assessmentyear) & assessmentmodel != lead(assessmentmodel)  ~ 3, 
+        assessmentyear == min(assessmentyear) & assessmentmodel == lead(assessmentmodel)  ~ 1,
+        assessmentyear == max(assessmentyear) & assessmentmodel != lag(assessmentmodel) ~ 3, 
+        assessmentyear == max(assessmentyear) & assessmentmodel == lag(assessmentmodel) ~ 2,
+        TRUE ~ assessmentmodelchange
+      ) 
+    ) %>%
+
+  # stock definition change
+  mutate(
+    stockdefinitionchange = ifelse(stockkeylabelold != lead(stockkeylabelold), 2, 0),
+    stockdefinitionchange = ifelse(stockkeylabelold != lag(stockkeylabelold), 1, stockdefinitionchange),
+    stockdefinitionchange = ifelse(is.na(stockdefinitionchange), 0, stockdefinitionchange),
+
+    # add first and last year to stock definitions
+    stockdefinitionchange = ifelse(assessmentyear == min(assessmentyear), 1, stockdefinitionchange),
+    stockdefinitionchange = ifelse(assessmentyear == max(assessmentyear), 2, stockdefinitionchange)
+
+  ) %>% 
+  #     ifelse(assessmentyear == min(assessmentyear), 1, stockdefinitionchange),
+  #   stockdefinitionchange = ifelse(assessmentyear == max(assessmentyear), 2, stockdefinitionchange)
+
+  # benchmarks
+  left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
+  mutate(benchmark = ifelse(is.na(benchmark), 0, benchmark))
+
+# dataset with stock definition changes
+d.stockdefinitionchange <-
+  d.assessments %>% 
+  filter(stockdefinitionchange >= 1) %>% 
+  distinct(stock, stockkeylabelold, assessmentyear, stockdefinitionchange, .keep_all = FALSE) %>%
+
+  group_by(stock, stockkeylabelold) %>% 
+  summarize(tmp = paste(assessmentyear, collapse="_")) %>% 
+  separate(tmp, into=c("y1","y2"), sep="_") %>% 
+  mutate_at(c("y1","y2"), funs(as.numeric)) %>% 
+  arrange(stock, y1) %>%
 
   group_by(stock) %>% 
-  arrange(stock, assessmentyear) %>% 
-  mutate(stockdefinitionchange = ifelse(stockkeylabelold != lag(stockkeylabelold), 1, 0),
-         stockdefinitionchange = ifelse(is.na(stockdefinitionchange), 0, stockdefinitionchange))
+  mutate(rank = as.character(row_number()))
+
+# dataset with assessment model periods
+d.assessmentmodels3 <-
+  d.assessments %>% 
+  filter(assessmentmodelchange == 3) %>% 
+  distinct(stock, assessmentyear, assessmentmodel, assessmentmodelchange, 
+           .keep_all = FALSE) %>% 
+  mutate(y1 = assessmentyear, y2 = assessmentyear+0.8) %>% 
+  ungroup() %>% 
+  select(-assessmentyear, -assessmentmodelchange)
+
+d.assessmentmodels <-
+  d.assessments %>% 
+  filter(assessmentmodelchange %in% 1:2) %>% 
+  distinct(stock, assessmentyear, assessmentmodel, assessmentmodelchange, 
+           .keep_all = FALSE) %>%
+
+  group_by(stock, assessmentmodel) %>% 
+  summarize(tmp = paste(assessmentyear, collapse="_")) %>% 
+  separate(tmp, into=c("y1","y2"), sep="_") %>% 
+  mutate_at(c("y1","y2"), funs(as.numeric)) %>% 
+  mutate(y2 = y2 +0.8) %>%
+
+  # add the single year assessments
+  bind_rows(d.assessmentmodels3) %>%
+
+  arrange(stock, y1) %>%
+
+  group_by(stock) %>% 
+  mutate(rank = as.character(row_number()))
 
@@
             mohn_rho     = mean(mohn_rho), 
-            firstyear    = min(year),
-            lastyear     = max(year))
+            firstyear    = min(year, na.rm=TRUE),
+            lastyear     = max(year, na.rm=TRUE))
 
@@
   mutate(logB     = log(value),
-         meanlogB = mean(log(value)),
+         meanlogB = mean(log(value), na.rm=TRUE),
          logBdev  = logB - meanlogB,
@@
   # now do summing over years
-  mutate(sumlogBdev2 = sum(logBdev2) ) %>%
+  mutate(sumlogBdev2 = sum(logBdev2), na.rm=TRUE ) %>%
 
@@
   summarize(
-    meanlogBdev = mean(logBdev),
-    sumlogBdev2 = sum(logBdev2),
-    n_1         = mean(n_1), 
+    meanlogBdev = mean(logBdev, na.rm=TRUE),
+    sumlogBdev2 = sum(logBdev2, na.rm=TRUE),
+    n_1         = mean(n_1, na.rm=TRUE), 
     ralston_sigma = sqrt(sumlogBdev2/n_1)) %>%
@@
     n_1            = sum(n_1), 
-    firstyear    = min(year),
-    lastyear     = max(year)
+    firstyear    = min(year, na.rm=TRUE),
+    lastyear     = max(year, na.rm=TRUE)
   ) %>%
@@
   select(-assessmentyearprior) %>% 
-  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
   mutate(label="last") %>% 
@@
   select(-assessmentyearlast) %>% 
-  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
   mutate(label="prior") %>% 
@@
 
-d.jumps <-
+# pairwise data roughly
+d.pairwise.roughly <-
   bind_rows(d.last, d.prior) %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears -5)
+
+# pairwise data
+d.pairwise <-
+  d.pairwise.roughly %>% 
   group_by(stock, assessmentpair, variable) %>%
@@
   mutate(n = n()) %>% 
-  filter(n() == 2) %>%
+  filter(n() == 2)
+
+d.jumps <-
+  d.pairwise %>%
 
@@
          ratio3       = paste(ratio2," (", n, ")", sep="") )
+
+# jumps by year
+d.jumps.byyear <-
+  d.jumps %>% 
+  group_by(variable, assessmentyear) %>% 
+  summarise(n         = n(),
+            n2        = sum(phi_abs > 0.1, na.rm=TRUE)) %>% 
+  mutate(ratio = n2 / n )
 
@@
 
-fig_nums(name    = "scalejumps", level = 1, display = FALSE, 
+fig_nums(name    = "scalecomparisons", level = 1, display = FALSE, 
+         caption = "Example of pairwise assessment comparisons for Northeast Atlantic mackerel  over a period of 10 years in common.
+The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
+
+fig_nums(name    = "scalejumps", level = 2, display = FALSE, 
          caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records.
-Results from pairwise comparisons over periods of 8 years.
+Results from pairwise comparisons over periods of 10 years.
 Large dots  refer to years when the scale difference is larger than 10%.
-Numbers at the bottom left of each panel refer to the number of assessments with a scale difference larger than 10% relative to the total number of assessments available for that stock.
 Colours indicate the decades.
-Black triangles and dashed lines indicate benchmark assessment years.
-Red triangles and dashed lines indicate years when assessment models were changed compared to the year before")
+Black triangles indicate benchmark assessment years.
+The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
 
@@
          caption = "Comparisons of assessments (SSB) where stock definitions changed. Green is old stock definition, red is new stock definition")
+
+fig_nums(name    = "jumpratio", level = 2, display = FALSE, 
+         caption = "Ratio of scale-jumps in assessment relative to the number of assessment by year.")
+
+getPalettePurples         <- colorRampPalette(brewer.pal(7, "Purples")[-(1:2)])
+myPurples                 <- getPalettePurples(10)
+
+getPaletteBlues         <- colorRampPalette(brewer.pal(7, "Blues")[-(1:2)])
+myBlues                 <- getPaletteBlues(10)
 
@@
 <!-- ################################################################################## -->
+<!-- Plot pairwise comparisons                                                          -->
+<!-- ################################################################################## -->
+
+```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
+
+myvar     <- "stocksize"
+myspecies <- "mac"
+firstyear <- 1991
+
+d.pairwise.means <-
+
+  d.pairwise %>%
+
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  # calculate means per period
+  group_by(stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(stock, assessmentpair, variable, assessmentyear, label)
+
+d.pairwise %>%
+
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  ggplot(aes(x=year, y=value, group=assessmentyear)) +
+  theme_publication() +
+  theme(
+    panel.spacing.x = unit(1, "mm"),
+    panel.spacing.y = unit(1, "mm"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(1,0,0,0, "mm")),
+    strip.text.y     = element_text(angle = 0),
+   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
+    legend.position  = "none"
+  ) +
+
+  geom_line(data=filter(d.pairwise.roughly, 
+                        variable == myvar, 
+                        grepl(myspecies, stockkeylabelold),
+                        as.numeric(substr(assessmentpair,1,4)) >= firstyear), size=0.5) +
+
+  geom_line(aes(colour=factor(label)), size=1) +
+
+  geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)), 
+            linetype="dashed", lineend="butt") +
+  geom_point(data=filter(d.pairwise.means, year==min(year)),
+             aes(colour=factor(label)), size=1) +
+  geom_point(data=filter(d.pairwise.means, year==max(year)),
+             aes(colour=factor(label)), size=1) +
+
+  scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
+  scale_x_continuous(breaks = pretty_breaks(n=2)) +
+  expand_limits(y=0) +
+  labs(x="", y="ssb (tonnes)") +
+  facet_wrap( ~ assessmentpair, scales = "free_x", ncol=5)
+
+```
+
+*`r fig_nums("scalecomparisons")`*
+
+#####Page break
+
+<!-- ################################################################################## -->
 <!-- Plot scale jumps                                                                   -->
@@
 
-  ggplot(aes(x=assessmentyear, y=value)) +
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
   theme_publication() +
@@
 
+  geom_rect(data=filter(d.stockdefinitionchange),
+            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
+            alpha=0.4, colour=myPurples[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.stockdefinitionchange),
+          aes(label  = stockkeylabelold, x = y1, y = -0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
+          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +
+
   geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
@@
 
-  # geom_point(data=d.benchmarks, 
-  #            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
-
-  # geom_point(data=d.assessmentmodelchange, 
-  #            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
-
-  # geom_segment(data=d.benchmarks, 
-  #              aes(x=assessmentyear+0.1, xend=assessmentyear+0.1, y=-1.0, yend=1.0), 
-  #              colour="gray20", size=0.5, linetype="dashed") +
-
-  # geom_segment(data=d.assessmentmodelchange, 
-  #              aes(x=assessmentyear-0.1, xend=assessmentyear-0.1, y=-1.0, yend=1.0), 
-  #              colour="red", size=0.5, linetype="dashed") +
-
   geom_line (aes(colour=factor(decade))) +
@@
 
-  # geom_text(data=filter(d.scale, variable == myvar), 
-  #           aes(label=ratio),  
-  #           x=-Inf, y=-Inf, hjust=-0.2, vjust=-0.8, inherit.aes=FALSE) +
-
-  coord_cartesian(ylim=c(-1, 1)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+  geom_point(data=filter(d.assessments, benchmark == 1),
+            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
+
+  geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
+            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
+
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  scale_fill_manual(values = brewer.pal(6, "Purples")[-(1:2)]) +
+
   scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks(), limits=c(1982, NA)) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
   labs(x=" ", y="") +
@@
 *`r fig_nums("scalejumps")`*
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot scale jumps between assessments vs assessment methods
+
+d.jumps %>% 
+  distinct(stock, variable, assessmentyear, phi) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_rect(data=filter(d.assessmentmodels),
+            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
+            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.assessmentmodels),
+          aes(label  = assessmentmodel, x = y1, y = -0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
+          inherit.aes = FALSE, colour=myBlues[6]) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
+
+  geom_point(data=filter(d.assessments, benchmark == 1),
+            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
+
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="") +
+  facet_wrap(~stock)
+
+```
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot scale jumps between assessments vs assessment methods
+
+d.jumps %>% 
+  distinct(stock, variable, assessmentyear, phi) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_rect(data=filter(d.stockdefinitionchange),
+            aes(xmin = y1, xmax=y2, ymin = 0.00, ymax=1.0), 
+            alpha=0.4, fill=myPurples[6], colour=myPurples[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.stockdefinitionchange),
+          aes(label  = stockkeylabelold, x = y1, y = 0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=1.0, vjust=0.9),
+          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +
+
+  geom_rect(data=filter(d.assessmentmodels),
+            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=0.0, fill=rank), 
+            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.assessmentmodels),
+          aes(label  = assessmentmodel, x = y1, y = -0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
+          inherit.aes = FALSE, colour=myBlues[6]) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
+
+  geom_point(data=filter(d.assessments, benchmark == 1),
+            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
+
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="") +
+  facet_wrap(~stock)
+
+```
 
@@
 *`r fig_nums("stockdefchange")`*
+
+<!-- ################################################################################## -->
+<!-- COmpare the number of scale jumps (as ratio) over time                             -->
+<!-- ################################################################################## -->
+
+```{r jumpratio, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}
+
+d.jumps.byyear %>% 
+  filter(assessmentyear >= 1990) %>%
+
+  ggplot(aes(x=assessmentyear, y=ratio)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.placement = "outside",
+    legend.title    = element_blank(),
+    legend.position  = "none"
+  ) +
+  geom_bar(stat="identity") +
+  labs(x="", y="ratio")
+
+```
+
+*`r fig_nums("jumpratio")`*

### 2019-01-21 R/Analyse scale differences.r

@@ -607,3 +607,11 @@ d.stockdefinitionschange %>%
   left_join(d.year, by=c("stockkeylabelold", "assessmentyear")) %>% 
   filter(!is.na(assessmentpair)) %>% 
   filter(variable == "stocksize") 
+
+
+# Checking
+
+d.jumps.byyear %>% 
+  filter(assessmentyear >= 1990) %>% 
+  ggplot(aes(x=assessmentyear, y=ratio)) +
+  geom_bar(stat="identity")

### 2019-01-21 R/iDatabases generator.r

@@ -192,22 +192,6 @@ qcsexcel <-
   
 save(qcsexcel, file=paste(advicedir, "/rdata/qcsexcel.RData",sep=""))
 
-# qcsexcel %>% 
-#   filter(grepl("cod-347d", stockkeylabelold)) %>% 
-#   filter(year >= 2000) %>% 
-#   select(stockkeylabelold, assessmentyear, year, stocksize) %>% 
-#   spread(key=year, value=stocksize) %>% 
-#   arrange(desc(assessmentyear)) %>% 
-#   View()
-
-# glimpse(qcsexcel)
-# unique(qcsexcel$assessmentdate)
-# qcsexcel %>% filter(grepl("had.27.7b", stockkeylabelnew)) %>% View()
-# qcsexcel %>% filter(is.na(stockkey)) %>% View()
-# qcsexcel %>% filter(!is.na(assessmentdate.x) & !is.na(assessmentdate.y) & assessmentdate.x != assessmentdate.y) %>% View()
-# iAdvice %>% filter(stockkey == 169048) %>% View()
-# iAdvice %>% filter(is.na(speciesfaocode)) %>% View()
-
 
 # -----------------------------------------------------------------------------------------
 # load Stock Database 
@@ -221,19 +205,6 @@ sd <-
   left_join(iRename[,c("stockkeylabel","stockkey")], by="stockkeylabel") %>%
   left_join(iStockkey, by="stockkey") %>% 
   rename(assessmentyear = activeyear) 
-
-# sd %>% 
-#   filter(stockkeylabel %in% c("cod.27.24-32","pra.27.3a4a","mon.27.78abd","rjn.27.678abd")) %>% 
-#   distinct(stockkey, stockkeylabel) %>% 
-#   View()
-# 
-# sd %>% 
-#   filter(grepl("mon.27.78abd", stockkeylabel)) %>% 
-#   View()
-# 
-# sd %>% 
-#   filter(grepl("smn-grl",stockkeylabel)) %>% 
-#   View()
 
 
 # -----------------------------------------------------------------------------------------
@@ -314,84 +285,13 @@ sag <-
   # remove duplicate assessments (Careful!)
   group_by(stockkey, stockkeylabel, assessmentyear, purpose, year) %>% 
   filter(row_number() == 1) %>% 
-  ungroup()
-glimpse(sag)
+  ungroup() %>% 
+  
+  # remove all the custom fiels (for now)
[…]

### 2019-03-23 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Upload and cleanup 23/3/2019

@@
 # 15/01/2019 added scale jump figure with explanatory variables
+# 22/02/2019 focussed on Pairwise Comparison of Means = scale jumps under different conditions
 # -----------------------------------------------------------------------------------------------
@@
 library(ggthemes)  # for themes
+library(ggExtra)   # for specific type of plots e.g. ggMarginal
 library(pander)    # for print tables
@@
   # benchmarks
-  left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
+  full_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
+  # left_join(d.benchmarks, by=c("stock", "stockkeylabelold", "assessmentyear")) %>% 
   mutate(benchmark = ifelse(is.na(benchmark), 0, benchmark))
@@
 
-# Calculations of scale-differences and scale jumps ==========================================
-
-removelastyears <- 0
-numberofyears   <- 10
-threshold       <- 0.1   # threshold for scale differences
+# Calculations of Pairwise Comparison of Mean and scale jumps ==========================================
+
+pairwise_comparisons <- function(removelastyears=0, numberofyears=10) {
 
@@
 
-# data series with most recent (last) estimation
-d.last <-
+  # data series with most recent (last) estimation of the pair
+  d.pairs.last <-
     d.pairs %>% 
@@
 
-# data series with previous (prior) estimation
-d.prior <-
+  # data series with previous (prior) estimation of the pair
+  d.pairs.prior <-
     d.pairs %>% 
@@
   d.pairwise.roughly <-
-  bind_rows(d.last, d.prior) %>% 
+    bind_rows(d.pairs.last, d.pairs.prior) %>% 
     group_by(stock, assessmentpair, variable) %>%
@@
 
-d.jumps <-
+  d.pcm <-
     d.pairwise %>%
@@
            assessmentyear = as.integer(substr(assessmentpair, 6,9)))
+
+  res <- list("d.pcm"= d.pcm, 
+              "d.pairwise.roughly"=d.pairwise.roughly,
+              "d.pairwise" =d.pairwise)
+
+  return(res)
+
+} # End of pairwise_comparisons function
+
+tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5)
+
+# REMOVE FROM HERE !!
+
+removelastyears <- 0
+numberofyears   <- 10
+threshold       <- 0.1   # threshold for scale differences
+
+# Create a set of assessment pairs: two subsequent assessments  
+d.pairs <-
+  d.selection %>% 
+  distinct(stock, assessmentyear, variable) %>% 
+  group_by(stock, variable) %>% 
+  arrange(stock, variable, assessmentyear) %>% 
+  mutate(assessmentyearprior = lag(assessmentyear),
+         assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
+  rename(assessmentyearlast  = assessmentyear) %>% 
+  filter(!is.na(assessmentyearprior))
+
+# data series with most recent (last) estimation of the pair
+d.pairs.last <-
+  d.pairs %>% 
+  select(-assessmentyearprior) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  mutate(label="last") %>% 
+  rename(assessmentyear = assessmentyearlast)
+
+# data series with previous (prior) estimation of the pair
+d.pairs.prior <-
+  d.pairs %>% 
+  select(-assessmentyearlast) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  mutate(label="prior") %>% 
+  rename(assessmentyear = assessmentyearprior)
+
+# pairwise data roughly
+d.pairwise.roughly <-
+  bind_rows(d.pairs.last, d.pairs.prior) %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears -5)
+
+# pairwise data
+d.pairwise <-
+  d.pairwise.roughly %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears) %>%
+
+  # sort
+  arrange(stock, assessmentpair, assessmentyear, year) %>%
+
+  # filter only years that are in common in last and prior
+  group_by(stock, assessmentpair, variable, year) %>% 
+  mutate(n = n()) %>% 
+  filter(n() == 2)
+
+d.jumps <-
+  d.pairwise %>%
+
+  # calculate means per period
+  group_by(stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(stock, assessmentpair, variable, assessmentyear, label) %>%
+
+  # calculate metrics
+  filter(tmp=="firstyear") %>% 
+  ungroup() %>% 
+  select(stock, assessmentpair, variable, label, value) %>% 
+  spread(key=label, value=value) %>% 
+  mutate(phi            = (prior-last)/last,
+         phi_abs        = abs(phi),
+         assessmentyear = as.integer(substr(assessmentpair, 6,9)))
+
+# CONTINUE FROM HERE
 
@@
 
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
+
+## NOT SHOWN
 
@@
 
-```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
 
@@
 myspecies <- "mac"
-firstyear <- 1991
+firstyear <- 2008
 
@@
 
-  d.pairwise %>%
+  res$d.pairwise %>%
 
@@
 
-d.pairwise %>%
+res$d.pairwise %>%
 
@@
 
-  geom_line(data=filter(d.pairwise.roughly, 
+  geom_line(data=filter(res$d.pairwise.roughly, 
                         variable == myvar, 
@@
 <!-- ################################################################################## -->
-<!-- Plot scale jumps                                                                   -->
-<!-- ################################################################################## -->
-
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+<!-- Plot pairwise comparison means                                                     -->
+<!-- ################################################################################## -->
+
+```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
+
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
+
+threshold <- 0.10
 
 # plot scale jumps between assessments
-d.jumps %>% 
+res$d.pcm %>% 
   distinct(stock, variable, assessmentyear, phi) %>%
@@
   geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
-
-  geom_point(data=filter(d.assessments, benchmark == 1),
+  # geom_point(aes(colour=factor(decade), size = abs(value) )) +
+
+  geom_point(data=filter(d.benchmarks, benchmark == 1),
             aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
@@
 
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
+
+## NOT SHOWN
+
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
+
+threshold <- 0.10
 
@@
 
-d.jumps %>% 
+res$d.pcm %>% 
   distinct(stock, variable, assessmentyear, phi) %>%
@@
 
-  geom_point(data=filter(d.assessments, benchmark == 1),
+  geom_point(data=filter(d.benchmarks, benchmark == 1),
             aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
@@
 
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
+
+threshold <- 0.10
+
 # plot scale jumps between assessments vs assessment methods
 
-d.jumps %>% 
+res$d.pcm %>% 
   distinct(stock, variable, assessmentyear, phi) %>%
@@
 
-  geom_point(data=filter(d.assessments, benchmark == 1),
+  geom_point(data=filter(d.benchmarks, benchmark == 1),
             aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
 
+  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
   coord_cartesian(ylim=c(-1, 1),
@@
 
+  # ggMarginal(data=d.assessments, type="histogram")
   scale_fill_manual(values = myBlues) +
@@
   facet_wrap(~stock)
+
+# filter(d.assessments, benchmark == 1) %>% View()
 

### 2019-03-23 R/Analyse reference points.r

@@ -32,6 +32,7 @@ iAdvice %>%
   
   # add the reference points from sagrefpoints
   bind_rows(sagrefpoints) %>% 
+  filter(assessmentyear >= 1997) %>% 
   
   group_by(assessmentyear) %>% 
   summarize(flim = sum(!is.na(flim)),

### 2019-03-23 R/Combine FAO and IHO areas.r

@@ -2,6 +2,8 @@
 # combine FAO and IHO areas.r
 #
 # 19/12/2018 based on SFrame code
+# 21/01/2019 now run the code with the original FAO files (includes more levels)
+# 22/01/2019 contains both the EEZ_FAO combinations, and the FAO codes as stand alone. 
 # --------------------------------------------------------------------------------------------
 
 # devtools::install_github("fishvice/tidyices", dependencies = FALSE)
@@ -22,55 +24,61 @@ library(viridis)
 source("../mptools/r/my_utils.r")
 # load("../prf/rdata/world.df.RData")
 
-load(file.path(onedrive,"rdata/iho_eez.RData"))
-
 # get dropbox directory
 dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
-# source("https://raw.githubusercontent.com/einarhjorleifsson/gisland/master/R/read_sf_ftp.R")
-# faolink <- read_csv("/net/www/export/home/ftp/pub/reiknid/einar/shapes/fao-areas_nocoastline.csv")
-# fao <- read_sf_ftp("fao-areas_nocoastline")
-# eez <- read_sf_ftp("eez")
-
-
+# set onedrive directory
+onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 
 
 # ----------------------------------------------------------------------------------------------------------
 # set world in sf format
 # ----------------------------------------------------------------------------------------------------------
 
-world1 <- sf::st_as_sf(
-  map('world', plot = FALSE, fill = TRUE, 
-               regions = c("netherlands","belgium","France(?!:Corsica)",
-                           "ireland","united kingdom", "UK", "Denmark","Germany",
-                           "norway","iceland", "greenland", "faroe islands", "spain","portugal",
-                           "sweden", "finland","poland", "luxembourg"))
-)
+world_sf <- 
+  sf::st_as_sf(map('world', 
+                   plot    = FALSE, 
+                   fill    = TRUE, 
+                   regions = c("netherlands","belgium","France(?!:Corsica)",
+                               "ireland","united kingdom", "UK", "Denmark","Germany",
+                               "norway","iceland", "greenland", "faroe islands", "spain","portugal",
+                               "sweden", "finland","poland", "luxembourg")) )
+
[…]

### 2019-03-23 R/DownloadDataFromSAG.r

@@ -25,73 +25,164 @@ options(icesSAG.use_token = TRUE)
 source("../mptools/r/my_utils.r")
 
 # Set dropbox folder
-assessdir <- paste(get_dropbox(), "/iAdvice", sep="")
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# Load old SAG download data
+iSAGstock_astext_complete     <- get(load(file=paste(dropboxdir,"/rdata/iSAGstock_astext_complete.RData", sep="")))
+iSAGrefpoints_astext_complete <- get(load(file=paste(dropboxdir,"/rdata/iSAGrefpoints_astext_complete.RData", sep="")))
+load(file=paste(dropboxdir, "/rdata/iRename.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iStockkey.RData",sep=""))
+
+# set date
+today <- format(Sys.time(), '%Y%m%d')
+
+# =====================================================================================
+# which year (set year = 0 for all years)
+# =====================================================================================
+
+myyear <- 2019
 
 # =====================================================================================
 # generate list of assessmentkeys
 # =====================================================================================
 
-assessmentkeys <- sort(findAssessmentKey())
+assessmentkeys <- sort(findAssessmentKey(year=myyear))
+
+# assessmentkeys <- sort(findAssessmentKey())
+# assessmentkeys[assessmentkeys == 10362]
+# assessmentkeys <- findAssessmentKey(year=2013, full=TRUE)
 # getStockDownloadData(7306)
 
 # =====================================================================================
-# Download all assessments in SAG in full
+# Download assessments from SAG
 # =====================================================================================
 
-# Download all the stock data - takes a long time
+# Download the stock data; use lapply on the assessmentkeys with method getStockDownloadData and then bind the rows. Does not work because of type mismatches for variables
+# t1 <-
+#   do.call(bind.rows, lapply(assessmentkeys, getStockDownloadData)) %>%
+#   lowcase() %>%
+#   dplyr::select(-starts_with('custom')) %>%
+#   distinct()
+
[…]

### 2019-03-23 R/Plot all assessments for all stocks.r

@@ -20,6 +20,8 @@ advicedir <- paste(get_dropbox(), "/iAdvice", sep="")
 # load the data
 load(file=paste(advicedir,"/rdata/iAssess.RData", sep=""))
 
+sort(unique(iAssess$stockkeylabelold))
+
 # ---------------------------------------------------------------------------------------------
 # Historic retros: plot stock data over different assessment years 
 # ---------------------------------------------------------------------------------------------
@@ -33,7 +35,8 @@ iAssess %>%
   filter(!is.na(stockkeylabelold)) %>%
   filter(!is.na(stocksize)) %>%
   
-  filter(!(stockkeylabelold %in% c("ghl-arct", "san-shet", "san-scow"))) %>% 
+  # filter(!(stockkeylabelold %in% c("ghl-arct", "san-shet", "san-scow"))) %>% 
+  filter(!(stockkeylabelold %in% c("cod-scow", "had-scow", "whg-scow", "her-vian"))) %>% 
   
   filter(year > 1980, year <= assessmentyear, year >= assessmentyear - 15) %>% 
   
@@ -148,3 +151,59 @@ iAssess %>%
   # xlim(2005,2020) +
   labs(x = NULL, y = NULL , title = "Recruitment")  +
   facet_wrap(~ stockkeylabelold, scales="free_y")
+
+
+# ---------------------------------------------------------------------------------------------
+# Historic retros: plot specific stock data over different assessment years 
+# ---------------------------------------------------------------------------------------------
+
+iAssess %>% 
+  mutate(stockkeylabelold = ifelse(is.na(stockkeylabelold), stockkeylabel, stockkeylabelold)) %>% 
+  mutate(decade = as.character(10*floor(assessmentyear/10))) %>% 
+  
+  filter(purpose              == "advice") %>%   # this ignores situations with "initial advice"
+  filter(stocksizeunits       == "tonnes") %>% 
+  filter(stocksizedescription == "ssb") %>% 
+  filter(!is.na(stockkeylabelold)) %>%
+  filter(!is.na(stocksize)) %>%
+  
+  # filter(!(stockkeylabelold %in% c("ghl-arct", "san-shet", "san-scow"))) %>% 
+  filter(stockkeylabelold %in% c("cod-scow", "had-scow", "whg-scow", "her-vian")) %>% 
+  
[…]

### 2019-03-23 R/Plot of number of assessments per stock.R

@@ -4,6 +4,7 @@
 # 25/09/2017 taken from ices SAG data plot
 # 16/10/2017 now includes datapublished
 # 15/11/2018 updated for new database layout
+# 23/03/2019 updated for new database again; only source is working so far
 # -----------------------------------------------------------------------------------------------
 
 library(tidyverse) # for piping and easy coding
@@ -20,11 +21,16 @@ library(lubridate)
 source("../mptools/r/my_utils.r")
 
 # Set working directory to dropbox folder
-advicedir <- paste(get_dropbox(), "/iAdvice", sep="")
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
 # Load dataset
-load(file=paste(advicedir, "/rdata/iAssess.RData",sep=""))
-load(file=paste(advicedir, "/rdata/iSpecies.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iSpecies.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iSAG.RData",sep=""))
+
+load(file=paste(dropboxdir, "/rdata/iRename.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iStockkey.RData",sep=""))
+load(file=paste(dropboxdir, "/rdata/iSpecies.RData",sep=""))
 
 # ---------------------------------------------------------------------------------------------
 # plots from iAssess by assessment year and purpose
@@ -35,12 +41,18 @@ splitter <- 80
 x <-
   # iAdvice %>% 
   iAssess %>% 
+  # iSAGstock %>% 
+  # dplyr::select(-stockkey, -icesareas) %>% 
+  # left_join(iRename[,c("stockkeylabel","stockkey")], by="stockkeylabel") %>%
+  # left_join(iStockkey, by="stockkey") %>% 
+  
   mutate_at(c("stockkeylabel","stockkeylabelold"), funs(tolower)) %>% 
   filter(!grepl("nep", stockkeylabel)) %>% 
   filter(!grepl("^rj", stockkeylabel)) %>% 
   
   # filter(speciesfaocode %in% c("cod","her","hom","mac","whb","ple","sol")) %>% 
   # filter(speciesfaocode %in% c("mac")) %>% 
+  # filter(speciesfaocode %in% c("her")) %>% 
   
[…]

### 2019-03-23 R/Table of assessment models.R

@@ -1,0 +1,80 @@
+# -----------------------------------------------------------------------------------------------
+# Table of assessment models
+#
+# 19/11/2018 first coding
+# 22/03/2019 adapted for assessment models
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(ggthemes)  # for themes
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(lubridate)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+advicedir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# Load dataset
+load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+
+# ---------------------------------------------------------------------------------------------
+# table of number of assessments by year and species
+# ---------------------------------------------------------------------------------------------
+
+iAdvice %>% 
+  ungroup() %>% 
+  mutate_at(c("stockkeylabel","stockkeylabelold"), funs(tolower)) %>% 
+  mutate(speciesfaocode   = substr(stockkeylabel, 1, 3) ) %>% 
+  mutate(assessmentmodel=tolower(assessmentmodel)) %>% 
+  distinct(stockkey, stockkeylabel, stockkeylabelold, assessmentyear, purpose, published, assessmentmodel, .keep_all = FALSE) %>% 
+  group_by(assessmentmodel) %>% 
+  summarize(n = n()) %>% 
+  filter(nchar(assessmentmodel) <= 15) %>% 
+  arrange(-n) %>% 
+  write.csv(., file="assessmentmodel.csv", row.names = FALSE)
+
+  pandoc.table(.,
+               style = "simple",
+               split.tables=400, 
+               justify = "left",
+               missing=".",
+               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)) 
+
+
+# ---------------------------------------------------------------------------------------------
+# table of number of assessments by year and species
+# ---------------------------------------------------------------------------------------------
+
+qcsexcel %>%
+  
+  ungroup() %>% 
[…]

### 2019-03-23 R/UploadDataIntoSAG.r

@@ -1,110 +1,0 @@
-# ============================================================================
-# UPload Assessment Data in SAG
-#
-# 21/03/2018 First coding of adding historical data to standardgraphs
-# 27/04/2018 Change to Historical in the Purpose Field
-# 15/11/2018 Change just the comments of an assessment (trial)
-# 19/03/2019 Updated for HAWG 2019
-# ============================================================================
-
-rm(list=ls())
-
-library(icesSAG)  # devtools::install_github("ices-tools-prod/icesSAG")
-library(tidyverse)
-
-# use token
-options(icesSAG.use_token = TRUE)
-
-# Load utils code
-source("../mptools/r/my_utils.r")
-
-# Set dropbox folder
-advicedir <- paste(get_dropbox(), "/iAdvice", sep="")
-
-# Get the assessment data and convert to dataframe
-load("//community.ices.dk@SSL/DavWWWRoot/ExpertGroups/HAWG/2019 Meeting Docs/06. Data/NSAS/NSH_HAWG2019_sf.Rdata")
-
-# rename the FLStock and FLSAM objects
-STK     <- NSH
-STK.sam <- NSH.sam
-
-# Set years and ranges
-FiY   <- as.numeric(min(dimnames(STK)$year))
-DtY   <- as.numeric(max(dimnames(STK)$year))
-LaY   <- as.numeric(max(dimnames(STK.sam)$year))
-nyrs  <- ((DtY)-(FiY))+1
-nyrs2 <- ((LaY)-(FiY))+1
-
-# Meta information
-stockkeylabel  <- "her.27.3a47d"
-assessmentyear <- 2019
-contactperson  <- "benoit.berges@wur.nl"
-
-# SSB in intermediate year 
-SSBint <- 1529000
-
-# Create the input data for uploading  
-info     <- stockInfo(
-  StockCode      = stockkeylabel, 
-  AssessmentYear = assessmentyear, 
-  ContactPerson  = contactperson)
-
-info$StockCategory             <- "1"
-info$MSYBtrigger               <- 1400000
-info$Blim                      <- 800000
-info$Bpa                       <- 900000
-info$Flim                      <- 0.34
-info$Fpa                       <- 0.30
[…]

### 2019-03-23 R/UploadHistoricalDataIntoSAG.r

@@ -1,0 +1,112 @@
+# ============================================================================
+# UPload Historical Data into SAG
+#
+# 21/03/2018 First coding of adding historical data to standardgraphs
+# 27/04/2018 Change to Historical in the Purpose Field
+# 15/11/2018 Change just the comments of an assessment (trial)
+# 19/03/2019 Updated for HAWG 2019
+# 22/03/2019 Updated after HAWG 2019
+# ============================================================================
+
+rm(list=ls())
+
+library(icesSAG)  # devtools::install_github("ices-tools-prod/icesSAG")
+library(tidyverse)
+
+# use token
+options(icesSAG.use_token = TRUE)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# load the data
+load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+
+# Select a stock and year
+STK   <- filter(iAssess, stockkeylabelold == "her-47d", assessmentyear == 1983)
+STK2  <-
+  distinct(STK, stockkeylabelold, assessmentyear, purpose) %>% 
+  left_join(iAdvice, by=c("stockkeylabelold", "assessmentyear", "purpose"))
+
+# Set years and ranges
+FiY   <- min(STK$year)
+LaY   <- max(STK$year)
+nyrs  <- ((LaY)-(FiY))+1
+
+# Meta information
+stockkeylabel  <- unique(STK$stockkeylabelold)
+assessmentyear <- unique(STK$assessmentyear)
+contactperson  <- "martin.pastoors@gmail.com"
+
+# ------------------------------------------------------------------------------------------
+# Create the input data for uploading  
+# ------------------------------------------------------------------------------------------
+
+info     <- stockInfo(
+  StockCode      = stockkeylabel, 
+  AssessmentYear = assessmentyear, 
+  ContactPerson  = contactperson)
+
+info$StockCategory             <- ""
+info$MSYBtrigger               <- STK2$msybtrigger
+info$Blim                      <- STK2$blim
+info$Bpa                       <- STK2$bpa
+info$Flim                      <- STK2$flim
+info$Fpa                       <- STK2$fpa
+info$FMSY                      <- STK2$fmsy
+info$Fage                      <- STK2$fage 
[…]

### 2019-03-23 R/dev/Save token.r

@@ -1,4 +1,4 @@
 cat("# Standard Graphs personal access token",
-    "SG_PAT=421de31d-f55a-432a-8968-6310172aa3ad",
+    "SG_PAT=2dcb38ff-e500-4865-8271-2e2a2d4a9701",
     sep = "\n",
     file = "~/.Renviron_SG")

### 2019-03-23 R/dev/calculate Fcv and Fphi.r

@@ -1,0 +1,40 @@
+# -----------------------------------------------------------------------------------------------
+# calculate Fcv and Fphi from retrospective data
+#
+# 27/02/2019 first coding
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse) # for piping and easy coding
+library(reshape2)  # reshaping data; e.g. dcast
+library(pander)    # for print tables
+library(readxl)    # read excel files
+library(cowplot)   # multiplots
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)    # scales and formatting
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set working directory to dropbox folder
+advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
+
+load(file=paste(advicedir, "/rdata/iAssess.RData",sep=""))
+
+
+# -----------------------------------------------------------------------------------------
+# Select stocks to work on
+# -----------------------------------------------------------------------------------------
+
+iAssess %>% filter(grepl("her", stockkeylabelold)) %>% distinct(stockkeylabelold)
+
+d <-
+  iAssess %>%
+  
+  filter(grepl("her-vian|her-67bc", stockkeylabelold)) %>% 
+  filter(assessmentyear >= 2000) %>% 
+  filter(!grepl("benchmark|exploratory", purpose)) %>% 
+  filter(fishingpressureunits %in% c("per year","year-1")) 
+
+## NEED TO BE FINISHED FROM HERE
+
+

### 2019-03-23 R/iAssess figure historic retro by stock.r

@@ -50,11 +50,14 @@ d <-
   # filter(grepl("ple-nsea", stockkeylabelold) ) %>% 
   # filter(grepl("cod-347d", stockkeylabelold) ) %>% 
   
-  # filter(grepl("spr-nsea", stockkeylabelold) | grepl("spr.27.3a4", stockkeylabelnew) ) %>%
-  # mutate(stockkeylabelold = "spr.27.3a4") %>% 
+  filter(grepl("spr-nsea", stockkeylabelold) | grepl("spr.27.3a4", stockkeylabelnew) ) %>%
+  mutate(stockkeylabelold = "spr.27.3a4") %>% 
+  filter(assessmentyear >= 2015) %>% 
+  mutate(recruitment = ifelse(year == assessmentyear, NA, recruitment)) %>% 
   
   # filter(grepl("her-3a22", stockkeylabelold) ) %>% 
-  filter(grepl("her-47d3", stockkeylabelold) ) %>% 
+  # filter(grepl("her-47d3", stockkeylabelold) ) %>% 
+  # filter(grepl("her-nirs", stockkeylabelold) ) %>% 
   # filter(grepl("hom-west", stockkeylabelold) ) %>% 
   
   # filter(grepl("mac-nea|hom-west|whb-comb|her-noss", stockkeylabelold) ) %>% 
@@ -66,7 +69,9 @@ d <-
   # filter(year             >  2000) %>%  
   # filter(year             <= assessmentyear) %>% 
   # filter(assessmentyear   >  1980) %>% 
-  filter(assessmentyear   >  2017) %>% 
+  filter(assessmentyear   >  2012) %>% 
+  
+  filter(purpose %in% c("advice", "update")) %>% 
   
   select(assessmentyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
          purpose, 
@@ -131,7 +136,8 @@ p1 <-
   filter(!is.na(stocksize)) %>%  
   # filter(grepl("2016", tyear)) %>% 
   # filter(purpose == "withdrawn") %>% 
-  
+  # View()
+
   ggplot(aes(year,stocksize, group=paste0(tyear,purpose))) +
   
   theme_publication() +
@@ -142,15 +148,15 @@ p1 <-
         # strip.background = element_blank(),
         legend.position = "null") +
   
-  geom_ribbon(aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill = purpose), alpha=0.3, inherit.aes = FALSE) +
[…]

### 2019-03-23 R/iAssess figure trends for selected stocks.r

@@ -1,0 +1,100 @@
+# -----------------------------------------------------------------------------------------------
+# ICES Stock Assessment Graph plotting
+#
+# 30/03/2017 first coding during HAWG
+# 14/07/2017 adapted during HERAS
+# 11/08/2017 adapter for R 3.4.1 and tidyverse
+# 14/08/2017 added plot for assessment methods
+# 04/10/2017 only plot for retro. Updated for iAssess
+# 07/03/2019 adapted for new database layout
+# -----------------------------------------------------------------------------------------------
+
+rm(list=ls())
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# load the data
+load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+
+# filter(iAssess, stockkeylabel=="her.27.6a7bc", assessmentyear==2017) %>% View()
+# filter(sagdata, stockkeylabel %in% c("her-noss","her.27.1-24a514a"), assessmentyear==2017) %>% 
+#   View()
+
+# ---------------------------------------------------------------------------------------------
+# Historic retros: plot stock data over different assessment years 
+# ---------------------------------------------------------------------------------------------
+
+# iAssess %>% filter(assessmentyear == 2019, stockkeylabelold=="her-67bc") %>% View()
+# d %>% filter(assessmentyear == 2019) %>% distinct(stockkeylabelold, purpose) %>% arrange(stockkeylabelold) %>% View()
+
+d <-
+  iAssess %>% 
+
+  # filter(grepl("her-irls|her-nirs", stockkeylabelold) ) %>%
+  filter(grepl("her-67bc|her-3a22|her-vian|her-47d3", stockkeylabelold) ) %>%
+  # filter(grepl("her-irls|her-67bc|her-3a22|her-47d3|her-nirs", stockkeylabelold) ) %>%
+  filter(assessmentyear   %in%  2014:2019) %>% 
+  filter(year             >   1992) %>% 
+  filter(purpose %in% c("advice", "update", "interbenchmark")) %>%
+  
+  ungroup() %>% 
+  
[…]

### 2019-03-23 R/iDatabases generator.r

@@ -35,6 +35,8 @@
 # 08/11/2018 updated files on qcsexcel and iadvice; no longer need iStock
 # 20/11/2018 small updates; making sure the database is consistently filled
 # 26/11/2018 added comparison to the data used by Esther Schuch
+# 20/03/2019 adapted with new data during HAWG
+# 23/03/2019 adapated after change to download data procedure
 # -----------------------------------------------------------------------------------------------
 
 # rm(list=ls())
@@ -51,7 +53,7 @@ library(directlabels)  # for printing labels at end of geom lines
 source("../mptools/r/my_utils.r")
 
 # Set working directory to dropbox folder
-advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
+dropboxdir  <- paste(get_dropbox(), "/iAdvice", sep="")
 
 # -----------------------------------------------------------------------------------------
 # load the Advice database
@@ -59,7 +61,7 @@ advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
 
 iAdvice <-
   readxl::read_excel(
-    path= paste(advicedir, "/Excel/ICES Scientific Advice database.xlsm", sep=""), 
+    path= paste(dropboxdir, "/Excel/ICES Scientific Advice database.xlsm", sep=""), 
     sheet     = "DATA",
     col_names = TRUE, 
     col_types = "text", 
@@ -89,7 +91,7 @@ iAdvice <-
     adviceonstock = as.logical(adviceonstock)
   )
   
-save(iAdvice, file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+save(iAdvice, file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
 
 # glimpse(iAdvice)
 
@@ -99,7 +101,7 @@ save(iAdvice, file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
 
 iSpecies <-
   readxl::read_excel(
-    path=paste(advicedir, "/excel/species_list.xlsx",sep=""), 
+    path=paste(dropboxdir, "/excel/species_list.xlsx",sep=""), 
     col_names=TRUE, 
     col_types="text") %>%
   mutate_at(c("speciescommonname","trophicguild","fisheriesguild","sizeguild"), 
@@ -107,7 +109,7 @@ iSpecies <-
[…]

### 2019-03-23 R/old/changeHistoricalDataInSAG.r

@@ -61,6 +61,7 @@ info$FishingPressureUnits      <- "Year-1"
 info$StockSizeDescription      <- "SSB"
 info$StockSizeUnits            <- unique(rby$stockSizeUnits[!is.na(rby$stockSizeUnits)])
 info$Purpose                   <- "Advice"
+info$ModelName                 <- "SAM"
 # info$CustomLimitName1         <- "MBAL"
 # info$CustomLimitValue1        <- 800000
 

### 2019-03-23 R/old/plot TAC area.r

@@ -31,53 +31,53 @@ dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 # set world in sf format
 # ----------------------------------------------------------------------------------------------------------
 
-world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE, 
+world_sf <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE, 
                            regions = c("netherlands","belgium","France(?!:Corsica)",
                                        "ireland","united kingdom", "UK", "Denmark","Germany",
                                        "norway","iceland", "greenland", "faroe islands", "spain","portugal",
                                        "sweden", "finland","poland")))
-# ggplot(data=world1) + geom_sf()
+# ggplot(data=world_sf) + geom_sf()
 
 
 # ----------------------------------------------------------------------------------------------------------
 # load the iAdvice data
 # ----------------------------------------------------------------------------------------------------------
 
-load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
-load(file.path(advicedir,"rdata/fao_iho.RData"))
-
-fao_iho %>% filter(grepl("27.2.a", name)) %>% View()
-
+load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+load(file.path(dropboxdir,"rdata/fao_iho.RData"))
 
 # ----------------------------------------------------------------------------------------------------------
-# generate the stk object: stock, assessmentyear and area(s) 
+# generate the tac object: stock, assessmentyear and area(s) 
 # ----------------------------------------------------------------------------------------------------------
-
 
 tac <-
   iAdvice %>% 
   
-  filter(speciesfaocode == "ple") %>%
-  filter(grepl("27.2|27.3|27.4", tacarea)) %>% 
+  # filter(speciesfaocode == "ple") %>%
+  # filter(grepl("27.2|27.3|27.4", tacarea)) %>% 
+  
+  filter(speciesfaocode == "her") %>%
+  filter(grepl("27.1", tacarea)) %>% 
+  
   filter(assessmentyear %in% 2010:2018) %>%
   distinct(fishstock = stockkeylabelold, stockkey, assessmentyear, tacarea) %>%
-  mutate(toupper(tacarea)) %>% 
+  mutate(tacarea = toupper(tacarea)) %>% 
   
[…]

### 2019-03-23 R/pdf analysis of advice.r

@@ -9,16 +9,24 @@ library(tidyverse)
 # loop over years
 for (y in 1981:2018) {
   assign(paste("results",y,sep=""), {
-    print(y)
-    keyword_directory(file.path("E:/PA",y),
-                      keyword = c('precautionary approach', 
-                                  'biological reference point',
-                                  "minimum biologically acceptable", "mbal", 
-                                  "safe biological limit",
-                                  "maximum sustainable yield", "msy",
-                                  "btrigger", "bpa", "blim", "fpa", "fmsy"),
-                      # file_pattern   = ".pdf"
-                      ignore_case    = TRUE, recursive = FALSE, surround_lines = 1, full_names     = TRUE) %>% 
+    
+    # print(y)
+    print(file.path("F:/ices/advice",y))
+    
+    keyword_directory(file.path("F:/ices/advice",y),
+                      keyword = c('robust'),
+                      file_pattern   = ".pdf",
+                      ignore_case    = TRUE, recursive = TRUE, 
+                      surround_lines = 1, full_names     = TRUE) %>% 
+      # keyword_directory(file.path("E:/PA",y),
+      #                 keyword = c('precautionary approach', 
+      #                             'biological reference point',
+      #                             "minimum biologically acceptable", "mbal", 
+      #                             "safe biological limit",
+      #                             "maximum sustainable yield", "msy",
+      #                             "btrigger", "bpa", "blim", "fpa", "fmsy"),
+      #                 # file_pattern   = ".pdf"
+      #                 ignore_case    = TRUE, recursive = FALSE, surround_lines = 1, full_names     = TRUE) %>% 
       mutate(year = y) 
     
   })

### 2019-03-23 R/plot herring areas.r

@@ -1,0 +1,257 @@
+# --------------------------------------------------------------------------------------------
+# plot herring areas.r
+#
+# plotting stock areas (over time)
+# 
+# 27/04/2018 cleaned up the code
+# 28/11/2018 updated the code to work with new iAdvice database
+# 19/12/2018 using new FAO_IHO dataset for international and EEZ waters
+# --------------------------------------------------------------------------------------------
+
+# devtools::install_github("fishvice/tidyices", dependencies = FALSE)
+
+# devtools::install_github("einarhjorleifsson/iceshape", dependencies = FALSE)
+# devtools::install_github("einarhjorleifsson/ggmisc", dependencies = FALSE)
+library(iceshape)   # ls("package:iceshape")
+
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(viridis)
+
+source("../mptools/r/my_utils.r")
+# load("../prf/rdata/world.df.RData")
+
+# set onedrive directory
+onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 
+
+# ----------------------------------------------------------------------------------------------------------
+# load spatial objects in sf format
+# ----------------------------------------------------------------------------------------------------------
+
+world.sf <- get(load(file.path(onedrive,"rdata/world.sf.RData"))) %>% 
+  rownames_to_column()
+
+world2.sf <- get(load(file.path(onedrive,"rdata/world2.sf.RData"))) %>% 
+  rownames_to_column()
+
+fao.eez.sf <-
+  get(load(file.path(onedrive,"rdata/fao.eez.sf.RData"))) 
+
+fao.sf <-
+  get(load(file.path(onedrive,"rdata/fao.sf.RData"))) 
+
+load(file.path(onedrive,"rdata/iho.eez.sf.RData"))
+
+# ----------------------------------------------------------------------------------------------------------
+# load herring data
+# ----------------------------------------------------------------------------------------------------------
+
+herring_areas <- 
+  readxl::read_excel(path="D:/HAWG/Herring areas.xlsx", col_names=TRUE, col_types="text") %>% 
+  # make "area" into a long table
[…]

### 2019-03-23 R/plot stock and TAC area.r

@@ -1,0 +1,203 @@
+# --------------------------------------------------------------------------------------------
+# plot Stock and TAC area.r
+#
+# plotting stock areas (over time)
+# 
+# 27/04/2018 cleaned up the code
+# 28/11/2018 updated the code to work with new iAdvice database
+# 19/12/2018 using new FAO_IHO dataset for international and EEZ waters
+# --------------------------------------------------------------------------------------------
+
+# devtools::install_github("fishvice/tidyices", dependencies = FALSE)
+
+# devtools::install_github("einarhjorleifsson/iceshape", dependencies = FALSE)
+# devtools::install_github("einarhjorleifsson/ggmisc", dependencies = FALSE)
+library(iceshape)   # ls("package:iceshape")
+
+library(sf)
+library(tidyverse)
+library(readxl)
+library(maps)
+library(ggplot2)
+library(viridis)
+
+source("../mptools/r/my_utils.r")
+# load("../prf/rdata/world.df.RData")
+
+# get dropbox directory
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# set onedrive directory
+onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 
+
+
+# ----------------------------------------------------------------------------------------------------------
+# load the iAdvice data
+# ----------------------------------------------------------------------------------------------------------
+
+load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+
+# ----------------------------------------------------------------------------------------------------------
+# load spatial objects in sf format
+# ----------------------------------------------------------------------------------------------------------
+
+world.sf <- get(load(file.path(onedrive,"rdata/world.sf.RData"))) %>% 
+  rownames_to_column()
+
+world2.sf <- get(load(file.path(onedrive,"rdata/world2.sf.RData"))) %>% 
+  rownames_to_column()
+
+fao.eez.sf <-
+  get(load(file.path(onedrive,"rdata/fao.eez.sf.RData"))) 
+
+fao.sf <-
+  get(load(file.path(onedrive,"rdata/fao.sf.RData"))) 
+
+load(file.path(onedrive,"rdata/iho.eez.sf.RData"))
+
[…]

### 2019-05-06 R/Analyse changes in pelagic stocks.Rmd
General commit of changes

@@
 # 10/01/2018 made small changes to the table
+# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
 # -----------------------------------------------------------------------------------------------
@@
                stockkeylabelold)) %>%
-  filter(!grepl("benchmark", purpose)) %>% 
+
+  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
+  group_by(stockkeylabelold, assessmentyear) %>% 
+  mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
+  # filter(n == 1 | ((n > 1 & adviceonstock == TRUE) | (n > 1 & benchmark == TRUE))) %>% 
+  filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
+  # filter(!grepl("benchmark", purpose)) %>%
+
   filter(assessmentyear >= 2008) %>% 
   ungroup()
+
+# d %>% filter(stockkeylabelold =="mac-nea", assessmentyear >= 2015) %>% View()
 
@@
   iAssess %>% 
+  group_by(assessmentyear) %>% 
   filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
-  filter(assessmentyear == 2018) %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+  filter(year <= assessmentyear) %>%
+
+  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
+  group_by(stockkeylabelold, assessmentyear) %>% 
+  mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
+  filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
   filter(year >= 2008) %>% 
-  filter(year <= assessmentyear) %>% 
-  filter(adviceonstock == TRUE)
+  ungroup()
+#   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) )
+
+# d %>%
+#   ungroup() %>%
+#   distinct(stockkeylabelold, assessmentyear,  benchmark, adviceonstock, n, .keep_all = FALSE) %>%
+#   arrange(stockkeylabelold, assessmentyear, benchmark, adviceonstock, n) %>%
+#   View()
 
@@
   gather(key=var, value=value, fpa:fmsy) %>% 
+  # View()
 
@@
   dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>%
+
   ggplot(aes(x=tacyear)) +
@@
   scale_color_manual(values = c("red", "blue")) +
-  scale_linetype_manual(values = c("solid", "blank")) +
+  # scale_linetype_manual(values = c("solid", "blank", "dotted")) +
   facet_wrap(~stockkeylabelold, scales="free_y")
@@
   rename(year = assessmentyear) %>% 
+  View()
 
@@
   scale_color_manual(values = c("red", "blue")) +
+  facet_wrap(~stockkeylabelold, scales="free_y")
+```
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+# now based on assessment
+t <-
+  e %>% 
+  filter(year == assessmentyear - 1 ) %>% 
+  mutate(colour  = paste0(adviceonstock,benchmark),
+         purpose = ifelse(grepl("bench", purpose), "benchmark", purpose),
+         purpose = ifelse(grepl("initial", purpose), "replaced", purpose))
+
+glimpse(d)
+
+e %>% 
+  group_by(stockkeylabelold) %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+  dplyr::select(stockkeylabelold, assessmentyear, year, adviceonstock, stocksize) %>% 
+  # rename(year = assessmentyear) %>% 
+  # View()
+
+  ggplot(aes(x=year)) +
+  theme_publication() +
+  theme(legend.title = element_blank()) +
+
+  # geom_point(aes(y=ssbay,
+  #                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
+  geom_line(aes(y = stocksize )) +
+  geom_point(aes(y = stocksize )) +
+
+  geom_point(data=t, aes(x=year, y=stocksize, colour=purpose)) +
+  scale_x_continuous(breaks=pretty_breaks()) +
+  expand_limits(y=0) + 
+  labs(y="SSB") +
+  # scale_color_manual(values = c("red", "blue")) +
   facet_wrap(~stockkeylabelold, scales="free_y")

### 2019-05-06 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
General commit of changes

@@
 
-pairwise_comparisons <- function(removelastyears=0, numberofyears=10) {
+pairwise_comparisons <- function(removelastyears=0, numberofyears=10, threshold=0.10) {
 
@@
            phi_abs        = abs(phi),
-           assessmentyear = as.integer(substr(assessmentpair, 6,9)))
-
-  res <- list("d.pcm"= d.pcm, 
-              "d.pairwise.roughly"=d.pairwise.roughly,
-              "d.pairwise" =d.pairwise)
+           assessmentyear = as.integer(substr(assessmentpair, 6,9)), 
+           scale_jump     = ifelse(phi_abs  > threshold, TRUE, NA),
+           scale_jump     = ifelse(phi_abs <= threshold, FALSE, scale_jump)) %>% 
+    group_by(stock, variable) %>% 
+    mutate(scale_jump     = ifelse(row_number() == 1, NA, scale_jump),
+           removelastyears= removelastyears, 
+           numberofyears  = numberofyears, 
+           threshold      = threshold)
+
+  res <- list("d.pcm"              = d.pcm, 
+              "d.pairwise.roughly" = d.pairwise.roughly,
+              "d.pairwise"         = d.pairwise)
 
@@
 
-tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5)
+tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5, threshold = 0.1)
+
+# tmp$d.pcm %>% View()
 
@@
 
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
-
-threshold <- 0.10
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)
+
+# threshold <- 0.10
 
@@
 res$d.pcm %>% 
-  distinct(stock, variable, assessmentyear, phi) %>%
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
   mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
@@
   geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
   # geom_point(aes(colour=factor(decade), size = abs(value) )) +
@@
 
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
-
-threshold <- 0.10
+res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold=0.1)
 
@@
   geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
 
@@
   facet_wrap(~stock)
+
+# filter(d.assessments, benchmark == 1) %>% View()
+
+```
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot scale jumps with different settings for number of years and thresholds
+
+d.pcm <- data.frame()
+
+for (ny in c(5, 10, 15)) {
+  for (nt in c(0.05, 0.1, 0.15, 0.2)) {
+    res <- pairwise_comparisons(removelastyears = 0, numberofyears=ny, threshold=nt)
+    d.pcm <- bind_rows(d.pcm, res$d.pcm)
+  }
+}
+
+d.pcm %>% 
+  filter(stock == "cod (north sea)") %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
+
+  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  # ggMarginal(data=d.assessments, type="histogram")
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="") +
+  facet_grid(threshold ~ numberofyears)
+
+# filter(d.assessments, benchmark == 1) %>% View()
+
+```
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot number of scale jumps with different settings for number of years and thresholds
+
+d.pcm <- data.frame()
+
+for (ny in c(5, 10, 15)) {
+  for (nt in c(0.05, 0.1, 0.15, 0.2)) {
+    res <- pairwise_comparisons(removelastyears = 0, numberofyears=ny, threshold=nt)
+    d.pcm <- bind_rows(d.pcm, res$d.pcm)
+  }
+}
+
+d.pcm %>% 
+  filter(stock == "cod (north sea)") %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
+
+  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  # ggMarginal(data=d.assessments, type="histogram")
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="") +
+  facet_grid(threshold ~ numberofyears)
 
@@
 d.jumps.byyear %>% 
-  filter(assessmentyear >= 1990) %>%
+  filter(assessmentyear >= 1982) %>%
 
@@
 *`r fig_nums("jumpratio")`*
+
+```{r jumpratio, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}
+
+d.assessments%>% 
+  filter(assessmentyear <= 2018) %>% 
+  group_by(assessmentyear) %>% 
+  summarize(ncpueseries = mean(ncpueseries, na.rm=TRUE),
+            nsurveyseries = mean(nsurveyseries, na.rm=TRUE)) %>% 
+  gather(key=type, value=n, ncpueseries:nsurveyseries) %>%
+
+  ggplot(aes(x=assessmentyear, y=n, group=type)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.placement = "outside",
+    legend.title    = element_blank()
+  ) +
+  # theme(legend.position  = "none") + 
+  geom_line(aes(colour=type)) +
+  labs(x="", y="avg")
+
+```

### 2019-05-06 R/Analyse reference points.r

@@ -1,7 +1,8 @@
 # -----------------------------------------------------------------------------------------------
-# analyse benchmarks
+# analyse reference points
 #
 # 15/11/2018 first coding
+# 03/04/2019 added reference points for specific stocks
 # -----------------------------------------------------------------------------------------------
 
 library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
@@ -20,19 +21,19 @@ advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
 
 # load(file=paste(advicedir, "/rdata/iStock.RData",sep=""))
 # load(file=paste(advicedir, "/rdata/qcsexcel.RData",sep=""))
-# load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
 
 # -----------------------------------------------------------------------------------------
 # Plot of number of reference points
 # -----------------------------------------------------------------------------------------
 
 iAdvice %>% 
-  filter(assessmentyear <= 2000) %>%
+  filter(assessmentyear >= 2000) %>%
   filter(!grepl("bench", purpose)) %>% 
   
   # add the reference points from sagrefpoints
-  bind_rows(sagrefpoints) %>% 
-  filter(assessmentyear >= 1997) %>% 
+  # bind_rows(sagrefpoints) %>% 
+  # filter(assessmentyear >= 1997) %>% 
   
   group_by(assessmentyear) %>% 
   summarize(flim = sum(!is.na(flim)),
@@ -87,3 +88,38 @@ iAdvice %>%
   facet_wrap(~type)
 
 
+# -----------------------------------------------------------------------------------------
+# Plot reference points by stock
+# -----------------------------------------------------------------------------------------
+
+mystock <- "hom-west"
+
+iAdvice %>% 
+  filter(stockkeylabelold == mystock) %>% 
+  filter(assessmentyear >= 2000) %>%
+  filter(!grepl("bench", purpose)) %>% 
+  
+  # add the reference points from sagrefpoints
+  # bind_rows(sagrefpoints) %>% 
+  # filter(assessmentyear >= 1997) %>% 
+  
+  select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, assessmentyear, 
+         flim, fpa, fmsy, blim, bpa, msybtrigger) %>% 
[…]

### 2019-05-06 R/UploadHistoricalDataIntoSAG.r

@@ -97,20 +97,21 @@ fishdata$Low_FishingPressure      <- STK$lowfishingpressure
 fishdata$FishingPressure          <- STK$fishingpressure
 fishdata$High_FishingPressure     <- STK$highfishingpressure
 
+# save(info, fishdata, file="testupload.RData")
+
 # upload to SAG
 key <- icesSAG::uploadStock(info, fishdata)
 
-# save(info, fishdata, file="testupload.RData")
-
-# Get SAG settings
-# getSAGSettingsForAStock(assessmentKey=key) %>% View()
-getSAGSettingsForAStock(assessmentKey=10364) %>% View()
-
 # Add comment to SAG settings
-setSAGSettingForAStock(assessmentKey=10364, 
+setSAGSettingForAStock(assessmentKey=key, 
                        chartKey=0,
                        settingKey=21,
                        settingValue="Martin Pastoors Historical data",
                        copyNextYear=FALSE) 
 
 
+# Get SAG settings
+# getSAGSettingsForAStock(assessmentKey=key) %>% View()
+
+assessmentkeys <- sort(findAssessmentKey(year=1983))
+t1  <- unlist(getStockDownloadData(10368))

### 2019-05-06 R/iAdvice advice tac and catch.r

@@ -1,0 +1,65 @@
+# ===========================================================================
+# iAdvice refpoints.r
+#
+# 10/10/2018 initial coding
+# ===========================================================================
+
+# rm(list=ls())
+
+library(readxl)
+library(tidyverse)
+library(stringr)
+library(pander)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+
+# Source the utils file
+source("../mptools/R/my_utils.r")
+
+# set dropboxdirs
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+
+load(file=paste(dropboxdir, "/rdata/iStockkey.RData", sep=""))
+load(file=paste(dropboxdir, "/rdata/iRename.RData", sep=""))
+
+load(file=paste(dropboxdir, "/rdata/iAdvice.RData", sep=""))
+
+# ---------------------------------------------------------------------------
+# overview of %change in reference points
+# ---------------------------------------------------------------------------
+
+compare <-
+  iAdvice %>% 
+  dplyr::select(year=tacyear, stockkeylabelold, advisedlandingsmax, tal, landings, catches) %>% 
+  rename(advice=advisedlandingsmax, tac = tal) %>% 
+  filter(stockkeylabelold %in% c("her-47d3", "mac-nea","whb-comb", "her-noss")) %>% 
+  # filter(stockkeylabelold %in% c("her-47d3", "mac-nea","whb-comb", "her-noss")) %>% 
+  filter(year >= 2000) %>% 
+  gather(key=variable, value=value, advice:catches)
+
+# plot of comparison of advice, tac, landings and catch
+compare %>% 
+  ungroup() %>% 
+  mutate(variable = factor(variable, levels=c("advice", "tac", "landings", "catches"))) %>% 
+  
+  ggplot(aes(x=year, y=value, group=variable)) +
+  theme_publication() +
+  geom_path(aes(colour=factor(variable)), size=1) +
+  geom_point(aes(colour=factor(variable)), size=2) +
+  # scale_size_manual    (values=c(advice=1.5 ,tac = 1.5, landings = 1.0, catches = 1.0)) +
+  expand_limits(y = 0) +
+  facet_wrap(~stockkeylabelold, scales="free_y") 
+
+# plot of difference between tac and advice
+compare %>% 
+  filter(variable %in% c("advice","tac")) %>% 
[…]

### 2019-05-06 R/iAssess figure historic retro by stock.r

@@ -24,6 +24,8 @@ dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
 # load the data
 load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iSAGrefpoints.RData", sep=""))
 
 # filter(iAssess, stockkeylabel=="her.27.6a7bc", assessmentyear==2017) %>% View()
 # filter(sagdata, stockkeylabel %in% c("her-noss","her.27.1-24a514a"), assessmentyear==2017) %>% 
@@ -43,39 +45,41 @@ d <-
   #        assessmentyear>=2017) %>% 
   
   # filter(grepl("mac-nea", stockkeylabelold),
-  #        purpose %in% c("advice","benchmark","interbenchmark","withdrawn"),
-  #        assessmentyear>=2017) %>% 
+  #        purpose %in% c("advice","interbenchmark","withdrawn"),
+  #        assessmentyear>=2016) %>%
   
   # filter(grepl("hom-west", stockkeylabelold),purpose %in% c("update","bench"),assessmentyear>=2016) %>% 
   # filter(grepl("ple-nsea", stockkeylabelold) ) %>% 
   # filter(grepl("cod-347d", stockkeylabelold) ) %>% 
   
-  filter(grepl("spr-nsea", stockkeylabelold) | grepl("spr.27.3a4", stockkeylabelnew) ) %>%
-  mutate(stockkeylabelold = "spr.27.3a4") %>%
-  filter(assessmentyear >= 2015) %>%
-  mutate(recruitment = ifelse(year == assessmentyear, NA, recruitment)) %>%
+  # filter(grepl("spr-nsea", stockkeylabelold) | grepl("spr.27.3a4", stockkeylabelnew) ) %>%
+  # mutate(stockkeylabelold = "spr.27.3a4") %>%
+  # filter(assessmentyear >= 2015) %>%
+  # mutate(recruitment = ifelse(year == assessmentyear, NA, recruitment)) %>%
   
   # filter(grepl("her-irls", stockkeylabelold) ) %>% 
   # filter(grepl("her-67bc", stockkeylabelold) ) %>% 
-  # filter(grepl("her-47d3", stockkeylabelold) ) %>%
[…]

### 2019-05-06 R/iDatabases generator.r

@@ -37,6 +37,17 @@
 # 26/11/2018 added comparison to the data used by Esther Schuch
 # 20/03/2019 adapted with new data during HAWG
 # 23/03/2019 adapated after change to download data procedure
+
+# Warning messages:
+#   1: funs() is soft deprecated as of dplyr 0.8.0
+# please use list() instead
+# 
+# # Before:
+# funs(name = f(.)
+#      
+#      # After: 
+#      list(name = ~f(.))
+
 # -----------------------------------------------------------------------------------------------
 
 # rm(list=ls())
@@ -314,8 +325,8 @@ qcsexcel_to_merge <-
 # iAdvice (only model specifications)
 iadvice_to_merge <-
   iAdvice %>% 
-  dplyr::select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, assessmentyear, purpose, 
-                stockarea, assessmentmodel, benchmark, assessmentscale, nsurveyseries, ncpueseries, 
+  dplyr::select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, assessmentyear, assessmentdate,
+                purpose,  stockarea, assessmentmodel, benchmark, assessmentscale, nsurveyseries, ncpueseries, 
                 adviceonstock, published) %>% 
   mutate_at(c("assessmentmodel"), funs(tolower))
 

### 2019-05-06 R/old/iAdvice advice tac and catch.r

@@ -1,88 +1,0 @@
-# ===========================================================================
-# iAdvice refpoints.r
-#
-# 10/10/2018 initial coding
-# ===========================================================================
-
-# rm(list=ls())
-
-library(readxl)
-library(tidyverse)
-library(stringr)
-library(pander)
-library(directlabels)  # for printing labels at end of geom lines
-library(scales)
-
-# Source the utils file
-source("../mptools/R/my_utils.r")
-
-# set dropboxdirs
-assessdir <- paste(get_dropbox(), "/ICES assessment database", sep="")
-advicedir <- paste(get_dropbox(), "/ICES advice database", sep="")
-
-
-load(file=paste(assessdir, "/rdata/iStockkey.RData", sep=""))
-load(file=paste(assessdir, "/rdata/iRename.RData", sep=""))
-
-iAdvice <- 
-  read_excel(paste0(advicedir, "/ICES Scientific Advice database 20181001.xlsx",sep=""), 
-             sheet     = "DATA",
-             col_names = TRUE,
-             col_types = "text",
-             skip      = 0)  %>% 
-  lowcase() %>% 
-  mutate_at( c("year", "assessyear", "ncpueseries","nsurveyseries"), 
-             funs(as.integer)) %>% 
-  mutate_at( c("advisedlandingsmax","advisedcatchmax","tal","tac",
-               "officiallandings","iceslandings","icesindustrialbycatch","icesdiscards","icescatch",
-               "fsqymin1","ssbymin1","fadvmax","fmax","f01","fmed","f35spr","flim","fpa","fmsy",
-               "blim","bpa","msybtrig"), funs(as.numeric)) %>% 
-  mutate( blim     = 1000 * blim, 
-          msybtrig = 1000 * msybtrig) %>% 
-  
-  mutate_at( c("stockices","speciestype", "assessmodel", 
-               "multispeciesmanagement", "stockmanagement", "adviceonstock", 
-               "aggregated"), funs(tolower)) %>% 
-  rename(stockkeylabelold = stockices, 
-         assessmentyear = assessyear) %>% 
-  arrange(stockkeylabelold, tacarea, year) %>%  
-  mutate(species    = substr(stockkeylabelold,1,3) ) %>% 
-  left_join(iStockkey, by="stockkeylabelold") 
-
-
-
-
-
-# ---------------------------------------------------------------------------
[…]

### 2019-05-06 R/plot herring areas.r

@@ -115,213 +115,3 @@ herring_areas %>%
   # facet_grid(variable~assessmentyear)
 
 
-# ----------------------------------------------------------------------------------------------------------
-# generate FAO overview
-# ----------------------------------------------------------------------------------------------------------
-
-fao.sf %>%
-  filter(grepl("27.6", F_CODE)) %>%
-
-  mutate(F_CODE = str_sub(F_CODE, start=4)) %>% 
-  
-  ggplot() +
-  theme_publication() +
-  theme(panel.spacing    = unit(0.1, "lines"),
-        text             = element_text(size=8),
-        axis.text        = element_blank(),
-        axis.ticks       = element_blank(),
-        strip.background = element_blank(),
-        strip.text       = element_text(size=8, face="bold", hjust=0.5, margin = margin(0.1,0,0,0, "mm")),
-        legend.title     = element_blank()) +
-  theme(legend.position = "none") +
-  geom_sf(aes(fill=F_CODE)) +
-  geom_sf_text(aes(label = F_CODE)) +
-  facet_wrap(~F_LEVEL)
-
-
-# ----------------------------------------------------------------------------------------------------------
-# generate FAO_EEZ overview
-# ----------------------------------------------------------------------------------------------------------
-
-fao.eez.sf %>%
-  filter(grepl("27.6|27.7", F_CODE)) %>%
-  data.frame() %>% 
-  mutate(F_CODE = str_sub(F_CODE, start=4)) %>% 
-
-  ggplot() +
-  theme_publication() +
-  theme(panel.spacing    = unit(0.1, "lines"),
-        text             = element_text(size=8),
-        axis.text        = element_blank(),
-        axis.ticks       = element_blank(),
-        strip.background = element_blank(),
-        strip.text       = element_text(size=8, face="bold", hjust=0.5, margin = margin(0.1,0,0,0, "mm")),
-        legend.title     = element_blank()) +
-  theme(legend.position = "none") +
-  geom_sf(aes(fill=F_CODE)) +
-  geom_sf_text(aes(label = F_CODE)) +
-  facet_wrap(~F_LEVEL)
-
-
-# world.sf %>%
[…]

### 2019-05-06 R/plot stock and TAC area.r

@@ -1,11 +1,12 @@
 # --------------------------------------------------------------------------------------------
-# plot Stock and TAC area.r
+# plot stock and tac areas.r
 #
-# plotting stock areas (over time)
+# plotting stock and tac areas (over time); also overview of FAO and EEZ areas
 # 
 # 27/04/2018 cleaned up the code
 # 28/11/2018 updated the code to work with new iAdvice database
 # 19/12/2018 using new FAO_IHO dataset for international and EEZ waters
+# 28/03/2019 updated after HAWG
 # --------------------------------------------------------------------------------------------
 
 # devtools::install_github("fishvice/tidyices", dependencies = FALSE)
@@ -24,17 +25,15 @@ library(viridis)
 source("../mptools/r/my_utils.r")
 # load("../prf/rdata/world.df.RData")
 
-# get dropbox directory
-dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+# source("r/iDatabases generator.r")
 
 # set onedrive directory
 onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 
 
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
-# ----------------------------------------------------------------------------------------------------------
-# load the iAdvice data
-# ----------------------------------------------------------------------------------------------------------
-
+# Load dataset
 load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
 
 # ----------------------------------------------------------------------------------------------------------
@@ -47,127 +46,11 @@ world.sf <- get(load(file.path(onedrive,"rdata/world.sf.RData"))) %>%
 world2.sf <- get(load(file.path(onedrive,"rdata/world2.sf.RData"))) %>% 
   rownames_to_column()
 
-fao.eez.sf <-
-  get(load(file.path(onedrive,"rdata/fao.eez.sf.RData"))) 
-
 fao.sf <-
   get(load(file.path(onedrive,"rdata/fao.sf.RData"))) 
 
-load(file.path(onedrive,"rdata/iho.eez.sf.RData"))
-
-# ----------------------------------------------------------------------------------------------------------
[…]

### 2019-05-16 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Updated paper; fixed errors in database generator

@@
   # Exclude prediction years (but keep the benchmark information that has no year attached to it)
-  filter(is.na(year) | (year <= assessmentyear & year >= 1980) ) %>% 
+  filter(is.na(year) | (year <= assessmentyear) ) %>% 
   filter(assessmentyear >= 1982) %>% 
+  filter(assessmentyear <= 2018) %>% 
   filter(is.na(year) | !(is.na(stocksize) & is.na(fishingpressure))) %>%
@@
   # Keep this variable
-  filter( variable %in% c("stocksize")) %>%
-
-  arrange(stock, variable, (assessmentyear)) %>% 
+  filter( variable %in% c("stocksize", "recruitment", "fishingpressure")) %>%
+
+  arrange(stock, assessmentyear, variable, year) %>% 
   ungroup() %>%
@@
   d.selection %>% 
-  filter(assessmentyear == 2018) %>% 
+  ungroup() %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
   ungroup() %>% 
@@
   d.selection %>% 
-  filter( (assessmentyear != 2018 & year == max(year)) ) %>%
+  ungroup() %>% 
+  filter( (assessmentyear != max(assessmentyear, na.rm=TRUE) ) ) %>%
+  group_by(stock, assessmentyear, purpose, variable) %>% 
+  filter(year == max(year)) %>%
 
@@
 d.points <-
+  d.segment %>% 
+  filter( (assessmentyear != 2018 & year == max(year)) )
+
+# Calculations of Pairwise Comparison of Mean and scale jumps ==========================================
+
+# Create a set of assessment pairs: two subsequent assessments  
+d.pairs <-
   d.selection %>% 
-  filter( (assessmentyear != 2018 & year == max(year)) )
+  distinct(stock, assessmentyear, variable) %>% 
+  group_by(stock, variable) %>% 
+  arrange(stock, variable, assessmentyear) %>% 
+  mutate(assessmentyearprior = lag(assessmentyear),
+         assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
+  rename(assessmentyearlast  = assessmentyear) %>% 
+  filter(!is.na(assessmentyearprior))
+
+# data series with most recent (last) estimation of the pair
+d.pairs.last <-
+  d.pairs %>% 
+  select(-assessmentyearprior) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  mutate(label="last") %>% 
+  rename(assessmentyear = assessmentyearlast)
+
+# data series with previous (prior) estimation of the pair
+d.pairs.prior <-
+  d.pairs %>% 
+  select(-assessmentyearlast) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  mutate(label="prior") %>% 
+  rename(assessmentyear = assessmentyearprior)
+
+# removelastyears=0; numberofyears=c(5,10); threshold=0.10
+
+pairwise_comparisons <- function(removelastyears=0, numberofyears=10, threshold=0.10) {
+
+  # rly <- 0
+  # ny <- 5
+  # th <- 0.20
+
+  d.pairwise.roughly <- data.frame()
+
+  for (rly in removelastyears) {
+    for (ny in numberofyears) {
+
+      print(paste("removelastyears:",rly, "numberofyears:",ny,sep=" "))
+
+      # pairwise data roughly
+      d.pairwise.roughly <-
+        bind_rows(d.pairs.prior, d.pairs.last) %>%
+
+        # filter the appropriate period
+        group_by(stock, assessmentpair, variable) %>% 
+        # filter(year <= as.numeric(substr(assessmentpair,1,4)) - rly) %>% 
+        filter(year >= as.numeric(substr(assessmentpair,1,4)) - rly - ny -5) %>% 
+        arrange(stock, assessmentpair, assessmentyear, variable) %>% 
+        mutate(removelastyears = rly, 
+               numberofyears   = ny) %>%
+
+        # add actual yearrange used
+        group_by(stock, assessmentpair, assessmentyear, variable, removelastyears, numberofyears) %>%
+        mutate(yearrange      = paste(as.character(min(year)),as.character(max(year)), sep="-"),
+               nyears         = n_distinct(year)) %>%
+
+        # add to dataframe
+        bind_rows(d.pairwise.roughly)
+
+    } # end of numberofyears
+  } # end of removelastyears
+
+  # pairwise data
+  d.pairwise <-
+    d.pairwise.roughly %>%
+    group_by(stock, assessmentpair, variable, removelastyears, numberofyears) %>%
+
+    # filter the appropriate period
+    filter(
+      year <= as.numeric(substr(assessmentpair,1,4)) - removelastyears,
+      year >= as.numeric(substr(assessmentpair,1,4)) - removelastyears - (numberofyears-1)) %>%
+
+    # sort
+    arrange(stock, assessmentpair, assessmentyear, removelastyears, numberofyears, year) %>%
+
+    # filter only years that are in common in last and prior
+    group_by(stock, assessmentpair, variable, removelastyears, numberofyears, year) %>%
+    mutate(n = n()) %>%
+    filter(n() == 2) %>%
+
+    # add actual yearrange used
+    group_by(stock, assessmentpair, variable, removelastyears, numberofyears) %>%
+    mutate(yearrange      = paste(as.character(min(year)),as.character(max(year)), sep="-"),
+           nyears         = n_distinct(year))
+
+  # start with empty data.frame
+  d.pcm <- data.frame()
+  # th <- 0.2
+
+  # loop over thresholds
+  for (th in threshold) {
+
+    print(paste("threshold:",th,sep=" "))
+
+    d.pcm <-
+      d.pairwise %>%
+
+      # calculate means per period
+      group_by(stock, assessmentpair, variable, removelastyears, numberofyears, yearrange, nyears) %>%
+      mutate(   firstyear      = min(year),
+                lastyear       = max(year)) %>%
+      group_by(stock, assessmentpair, variable, assessmentyear, label, yearrange, nyears,
+               removelastyears, numberofyears, firstyear, lastyear) %>%
+      summarise(value = mean(value, na.rm=TRUE)) %>%
+      gather(key=tmp, value=year, firstyear:lastyear) %>%
+      arrange(stock, assessmentpair, variable, assessmentyear, yearrange, label) %>%
+
+      # calculate metrics
+      filter(tmp=="firstyear") %>%
+      ungroup() %>%
+      dplyr::select(-tmp, -assessmentyear) %>%
+      spread(key=label, value=value) %>%
+      mutate(phi            = (prior-last)/last,
+             phi_abs        = abs(phi),
+             assessmentyear = as.integer(substr(assessmentpair, 6,9)),
+             scale_jump     = ifelse(phi_abs  > th, TRUE, NA),
+             scale_jump     = ifelse(phi_abs <= th, FALSE, scale_jump),
+             threshold      = th) %>%
+      group_by(stock, variable, removelastyears, numberofyears, threshold) %>%
+      mutate(scale_jump     = ifelse(row_number() == 1, NA, scale_jump)) %>%
+
+      bind_rows(d.pcm)
+
+  } # end of threshold
+
+  res <- list("d.pcm"              = d.pcm,
+              "d.pairwise.roughly" = d.pairwise.roughly,
+              "d.pairwise"         = d.pairwise)
+
+  return(res)
+
+} # End of pairwise_comparisons function
+
+res <- pairwise_comparisons(removelastyears = c(0,1), numberofyears=c(2, 5,10), threshold = c(0.1, 0.2))
+
+# res$d.pairwise.roughly %>% 
+#   filter(grepl("mac", stock), variable=="stocksize", numberofyears==5, removelastyears==1,
+#          assessmentpair=="2017-2018") %>% View()
+
+# res$d.pairwise %>% 
+#   filter(grepl("mac", stock), variable=="stocksize", numberofyears==10, removelastyears==1,
+#          assessmentpair=="2017-2018") %>% View()
+
+# res$d.pcm %>% View()
+# res$d.pairwise.roughly %>% View()
+# res$d.pairwise %>% View()
+
+# d.pairwise %>% 
+#   filter(grepl("mac", stock), assessmentpair == "2017-2018") %>% 
+#   filter(removelastyears==myremoveyears, numberofyears==mynyears) %>%  
+#   filter(variable == myvar) %>% 
+#   View()
+
+# res$d.pcm %>% ungroup() %>% distinct(threshold)
+
+# Captions used in this document ==========================================================
+
+# section 1 -----------------------------------------------------
+
+tab_nums(name    = "section1", level = 1, display = FALSE)
+
+fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
+         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")
+
+fig_nums(name    = "differentmethods", level = 2, display = FALSE, 
+         caption = "Comparison of three different types of retrospective metrics.
+t refers to the year for a stock metric X, i refers to the assessment year.
+Diagram A is applicable to Mohn’s rho and Jonsson and Hjorleifsson’s retrospective metrics while Ap refers to the same metrics calculated in peels (in this example of 4 years), B refers to Ralston’s et al metrics and Bp to the same metrics in peels and C refers to the scale differences calculated in this example over 4 years (this paper)")
+
+fig_nums(name    = "retroplots", level = 2, display = FALSE, 
+         caption = "Historic retrospective analysis of the nine stocks analysed.
+For each stock, the most recent assessment is in red and the previous assessments in black.
+The end point of each assessment is indicated by a dot.
+Stock definitions may have changed over time, but this is not included in the graphs")
+
+# section2 -----------------------------------------------------
+
+tab_nums(name    = "section2", level = 1, display = FALSE)
+
+fig_nums(name    = "retrometrics", level = 1, display = FALSE, 
+         caption = " Retrospective metrics for 9 species derived from historical assessment records.
+Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
+In red: calculated values based on the whole time-series of historical data.
+In blue: based on retrospective peels of 5 years.
+Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
+
+fig_nums(name    = "retrometrics2", level = 2, display = FALSE, 
+         caption = "Summary of retrospective metrics for 9 species derived from historical assessment records.
+Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
+In red: calculated values based on the whole time-series of historical data.
+In blue: based on retrospective peels of 5 years.
+Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
+
+# section 3 -----------------------------------------------------
+
+tab_nums(name    = "scalejumps", level = 1, display = FALSE, 
+         caption = "Percentage assessments with scale jumps of more than 10% expressed by decade. Between brackets: number of assessments within the decade.")
+
+fig_nums(name    = "scalecomparisons", level = 1, display = FALSE, 
+         caption = "Example of pairwise assessment comparisons for Northeast Atlantic mackerel  over a period of 10 years in common.
+The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
+
+fig_nums(name    = "scalejumps", level = 2, display = FALSE, 
+         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records.
+Results from pairwise comparisons over periods of 10 years.
+Large dots  refer to years when the scale difference is larger than 10%.
+Colours indicate the decades.
+Black triangles indicate benchmark assessment years.
+The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
+
+fig_nums(name    = "metricsinpeels", level = 2, display = FALSE, 
+         caption = "Comparison of retrospective uncertainty metrics from historical assessment records, calculated in peels, normalized between 0 and 1.
+Asd from Jonsson & Hjorleifsson 2000, rhoabs as absolute value of rho from Mohn 1999, sigma from Ralston et al 2011, percjumps as the percentage of jumps (this paper) and phiabs as the absolute value of phi (this paper).")
+
+# section 4 -----------------------------------------------------
+
+fig_nums(name    = "stockdefchange", level = 1, display = FALSE, 
+         caption = "Comparisons of assessments (SSB) where stock definitions changed. Green is old stock definition, red is new stock definition")
+
+fig_nums(name    = "jumpratio", level = 2, display = FALSE, 
+         caption = "Ratio of scale-jumps in assessment relative to the number of assessment by year.")
+
+getPalettePurples         <- colorRampPalette(brewer.pal(7, "Purples")[-(1:2)])
+myPurples                 <- getPalettePurples(10)
+
+getPaletteBlues         <- colorRampPalette(brewer.pal(7, "Blues")[-(1:2)])
+myBlues                 <- getPaletteBlues(10)
+
+```
+
+**Scale-jumps: a new metric to describe historical retrospective patterns in ICES stock assessments**
+
+<!-- Tables ------------------------------------------------------ -->
+
+**Tables**
+
+*`r tab_nums("scalejumps")`*
+
+```{r}
+
+# print table
+d.jumps.bydecade %>% 
+  filter(variable == "stocksize") %>% 
+  ungroup() %>% 
+  select(stock, decade, ratio3) %>% 
+  spread(key=decade, value=ratio3) %>% 
+  pandoc.table(.,
+               style = "simple",
+               split.tables=400, 
+               justify = "left",
+               missing=".",
+               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
+
+```
+
+<!-- Figures ------------------------------------------------------ -->
+
+#####page break
+
+**Figures**
+
+<!-- ################################################################################## -->
+<!-- Plot Stock definitions      ------------------------------------------------------ -->
+<!-- ################################################################################## -->
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+d %>% 
+  select(assessmentyear, stock, stockkeylabelold) %>% 
+  distinct() %>% 
+  mutate(stock = factor(stock),
+         stock = factor(stock, levels=rev(levels(stock)))) %>% 
+  group_by(stock, stockkeylabelold) %>% 
+  summarize(firstyear = min(assessmentyear),
+            lastyear = max(assessmentyear)) %>%
+
+  ggplot(aes(x=assessmentyear, y=stock)) +
+  theme_publication() +
+  ggalt::geom_dumbbell(aes(xend = lastyear, x=firstyear, group=stock)) +
+  geom_text(aes(x=firstyear, label=stockkeylabelold), hjust=0, nudge_y=0.15, size=3)
+
+```
+
+*`r fig_nums("stockdefinitions")`*
+
+#####page break
+
+<!-- ################################################################################## -->
+<!-- Plot different methods of creating historical retrospectives --------------------- -->
+<!-- ################################################################################## -->
+
+```{r echo=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE}
+
+knitr::include_graphics("fig1.2_retromethods.jpg")
+
+```
+
+*`r fig_nums("differentmethods")`*
+
+#####Page break
+
+<!-- Plot Retroplots --------------------------- -->
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+myvar <- "stocksize"
+
+# plot
+d.selection %>%  
+  filter(variable==myvar) %>% 
+  ggplot(aes(x=year, y=value)) +
+  theme_publication() +
+  theme(
+    panel.spacing.x = unit(1, "mm"),
+    panel.spacing.y = unit(1, "mm"),
+    strip.background = element_blank(),
+    strip.text       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    legend.position  = "none"
+  ) +
+
+  # geom_boxplot(aes(group=year), outlier.shape=NA, colour="gray50", fill=NA) +
+  geom_segment(data=filter(d.segment, variable==myvar), 
+               aes(x=year, xend=year,y=value,yend=value_last), colour="gray", linetype="dashed", size=0.5) +
+  geom_line(aes(colour = plot, size=plot, group=assessmentyear)) +
+  # geom_point(data=d.points) +
+  geom_point(data=filter(d.segment, variable==myvar), 
+               aes(x=year, y=value), colour="black", size=1) +
+
+  scale_colour_manual(values=c(last = "red", other="black")) +
+  scale_size_manual (values=c(last=1, other=0.5)) +
+  expand_limits(y=0) +
+  scale_y_continuous(breaks = pretty_breaks(),  labels = scales::scientific_format(digits=2)) +
+  scale_x_continuous(breaks = pretty_breaks(n=3)) +
+  labs(y=myvar) +
+  facet_wrap(~stock, scales="free")
+
+```
+
+*`r fig_nums("retroplots")`*
+
+#####Page break
+
+#####Page break
+
+<!-- ################################################################################## -->
+<!-- Plot pairwise comparisons                                                          -->
+<!-- ################################################################################## -->
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+myvar         <- "stocksize"
+myspecies     <- "her"
+firstyear     <- 2008
+firstplotyear <- 2000
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", number of years: ",mynyears, 
+                        ", removed years: ", myremoveyears)
+
+d.pairwise.means <-
+
+  res$d.pairwise %>% 
+  # ungroup() %>% 
+  # distinct(removelastyears, numberofyears, threshold) %>% 
+  # View()
+
+  filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  # calculate means per period
+  group_by(stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(stock, assessmentpair, variable, assessmentyear, label)
+
+res$d.pairwise %>%
+
+  filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  ggplot(aes(x=year, y=value, group=assessmentyear)) +
+  theme_publication() +
+  theme(
+    panel.spacing.x = unit(1, "mm"),
+    panel.spacing.y = unit(1, "mm"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(1,0,0,0, "mm")),
+    strip.text.y     = element_text(angle = 0),
+   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
+    legend.position  = "none"
+  ) +
+
+  geom_line(data=filter(res$d.pairwise.roughly, 
+                        variable == myvar, 
+                        grepl(myspecies, stockkeylabelold),
+                        as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+                        year >= firstplotyear), 
+            size=0.5) +
+
+  geom_line(aes(colour=factor(label)), size=1) +
+
+  geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)),
+            linetype="dashed", lineend="butt", inherit.aes = FALSE) +
+  geom_point(data=filter(d.pairwise.means, year==min(year)),
+             aes(colour=factor(label)), size=1) +
+  geom_point(data=filter(d.pairwise.means, year==max(year)),
+             aes(colour=factor(label)), size=1) +
+  geom_text (data=filter(res$d.pcm,
+                         variable == myvar,
+                         grepl(myspecies, stock),
+                         as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+                         numberofyears == mynyears,
+                         removelastyears == myremoveyears,
+                         threshold == mythreshold),
+             aes(x=firstplotyear, y=0, label = paste0("phi=", round(phi_abs, digits=2))), 
+             hjust=0, vjust=0, nudge_x = -0.1, nudge_y=0.1,  inherit.aes = FALSE) +
+
+  scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
+  scale_x_continuous(breaks = pretty_breaks(n=2)) +
+  expand_limits(y=0) +
+  labs(x="", y=myvar, title=mytitle) +
+  facet_wrap( ~ assessmentpair, ncol=5)
+
+# names(res)
+# filter(res$d.pcm,
+#                          variable = myvar,
+#                          grepl(myspecies, stock),
+#                          as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+#                          numberofyears == mynyears,
+#                          removelastyears == myremoveyears,
+#                          threshold == mythreshold)
+```
+
+*`r fig_nums("scalecomparisons")`*
+
+#####Page break
+
+<!-- ################################################################################## -->
+<!-- Plot scale jumps                                                     -->
+<!-- ################################################################################## -->
+
+```{r echo=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE}
+
+# res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)
+
+# threshold <- 0.10
+mynyears      <- 10
+myremoveyears <- 0
+mythreshold   <- 0.1
+myvar         <- "stocksize"
+mytitle       <- paste0(myvar, ", number of years: ",mynyears, ", threshold: ", 
+                        comma_format(digits=1, suffix="%")(100*mythreshold))
+
+# res$d.pcm %>% View()
+# res$d.pcm %>% ungroup() %>% distinct(threshold)
+
+# plot scale jumps between assessments
+t <-
+  res$d.pcm %>% 
+  filter(variable        == myvar,
+         numberofyears   == mynyears,
+         removelastyears == myremoveyears,
+         threshold       == mythreshold) %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi))
+
+t %>%   
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  # stock definition change
+  geom_rect(data=filter(d.stockdefinitionchange),
+            aes(xmin = y1, xmax=y2, ymin = 0.00, ymax=1.0), 
+            alpha=0.4, fill=myPurples[6], colour=myPurples[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.stockdefinitionchange),
+          aes(label  = stockkeylabelold, x = y1, y = 0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=1.0, vjust=0.9),
+          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +
+
+  # assessmentmodel change
+  geom_rect(data=filter(d.assessmentmodels),
+            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=0.0, fill=rank), 
+            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +
+
+  geom_dl(data=filter(d.assessmentmodels),
+          aes(label  = assessmentmodel, x = y1, y = -0.95),
+          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
+          inherit.aes = FALSE, colour=myBlues[6]) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
+  # geom_point(aes(colour=factor(decade), size = abs(value) )) +
+
+  geom_point(data=filter(d.benchmarks, benchmark == 1),
+            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
+
+  geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
+            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
+
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="", title=mytitle) +
+  facet_wrap(~stock)
+
+```
+
+*`r fig_nums("scalejumps")`*
+
+```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot scale jumps with different settings for number of years and thresholds
+
+myremoveyears <- 0
+myvar         <- "stocksize"
+myspecies     <- "cod"
+mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
+
+res$d.pcm %>%
+
+  filter(grepl(myspecies, stock)) %>% 
+  filter(variable        == myvar) %>% 
+  filter(removelastyears == myremoveyears) %>%
+
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
+
+  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  # ggMarginal(data=d.assessments, type="histogram")
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="", title=mytitle) +
+  facet_grid(threshold ~ numberofyears)
+
+# filter(d.assessments, benchmark == 1) %>% View()
+
+```
+
+*`r fig_nums("metricsinpeels")`*
+
+<!-- ################################################################################## -->
+<!-- COmpare scale jumps when stock definition changed                                  -->
+<!-- ################################################################################## -->
+
+```{r metricsatstockchange, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot scale jumps between assessments when stock definition changed
+t3 %>%
+
+  ggplot(aes(x=year, y=value, group=label)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.placement = "outside",
+    legend.title    = element_blank(),
+    legend.position  = "none"
+  ) +
+
+  geom_line(aes(colour=label), size=0.5) +
+
+  geom_dl(aes(label  = stockkeylabelold, colour = label),
+          method = list(dl.combine("first.bumpup"), cex = 0.8, hjust=0)) +
+
+  scale_y_continuous(labels=scientific_format(digits=2)) +
+  scale_x_continuous(breaks = pretty_breaks(n=3)) +
+  expand_limits(y=0) +
+  labs(x="", y="ssb") +
+  facet_wrap( ~ header, scales="free")
+
+```
+
+*`r fig_nums("stockdefchange")`*
+
+<!-- ################################################################################## -->
+<!-- COmpare the number of scale jumps (as ratio) over time                             -->
+<!-- ################################################################################## -->
+
+```{r jumpratio, eval=FALSE, fig.align="center", fig.asp=0.5, message=FALSE, warning=FALSE, include=FALSE}
+
+# NOT SHOWN, NOT USED
+
+d.jumps.byyear %>% 
+  filter(assessmentyear >= 1982) %>%
+
+  ggplot(aes(x=assessmentyear, y=ratio)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.placement = "outside",
+    legend.title    = element_blank(),
+    legend.position  = "none"
+  ) +
+  geom_bar(stat="identity") +
+  labs(x="", y="ratio")
+
+```
+
+*`r fig_nums("jumpratio")`*
+
+```{r jumpratio, eval=FALSE, fig.align="center", fig.asp=0.5, message=FALSE, warning=FALSE, include=FALSE}
+
+# NOT SHOWN, NOT USED
+
+d.assessments%>% 
+  filter(assessmentyear <= 2018) %>% 
+  group_by(assessmentyear) %>% 
+  summarize(ncpueseries = mean(ncpueseries, na.rm=TRUE),
+            nsurveyseries = mean(nsurveyseries, na.rm=TRUE)) %>% 
+  gather(key=type, value=n, ncpueseries:nsurveyseries) %>%
+
+  ggplot(aes(x=assessmentyear, y=n, group=type)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.placement = "outside",
+    legend.title    = element_blank()
+  ) +
+  # theme(legend.position  = "none") + 
+  geom_line(aes(colour=type)) +
+  labs(x="", y="avg")
+
+```
+
+```{r eval=FALSE, include=FALSE}
+
+# NOT SHOWN, NOT USED
+
+# REMOVE FROM HERE !!
+
+removelastyears <- 0
+numberofyears   <- 10
+threshold       <- 0.1   # threshold for scale differences
+
+# Create a set of assessment pairs: two subsequent assessments  
+d.pairs <-
+  d.selection %>% 
+  distinct(stock, assessmentyear, variable) %>% 
+  group_by(stock, variable) %>% 
+  arrange(stock, variable, assessmentyear) %>% 
+  mutate(assessmentyearprior = lag(assessmentyear),
+         assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
+  rename(assessmentyearlast  = assessmentyear) %>% 
+  filter(!is.na(assessmentyearprior))
+
+# data series with most recent (last) estimation of the pair
+d.pairs.last <-
+  d.pairs %>% 
+  select(-assessmentyearprior) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  mutate(label="last") %>% 
+  rename(assessmentyear = assessmentyearlast)
+
+# data series with previous (prior) estimation of the pair
+d.pairs.prior <-
+  d.pairs %>% 
+  select(-assessmentyearlast) %>% 
+  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  mutate(label="prior") %>% 
+  rename(assessmentyear = assessmentyearprior)
+
+# pairwise data roughly
+d.pairwise.roughly <-
+  bind_rows(d.pairs.last, d.pairs.prior) %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears -5)
+
+# pairwise data
+d.pairwise <-
+  d.pairwise.roughly %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears) %>%
+
+  # sort
+  arrange(stock, assessmentpair, assessmentyear, year) %>%
+
+  # filter only years that are in common in last and prior
+  group_by(stock, assessmentpair, variable, year) %>% 
+  mutate(n = n()) %>% 
+  filter(n() == 2)
+
+d.jumps <-
+  d.pairwise %>%
+
+  # calculate means per period
+  group_by(stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(stock, assessmentpair, variable, assessmentyear, label) %>%
+
+  # calculate metrics
+  filter(tmp=="firstyear") %>% 
+  ungroup() %>% 
+  select(stock, assessmentpair, variable, label, value) %>% 
+  spread(key=label, value=value) %>% 
+  mutate(phi            = (prior-last)/last,
+         phi_abs        = abs(phi),
+         assessmentyear = as.integer(substr(assessmentpair, 6,9)))
+
+# CONTINUE FROM HERE
+
+# calculate average jump per stock
+d.jumps.bystock <-
+  d.jumps %>% 
+  group_by(stock, variable) %>% 
+  summarize(
+    pastoors_njumps    = sum(phi_abs > threshold),
+    pastoors_percjumps = pastoors_njumps / n(),
+    pastoors_phi       = mean(phi),
+    pastoors_phiabs    = mean(phi_abs), 
+    firstyear          = min(assessmentyear),
+    lastyear           = max(assessmentyear),
+    calctype           = "in_peels")
+
+# jumps by decade
+d.jumps.bydecade <-
+  d.jumps %>% 
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  group_by(stock, variable, decade) %>% 
+  summarise(n         = n(),
+            n2        = sum(phi_abs > 0.1, na.rm=TRUE),
+            phi     = mean(phi, na.rm=TRUE),
+            phi_abs = mean(phi_abs, na.rm=TRUE)
+  ) %>%
+  mutate(phi_abs    = percent_format(accuracy=1)(phi_abs),
+         ratio        = paste(n2,n,sep="/"),
+         ratio2       = percent_format(accuracy=1) (n2/n),
+         ratio3       = paste(ratio2," (", n, ")", sep="") )
+
+# jumps by year
+d.jumps.byyear <-
+  d.jumps %>% 
+  group_by(variable, assessmentyear) %>% 
+  summarise(n         = n(),
+            n2        = sum(phi_abs > 0.1, na.rm=TRUE)) %>% 
+  mutate(ratio = n2 / n )
+
+# update d.metrics_bystock
+d.metrics_bystock2 <-
+  bind_rows(d.mohnjonsson.bystock, d.ralston.bystock) %>% 
+  bind_rows(select(d.mohnjonsson_inpeels.bystock, 
+                   stock, variable, calctype, firstyear, lastyear, 
+                   jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs)) %>% 
+  bind_rows(select(d.ralston_inpeels.bystock,
+                   stock, variable, calctype, firstyear, lastyear, 
+                   ralston_sigma)) %>% 
+  bind_rows(select(d.jumps.bystock,
+                   stock, variable, calctype, firstyear, lastyear, 
+                   pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs)) %>% 
+  ungroup() %>% 
+  select(calctype, stock, firstyear, lastyear, jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs, ralston_sigma,
+         pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs) %>% 
+  gather(key=metric, value=value, jonsson_ab:pastoors_phiabs) %>%
+  filter(!is.na(value)) %>% 
+  separate(metric, into=c("author", "metric"), by="_") %>%
+  group_by(calctype, stock, author, metric) %>% 
+  mutate(stock2 = substr(stock,1,3))
+
+# Compare scale jumps when stock definition has changed.
+
+t <-
+  d %>% 
+  select(assessmentyear, stock, stockkeylabelold) %>% 
+  distinct() %>% 
+  mutate(
+    stock = factor(stock),
+    stock = factor(stock, levels=rev(levels(stock)))) %>% 
+  group_by(stock, stockkeylabelold) %>% 
+  summarize(
+    firstyear = min(assessmentyear),
+    lastyear  = max(assessmentyear)) %>% 
+  arrange(stock, firstyear) %>% 
+  group_by(stock) %>% 
+  mutate(
+    previous       = lag(lastyear),
+    assessmentpair = paste(previous, firstyear, sep="-"),
+    stockold       = lag(stockkeylabelold)) %>% 
+  filter(!is.na(previous))  %>% 
+  select(stock, assessmentpair)
+
+t1 <-
+  t %>% 
+  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
+  select(-assessmentyearprior) %>% 
+  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  mutate(label="after") %>% 
+  rename(assessmentyear = assessmentyearlast)
+
+# data series with previous (prior) estimation
+t2 <-
+  t %>% 
+  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
+  select(-assessmentyearlast) %>% 
+  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  mutate(label="before") %>% 
+  rename(assessmentyear = assessmentyearprior)
+
+# combine
+t3 <-
+  bind_rows(t2, t1) %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - removelastyears,
+         year >= max(assessmentyear) - removelastyears - numberofyears) %>%
+
+  # sort
+  arrange(stock, assessmentpair, assessmentyear, year) %>%
+
+  # filter only years that are in common in last and prior
+  group_by(stock, assessmentpair, variable, year) %>% 
+  mutate(
+    n = n(),
+    header = paste(stock, assessmentpair, sep=" ")
+  ) %>% 
+  filter(n() == 2)
 
@@
 
-# Calculations of Pairwise Comparison of Mean and scale jumps ==========================================
-
-pairwise_comparisons <- function(removelastyears=0, numberofyears=10, threshold=0.10) {
-
-  # Create a set of assessment pairs: two subsequent assessments  
-  d.pairs <-
-    d.selection %>% 
-    distinct(stock, assessmentyear, variable) %>% 
-    group_by(stock, variable) %>% 
-    arrange(stock, variable, assessmentyear) %>% 
-    mutate(assessmentyearprior = lag(assessmentyear),
-           assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
-    rename(assessmentyearlast  = assessmentyear) %>% 
-    filter(!is.na(assessmentyearprior))
-
-  # data series with most recent (last) estimation of the pair
-  d.pairs.last <-
-    d.pairs %>% 
-    select(-assessmentyearprior) %>% 
-    left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
-    mutate(label="last") %>% 
-    rename(assessmentyear = assessmentyearlast)
-
-  # data series with previous (prior) estimation of the pair
-  d.pairs.prior <-
-    d.pairs %>% 
-    select(-assessmentyearlast) %>% 
-    left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
-    mutate(label="prior") %>% 
-    rename(assessmentyear = assessmentyearprior)
-
-  # pairwise data roughly
-  d.pairwise.roughly <-
-    bind_rows(d.pairs.last, d.pairs.prior) %>% 
-    group_by(stock, assessmentpair, variable) %>%
-
-    # filter the appropriate period
-    filter(year <= max(assessmentyear) - removelastyears,
-           year >= max(assessmentyear) - removelastyears - numberofyears -5)
-
-  # pairwise data
-  d.pairwise <-
-    d.pairwise.roughly %>% 
-    group_by(stock, assessmentpair, variable) %>%
-
-    # filter the appropriate period
-    filter(year <= max(assessmentyear) - removelastyears,
-           year >= max(assessmentyear) - removelastyears - numberofyears) %>%
-
-    # sort
-    arrange(stock, assessmentpair, assessmentyear, year) %>%
-
-    # filter only years that are in common in last and prior
-    group_by(stock, assessmentpair, variable, year) %>% 
-    mutate(n = n()) %>% 
-    filter(n() == 2)
-
-  d.pcm <-
-    d.pairwise %>%
-
-    # calculate means per period
-    group_by(stock, assessmentpair, variable) %>% 
-    mutate(   firstyear      = min(year),
-              lastyear       = max(year)) %>% 
-    group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
-    summarise(value = mean(value, na.rm=TRUE)) %>% 
-    gather(key=tmp, value=year, firstyear:lastyear) %>% 
-    arrange(stock, assessmentpair, variable, assessmentyear, label) %>%
-
-    # calculate metrics
-    filter(tmp=="firstyear") %>% 
-    ungroup() %>% 
-    select(stock, assessmentpair, variable, label, value) %>% 
-    spread(key=label, value=value) %>% 
-    mutate(phi            = (prior-last)/last,
-           phi_abs        = abs(phi),
-           assessmentyear = as.integer(substr(assessmentpair, 6,9)), 
-           scale_jump     = ifelse(phi_abs  > threshold, TRUE, NA),
-           scale_jump     = ifelse(phi_abs <= threshold, FALSE, scale_jump)) %>% 
-    group_by(stock, variable) %>% 
-    mutate(scale_jump     = ifelse(row_number() == 1, NA, scale_jump),
-           removelastyears= removelastyears, 
-           numberofyears  = numberofyears, 
-           threshold      = threshold)
-
-  res <- list("d.pcm"              = d.pcm, 
-              "d.pairwise.roughly" = d.pairwise.roughly,
-              "d.pairwise"         = d.pairwise)
-
-  return(res)
-
-} # End of pairwise_comparisons function
-
-tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5, threshold = 0.1)
-
-# tmp$d.pcm %>% View()
-
-# REMOVE FROM HERE !!
-
-removelastyears <- 0
-numberofyears   <- 10
-threshold       <- 0.1   # threshold for scale differences
-
-# Create a set of assessment pairs: two subsequent assessments  
-d.pairs <-
-  d.selection %>% 
-  distinct(stock, assessmentyear, variable) %>% 
-  group_by(stock, variable) %>% 
-  arrange(stock, variable, assessmentyear) %>% 
-  mutate(assessmentyearprior = lag(assessmentyear),
-         assessmentpair      = paste(assessmentyearprior, assessmentyear, sep="-")) %>% 
-  rename(assessmentyearlast  = assessmentyear) %>% 
-  filter(!is.na(assessmentyearprior))
-
-# data series with most recent (last) estimation of the pair
-d.pairs.last <-
-  d.pairs %>% 
-  select(-assessmentyearprior) %>% 
-  left_join(d.selection, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
-  mutate(label="last") %>% 
-  rename(assessmentyear = assessmentyearlast)
-
-# data series with previous (prior) estimation of the pair
-d.pairs.prior <-
-  d.pairs %>% 
-  select(-assessmentyearlast) %>% 
-  left_join(d.selection, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
-  mutate(label="prior") %>% 
-  rename(assessmentyear = assessmentyearprior)
-
-# pairwise data roughly
-d.pairwise.roughly <-
-  bind_rows(d.pairs.last, d.pairs.prior) %>% 
-  group_by(stock, assessmentpair, variable) %>%
-
-  # filter the appropriate period
-  filter(year <= max(assessmentyear) - removelastyears,
-         year >= max(assessmentyear) - removelastyears - numberofyears -5)
-
-# pairwise data
-d.pairwise <-
-  d.pairwise.roughly %>% 
-  group_by(stock, assessmentpair, variable) %>%
-
-  # filter the appropriate period
-  filter(year <= max(assessmentyear) - removelastyears,
-         year >= max(assessmentyear) - removelastyears - numberofyears) %>%
-
-  # sort
-  arrange(stock, assessmentpair, assessmentyear, year) %>%
-
-  # filter only years that are in common in last and prior
-  group_by(stock, assessmentpair, variable, year) %>% 
-  mutate(n = n()) %>% 
-  filter(n() == 2)
-
-d.jumps <-
-  d.pairwise %>%
-
-  # calculate means per period
-  group_by(stock, assessmentpair, variable) %>% 
-  mutate(   firstyear      = min(year),
-            lastyear       = max(year)) %>% 
-  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
-  summarise(value = mean(value, na.rm=TRUE)) %>% 
-  gather(key=tmp, value=year, firstyear:lastyear) %>% 
-  arrange(stock, assessmentpair, variable, assessmentyear, label) %>%
-
-  # calculate metrics
-  filter(tmp=="firstyear") %>% 
-  ungroup() %>% 
-  select(stock, assessmentpair, variable, label, value) %>% 
-  spread(key=label, value=value) %>% 
-  mutate(phi            = (prior-last)/last,
-         phi_abs        = abs(phi),
-         assessmentyear = as.integer(substr(assessmentpair, 6,9)))
-
-# CONTINUE FROM HERE
-
-# calculate average jump per stock
-d.jumps.bystock <-
-  d.jumps %>% 
-  group_by(stock, variable) %>% 
-  summarize(
-    pastoors_njumps    = sum(phi_abs > threshold),
-    pastoors_percjumps = pastoors_njumps / n(),
-    pastoors_phi       = mean(phi),
-    pastoors_phiabs    = mean(phi_abs), 
-    firstyear          = min(assessmentyear),
-    lastyear           = max(assessmentyear),
-    calctype           = "in_peels")
-
-# jumps by decade
-d.jumps.bydecade <-
-  d.jumps %>% 
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  group_by(stock, variable, decade) %>% 
-  summarise(n         = n(),
-            n2        = sum(phi_abs > 0.1, na.rm=TRUE),
-            phi     = mean(phi, na.rm=TRUE),
-            phi_abs = mean(phi_abs, na.rm=TRUE)
-  ) %>%
-  mutate(phi_abs    = percent_format(accuracy=1)(phi_abs),
-         ratio        = paste(n2,n,sep="/"),
-         ratio2       = percent_format(accuracy=1) (n2/n),
-         ratio3       = paste(ratio2," (", n, ")", sep="") )
-
-# jumps by year
-d.jumps.byyear <-
-  d.jumps %>% 
-  group_by(variable, assessmentyear) %>% 
-  summarise(n         = n(),
-            n2        = sum(phi_abs > 0.1, na.rm=TRUE)) %>% 
-  mutate(ratio = n2 / n )
-
-# update d.metrics_bystock
-d.metrics_bystock2 <-
-  bind_rows(d.mohnjonsson.bystock, d.ralston.bystock) %>% 
-  bind_rows(select(d.mohnjonsson_inpeels.bystock, 
-                   stock, variable, calctype, firstyear, lastyear, 
-                   jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs)) %>% 
-  bind_rows(select(d.ralston_inpeels.bystock,
-                   stock, variable, calctype, firstyear, lastyear, 
-                   ralston_sigma)) %>% 
-  bind_rows(select(d.jumps.bystock,
-                   stock, variable, calctype, firstyear, lastyear, 
-                   pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs)) %>% 
-  ungroup() %>% 
-  select(calctype, stock, firstyear, lastyear, jonsson_ab, jonsson_asd, mohn_rho, mohn_rhoabs, ralston_sigma,
-         pastoors_njumps, pastoors_percjumps, pastoors_phi, pastoors_phiabs) %>% 
-  gather(key=metric, value=value, jonsson_ab:pastoors_phiabs) %>%
-  filter(!is.na(value)) %>% 
-  separate(metric, into=c("author", "metric"), by="_") %>%
-  group_by(calctype, stock, author, metric) %>% 
-  mutate(stock2 = substr(stock,1,3))
-
-# Compare scale jumps when stock definition has changed.
-
-t <-
-  d %>% 
-  select(assessmentyear, stock, stockkeylabelold) %>% 
-  distinct() %>% 
-  mutate(
-    stock = factor(stock),
-    stock = factor(stock, levels=rev(levels(stock)))) %>% 
-  group_by(stock, stockkeylabelold) %>% 
-  summarize(
-    firstyear = min(assessmentyear),
-    lastyear  = max(assessmentyear)) %>% 
-  arrange(stock, firstyear) %>% 
-  group_by(stock) %>% 
-  mutate(
-    previous       = lag(lastyear),
-    assessmentpair = paste(previous, firstyear, sep="-"),
-    stockold       = lag(stockkeylabelold)) %>% 
-  filter(!is.na(previous))  %>% 
-  select(stock, assessmentpair)
-
-t1 <-
-  t %>% 
-  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
-  select(-assessmentyearprior) %>% 
-  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
-  mutate(label="after") %>% 
-  rename(assessmentyear = assessmentyearlast)
-
-# data series with previous (prior) estimation
-t2 <-
-  t %>% 
-  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
-  select(-assessmentyearlast) %>% 
-  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
-  mutate(label="before") %>% 
-  rename(assessmentyear = assessmentyearprior)
-
-# combine
-t3 <-
-  bind_rows(t2, t1) %>% 
-  group_by(stock, assessmentpair, variable) %>%
-
-  # filter the appropriate period
-  filter(year <= max(assessmentyear) - removelastyears,
-         year >= max(assessmentyear) - removelastyears - numberofyears) %>%
-
-  # sort
-  arrange(stock, assessmentpair, assessmentyear, year) %>%
-
-  # filter only years that are in common in last and prior
-  group_by(stock, assessmentpair, variable, year) %>% 
-  mutate(
-    n = n(),
-    header = paste(stock, assessmentpair, sep=" ")
-  ) %>% 
-  filter(n() == 2)
-
-# Captions used in this document ==========================================================
-
-# section 1 -----------------------------------------------------
-
-tab_nums(name    = "section1", level = 1, display = FALSE)
-
-fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
-         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")
-
-fig_nums(name    = "differentmethods", level = 2, display = FALSE, 
-         caption = "Comparison of three different types of retrospective metrics.
-t refers to the year for a stock metric X, i refers to the assessment year.
-Diagram A is applicable to Mohn’s rho and Jonsson and Hjorleifsson’s retrospective metrics while Ap refers to the same metrics calculated in peels (in this example of 4 years), B refers to Ralston’s et al metrics and Bp to the same metrics in peels and C refers to the scale differences calculated in this example over 4 years (this paper)")
-
-fig_nums(name    = "retroplots", level = 2, display = FALSE, 
-         caption = "Historic retrospective analysis of the nine stocks analysed.
-For each stock, the most recent assessment is in red and the previous assessments in black.
-The end point of each assessment is indicated by a dot.
-Stock definitions may have changed over time, but this is not included in the graphs")
-
-# section2 -----------------------------------------------------
-
-tab_nums(name    = "section2", level = 1, display = FALSE)
-
-fig_nums(name    = "retrometrics", level = 1, display = FALSE, 
-         caption = " Retrospective metrics for 9 species derived from historical assessment records.
-Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
-In red: calculated values based on the whole time-series of historical data.
-In blue: based on retrospective peels of 5 years.
-Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
-
-fig_nums(name    = "retrometrics2", level = 2, display = FALSE, 
-         caption = "Summary of retrospective metrics for 9 species derived from historical assessment records.
-Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
-In red: calculated values based on the whole time-series of historical data.
-In blue: based on retrospective peels of 5 years.
-Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
-
-# section 3 -----------------------------------------------------
-
-tab_nums(name    = "scalejumps", level = 1, display = FALSE, 
-         caption = "Percentage assessments with scale jumps of more than 10% expressed by decade. Between brackets: number of assessments within the decade.")
-
-fig_nums(name    = "scalecomparisons", level = 1, display = FALSE, 
-         caption = "Example of pairwise assessment comparisons for Northeast Atlantic mackerel  over a period of 10 years in common.
-The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
-
-fig_nums(name    = "scalejumps", level = 2, display = FALSE, 
-         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records.
-Results from pairwise comparisons over periods of 10 years.
-Large dots  refer to years when the scale difference is larger than 10%.
-Colours indicate the decades.
-Black triangles indicate benchmark assessment years.
-The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
-
-fig_nums(name    = "metricsinpeels", level = 2, display = FALSE, 
-         caption = "Comparison of retrospective uncertainty metrics from historical assessment records, calculated in peels, normalized between 0 and 1.
-Asd from Jonsson & Hjorleifsson 2000, rhoabs as absolute value of rho from Mohn 1999, sigma from Ralston et al 2011, percjumps as the percentage of jumps (this paper) and phiabs as the absolute value of phi (this paper).")
-
-# section 4 -----------------------------------------------------
-
-fig_nums(name    = "stockdefchange", level = 1, display = FALSE, 
-         caption = "Comparisons of assessments (SSB) where stock definitions changed. Green is old stock definition, red is new stock definition")
-
-fig_nums(name    = "jumpratio", level = 2, display = FALSE, 
-         caption = "Ratio of scale-jumps in assessment relative to the number of assessment by year.")
-
-getPalettePurples         <- colorRampPalette(brewer.pal(7, "Purples")[-(1:2)])
-myPurples                 <- getPalettePurples(10)
-
-getPaletteBlues         <- colorRampPalette(brewer.pal(7, "Blues")[-(1:2)])
-myBlues                 <- getPaletteBlues(10)
-
 ```
-
-**Scale-jumps: a new metric to describe historical retrospective patterns in ICES stock assessments**
-
-<!-- Tables ------------------------------------------------------ -->
-
-**Tables**
-
-*`r tab_nums("scalejumps")`*
-
-```{r}
-
-# print table
-d.jumps.bydecade %>% 
-  filter(variable == "stocksize") %>% 
-  ungroup() %>% 
-  select(stock, decade, ratio3) %>% 
-  spread(key=decade, value=ratio3) %>% 
-  pandoc.table(.,
-               style = "simple",
-               split.tables=400, 
-               justify = "left",
-               missing=".",
-               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-
-```
-
-<!-- Figures ------------------------------------------------------ -->
-
-#####page break
-
-**Figures**
-
-<!-- ################################################################################## -->
-<!-- Plot Stock definitions      ------------------------------------------------------ -->
-<!-- ################################################################################## -->
-
-```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
-
-d %>% 
-  select(assessmentyear, stock, stockkeylabelold) %>% 
-  distinct() %>% 
-  mutate(stock = factor(stock),
-         stock = factor(stock, levels=rev(levels(stock)))) %>% 
-  group_by(stock, stockkeylabelold) %>% 
-  summarize(firstyear = min(assessmentyear),
-            lastyear = max(assessmentyear)) %>%
-
-  ggplot(aes(x=assessmentyear, y=stock)) +
-  theme_publication() +
-  ggalt::geom_dumbbell(aes(xend = lastyear, x=firstyear, group=stock)) +
-  geom_text(aes(x=firstyear, label=stockkeylabelold), hjust=0, nudge_y=0.15, size=3)
-
-```
-
-*`r fig_nums("stockdefinitions")`*
-
-#####page break
-
-<!-- ################################################################################## -->
-<!-- Plot different methods of creating historical retrospectives --------------------- -->
-<!-- ################################################################################## -->
-
-```{r echo=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE}
-
-knitr::include_graphics("fig1.2_retromethods.jpg")
-
-```
-
-*`r fig_nums("differentmethods")`*
-
-#####Page break
-
-<!-- Plot Retroplots --------------------------- -->
-
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
-
-# plot
-d.selection %>%   
-  ggplot(aes(x=year, y=value)) +
-  theme_publication() +
-  theme(
-    panel.spacing.x = unit(1, "mm"),
-    panel.spacing.y = unit(1, "mm"),
-    strip.background = element_blank(),
-    strip.text       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
-    legend.position  = "none"
-  ) +
-
-  # geom_boxplot(aes(group=year), outlier.shape=NA, colour="gray50", fill=NA) +
-  geom_segment(data=d.segment, 
-               aes(x=year, xend=year,y=value,yend=value_last), colour="gray", linetype="dashed", size=0.5) +
-  geom_line(aes(colour = plot, size=plot, group=assessmentyear)) +
-  geom_point(data=d.points) +
-  scale_colour_manual(values=c(last = "red", other="black")) +
-  scale_size_manual (values=c(last=1, other=0.5)) +
-  expand_limits(y=0) +
-  scale_y_continuous(breaks = pretty_breaks(),  labels = scales::scientific_format(digits=2)) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-  labs(y="ssb (tonnes)") +
-  facet_wrap(~stock, scales="free")
-
-```
-
-*`r fig_nums("retroplots")`*
-
-#####Page break
 
@@
 
+*`r fig_nums("retrometrics")`*
+
+#####Page break
+
 ```
-
-*`r fig_nums("retrometrics")`*
-
-#####Page break
 
@@
 
+*`r fig_nums("retrometrics2")`*
+
 ```
 
-*`r fig_nums("retrometrics2")`*
-
-#####Page break
-
-<!-- ################################################################################## -->
-<!-- Plot pairwise comparisons                                                          -->
-<!-- ################################################################################## -->
-
-```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
-
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
-
-myvar     <- "stocksize"
-myspecies <- "mac"
-firstyear <- 2008
-
-d.pairwise.means <-
-
-  res$d.pairwise %>%
-
-  filter(variable == myvar) %>% 
-  filter(grepl(myspecies, stockkeylabelold)) %>% 
-  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
-
-  # calculate means per period
-  group_by(stock, assessmentpair, variable) %>% 
-  mutate(   firstyear      = min(year),
-            lastyear       = max(year)) %>% 
-  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
-  summarise(value = mean(value, na.rm=TRUE)) %>% 
-  gather(key=tmp, value=year, firstyear:lastyear) %>% 
-  arrange(stock, assessmentpair, variable, assessmentyear, label)
-
-res$d.pairwise %>%
-
-  filter(variable == myvar) %>% 
-  filter(grepl(myspecies, stockkeylabelold)) %>% 
-  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
-
-  ggplot(aes(x=year, y=value, group=assessmentyear)) +
-  theme_publication() +
-  theme(
-    panel.spacing.x = unit(1, "mm"),
-    panel.spacing.y = unit(1, "mm"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(1,0,0,0, "mm")),
-    strip.text.y     = element_text(angle = 0),
-   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
-    legend.position  = "none"
-  ) +
-
-  geom_line(data=filter(res$d.pairwise.roughly, 
-                        variable == myvar, 
-                        grepl(myspecies, stockkeylabelold),
-                        as.numeric(substr(assessmentpair,1,4)) >= firstyear), size=0.5) +
-
-  geom_line(aes(colour=factor(label)), size=1) +
-
-  geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)), 
-            linetype="dashed", lineend="butt") +
-  geom_point(data=filter(d.pairwise.means, year==min(year)),
-             aes(colour=factor(label)), size=1) +
-  geom_point(data=filter(d.pairwise.means, year==max(year)),
-             aes(colour=factor(label)), size=1) +
-
-  scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
-  scale_x_continuous(breaks = pretty_breaks(n=2)) +
-  expand_limits(y=0) +
-  labs(x="", y="ssb (tonnes)") +
-  facet_wrap( ~ assessmentpair, scales = "free_x", ncol=5)
-
-```
-
-*`r fig_nums("scalecomparisons")`*
-
-#####Page break
-
-<!-- ################################################################################## -->
-<!-- Plot pairwise comparison means                                                     -->
-<!-- ################################################################################## -->
-
 ```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
 
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)
-
-# threshold <- 0.10
-
-# plot scale jumps between assessments
-res$d.pcm %>% 
-  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  gather(key=metric, value=value, c(phi)) %>%
-
-  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
-    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
-    strip.placement = "outside",
-    legend.position  = "none"
-  ) +
-
-  geom_rect(data=filter(d.stockdefinitionchange),
-            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
-            alpha=0.4, colour=myPurples[6], inherit.aes = FALSE) +
-
-  geom_dl(data=filter(d.stockdefinitionchange),
-          aes(label  = stockkeylabelold, x = y1, y = -0.95),
-          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
-          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +
-
-  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
-  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
-  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
-
-  geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
-  # geom_point(aes(colour=factor(decade), size = abs(value) )) +
-
-  geom_point(data=filter(d.benchmarks, benchmark == 1),
-            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
-
-  geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
-            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
-
-  coord_cartesian(ylim=c(-1, 1),
-                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-
-  scale_fill_manual(values = brewer.pal(6, "Purples")[-(1:2)]) +
-
-  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-
-  labs(x=" ", y="") +
-  facet_wrap(~stock)
-
-```
-
-*`r fig_nums("scalejumps")`*
-
-```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
-
-## NOT SHOWN
-
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10)
-
-threshold <- 0.10
-
-# plot scale jumps between assessments vs assessment methods
-
-res$d.pcm %>% 
-  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  gather(key=metric, value=value, c(phi)) %>%
-
-  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
-    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
-    strip.placement = "outside",
-    legend.position  = "none"
-  ) +
-
-  geom_rect(data=filter(d.assessmentmodels),
-            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=1.0, fill=rank), 
-            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +
-
-  geom_dl(data=filter(d.assessmentmodels),
-          aes(label  = assessmentmodel, x = y1, y = -0.95),
-          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
-          inherit.aes = FALSE, colour=myBlues[6]) +
-
-  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
-  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
-  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
-
-  geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
-
-  geom_point(data=filter(d.benchmarks, benchmark == 1),
-            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
-
-  coord_cartesian(ylim=c(-1, 1),
-                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-
-  scale_fill_manual(values = myBlues) +
-
-  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-
-  labs(x=" ", y="") +
-  facet_wrap(~stock)
-
-```
-
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
-
-res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold=0.1)
-
-# plot scale jumps between assessments vs assessment methods
-
-res$d.pcm %>% 
-  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  gather(key=metric, value=value, c(phi)) %>%
-
-  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
-    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), angle = 180),
-    strip.placement = "outside",
-    legend.position  = "none"
-  ) +
-
-  geom_rect(data=filter(d.stockdefinitionchange),
-            aes(xmin = y1, xmax=y2, ymin = 0.00, ymax=1.0), 
-            alpha=0.4, fill=myPurples[6], colour=myPurples[6], inherit.aes = FALSE) +
-
-  geom_dl(data=filter(d.stockdefinitionchange),
-          aes(label  = stockkeylabelold, x = y1, y = 0.95),
-          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=1.0, vjust=0.9),
-          inherit.aes = FALSE, colour=brewer.pal(6, "Purples")[6]) +
-
-  geom_rect(data=filter(d.assessmentmodels),
-            aes(xmin = y1, xmax=y2, ymin = -1.0, ymax=0.0, fill=rank), 
-            alpha=0.4, colour=myBlues[6], inherit.aes = FALSE) +
-
-  geom_dl(data=filter(d.assessmentmodels),
-          aes(label  = assessmentmodel, x = y1, y = -0.95),
-          method = list(dl.combine("first.points"), rot = 90, cex = 0.6, hjust=0.0, vjust=0.9),
-          inherit.aes = FALSE, colour=myBlues[6]) +
-
-  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
-  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
-  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
-
-  geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
-
-  geom_point(data=filter(d.benchmarks, benchmark == 1),
-            aes(x=assessmentyear+0.1, y=1.0), colour="gray20", fill="gray20", shape=25, size=2) +
-
-  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-  coord_cartesian(ylim=c(-1, 1),
-                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-
-  # ggMarginal(data=d.assessments, type="histogram")
-  scale_fill_manual(values = myBlues) +
-
-  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-
-  labs(x=" ", y="") +
-  facet_wrap(~stock)
-
-# filter(d.assessments, benchmark == 1) %>% View()
-
-```
-
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
-
-# plot scale jumps with different settings for number of years and thresholds
-
-d.pcm <- data.frame()
-
-for (ny in c(5, 10, 15)) {
-  for (nt in c(0.05, 0.1, 0.15, 0.2)) {
-    res <- pairwise_comparisons(removelastyears = 0, numberofyears=ny, threshold=nt)
-    d.pcm <- bind_rows(d.pcm, res$d.pcm)
-  }
-}
-
-d.pcm %>% 
-  filter(stock == "cod (north sea)") %>% 
-  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  gather(key=metric, value=value, c(phi)) %>%
-
-  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
-    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
-    strip.placement = "outside",
-    legend.position  = "none"
-  ) +
-
-  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
-  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
-  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
-
-  geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
-
-  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-  coord_cartesian(ylim=c(-1, 1),
-                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-
-  # ggMarginal(data=d.assessments, type="histogram")
-  scale_fill_manual(values = myBlues) +
-
-  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-
-  labs(x=" ", y="") +
-  facet_grid(threshold ~ numberofyears)
-
-# filter(d.assessments, benchmark == 1) %>% View()
-
-```
-
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+# NOT SHOWN
 
@@
 ```
-
-*`r fig_nums("metricsinpeels")`*
-
-<!-- ################################################################################## -->
-<!-- COmpare scale jumps when stock definition changed                                  -->
-<!-- ################################################################################## -->
-
-```{r metricsatstockchange, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
-
-# plot scale jumps between assessments when stock definition changed
-t3 %>%
-
-  ggplot(aes(x=year, y=value, group=label)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
-    strip.placement = "outside",
-    legend.title    = element_blank(),
-    legend.position  = "none"
-  ) +
-
-  geom_line(aes(colour=label), size=0.5) +
-
-  geom_dl(aes(label  = stockkeylabelold, colour = label),
-          method = list(dl.combine("first.bumpup"), cex = 0.8, hjust=0)) +
-
-  scale_y_continuous(labels=scientific_format(digits=2)) +
-  scale_x_continuous(breaks = pretty_breaks(n=3)) +
-  expand_limits(y=0) +
-  labs(x="", y="ssb") +
-  facet_wrap( ~ header, scales="free")
-
-```
-
-*`r fig_nums("stockdefchange")`*
-
-<!-- ################################################################################## -->
-<!-- COmpare the number of scale jumps (as ratio) over time                             -->
-<!-- ################################################################################## -->
-
-```{r jumpratio, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}
-
-d.jumps.byyear %>% 
-  filter(assessmentyear >= 1982) %>%
-
-  ggplot(aes(x=assessmentyear, y=ratio)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
-    strip.placement = "outside",
-    legend.title    = element_blank(),
-    legend.position  = "none"
-  ) +
-  geom_bar(stat="identity") +
-  labs(x="", y="ratio")
-
-```
-
-*`r fig_nums("jumpratio")`*
-
-```{r jumpratio, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}
-
-d.assessments%>% 
-  filter(assessmentyear <= 2018) %>% 
-  group_by(assessmentyear) %>% 
-  summarize(ncpueseries = mean(ncpueseries, na.rm=TRUE),
-            nsurveyseries = mean(nsurveyseries, na.rm=TRUE)) %>% 
-  gather(key=type, value=n, ncpueseries:nsurveyseries) %>%
-
-  ggplot(aes(x=assessmentyear, y=n, group=type)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
-    strip.placement = "outside",
-    legend.title    = element_blank()
-  ) +
-  # theme(legend.position  = "none") + 
-  geom_line(aes(colour=type)) +
-  labs(x="", y="avg")
-
-```

### 2019-05-16 R/DownloadDataFromSAG.r

@@ -108,6 +108,8 @@ iSAGstock_astext_complete <-
 
 save(iSAGstock_astext_complete, file=paste0(dropboxdir, "/rdata/iSAGstock_astext_complete", ".RData"))
 
+
+
 # Convert to dataset with appropriate field types
 iSAGstock <-
   iSAGstock_astext_complete %>% 
@@ -163,6 +165,11 @@ iSAGstock <-
     stockkeylabel    = ifelse(stockkeylabel == "ple-nsea" & assessmentyear %in% 2013:2014, "ple-nsea2", stockkeylabel)
   ) %>% 
   
+  # keep only the most recent assessment (in case of more than one assessment)
+  group_by(stockkey, stockkeylabel, assessmentyear, purpose) %>% 
+  mutate(assessmentkey = as.numeric(assessmentkey)) %>% 
+  filter(assessmentkey == max(assessmentkey, na.rm=TRUE) ) %>% 
+  
   # now do the stockkey transformations
   dplyr::select(-stockkey, -icesareas) %>% 
   left_join(iRename[,c("stockkeylabel","stockkey")], by="stockkeylabel") %>%
@@ -176,6 +183,7 @@ iSAGstock <-
 
 # Save dataset to file
 save(iSAGstock, file=paste0(dropboxdir, "/rdata/iSAGstock", ".RData"))
+
 
 
 # =====================================================================================

### 2019-05-16 R/iDatabases generator.r

@@ -37,6 +37,8 @@
 # 26/11/2018 added comparison to the data used by Esther Schuch
 # 20/03/2019 adapted with new data during HAWG
 # 23/03/2019 adapated after change to download data procedure
+# 06/05/2019 dplyr update. Changed "funs" to "list" in mutate_at functions 
+# 15/05/2019 removed double assessments from SAG
 
 # Warning messages:
 #   1: funs() is soft deprecated as of dplyr 0.8.0
@@ -88,13 +90,13 @@ iAdvice <-
               "fmax", "f01", "fmed", "f35spr", "flim", "fpa", "fmsy",
               "blim", "bpa", "msybtrigger",
               "m1", "m5"),
-            funs(as.numeric)) %>%
+            list(as.numeric)) %>%
   mutate_at(c("tacyear", "assessmentyear", "stockkey", "firstyearofdata", "ncpueseries","nsurveyseries"),
-            funs(as.integer)) %>% 
+            list(as.integer)) %>% 
   mutate_at(c("purpose", "speciescommonname", "stockkeylabel","stockkeylabelold","stockkeylabelnew",
               "stocksizeunits","fishingpressureunits"), 
-            funs(tolower)) %>% 
-  mutate_at(c("benchmark", "published"), funs(as.logical)) %>% 
+            list(tolower)) %>% 
+  mutate_at(c("benchmark", "published"), list(as.logical)) %>% 
   
   mutate(
     adviceonstock = ifelse(toupper(adviceonstock) == "Y", TRUE, adviceonstock),
@@ -116,7 +118,7 @@ iSpecies <-
     col_names=TRUE, 
     col_types="text") %>%
   mutate_at(c("speciescommonname","trophicguild","fisheriesguild","sizeguild"), 
-            funs(tolower)) %>%
+            list(tolower)) %>%
   group_by(speciesfaocode, speciesscientificname, speciescommonname) %>%
   arrange(speciesfaocode) 
 
@@ -181,17 +183,17 @@ qcsexcel <-
               "catches", "landings","discards","ibc","unallocatedremovals",
               "fishingpressure", "lowfishingpressure","highfishingpressure",
               "fdiscards","flandings","fibc","funallocated"), 
-            funs(as.numeric)) %>% 
+            list(as.numeric)) %>% 
   
   # make integer
[…]

### 2019-05-16 R/Analyse changes in mackerel.Rmd
Changes in mackerel

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Analyse changes in mackerel.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
# 06/05/2019 now only for mackerel
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)
library(directlabels)  # for printing labels at end of geom lines

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# create subset for advice
adv <-
  iAdvice %>%

  filter(grepl("mac-67|mac-nea", stockkeylabelold)) %>%
  filter(assessmentyear >= 2008) %>%

  # remove 2019 assessment (not yet carried out)
  filter(assessmentdate != 43711) %>%

  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
  group_by(stockkeylabelold, assessmentyear) %>% 
  mutate(n       = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>%

  mutate(purpose = ifelse(grepl("interbench", purpose), "benchmark", purpose)) %>% 
  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>% 
  mutate(ayear   = substr(as.character(assessmentyear), 3,4)) %>% 
  mutate(source  = "advice") %>% 
  arrange(assessmentyear, assessmentdate, purpose) %>% 
  ungroup()

# adv %>% distinct(stockkeylabelold, assessmentyear, purpose) %>% View()

adv_ssb <-
  adv %>% 
  dplyr::select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, id, assessmentyear, ayear, purpose, ssbay, source) %>% 
  rename(stocksize = ssbay) %>% 
  mutate(year = assessmentyear)

# create subset for assessments
ass <-
  iAssess %>% 
  filter(stockkeylabelold %in% unique(adv$stockkeylabelold)) %>% 
  filter(assessmentyear %in% unique(adv$assessmentyear)) %>% 
  filter(year <  assessmentyear) %>% 
  filter(year >= 2000) %>% 
  # remove 2019 assessment (not carried out yet)
  filter(!(purpose=="advice" & assessmentyear == 2019)) %>%

  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
  group_by(stockkeylabelold, assessmentyear) %>% 
  mutate(n       = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
  mutate(purpose = ifelse(grepl("interbench", purpose), "benchmark", purpose)) %>% 
  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>% 
  mutate(ayear   = substr(as.character(assessmentyear), 3,4)) %>% 
  mutate(source  = "assessment") %>% 
  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
  ungroup()

ass_final <-
  ass %>% 
  filter(assessmentyear == 2018, purpose == "advice")

ass_bench <-
  ass %>% 
  filter(assessmentyear == 2019, purpose == "benchmark")

ass_5year <-
  ass %>% 
  group_by(assessmentyear, purpose) %>% 
  filter(year >= (max(year, na.rm=TRUE)-4) ) %>%

  # add the current year estimate from adv
  bind_rows(adv_ssb) %>% 
  arrange(assessmentyear, assessmentdate, purpose, year)

ass_10year <-
  ass %>% 
  group_by(assessmentyear, purpose) %>% 
  filter(year >= (max(year, na.rm=TRUE)-9) ) %>%

  # add the current year estimate from adv
  bind_rows(adv_ssb) %>% 
  arrange(assessmentyear, assessmentdate, purpose, year)

# ass %>% distinct(stockkeylabelold, assessmentyear, purpose) %>% View()

```

**Analysis of assessment and advice changes in Northeast Atlantic Mackerel**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**Introduction**

In this document an overview is presented of the ICES assessments and advice for Northeast Atlantic Mackerel. 
The analysis is based on ICES advice database (Pastoors et al, in prep.) and the ICES stock assessment graph database ( [ICES SAG](www.http://ices.dk/marine-data/tools/Pages/stock-assessment-graphs.aspx)). The document outlines changes in reference points, changes in advice and a comparison between stock size assumed as the basis for the forecast and the most recent estimates of stock size from the 2018 assessments.

**Changes in reference points**

Fpa (blue) and Fmsy (green) reference points points as estimated in the different assessment years. Note that for a number of years it was no longer deemed necessary to provide Fpa estimates, and they were consequently removed from the advice documents. In the more recent years, Fpa values have been added again on the request of ICES clients. Blim (red) and MSY Btrigger (green) reference points points as estimated in the different assessment years.

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

pf <-
  adv %>% 
  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>% 
  # View()

  ggplot(aes(x=assessmentyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "right") +
  theme(legend.direction = "vertical") +

  geom_line(aes(y=value, group=var, colour=factor(var))) +
  geom_point(aes(y=value, group=var, colour=factor(var))) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="F", title="F reference points") +
  scale_color_manual(values = c("green", "blue"))

pb <-
  adv %>% 
  dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
  gather(key=var, value=value, blim:msybtrigger) %>%

  ggplot(aes(x=assessmentyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "right") +
  theme(legend.direction = "vertical") +

  geom_line(aes(y=value, group=var, colour=factor(var))) +
  geom_point(aes(y=value, group=var, colour=factor(var))) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="SSB", title="Biomass reference points") +
  scale_color_manual(values = c("red", "green"))

plot_grid(pf + theme(axis.title       = element_blank(),
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)), 
          pb + theme(axis.title       = element_blank(), 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Comparing perceived stock sizes with most recent estimates**

This section is devoted to the comparison of assessments and SSB estimates from forecasts. Because there have been a number of revisions and benchmarks for mackerel, the overall picture will be gradually built up. First the assessments (top) and the assessments including the SSB estimated in the forecast (bottom). The 2018 estimates of stock size (thick line) with the previous assessments (black). The year of the assessment is indicated. For each assessment, except for the 2018 assessment, only 5 years are shown. The SSB values estimated as part of the forecasts in each of the years is shown as a small dot.

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# adv %>% 
#   dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, purpose, ssbay) %>% 
#   rename(year = assessmentyear) %>%

  # View()
t <-
  ass_5year %>% 
  dplyr::select(stockkeylabelold, assessmentyear, year, ayear, purpose, stocksize, fishingpressure, id, source) %>%

  # need to generate unique labels, otherwise geom_dl will not work
  mutate(ayear = ifelse(purpose == "benchmark", paste0(ayear,"b"), ayear),
         ayear = ifelse(purpose == "replaced",  paste0(ayear,"r"), ayear)) %>% 
  filter(!is.na(stocksize)) %>% 
  arrange(id, year)

byears <- 
  t %>% 
  filter(purpose=="benchmark") %>% 
  distinct(assessmentyear)

ryears <- 
  t %>% 
  filter(purpose=="replaced") %>% 
  distinct(assessmentyear)

p1 <-
  t %>% 
  filter(purpose == "advice") %>% 
  filter(year < assessmentyear) %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  # geom_point(data=adv_ssb, aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="assessments")

p2 <-
  t %>% 
  filter(purpose == "advice") %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  geom_point(data=filter(adv_ssb, purpose=="advice"), aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="assessments + forecasts")

plot_grid(p1 + theme(legend.position  = "none", 
                     axis.title       = element_blank(),
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)), 
          p2 + theme(legend.position  = "none", 
                     axis.title       = element_blank(), 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Benchmarks and revisions**

Benchmark assessments have been carried out in 2014, 2017 and 2019 (blue) and are here compared with the assessment that were carried out in the same years. For benchmark assessments a "b" has been added to the year. Replaced assessments (due to errors in the assessment) have been carried out in 2015 and 2016 (red). For replaced assessments a "r" is added to the year

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# Now the benchmarks and the replaced assessments

p3 <-
  t %>% 
  filter(assessmentyear %in% byears$assessmentyear) %>% 
  filter(year < assessmentyear) %>% 
  filter(purpose %in% c("advice", "benchmark")) %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  # geom_point(data=filter(adv_ssb, purpose %in% c("advice", "benchmark")), aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="assessments + benchmarks")

p4 <-
  t %>% 
  filter(assessmentyear %in% ryears$assessmentyear) %>% 
  # filter(year < assessmentyear) %>% 
  filter(purpose %in% c("advice", "replaced")) %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  geom_point(data=filter(adv_ssb, purpose %in% c("advice", "replaced") & assessmentyear %in% ryears$assessmentyear), 
             aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="assessments + forecast + replaced assessments")

plot_grid(p3 + theme(legend.position  = "none", 
                     axis.title       = element_blank(),
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)), 
          p4 + theme(legend.position  = "none", 
                     axis.title       = element_blank(), 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Assessments, forecasts, benchmarks and replaced assessments all combined**

The top plot shows a comparison of the 2018 assessment, the 2019 benchmark and the implicit or perceived stock trend by connecting the SSB estimated in each of the forecasts. The bottom plot shows all information combined. Superimposed are the periods using the ICA assessment with egg survey only (2008-2012) and the SAM assessment with multiple tuning series (2013-2019).

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# Now the full plot

p5 <-
  t %>% 
  filter(purpose == "advice") %>%

  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize),
            inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  geom_point(data=filter(adv_ssb, purpose=="advice"), aes(colour = purpose)) +

  # geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_line(data=ass_bench, aes(colour=purpose ), size=1.2) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +

  expand_limits(y=0) + 
  labs(y="SSB",y="", title="Assessment, benchmark & perception")

p6 <-
  t %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  geom_point(data=adv_ssb, aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  geom_segment(aes(x=2008, xend=2012, y=2000000, yend=2000000), colour="red", size=2) +
  geom_segment(aes(x=2013, xend=2019, y=2000000, yend=2000000), colour="pink", size=2) +
  geom_dl(aes(x=2008, y=1800000, label="ica"), colour="red", method = list(dl.combine("last.points") ) )  +
  geom_dl(aes(x=2013, y=1800000, label="sam"), colour="pink", method = list(dl.combine("last.points") )) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +

  scale_x_continuous(breaks=pretty_breaks()) +

  expand_limits(y=0) + 
  labs(y="SSB",y="", title="Northeast Atlantic mackerel")

plot_grid(p5 + theme(legend.position  = "none", 
                     axis.title       = element_blank(),
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)), 
          p6+ theme(legend.position  = "none", 
                     axis.title       = element_blank(), 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Uncertainty estimates**

Mean + or - SD

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# Now the plot with uncertanty
# SE seem to be very very large. Plotted just +- SD

t1 <-
  t %>% 
  group_by(year) %>% 
  summarize(stocksize_mean = mean(stocksize, na.rm=TRUE),
            stocksize_n = n(),
            stocksize_sd = sd(stocksize, na.rm=TRUE)) %>%
  mutate(stocksize_se = 1.645 * stocksize_mean/sqrt(stocksize_n),
         stocksize_lb = stocksize_mean-stocksize_se,
         stocksize_ub = stocksize_mean+stocksize_se)

t %>% 
  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position="none") +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize),
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  # geom_point(data=filter(adv_ssb, purpose=="advice"), aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  # geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
  # geom_line(data=ass_bench, aes(colour=purpose ), size=1.2) +
  # geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  geom_point(data=t1, aes(x=year, y=stocksize_mean), inherit.aes = FALSE, size=1.5) +
  geom_errorbar(data=t1, aes(x=year, 
                             ymin=stocksize_mean-stocksize_sd, 
                             ymax=stocksize_mean+stocksize_sd ), inherit.aes = FALSE, size=0.9) +
  # geom_errorbar(data=t1, aes(x=year, 
  #                            ymin=stocksize_lb, 
  #                            ymax=stocksize_ub ), inherit.aes = FALSE) +
  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.6, replaced = 0.6)) +

  expand_limits(y=0) + 
  labs(y="SSB",y="", title="Assessments and uncertainty (SD)")

```

**Retrospective stock trends relative to MSY Btrigger and Fmsy**

The plots below show the retrospective stock trends relative to the MSY Btrigger and Fmsy values as used for the different assessment year.  
```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# Assessments relative to and Fmsy and MSY Btrigger (defined for the assessments) 
f <- 
  ass_final %>% 
  left_join(dplyr::select(adv, stockkeylabelold, assessmentyear, purpose, msybtrigger, fmsy)) %>% 
  mutate(relstocksize = stocksize / msybtrigger,
         relf         = fishingpressure / fmsy)

p1b <-
  t %>% 
  filter(purpose == "advice") %>% 
  filter(year < assessmentyear) %>% 
  left_join(dplyr::select(adv, stockkeylabelold, assessmentyear, purpose, msybtrigger, fmsy)) %>% 
  mutate(relstocksize = stocksize/msybtrigger) %>%

  ggplot(aes(x=year, y=relstocksize, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_path(data=ungroup(filter(adv_ssb, purpose == "advice")), aes(x=year, y=stocksize), 
  #           inherit.aes = FALSE, size=0.8, linetype="dashed", colour="gray40") +
  # geom_point(data=adv_ssb, aes(colour = purpose)) +

  geom_line(aes(size=purpose)) +
  geom_line(data=f, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  geom_hline(aes(yintercept=1), size=1.5, linetype="dashed") +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="ssb relative to msy btrigger")

p1f <-
  t %>% 
  filter(purpose == "advice") %>% 
  filter(year < assessmentyear) %>% 
  left_join(dplyr::select(adv, stockkeylabelold, assessmentyear, purpose, msybtrigger, fmsy)) %>% 
  mutate(relf = fishingpressure/fmsy) %>%

  ggplot(aes(x=year, y=relf, group=id, colour=purpose)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  geom_line(aes(size=purpose)) +
  geom_line(data=f, aes(colour=purpose ), size=1.5) +
  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +

  geom_hline(aes(yintercept=1), size=1.5, linetype="dashed") +

  scale_x_continuous(breaks=pretty_breaks()) +

  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
  expand_limits(y=0) + 
  labs(y="SSB", x="", title="F relative to Fmsy")

plot_grid(p1b + theme(legend.position  = "none", 
                     axis.title       = element_blank(),
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)), 
          p1f + theme(legend.position  = "none", 
                     axis.title       = element_blank(), 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Changes in advice**

Recommended catches (in fact: max of advised landings) presented for each of the different TAC years and relative change between years.  The light blue bars refer to replaced advice, the dark blue bars are the final advice.

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of advised landings
p7 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, purpose, advisedlandingsmax) %>% 
  filter(purpose %in% c("advice","replaced")) %>%
  mutate(purpose = factor(purpose, levels=c("replaced","advice"))) %>%

  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none") +

  # geom_hline(aes(yintercept=0), colour="black") +
  # geom_hline(aes(yintercept=20, label="20%"), colour="blue", linetype="dashed") +
  # geom_hline(aes(yintercept=-20, label="-20%"), colour="red", linetype="dashed") +
  geom_bar(aes(y  = advisedlandingsmax, fill=purpose), stat="identity", position=position_dodge2(width = 0.9, preserve = "single")) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advised landings (max)") +
  scale_fill_manual(values=c(advice = "blue",replaced = "lightblue"))

# plot of change in advice
p8 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>% 
  filter(adviceonstock==TRUE) %>% 
  mutate(previousadvice = lag(advisedlandingsmax),
         percchange     = 100*((advisedlandingsmax / previousadvice) -1),
         sign           = ifelse(percchange > 0, "pos", "neg")) %>%

  ggplot(aes(x=tacyear, y=percchange)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none") +

  geom_hline(aes(yintercept=0), colour="black") +
  geom_hline(aes(yintercept=20, label="20%"), colour="blue", linetype="dashed") +
  geom_hline(aes(yintercept=-20, label="-20%"), colour="red", linetype="dashed") +
  geom_bar(aes(fill=sign), stat="identity") +
  geom_text(aes(label=as.integer(percchange), colour=sign, vjust=ifelse(percchange >= 0, -0.1, 1.1)), 
            position=position_dodge(width=0.9) ) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=c(-100,100)) + 
  labs(y="Percentage advice change") +
  scale_fill_manual(values=c(pos = "blue",neg = "red")) +
  scale_colour_manual(values=c(pos = "blue",neg = "red"))

plot_grid(p7 + theme(legend.position  = "none", 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA),
                     axis.title.x     = element_blank() ), 
          p8 + theme(legend.position  = "none", 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**'TAC' over advice and Catch over 'TAC'**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of TAC over advice
p9 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, purpose, advisedlandingsmax, tal) %>% 
  filter(purpose %in% c("advice")) %>%
  rename(TAC=tal) %>% 
  gather(key=variable, value=value, advisedlandingsmax:TAC) %>%

  ggplot(aes(x=tacyear, y=value, group=variable)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  # theme(legend.position = "right") +
  # theme(legend.direction = "vertical") +

  # geom_bar(aes(fill=variable), stat="identity", position = position_dodge2()) + 
  geom_line(aes(colour=variable), size=1) + geom_point(aes(colour=variable), size=2) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="TAC and advice") +
  scale_fill_manual(values=c(advisedlandingsmax = "blue",TAC = "red")) +
  scale_colour_manual(values=c(advisedlandingsmax = "blue",TAC = "red"))

# plot of relative TAC over advice
p10 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, purpose, advisedlandingsmax, tal) %>% 
  filter(purpose %in% c("advice")) %>%
  rename(TAC=tal) %>% 
  mutate(relTAC = 100 * ((TAC/advisedlandingsmax)-1),
         sign   = ifelse(relTAC > 0, "pos", "neg")) %>%

  ggplot(aes(x=tacyear, y=relTAC)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none") +

  geom_hline(aes(yintercept=0), colour="black") +
  geom_hline(aes(yintercept=20, label="20%"), colour="blue", linetype="dashed") +
  geom_hline(aes(yintercept=-20, label="-20%"), colour="red", linetype="dashed") +
  geom_bar(aes(fill=sign), stat="identity") +
  geom_text(aes(label=as.integer(relTAC), colour=sign, vjust=ifelse(relTAC >= 0, -0.1, 1.1)), 
            position=position_dodge(width=0.9) ) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=c(-100,100)) + 
  labs(y="TAC / advice") +
  scale_fill_manual(values=c(pos = "blue",neg = "red")) +
  scale_colour_manual(values=c(pos = "blue",neg = "red"))

plot_grid(p9 + theme(strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA),
                     axis.title.x     = element_blank() ), 
          p10 + theme(legend.position  = "none", 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**'TAC' over advice and Catch over 'TAC'**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of landings over TAC
p11 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, purpose, tal, catches) %>% 
  filter(purpose %in% c("advice")) %>%
  rename(TAC=tal) %>% 
  gather(key=variable, value=value, TAC:catches) %>%

  ggplot(aes(x=tacyear, y=value, group=variable)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

  # geom_bar(aes(fill=variable), stat="identity", position = position_dodge2()) + 
  geom_line(aes(colour=variable), size=1) + geom_point(aes(colour=variable), size=2) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="TAC and advice") +
  scale_fill_manual(values=c(TAC = "red", catches="seagreen4")) + 
  scale_colour_manual(values=c(TAC = "red", catches="seagreen4"))

# plot of relative landings over TAC
p12 <-
  adv %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, purpose, tal, catches) %>% 
  filter(purpose %in% c("advice")) %>%
  rename(TAC=tal, catch=catches) %>% 
  mutate(relcatch = 100 * ((catch/TAC)-1),
         sign   = ifelse(relcatch > 0, "pos", "neg")) %>%

  ggplot(aes(x=tacyear, y=relcatch)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  theme(legend.position = "none") +

  geom_hline(aes(yintercept=0), colour="black") +
  geom_hline(aes(yintercept=20, label="20%"), colour="blue", linetype="dashed") +
  geom_hline(aes(yintercept=-20, label="-20%"), colour="red", linetype="dashed") +
  geom_bar(aes(fill=sign), stat="identity") +
  geom_text(aes(label=as.integer(relcatch), colour=sign, vjust=ifelse(relcatch >= 0, -0.1, 1.1)), 
            position=position_dodge(width=0.9) ) +

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=c(-100,100)) + 
  labs(y="catch / TAC") +
  scale_fill_manual(values=c(pos = "blue",neg = "red")) +
  scale_colour_manual(values=c(pos = "blue",neg = "red"))

plot_grid(p11 + theme(strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA),
                     axis.title.x     = element_blank() ), 
          p12 + theme(legend.position  = "none", 
                     strip.background = element_rect(colour=NA, fill=NA),
                     strip.text       = element_text(colour=NA)),
          ncol=1, align = 'hv')

```

**Conclusions**

* Historical advice and assessment data in preparation for scientific publication
* Since 2010: four benchmark assessments (of which three available in the database) and two revisions in advice due to errors. 
* Substantial volatility in assessment results from 2013 onwards. This is the year that the assessment was changed to SAM and using multiple tuning series. 
* Large revision in stock perception from 2018 assessment to 2019 benchmark
* 3 out of 10 years with a change in advice for more than 20%
* 10 out of 10 year with the sum of unilateral TACs higher than the scientific advice !
* Catches largely in line with the sum of the unilateral TACs.

**References**

Pastoors, M.A. et al (*in prep*) Documenting the past of scientific advice: the ICES advice database.

**Advice tables**

The tables below are the summaries of advice for the stocks mentioned in this paper. Most of the columns are self-explanatory. However, the columns 'purpose' and 'advice' have a specific meaning: purpose refers to the purpose of the advice, either used for advice (purpose is 'advice' and advice is TRUE) or the advice has been replaced or withdrawn (purpose is 'replaced' and advice is FALSE). The column SSB_STF refers to the SSB as estimated in the short term forecasts in the diferent years. All quantities are in tonnes, except for fishing mortality which is in 1/year.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

adv %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  pandoc.table(., 
             style        = "simple",
             split.tables = 180, 
             justify      = "right",
             missing      =".",
             big.mark     = '', 
             round        = c(0,0,0,2,2,0,0,0,0,2,0,0,0,0,0,0,0,0,0,0) )

```

```{r eval=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}

d %>% 
  dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,
                msybtrig=msybtrigger, ssb_stf=ssbay, tacyear, fadv=fadvmax,
                advlandings=advisedlandingsmax)  %>% 
  mutate(msybtrig = comma_format(big.mark ="") (msybtrig)) %>% 
  arrange(stock, assessyear) %>% 
  write.csv(., file="advice.csv")

```

### 2019-06-10 R/Analyse changes in pelagic stocks.Rmd
Updates on several codes

--- 2019-01-15 R/Analyse changes in pelagic stocks.Rmd
@@
 # 10/01/2018 made small changes to the table
+# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
+# 06/06/2019 changed the data frame d; now also has purpose
 # -----------------------------------------------------------------------------------------------
@@
 # create subset for advice
-d <-
+adv <-
   iAdvice %>%
@@
                stockkeylabelold)) %>%
-  filter(!grepl("benchmark", purpose)) %>% 
+  # filter(grepl("mac-67|mac-nea", 
+  #              stockkeylabelold)) %>%
+
+  mutate(purpose = ifelse(grepl("bench", purpose), "benchmark", purpose)) %>% 
+  mutate(purpose = ifelse(grepl("initial", purpose), "replaced", purpose)) %>%
+
+  filter(purpose %in% c("advice","replaced", "benchmark")) %>% 
+  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>%
+
   filter(assessmentyear >= 2008) %>% 
@@
 
+adv_ssb <-
+  adv %>% 
+  dplyr::select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, 
+                id, assessmentyear, purpose, ssbay) %>% 
+  rename(stocksize = ssbay) %>% 
+  mutate(year = assessmentyear)
+
+# adv %>% filter(stockkeylabelold =="mac-nea", assessmentyear >= 2015) %>% View()
+# adv %>% distinct(purpose) %>% View()
 # create subset for assessments
-e <-
+
+ass <-
   iAssess %>% 
-  filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
-  filter(assessmentyear == 2018) %>% 
-  filter(year >= 2008) %>% 
-  filter(year <= assessmentyear) %>% 
-  filter(adviceonstock == TRUE)
+  filter(stockkeylabelold %in% unique(adv$stockkeylabelold)) %>% 
+  filter(assessmentyear %in% unique(adv$assessmentyear)) %>% 
+  filter(year <  assessmentyear) %>% 
+  filter(year >= 2000) %>%
+
+  group_by(stockkeylabelold, assessmentyear) %>% 
+  mutate(n       = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>%
+
+  mutate(purpose = ifelse(grepl("bench", purpose), "benchmark", purpose)) %>% 
+  mutate(purpose = ifelse(grepl("initial", purpose), "replaced", purpose)) %>%
+
+  mutate(id      = paste0(stockkeylabelold, assessmentyear, assessmentdate, purpose)) %>% 
+  mutate(ayear   = substr(as.character(assessmentyear), 3,4)) %>% 
+  mutate(source  = "assessment") %>% 
+  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
+  ungroup()
+
+ass_final <-
+  ass %>% 
+  group_by(stockkeylabelold) %>%
+  filter(purpose == "advice") %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+  ungroup()
+
+ass_5year <-
+  ass %>% 
+  group_by(assessmentyear, purpose) %>% 
+  filter(year >= (max(year, na.rm=TRUE)-4) ) %>%
+
+  # add the current year estimate from adv
+  bind_rows(adv_ssb) %>% 
+  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
+  ungroup()
+
+ass_10year <-
+  ass %>% 
+  group_by(assessmentyear, purpose) %>% 
+  filter(year >= (max(year, na.rm=TRUE)-9) ) %>%
+
+  # add the current year estimate from adv
+  bind_rows(adv_ssb) %>% 
+  arrange(assessmentyear, assessmentdate, purpose, year) %>% 
+  ungroup()
+
+# ass <-
+#   iAssess %>% 
+#   group_by(assessmentyear) %>% 
+#   filter(stockkeylabelold %in% unique(adv$stockkeylabelold)) %>% 
+#   filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+#   filter(year <= assessmentyear) %>% 
+#   
+#   # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
+#   group_by(stockkeylabelold, assessmentyear) %>% 
+#   mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
+#   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
+#   filter(year >= 2008) %>% 
+#   ungroup()
+# #   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) )
+
+# adv %>%
+#   ungroup() %>%
+#   distinct(stockkeylabelold, assessmentyear,  benchmark, adviceonstock, n, .keep_all = FALSE) %>%
+#   arrange(stockkeylabelold, assessmentyear, benchmark, adviceonstock, n) %>%
+#   View()
 
@@
 
-d %>% 
+adv %>% 
+  filter(purpose == "advice") %>%
+
   dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
   gather(key=var, value=value, fpa:fmsy) %>% 
+  # View()
 
@@
 
-d %>% 
+adv %>% 
+  filter(purpose == "advice") %>%
+
   dplyr::select(stockkeylabelold, assessmentyear, blim, msybtrigger) %>% 
@@
 
-d %>% 
-  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>% 
+adv %>%
+
+  filter(purpose %in% c("advice","replaced")) %>%
+
+  dplyr::select(stockkeylabelold, tacyear, purpose, advisedlandingsmax) %>%
+
   ggplot(aes(x=tacyear)) +
@@
 
-  geom_point(aes(y       = advisedlandingsmax, 
-                 colour  = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
+  geom_point(aes(y       = advisedlandingsmax, colour=purpose)) +
   geom_line(aes(y        = advisedlandingsmax, 
-                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'),
-                linetype = ifelse(adviceonstock==TRUE,'advice',"replaced") )) +
+                colour   = purpose,
+                linetype = purpose )) +
 
@@
   scale_color_manual(values = c("red", "blue")) +
-  scale_linetype_manual(values = c("solid", "blank")) +
+  scale_linetype_manual(values = c("solid", "blank", "dotted")) +
   facet_wrap(~stockkeylabelold, scales="free_y")
@@
 
-d %>% 
-  dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, ssbay) %>% 
+# subset of ass
+t <-
+  ass %>% 
+  filter(purpose == "advice") %>% 
+  group_by(stockkeylabelold) %>% 
+  filter(assessmentyear == max(assessmentyear)  )
+
+adv %>%
+
+  filter(purpose %in% c("advice","replaced")) %>%
+
+  dplyr::select(stockkeylabelold, assessmentyear, purpose, ssbay) %>% 
   rename(year = assessmentyear) %>% 
+  # View()
 
@@
   geom_point(aes(y=ssbay,
-                 colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
-  geom_line(data=e,
+                 colour   = purpose)) +
+  geom_line(data=t,
             aes(y = stocksize )) +
@@
   facet_wrap(~stockkeylabelold, scales="free_y")
+```
+
+```{r eval=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}
+
+# NOT SURE WHAT THIS IS ADDING; NOT SHOWN
+
+# now based on assessment
+t <-
+  ass %>% 
+  filter(year == assessmentyear - 1 ) %>% 
+  mutate(colour  = paste0(adviceonstock,benchmark),
+         purpose = ifelse(grepl("bench", purpose), "benchmark", purpose),
+         purpose = ifelse(grepl("initial", purpose), "replaced", purpose))
+
+# glimpse(adv)
+
+ass %>% 
+  group_by(stockkeylabelold) %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+  dplyr::select(stockkeylabelold, assessmentyear, year, adviceonstock, stocksize) %>% 
+  # rename(year = assessmentyear) %>% 
+  # View()
+
+  ggplot(aes(x=year)) +
+  theme_publication() +
+  theme(legend.title = element_blank()) +
+
+  # geom_point(aes(y=ssbay,
+  #                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
+  geom_line(aes(y = stocksize )) +
+  geom_point(aes(y = stocksize )) +
+
+  geom_point(data=t, aes(x=year, y=stocksize, colour=purpose)) +
+  scale_x_continuous(breaks=pretty_breaks()) +
+  expand_limits(y=0) + 
+  labs(y="SSB") +
+  # scale_color_manual(values = c("red", "blue")) +
+  facet_wrap(~stockkeylabelold, scales="free_y")
+```
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+# Historical retro with all the variants
+
+ass_5year %>%
+
+  group_by(stockkeylabelold, assessmentyear) %>% 
+  filter(year < max(assessmentyear, na.rm=TRUE) ) %>%
+
+  ggplot(aes(x=year, y=stocksize, group=id, colour=purpose)) +
+  theme_publication() +
+  theme(legend.title = element_blank()) +
+
+  # geom_point(data=adv_ssb, aes(colour = purpose)) +
+
+  geom_line(aes(size=purpose)) +
+  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
+  # geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +
+
+  scale_colour_manual(values=c(advice = "black",benchmark = "blue",replaced = "red")) +
+  scale_size_manual  (values=c(advice = 0.6, benchmark = 0.9, replaced = 0.9)) +
+
+  scale_x_continuous(breaks=pretty_breaks()) +
+
+  expand_limits(y=0) + 
+  labs(y="SSB",y="") +
+  facet_wrap(~stockkeylabelold, scales="free_y")
+
+# distinct(ass_5year, purpose) %>% View()
+# ass %>% filter(grepl("bench", purpose)) %>% View()
+
 ```
@@
 
-d %>% 
+adv %>% 
   dplyr::select(stock=stockkeylabelold, assessyear=assessmentyear, purpose, fpa, fmsy, blim,

### 2019-06-10 R/ICES advice and assessment summaries.Rmd
Updates 15/1/2019
Updates on several codes

@@
 # 14/12/2018 updated for the new database structure
+# 01/06/2019 updated for herring advice round
 # -----------------------------------------------------------------------------------------------
@@
 
-# =====================================================================================
-# Load Stockkey and Rename data
-# =====================================================================================
+fay <- 2011
+fy  <- 2000
+lay <- 2019
+ly  <- 2019
 
-```
+# stock=c("spr-nsea","spr.27.3a4"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark==TRUE
+# stock=c("mac-nea"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark=TRUE; plot=c("stocksize","fishingpressure","recruitment")
 
-```{r functions,  message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-# stock="her-47d3"
-# firstassessyear=2014
-# lastassessyear=2018
-# firstyear=2000
-# lastyear=2020
+# d %>% filter(stockkeylabelold == "mac-nea", assessmentyear == 2019) %>% View()
+# unique(iAssess$purpose)
 
 # Function to plot historical assessments
-plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear) {
+plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, include.benchmark = FALSE, 
+                         plot=c("stocksize","fishingpressure")) {
 
@@
     ungroup() %>%
-    filter(purpose == "advice") %>% 
-    filter(stockkeylabelold == stock,
-           assessmentyear   %in% c(firstassessyear:lastassessyear), 
+
+    {if (include.benchmark == TRUE) {
+      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
+    } else {
+      filter(., purpose %in% c("advice")) 
+    }} %>%
+
+    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
+
+    filter(assessmentyear   %in% c(firstassessyear:lastassessyear), 
            year             %in% firstyear:lastyear) %>%
 
-    mutate(tyear     = substr(as.character(assessmentyear), 3,4),
-           type      = ifelse(assessmentyear == lastassessyear, "last", "other")) %>%
+    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>%
+
+    #set type
+    mutate(
+      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"       , "last"      , NA),
+      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose),  "last_bench", type),  
+      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"       , "advice"    , type),
+      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose),  "benchmark" , type),
+      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose) , "replaced"  , type),
+      type = ifelse(is.na(type), "other", type)) %>%
+
+    # set group
+    mutate(group = paste0(tyear, type)) %>%
+
     select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
-           type, source,
+           type, group, source,
            lowrecruitment, recruitment, highrecruitment, 
@@
 
+  # find out what is the last assessment
+  last <- filter(d, assessmentyear == lastassessyear) %>% distinct(type) %>% as.character()
+
   # plot stocksize
-  p1 <-
+  stocksize <-
     d %>% 
@@
 
-    ggplot(aes(year,stocksize, group=tyear)) +
+    ggplot(aes(year,stocksize, group=group)) +
 
@@
           plot.title  = element_text(size=10), 
+          plot.margin = margin(2, 2, 2, 2, "mm"),
           # strip.background = element_blank(),
@@
 
-    geom_ribbon(data=filter(d, type=="last"), 
+    geom_ribbon(data=filter(d, type==last), 
                 aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill=type), 
@@
 
-    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
                colour="red", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
@@
 
-    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes(x=assessmentyear, yintercept=msybtrigger), 
                colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes( firstyear, msybtrigger, label = "msy btrigger", vjust = -0.3, hjust = 0),
@@
 
-    scale_colour_manual  (values=c(last = "red",other="black")) +
-    scale_fill_manual    (values=c(last = "red",other="black")) +
-    scale_linetype_manual(values=c(last="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.0,   other=0.8)) +
+    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
+    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
+    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
+    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
 
@@
   # fishingpressure
-  p2 <-
+  fishingpressure <-
     d %>% 
@@
 
-    ggplot(aes(year,fishingpressure, group=tyear)) +
+    ggplot(aes(year,fishingpressure, group=group)) +
 
@@
 
-    geom_ribbon(data=filter(d, type=="last"), 
+    geom_ribbon(data=filter(d, type==last), 
                 aes(x=year, ymin=lowfishingpressure, ymax=highfishingpressure, fill=type), 
@@
 
-    geom_hline(data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes(x=assessmentyear, yintercept=fmsy), 
                colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold == stock, assessmentyear == lastassessyear),
+    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
                aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
@@
   # Recruitment
-  p3 <-
+  recruitment <-
     d %>% 
@@
 
-    ggplot(aes(year,recruitment, group=tyear)) +
+    ggplot(aes(year,recruitment, group=group)) +
 
@@
           axis.text.y = element_text(size=9),
+          plot.title  = element_text(size=10), 
+          plot.margin = margin(2, 2, 2, 2, "mm"),
           # strip.background = element_blank(),
@@
 
-    geom_ribbon(data=filter(d, type=="last"), 
+    geom_ribbon(data=filter(d, type==last), 
                 aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
@@
 
-  # Combined plot
-  # plot_grid(p1 + theme(legend.position  = "none", 
-  #                    axis.title       = element_blank(),
-  #                    strip.background = element_rect(colour=NA, fill=NA),
-  #                    strip.text       = element_text(colour=NA)), 
-  #         p2 + theme(axis.title       = element_blank()),
-  #         p3 + theme(axis.title       = element_blank()),
-  #         ncol=3, align = 'h')
+    if (length(plot) == 1) {
+      if ("stocksize" %in% plot ) print(stocksize)
+      if ("fishingpressure" %in% plot ) print(fishingpressure)
+      if ("recruitment" %in% plot ) print(recruitment)
 
-    plot_grid(p1, p2,  ncol=2, align = 'hv')
+    } else if (length(plot) == 2) {
+
+      if ("stocksize" %in% plot & "fishingpressure" %in% plot) plot_grid(stocksize, fishingpressure,  ncol=2, align = 'hv') 
+      if ("stocksize" %in% plot & "recruitment" %in% plot) plot_grid(stocksize, recruitment,  ncol=2, align = 'hv') 
+      if ("recruitment" %in% plot & "fishingpressure" %in% plot) plot_grid(fishingpressure, recruitment,  ncol=2, align = 'hv')
+
+    } else if (length(plot) == 3) {
+        plot_grid(stocksize, fishingpressure,  recruitment, ncol=length(plot), align = 'hv') 
+    }
 
@@
 
-# firstadviceyear <- 2016
-# lastadviceyear  <- 2019
-# stock           <- "her-47d3"
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"), include.benchmark = TRUE)
+
+# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("spr-nsea","spr.27.3a4"); include.replaced <- TRUE
+# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE
 
 # Function to plot advice overviews
-table_advice <- function(stock,firstadviceyear,lastadviceyear) {
+table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {
 
@@
     iAdvice %>%
-    filter(stockkeylabelold == stock, 
-           tacyear %in% firstadviceyear:lastadviceyear,
-           adviceonstock == TRUE) %>%
+
+    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
+
+    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%
+
+    {if (include.replaced == TRUE) {
+      filter(., purpose %in% c("advice","replaced"))
+    } else {
+      filter(., purpose %in% c("advice")) 
+    }} %>%
+
     mutate(advicebasis= substr(advicebasis,1,25),
@@
 
-    select(tacyear, advicebasis, advice, fadvmax, advicemax, tac, catch, 
+    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
            adv_c, tac_c, cat_c,
@@
   d %>% 
-  select(1, 2, 4, 5, 6, 7, 8, 9, 10) %>%
+  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
+  arrange(tacyear, desc(purpose)) %>% 
   pandoc.table(., 
@@
                big.mark     = '', 
-               round        = c(0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
+               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
 }
+
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
 
@@
 
-**Summary of ICES 2018 advice and assessments**
+**Summary of ICES 2019 assessments and advice**
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
 s   <- "her-47d3"
-fay <- 2010; lay <- 2018
-fy  <- 2000; ly  <- 2018
 
@@
 
-```
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               
+             plot = c("stocksize", "fishingpressure", "recruitment"))
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
 
 ```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-
-```
-
-Two weak year classes of North Sea herring are causing a decline in stock size and short-term prospects.
-The TAC and catches in 2018 have been substantially higher than the recommendations from ICES (600 000 instead of 518 000t).
-Due to the benchmark assessment of North Sea herring, the reference points for fishing mortality have been revised downwards.
-The combination of low recruitment, fishing harder than recommended and a lower Fmsy, lead to a substantially lower advice for 2019 (291 000 tonnes).
 
@@
 During the benchmark, a new method to standardize natural mortality estimates has been introduced which would hopefully circumvent re-scaling in the future?).
-However, this year there has been a revision of reference points again, leading to a lower Fmsy (0.26) compared to the previous estimate (0.33).
-Weak 2014 year class leads to substantially lower advice, which was already foreseen in 2017.
 
-[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.3a47d.pdf)
+Last year, the combination of low recruitment, fishing harder than recommended and a lower Fmsy, lead to a substantially lower advice for 2019 (291 000 tonnes).
+This year, the assessments estimates - overall - a somewhat higher stock size compared to the estimate in 2018.
+This is due to the 2018 herring acoustic survey, although the exact reason is not well understood.
+The advice for 2020 is higher than the advice that was given for 2019, despite that recruitment remains low and that the overall stock biomass is declining.
+The reason the advice is higher is because the catch forecast takes into account al the ages and weights of fish the make up the fishery.
+This is called the selectivity pattern.
+The fishery is heavily dependent on the abundance of the last strong year class (2013), which will be 7 year olds next year.
+The TAC advice uses only F on ages 2-6 years old, and this, together with the abundance and higher weight of 7 year olds is what is leading to a higher TAC for 2020.
+
+[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.3a47d.pdf)
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-s   <- "her-noss"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-
-```
-
-A new long-term management strategy for Atlanto-scandian herring has been evaluated and agreed.
+# s   <- "her-noss"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# A new long-term management strategy for Atlanto-scandian herring has been evaluated and agreed.
 Fmsy and the fishing mortality target are now higher than the previous values being used.
 The advice for 2019 is based on following the new management plan, with a TAC of around 590 000 tonnes (compared to an advice of 384 000 tonnes that was advised for 2018).
-
-The stock assessment is consistent with last year's assessment.
+# 
+# The stock assessment is consistent with last year's assessment.
 The stock appears to be slowly declining.
@@
 Last year, the advice was based on a target fishing mortality of 0.086, this year it is based on 0.14.
+# 
+# [her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.1-24a514a.pdf)
 
-[her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.1-24a514a.pdf)
+```
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
 s   <- "mac-nea"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
 
@@
 
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE, 
+             plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+
+# [mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/mac.27.nea.pdf)
+
 ```
+
+The ICES advice on Northeast Atlantic mackerel that was released in October 2018 has been received with a lot of scepticism and unbelief.
+According to that advice, the stock had been declining since 2011 and was estimated to be just below the MSY Btrigger.
+At the time, it was already found that the model results were very sensitive to the inclusion of the RFID tags.
+Leaving out only the RFID tags, gave a much higher stock estimate.
+
+To deal with this assessment problem, an inter-benchmark has been set up which is lead to a new assessment configuration in March 2019 and an updated advice for 2019 in May 2019.
+The new assessment configuration uses a more limited tagging dataset (restricted in ages, years and years out at sea) and a different way of estimating recruitment in the model.
+This lead to a substantially higher stock estimate and a substantial revision of the advice for 2019.
+
+[mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/special_requests/no.2019.09.pdf)
 
@@
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-
-```
-
-During the WGWIDE a press release came out on the Nordic swept-area survey that had shown that the biomass was 40% lower than the previous year.
-This survey result has an important effect on the stock assessment, as all the available indices now point to a lower stock size.
-This provides a much lower overall stock size compared to last year.
-Below is a comparison of the stock estimates of the current assessment (red line and uncertainty bands) in comparison the estimates of previous years.
-
-A major finding has been how sensitive the model results are to the inclusion of the RFID tags.
-Leaving out only the RFID tags, gives the perception of a much higher stock (see below).
-The RFID tags only  cover the most recent period.
-Compared to the old steel tags they exhibit a very strange phenomenon that the tag survival appears to be very low (~10%) compared to around 40% for the steel tags.
-Could it be that there is something wrong with the RFID tags or scanners in that they are missing fish?
-The problem is that the RFID tagging data are currently not assembled in a working group and are therefore not well evaluated.
-A proposal has been made to set up such a group.
-
-To deal with this problem of RFID tagging, an inter-benchmark has been set up which is leading up to a final meeting in the beginning of March 2019.
-
-[mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/mac.27.nea.pdf)
-
 ## Western horse mackerel
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-s   <- "mac-nea"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-
-```
-
-The stock assessment of Western horse mackerel is comparable to last year but there has been another revision upwards of the stock.
+# s   <- "hom-west"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# The stock assessment of Western horse mackerel is comparable to last year but there has been another revision upwards of the stock.
 This is especially relevant in conjunction with the biomass reference points that have been set after the benchmark in 2017 and where the biomass in 2015 (the last year in the time series) was selected as Blim.
@@
 The advice is based on the application of Fmsy.
-
-The assessment of western horse mackerel is very uncertain and there has been frequent re-scaling of the assessment during every new assessment.
+# 
+# The assessment of western horse mackerel is very uncertain and there has been frequent re-scaling of the assessment during every new assessment.
 This re-scaling is still happening.
@@
 The resulting advice is a direct application of the ICES advice rule to an assessment that could very well be too optimistic.
-
-[hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hom.27.2a4a5b6a7a-ce-k8.pdf)
-
-## Western horse mackerel
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-s   <- "whb-comb"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+#  
+# [hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hom.27.2a4a5b6a7a-ce-k8.pdf)
 
@@
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly)
+## Blue whiting
 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-
-```
-
-The stock size of blue whiting was high in 2018 but is predicted to be substantially lower in 2019 due to low recruitments coming in.
+# s   <- "whb-comb"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# The stock size of blue whiting was high in 2018 but is predicted to be substantially lower in 2019 due to low recruitments coming in.
 The stock size of blue whiting appears to have been overestimated.
@@
 The advice for 2019 is 1 144 000 tonnes, which is 18% lower than the advice for 2018.
-
-The blue whiting survey went well in 2018 and showed the highest biomass of the time series.
+# 
+# The blue whiting survey went well in 2018 and showed the highest biomass of the time series.
 Despite the high catches, the stock is still at a very high level although a decline does appear to have started.
 The most recent year classes of blue whiting are very weak so that decline in catching opportunities in the further years is foreseen.
+# 
+# [whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/whb.27.1-91214.pdf)
 
-[whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/whb.27.1-91214.pdf)
+```
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 s   <- "her-3a22"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
 
 cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+
+```
+
+Assessment was benchmarked in 2018 when reference points were also re-estimated.
+Redefinition of Blim from 90 kT to 120 kT which lead to stock having been below Blim for a number of years already.
+No possibility of rebuilding stock to Blim within one year (which is the unwritten new ICES rule), and therefore a zero catch advice.
+
+The new assessment has estimated the stock to be lower compared to the assessment of last year. This lead to a repeat of the zeor advice.
+
+[her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.20-24.pdf)
+
+## Baltic Sea herring
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+s   <- "her-2532-gor"
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               
+             plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+
+```
+
+Assessment largely consistent with last year's assessment. The 2014 year class is estimated to be strong.
+
+[her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.25-2932.pdf)
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+## Herring in 6a and 7bc
+
+# s   <- "her-67bc"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+# 
+# The advice for a zero TAC is the same as in the previous years.
+Unfortunately ICES has not presented advice on the proposed rebuilding plan that has been developed by the PELAC and that has been submitted for review to ICES and that had been evaluated by the herring working group in March 2018.
+In previous years the monitoring TAC for herring in 6a and 7bc had been suggested by ICES as a means to maintain the knowledge base for this herring stock.
+The monitoring TAC is still mentioned under issues relevant for the advice.
+# 
+# [her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.6a7bc.pdf)
+
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
+## Herring in the Irish Sea
 
-Assessment benchmarked in 2018.
-Revision of reference points.
-Redefinition of Blim from 90 kT to 120 kT which leads to stock having been below Blim for a number of years already.
-No possibility of rebuilding stock to Blim within one year (which is the unwritten new ICES rule), and therefore a zero catch advice.
-An additional analysis will be undertaken (by MP) to evaluate alternative options for the zero catch advice.
+# s   <- "her-nirs"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea. 
+# 
+# [her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.nirs.pdf)
 
-[her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.20-24.pdf)
-
-## Baltic Sea herring
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "her-2532-gor"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Substantially lower estimate of SSB and higher estimate of fishing mortality in the most recent assessment is probably due to the variability in the survey index.
-The 2014 year class is estimated to be strong.
-
-[her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.25-2932.pdf)
-
-## Herring in 6a and 7bc
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "her-67bc"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-The advice for a zero TAC is the same as in the previous years.
-Unfortunately ICES has not presented advice on the proposed rebuilding plan that has been developed by the PELAC and that has been submitted for review to ICES and that had been evaluated by the herring working group in March 2018.
-In previous years the monitoring TAC for herring in 6a and 7bc had been suggested by ICES as a means to maintain the knowledge base for this herring stock.
-The monitoring TAC is still mentioned under issues relevant for the advice.
-
-[her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.6a7bc.pdf)
-
-## Herring in the Irish Sea
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "her-nirs"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea.
-
-[her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.nirs.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "her-irls"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year.
+# s   <- "her-irls"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year.
 Evaluations conducted in 2018 showed that the agreed management strategy is no longer precautionary.
 Therefore the advice is according to the MSY approach.
+# 
+# [her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.irls.pdf)
 
-[her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/her.27.irls.pdf)
+```
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "spr-nsea"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
+s   <- c("spr-nsea", "spr.27.3a4")
 
 cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+
 ```
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
+Assessment was benchmarked in 2019, leading to a combination of sprat in 4 and sprat in 3a into a single stock.
+Some changes to the model al well, leading to a different perception of the history of the stock and lower stocksize in recent years.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-The change in recruitment level between this year’s assessment and the previous years is caused by a recent update of natural mortality of sprat in the first six months of their life.
-The 2018 assessment also revised the 2017 SSB downwards.
-The assessment has a tendency to overestimate above-average recruitments.
-
-[spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.4.pdf)
+[spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.4.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+
 s   <- "spr-2232"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
 
 cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-```
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+
 ```
@@
 
-[spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/spr.27.22-32.pdf)
+[spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.22-32.pdf)
+
+```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "arg-5b6a"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+# s   <- "arg-5b6a"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019. 
+# 
+# [aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019.
-
-[aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "hke-nrtn"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+# s   <- "hke-nrtn"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
+In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
+# 
+# [hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
-In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
-
-[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/hke.27.3a46-8abd.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "cod-347d"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+# s   <- "cod-347d"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
+According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey.
+The 2017 SSB is 32% lower compared to the estimate that was presented last year.
+This also leads to a substantially lower advice for 2019 compared to the advice for 2018.
+# 
+# [cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
-According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey.
-The 2017 SSB is 32% lower compared to the estimate that was presented last year.
-This also leads to a substantially lower advice for 2019 compared to the advice for 2018.
-
-[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/cod.27.47d20.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "ple-nsea"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+# s   <- "ple-nsea"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
+An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016).
+This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock.
+The advice is based on the MSY approach.
+# 
+# [ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
-An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016).
-This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock.
-The advice is based on the MSY approach.
-
-[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/ple.27.420.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "sol-nsea"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
+# s   <- "sol-nsea"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
+# 
+# [sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
 ```
@@
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea.
-
-[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/sol.27.4.pdf)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-s   <- "bss-47"
-fay <- 2014; lay <- 2018
-fy  <- 2000; ly  <- 2018
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly) 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1)
-```
-
-ICES has reissued the advice for 2018 and issued an advice for 2019.
+# s   <- "bss-47"
+# 
+# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# 
+# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
+# 
+# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+# 
+# ICES has reissued the advice for 2018 and issued an advice for 2019.
 ICES was requested by the EU to provide updated advice on fishing opportunities for 2018 because a benchmark of the assessment had been conducted in early 2018.
@@
 However, given the current low fishing mortality and the above-average recruitment in 2013 and 2014, SSB in 2019 and 2020 is predicted to be above Blim, allowing for some catches in 2018 and 2019.
+# 
+# [bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/bss.27.4bc7ad-h.pdf)
 
-[bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/bss.27.4bc7ad-h.pdf)
+```
 

### 2019-06-10 R/DownloadDataFromSAG.r

@@ -42,7 +42,7 @@ today <- format(Sys.time(), '%Y%m%d')
 # which year (set year = 0 for all years)
 # =====================================================================================
 
-myyear <- 2019
+myyear <- 2018
 
 # =====================================================================================
 # generate list of assessmentkeys
@@ -110,6 +110,7 @@ save(iSAGstock_astext_complete, file=paste0(dropboxdir, "/rdata/iSAGstock_astext
 
 
 
+
 # Convert to dataset with appropriate field types
 iSAGstock <-
   iSAGstock_astext_complete %>% 
@@ -125,19 +126,19 @@ iSAGstock <-
               "fdiscards","flandings","fibc","funallocated",
               "fpa","bpa", "flim", "blim", "fmsy", "msybtrigger",
               "custom1", "custom2", "custom3", "custom4", "custom5"), 
-            funs(as.numeric)) %>% 
+            list(as.numeric)) %>% 
   
   # make integer
-  mutate_at(c("year", "assessmentyear", "recruitmentage","stockdatabaseid"), funs(as.integer)) %>%
+  mutate_at(c("year", "assessmentyear", "recruitmentage","stockdatabaseid"), list(as.integer)) %>%
   
   # make logical
-  mutate_at(c("published"),  funs(as.logical)) %>% 
+  mutate_at(c("published"),  list(as.logical)) %>% 
   
   # make lowercase
   mutate_at(c("purpose", "stockkeylabel", "unitofrecruitment", "stocksizeunits", "stocksizedescription",
               "catcheslandingsunits", "fishingpressureunits",
               "customname1", "customname2", "customname3","customname4","customname5"), 
-            funs(tolower)) %>% 
+            list(tolower)) %>% 
   
   # change -alt for stock assessments to purpose "alternative"
   mutate(purpose       = ifelse(grepl("\\-alt", stockkeylabel), "alternative", purpose), 
@@ -169,16 +170,17 @@ iSAGstock <-
   group_by(stockkey, stockkeylabel, assessmentyear, purpose) %>% 
   mutate(assessmentkey = as.numeric(assessmentkey)) %>% 
   filter(assessmentkey == max(assessmentkey, na.rm=TRUE) ) %>% 
-  
+
+  ungroup() %>% 
+
   # now do the stockkey transformations
   dplyr::select(-stockkey, -icesareas) %>% 
[…]

### 2019-06-10 R/iAdvice advice tac and catch.r

@@ -33,7 +33,8 @@ compare <-
   iAdvice %>% 
   dplyr::select(year=tacyear, stockkeylabelold, advisedlandingsmax, tal, landings, catches) %>% 
   rename(advice=advisedlandingsmax, tac = tal) %>% 
-  filter(stockkeylabelold %in% c("her-47d3", "mac-nea","whb-comb", "her-noss")) %>% 
+  filter(stockkeylabelold %in% c("her-47d3")) %>% 
+  # filter(stockkeylabelold %in% c("her-47d3", "mac-nea","whb-comb", "her-noss")) %>% 
   # filter(stockkeylabelold %in% c("her-47d3", "mac-nea","whb-comb", "her-noss")) %>% 
   filter(year >= 2000) %>% 
   gather(key=variable, value=value, advice:catches)
@@ -41,7 +42,11 @@ compare <-
 # plot of comparison of advice, tac, landings and catch
 compare %>% 
   ungroup() %>% 
-  mutate(variable = factor(variable, levels=c("advice", "tac", "landings", "catches"))) %>% 
+  
+  filter(variable %in% c("advice", "catches")) %>% 
+  filter(!is.na(value)) %>% 
+  mutate(variable = factor(variable, levels=c("advice","catches"))) %>% 
+  # mutate(variable = factor(variable, levels=c("advice", "tac", "landings", "catches"))) %>% 
   
   ggplot(aes(x=year, y=value, group=variable)) +
   theme_publication() +
@@ -57,6 +62,24 @@ compare %>%
   spread(key=variable, value=value) %>% 
   mutate(deviance = as.integer(100*(tac/advice-1)) ) %>% 
   filter(!is.na(deviance)) %>% 
+
+  ggplot(aes(x=year, y=deviance)) +
+  theme_publication() +
+  geom_hline(aes(yintercept=0), colour="black") +
+  geom_line(colour="blue", size=1.5) +
+  facet_wrap(~stockkeylabelold) 
+
+# plot of difference between catch and advice
+compare %>% 
+  filter(variable %in% c("advice","catches")) %>% 
+  filter(!is.na(value)) %>% 
+  
+  spread(key=variable, value=value) %>% 
+  mutate(deviance = as.integer(100*(catches/advice-1)) ) %>% 
+  filter(!is.na(deviance)) %>%
+  
+  ungroup() %>% 
+  mutate(meandev = mean(deviance)) %>% 
 
   ggplot(aes(x=year, y=deviance)) +
   theme_publication() +

### 2019-06-10 R/iDatabases generator.r

@@ -457,3 +457,34 @@ save(iAssess, file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
 
 # unique(iAssess$purpose)
 # filter(iAssess, purpose=="historical") %>% View()
+
+# create overviews
+t1 <-
+  sag_unique %>% 
+  ungroup() %>% 
+  mutate(purpose = toupper(substr(purpose,1,1))) %>% 
+  group_by(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear) %>% 
+  arrange(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear, purpose) %>% 
+  summarize(purpose= paste(purpose, collapse="")) 
+
+t2 <-
+  qcsexcel_unique %>% 
+  ungroup() %>% 
+  mutate(purpose = ifelse(grepl("bench",purpose), "bench", purpose)) %>% 
+  mutate(purpose = ifelse(grepl("exploratory|trendsonly",purpose), "unofficial", purpose)) %>% 
+  mutate(purpose = tolower(substr(purpose,1,1))) %>% 
+  group_by(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear) %>% 
+  arrange(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear, purpose) %>% 
+  summarize(purpose= paste(purpose, collapse="")) 
+
+bind_rows(t1,t2) %>% 
+  filter(!is.na(stockkey)) %>% 
+  group_by(stockkeylabelold, stockkeylabelnew, stockkey, assessmentyear) %>% 
+  arrange(stockkeylabelold, stockkeylabelnew, stockkey, assessmentyear, purpose) %>% 
+  summarize(purpose= paste(purpose, collapse="")) %>% 
+  spread(key=assessmentyear, value=purpose) %>% 
+  write.csv(., file="iDatabases overview.csv", row.names=FALSE)
+
+
+
+  

### 2019-06-15 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Update 10 june 2019

@@
     reference_docx: ../report_template_v1.5.dotx
-bibliography: paper_02_scalejumps.bib
-csl: ices-journal-of-marine-science.csl
-link-citations: yes
 ---
@@
 
-fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
-         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")
-
 fig_nums(name    = "differentmethods", level = 2, display = FALSE, 
@@
 
-fig_nums(name    = "retroplots", level = 2, display = FALSE, 
-         caption = "Historic retrospective analysis of the nine stocks analysed.
-For each stock, the most recent assessment is in red and the previous assessments in black.
-The end point of each assessment is indicated by a dot.
-Stock definitions may have changed over time, but this is not included in the graphs")
-
 # section2 -----------------------------------------------------
@@
 
-fig_nums(name    = "retrometrics", level = 1, display = FALSE, 
-         caption = " Retrospective metrics for 9 species derived from historical assessment records.
-Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
-In red: calculated values based on the whole time-series of historical data.
-In blue: based on retrospective peels of 5 years.
-Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
-
-fig_nums(name    = "retrometrics2", level = 2, display = FALSE, 
-         caption = "Summary of retrospective metrics for 9 species derived from historical assessment records.
-Metrics ab (average bias) and asd (average variation) from Jonsson & Hjorleifsson 2000, rho from Mohn 1999 and sigma from Ralston et al 2011.
-In red: calculated values based on the whole time-series of historical data.
-In blue: based on retrospective peels of 5 years.
-Metrics ab, asd and sigma on a log scale, rho on a ratio scale.")
-
 # section 3 -----------------------------------------------------
-
-tab_nums(name    = "scalejumps", level = 1, display = FALSE, 
-         caption = "Percentage assessments with scale jumps of more than 10% expressed by decade. Between brackets: number of assessments within the decade.")
-
-fig_nums(name    = "scalecomparisons", level = 1, display = FALSE, 
-         caption = "Example of pairwise assessment comparisons for Northeast Atlantic mackerel  over a period of 10 years in common.
-The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
-
-fig_nums(name    = "scalejumps", level = 2, display = FALSE, 
-         caption = "Retrospective scale difference (sigma) for 9 species derived from historical assessment records.
-Results from pairwise comparisons over periods of 10 years.
-Large dots  refer to years when the scale difference is larger than 10%.
-Colours indicate the decades.
-Black triangles indicate benchmark assessment years.
-The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
-
-fig_nums(name    = "metricsinpeels", level = 2, display = FALSE, 
-         caption = "Comparison of retrospective uncertainty metrics from historical assessment records, calculated in peels, normalized between 0 and 1.
-Asd from Jonsson & Hjorleifsson 2000, rhoabs as absolute value of rho from Mohn 1999, sigma from Ralston et al 2011, percjumps as the percentage of jumps (this paper) and phiabs as the absolute value of phi (this paper).")
-
-# section 4 -----------------------------------------------------
 
@@
 <!-- Tables ------------------------------------------------------ -->
-
-**Tables**
-
-*`r tab_nums("scalejumps")`*
-
-```{r}
-
-# print table
-d.jumps.bydecade %>% 
-  filter(variable == "stocksize") %>% 
-  ungroup() %>% 
-  select(stock, decade, ratio3) %>% 
-  spread(key=decade, value=ratio3) %>% 
-  pandoc.table(.,
-               style = "simple",
-               split.tables=400, 
-               justify = "left",
-               missing=".",
-               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-
-```
 
@@
 ```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
+         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")
 
@@
 
-<!-- ################################################################################## -->
-<!-- Plot different methods of creating historical retrospectives --------------------- -->
-<!-- ################################################################################## -->
-
-```{r echo=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE}
-
-knitr::include_graphics("fig1.2_retromethods.jpg")
-
-```
-
-*`r fig_nums("differentmethods")`*
-
-#####Page break
-
 <!-- Plot Retroplots --------------------------- -->
@@
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+
+fig_nums(name    = "retroplots", level = 2, display = FALSE, 
+         caption = "Historic retrospective analysis of the nine stocks analysed.
+For each stock, the most recent assessment is in red and the previous assessments in black.
+The end point of each assessment is indicated by a dot.
+Stock definitions may have changed over time, but this is not included in the graphs")
 
@@
 
-#####Page break
-
 <!-- ################################################################################## -->
@@
 
-```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
+
+fig_nums(name    = "pairwisecomparisons", level = 1, display = FALSE, 
+         caption = "Pairwise assessment comparisons for the five most recent assessment pairs over a period of 5  years in common.
+The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
 
 myvar         <- "stocksize"
-myspecies     <- "her"
-firstyear     <- 2008
+# myvar         <- "fishingpressure"
+myspecies     <- c("cod", "mac", "hom", "had", "her","hke","ple","sol","whb")
+firstyear     <- 2013
 firstplotyear <- 2000
+lastplotyear  <- 2020
 mynyears      <- 5
 myremoveyears <- 0
-mythreshold   <- 0.1
-mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", number of years: ",mynyears, 
+mythreshold   <- 0.2
+mytitle       <- paste0(myvar, ", number of years: ",mynyears, 
                         ", removed years: ", myremoveyears)
 
+# selection of pairwise means
 d.pairwise.means <-
@@
   # View()
+  mutate(species = substr(stockkeylabelold,1,3)) %>%
 
@@
   filter(variable == myvar) %>% 
-  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(species %in% myspecies) %>% 
   filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
@@
   # calculate means per period
-  group_by(stock, assessmentpair, variable) %>% 
+  group_by(stock, species, assessmentpair, variable) %>% 
   mutate(   firstyear      = min(year),
             lastyear       = max(year)) %>% 
-  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  group_by(stock, species, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
   summarise(value = mean(value, na.rm=TRUE)) %>% 
   gather(key=tmp, value=year, firstyear:lastyear) %>% 
-  arrange(stock, assessmentpair, variable, assessmentyear, label)
-
-res$d.pairwise %>%
-
+  arrange(stock, species, assessmentpair, variable, assessmentyear, label)
+
+# selection of pairwise long timeseries
+d.pairwise.roughly <-
+
+  res$d.pairwise.roughly %>% 
+  mutate(species = substr(stockkeylabelold,1,3)) %>% 
   filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
   filter(variable == myvar) %>% 
-  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(species %in% myspecies) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear)
+
+# selection of pairwise phi texts
+d.pcm <-
+  res$d.pcm %>% 
+  mutate(species = substr(stock,1,3)) %>% 
+  filter(variable == myvar) %>% 
+  filter(species %in% myspecies) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>% 
+  filter(numberofyears == mynyears) %>% 
+  filter(removelastyears == myremoveyears) %>% 
+  filter(threshold == mythreshold)
+
+# start plot
+res$d.pairwise %>%
+
+  mutate(species = substr(stockkeylabelold,1,3)) %>%
+
+  filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
+  filter(variable == myvar) %>% 
+  filter(species %in% myspecies) %>% 
   filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
@@
     strip.text.y     = element_text(angle = 0),
-   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
+    axis.text.x      = element_text(angle = 90, vjust = 0.5), 
     legend.position  = "none"
@@
 
-  geom_line(data=filter(res$d.pairwise.roughly, 
-                        variable == myvar, 
-                        grepl(myspecies, stockkeylabelold),
-                        as.numeric(substr(assessmentpair,1,4)) >= firstyear,
-                        year >= firstplotyear), 
-            size=0.5) +
-
+  # long black lines
+  geom_line(data=d.pairwise.roughly,size=0.5) +
+
+  # short, thick, coloured comparison lines
   geom_line(aes(colour=factor(label)), size=1) +
 
+  # coloured means segments with start and end points
   geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)),
@@
              aes(colour=factor(label)), size=1) +
-  geom_text (data=filter(res$d.pcm,
-                         variable == myvar,
-                         grepl(myspecies, stock),
-                         as.numeric(substr(assessmentpair,1,4)) >= firstyear,
-                         numberofyears == mynyears,
-                         removelastyears == myremoveyears,
-                         threshold == mythreshold),
-             aes(x=firstplotyear, y=0, label = paste0("phi=", round(phi_abs, digits=2))), 
-             hjust=0, vjust=0, nudge_x = -0.1, nudge_y=0.1,  inherit.aes = FALSE) +
-
+
+  # phi text
+  geom_text (data=d.pcm,
+             aes(label = paste0("phi=", round(phi_abs, digits=2))), 
+             x=firstplotyear, y=0, hjust=0, vjust=0,  inherit.aes = FALSE) +
+
+  xlim(firstplotyear,lastplotyear) +
   scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
-  scale_x_continuous(breaks = pretty_breaks(n=2)) +
+  # scale_x_continuous(breaks = pretty_breaks(n=2)) +
   expand_limits(y=0) +
   labs(x="", y=myvar, title=mytitle) +
-  facet_wrap( ~ assessmentpair, ncol=5)
+  facet_grid(species~assessmentpair, scales = "free_y")
 
@@
 
+fig_nums(name    = "scalejumps", level = 1, display = FALSE, 
+         caption = "Retrospective scale difference (phi) for 9 species derived from historical assessment records.
+Results from pairwise comparisons over periods of 5 years.
+Large dots  refer to years when the scale difference is larger than 20%.
+Colours indicate the decades.
+Black triangles indicate benchmark assessment years.
+The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
+
 # res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)
@@
 # threshold <- 0.10
-mynyears      <- 10
+mynyears      <- 5
 myremoveyears <- 0
-mythreshold   <- 0.1
+mythreshold   <- 0.2
 myvar         <- "stocksize"
@@
 
+tab_nums(name    = "scalejumps2", level = 1, display = FALSE, 
+         caption = "Example (mackerel) of sensivity of scale-jumps to the number of years included and the threshold for defining a scale-jump between two assessments.")
+
 # plot scale jumps with different settings for number of years and thresholds
@@
 myvar         <- "stocksize"
-myspecies     <- "cod"
+myspecies     <- "mac"
 mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
@@
 
-```{r metricsatstockchange, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+```{r echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}
+
+# Compare scale jumps when stock definition has changed.
+
+mynyears      <- 5
+myremoveyears <- 0
+myvar         <- "stocksize"
+mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
+
+# select combinations of stock and assessmentyear when stock definition has changed
+t <-
+  d %>% 
+  select(assessmentyear, stock, stockkeylabelold) %>% 
+  distinct() %>% 
+  mutate(
+    stock = factor(stock),
+    stock = factor(stock, levels=rev(levels(stock)))) %>% 
+  group_by(stock, stockkeylabelold) %>% 
+  summarize(
+    firstyear = min(assessmentyear),
+    lastyear  = max(assessmentyear)) %>% 
+  arrange(stock, firstyear) %>% 
+  group_by(stock) %>% 
+  mutate(
+    previous       = lag(lastyear),
+    assessmentpair = paste(previous, firstyear, sep="-"),
+    stockold       = lag(stockkeylabelold)) %>% 
+  filter(!is.na(previous))  %>% 
+  select(stock, assessmentpair)
+
+# select the assessment information after the change
+t1 <-
+  t %>% 
+  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
+  select(-assessmentyearprior) %>% 
+  left_join(d, by=c("stock", "variable", "assessmentyearlast" = "assessmentyear")) %>% 
+  mutate(label="last") %>% 
+  rename(assessmentyear = assessmentyearlast)
+
+# data series with previous (prior) estimation
+t2 <-
+  t %>% 
+  left_join(d.pairs, by=c("stock", "assessmentpair")) %>% 
+  select(-assessmentyearlast) %>% 
+  left_join(d, by=c("stock", "variable", "assessmentyearprior" = "assessmentyear")) %>% 
+  mutate(label="prior") %>% 
+  rename(assessmentyear = assessmentyearprior)
+
+# combine before and after the change
+t3 <-
+  bind_rows(t2, t1) %>% 
+  group_by(stock, assessmentpair, variable) %>%
+
+  # filter the appropriate period
+  filter(year <= max(assessmentyear) - myremoveyears,
+         year >= max(assessmentyear) - myremoveyears - mynyears) %>%
+
+  # sort
+  arrange(stock, assessmentpair, assessmentyear, year) %>%
+
+  # filter only years that are in common in last and prior
+  group_by(stock, assessmentpair, variable, year) %>% 
+  mutate(
+    n = n(),
+    header = paste(stock, assessmentpair, sep=" ")
+  ) %>% 
+  filter(n() == 2)
+
+# calculate pairwise means
+t3.pairwise.means <-
+
+  t3 %>% 
+  group_by(header, stock, assessmentpair, variable) %>% 
+  distinct(header, stock, assessmentpair, variable) %>% 
+  left_join(res$d.pairwise, by=c("stock", "assessmentpair", "variable")) %>%
+
+  filter(
+    removelastyears == myremoveyears,
+    numberofyears   == mynyears,
+    variable        == myvar) %>%
+
+  # calculate means per period
+  group_by(header, stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(header, stock, assessmentpair, variable, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(header, stock, assessmentpair, variable, label)
+
+# selection of pairwise long timeseries
+t3.pairwise.roughly <-
+
+  t3 %>% 
+  group_by(header, stock, assessmentpair, variable) %>% 
+  distinct(header, stock, assessmentpair, variable) %>% 
+  left_join(res$d.pairwise.roughly, by=c("stock", "assessmentpair", "variable")) %>%
+
+  filter(
+    removelastyears == myremoveyears,
+    numberofyears   == mynyears,
+    variable        == myvar)
+
+# get phi for selected assessment pairs
+t3.pcm <-
+
+  t3 %>% 
+  group_by(header, stock, assessmentpair, variable) %>% 
+  distinct(header, stock, assessmentpair, variable) %>% 
+  left_join(res$d.pcm, by=c("stock", "assessmentpair", "variable")) %>%
+
+  filter(
+    removelastyears == myremoveyears,
+    numberofyears   == mynyears,
+    variable        == myvar) %>% 
+  ungroup() %>% 
+  distinct(header, stock, assessmentpair, variable, .keep_all = TRUE)
 
@@
 t3 %>%
+
+  filter(variable == myvar) %>%
 
@@
 
-  geom_line(aes(colour=label), size=0.5) +
-
-  geom_dl(aes(label  = stockkeylabelold, colour = label),
-          method = list(dl.combine("first.bumpup"), cex = 0.8, hjust=0)) +
+  # geom_dl(aes(label  = stockkeylabelold, colour = label),
+  #         method = list(dl.combine("first.bumpup"), cex = 0.8, hjust=0)) +
+
+  # long black lines
+  geom_line(data=t3.pairwise.roughly,size=0.5) +
+
+  # This coloured line segments
+  geom_line(aes(colour=label), size=1.5) +
+
+  # coloured means segments with start and end points
+  geom_line(data=t3.pairwise.means, aes(x=year, y=value, colour=label),
+            linetype="dashed", lineend="butt", inherit.aes = FALSE) +
+  geom_point(data=filter(t3.pairwise.means, year==min(year)),
+             aes(colour=label), size=1) +
+  geom_point(data=filter(t3.pairwise.means, year==max(year)),
+             aes(colour=label), size=1) +
+
+  # phi text
+  geom_text (data=t3.pcm,
+             aes(label = paste0("phi=", round(phi_abs, digits=2)), x= min(t3.pairwise.roughly$year)), 
+             y=0, 
+             hjust=0, vjust=0,  inherit.aes = FALSE) +
+
+  # annotate("text", label = "left", 
+  #           x =  min(t3.pairwise.roughly$year), y = 0, hjust = 0) +
 
@@
 
+# unique(d.pairwise.means$label)
+# unique(d.pairwise.roughly$label)
+# unique(t3$label)
+
 ```
@@
 
-```{r jumpratio, eval=FALSE, fig.align="center", fig.asp=0.5, message=FALSE, warning=FALSE, include=FALSE}
+```{r, eval=FALSE, fig.align="center", fig.asp=0.5, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 
-```{r metricsinpeels, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
+```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}
 
@@
 ```
+
+```{r eval=FALSE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}
+
+myvar         <- "stocksize"
+myspecies     <- "mac"
+firstyear     <- 2008
+firstplotyear <- 2000
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", number of years: ",mynyears, 
+                        ", removed years: ", myremoveyears)
+
+d.pairwise.means <-
+
+  res$d.pairwise %>% 
+  # ungroup() %>% 
+  # distinct(removelastyears, numberofyears, threshold) %>% 
+  # View()
+
+  filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  # calculate means per period
+  group_by(stock, assessmentpair, variable) %>% 
+  mutate(   firstyear      = min(year),
+            lastyear       = max(year)) %>% 
+  group_by(stock, assessmentpair, variable, assessmentyear, label, firstyear, lastyear) %>% 
+  summarise(value = mean(value, na.rm=TRUE)) %>% 
+  gather(key=tmp, value=year, firstyear:lastyear) %>% 
+  arrange(stock, assessmentpair, variable, assessmentyear, label)
+
+res$d.pairwise %>%
+
+  filter(removelastyears==myremoveyears, numberofyears==mynyears) %>% 
+  filter(variable == myvar) %>% 
+  filter(grepl(myspecies, stockkeylabelold)) %>% 
+  filter(as.numeric(substr(assessmentpair,1,4)) >= firstyear) %>%
+
+  ggplot(aes(x=year, y=value, group=assessmentyear)) +
+  theme_publication() +
+  theme(
+    panel.spacing.x = unit(1, "mm"),
+    panel.spacing.y = unit(1, "mm"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(1,0,0,0, "mm")),
+    strip.text.y     = element_text(angle = 0),
+   # axis.text.x      = element_text(angle = 90, vjust = 0.5), 
+    legend.position  = "none"
+  ) +
+
+  geom_line(data=filter(res$d.pairwise.roughly, 
+                        variable == myvar, 
+                        grepl(myspecies, stockkeylabelold),
+                        as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+                        year >= firstplotyear), 
+            size=0.5) +
+
+  geom_line(aes(colour=factor(label)), size=1) +
+
+  geom_line(data=d.pairwise.means, aes(x=year, y=value, colour=factor(label)),
+            linetype="dashed", lineend="butt", inherit.aes = FALSE) +
+  geom_point(data=filter(d.pairwise.means, year==min(year)),
+             aes(colour=factor(label)), size=1) +
+  geom_point(data=filter(d.pairwise.means, year==max(year)),
+             aes(colour=factor(label)), size=1) +
+  geom_text (data=filter(res$d.pcm,
+                         variable == myvar,
+                         grepl(myspecies, stock),
+                         as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+                         numberofyears == mynyears,
+                         removelastyears == myremoveyears,
+                         threshold == mythreshold),
+             aes(x=firstplotyear, y=0, label = paste0("phi=", round(phi_abs, digits=2))), 
+             hjust=0, vjust=0, nudge_x = -0.1, nudge_y=0.1,  inherit.aes = FALSE) +
+
+  scale_y_continuous(breaks = pretty_breaks(n=3), labels=scientific_format(digits=1)) +
+  scale_x_continuous(breaks = pretty_breaks(n=2)) +
+  expand_limits(y=0) +
+  labs(x="", y=myvar, title=mytitle) +
+  facet_wrap( ~ assessmentpair, ncol=5)
+
+# names(res)
+# filter(res$d.pcm,
+#                          variable = myvar,
+#                          grepl(myspecies, stock),
+#                          as.numeric(substr(assessmentpair,1,4)) >= firstyear,
+#                          numberofyears == mynyears,
+#                          removelastyears == myremoveyears,
+#                          threshold == mythreshold)
+```

### 2019-08-15 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Updates 15/8

@@
     stock = ifelse(grepl("her-4",stockkeylabelold), "her (north sea)",stock),
-    stock = ifelse(grepl("mac-67|mac-nea|mac-west",stockkeylabelold), "mac (west)",stock),
+    stock = ifelse(grepl("mac-67|mac-nea|mac-west",stockkeylabelold), "mac (nea)",stock),
     stock = ifelse(grepl("cod-34|cod-nsea",stockkeylabelold), "cod (north sea)",stock),
-    stock = ifelse(grepl("whb-c|whb-n",stockkeylabelold), "whb (comb)",stock),
+    stock = ifelse(grepl("whb-c|whb-n",stockkeylabelold), "whb (nea)",stock),
     stock = ifelse(grepl("sol-ns",stockkeylabelold), "sol (north sea)",stock),
-    stock = ifelse(grepl("hom-c|hom-w",stockkeylabelold), "hom (west)",stock),
+    stock = ifelse(grepl("hom-c|hom-w",stockkeylabelold), "hom (nea)",stock),
     stock = ifelse(grepl("had-3|had-n",stockkeylabelold), "had (north sea)",stock),
-    stock = ifelse(grepl("hke-n",stockkeylabelold), "hke (north)",stock)
+    stock = ifelse(grepl("hke-n",stockkeylabelold), "hke (nea)",stock)
   ) %>%
@@
 
-#####page break
-
-**Figures**
-
 <!-- ################################################################################## -->
@@
 ```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
-
-fig_nums(name    = "stockdefinitions", level = 1, display = FALSE,
-         caption = "Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series")
 
@@
 
-*`r fig_nums("stockdefinitions")`*
+*Figure 1: Definitions of stocks (by year) that have been ‘lumped’ into meta-stocks for long time-series*
 
@@
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}
-
-fig_nums(name    = "retroplots", level = 2, display = FALSE, 
-         caption = "Historic retrospective analysis of the nine stocks analysed.
-For each stock, the most recent assessment is in red and the previous assessments in black.
-The end point of each assessment is indicated by a dot.
-Stock definitions may have changed over time, but this is not included in the graphs")
 
@@
 
-*`r fig_nums("retroplots")`*
+*Figure 2: Historic retrospective patterns in stock size (SSB) of the nine meta-stocks consid-ered.
+For each meta-stock, the most recent assessment is in red and the previous assess-ments in black.
+The end point of each assessment is indicated by a dot.*
 
@@
 ```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
-
-fig_nums(name    = "pairwisecomparisons", level = 1, display = FALSE, 
-         caption = "Pairwise assessment comparisons for the five most recent assessment pairs over a period of 5  years in common.
-The dotted lines indicate the average values of the different assessments which are used to calculate scale differences between assessments.")
 
@@
 
-*`r fig_nums("scalecomparisons")`*
+*Figure 3: Pairwise Comparison of Means for the assessment pairs between 2013 and 2018 based on historic retrospective estimates of stock size (SSB) of the nine meta-stocks consid-ered.
+The coloured segments indicate the years for which the PCM is calculated (blue = first year, red = second year) and the coloured horizontal dotted lines indicated the means.*
 
 #####Page break
+
+```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}
+
+tab_nums(name    = "scalejumps2", level = 1, display = FALSE, 
+         caption = "Example (mackerel) of sensivity of scale-jumps to the number of years included and the threshold for defining a scale-jump between two assessments.")
+
+# plot scale jumps with different settings for number of years and thresholds
+
+myremoveyears <- 0
+myvar         <- "stocksize"
+myspecies     <- "mac"
+mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
+
+res$d.pcm %>%
+
+  filter(grepl(myspecies, stock)) %>% 
+  filter(variable        == myvar) %>% 
+  filter(removelastyears == myremoveyears) %>%
+
+  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  gather(key=metric, value=value, c(phi)) %>%
+
+  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
+  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
+  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
+
+  geom_line (aes(colour=factor(decade))) +
+  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
+
+  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+  coord_cartesian(ylim=c(-1, 1),
+                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
+
+  # ggMarginal(data=d.assessments, type="histogram")
+  scale_fill_manual(values = myBlues) +
+
+  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
+  scale_x_continuous(breaks = pretty_breaks()) +
+
+  labs(x=" ", y="", title=mytitle) +
+  facet_grid(threshold ~ numberofyears)
+
+# filter(d.assessments, benchmark == 1) %>% View()
+
+```
+
+*Figure 4: Example of an evaluation of the sensitivity of scale-jumps to the number of years over which the means are calculated (2, 5 or 10 years; columns) and the threshold for defin-ing a scale-jump (0.1, 0.2; rows) based on the historical assessments of NEA Mackerel.
+Dot-ted horizontal lines indicate the thresholds.
+Large dots refer to years when the scale differ-ence is larger than the threshold.
+Colours indicate the decades.*
+
+#####page break
 
@@
 
-fig_nums(name    = "scalejumps", level = 1, display = FALSE, 
-         caption = "Retrospective scale difference (phi) for 9 species derived from historical assessment records.
-Results from pairwise comparisons over periods of 5 years.
-Large dots  refer to years when the scale difference is larger than 20%.
-Colours indicate the decades.
-Black triangles indicate benchmark assessment years.
-The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.")
-
 #  res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)
@@
 myremoveyears <- 0
-mythreshold   <- 0.2
+mythreshold   <- 0.1
 myvar         <- "stocksize"
@@
 
-# plot scale jumps between assessments
+# plot PCM by year with meta-data included
 t <-
@@
 
-  geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
-            aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
+  # geom_point(data=filter(d.assessments, assessmentmodelchange == 1),
+  #           aes(x=assessmentyear-0.1, y=-1.0), colour="red", fill="red", shape=24, size=2) +
 
@@
 
-*`r fig_nums("scalejumps")`*
-
-```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}
-
-tab_nums(name    = "scalejumps2", level = 1, display = FALSE, 
-         caption = "Example (mackerel) of sensivity of scale-jumps to the number of years included and the threshold for defining a scale-jump between two assessments.")
-
-# plot scale jumps with different settings for number of years and thresholds
-
-myremoveyears <- 0
-myvar         <- "stocksize"
-myspecies     <- "mac"
-mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
-
-res$d.pcm %>%
-
-  filter(grepl(myspecies, stock)) %>% 
-  filter(variable        == myvar) %>% 
-  filter(removelastyears == myremoveyears) %>%
-
-  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
-  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
-  gather(key=metric, value=value, c(phi)) %>%
-
-  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
-  theme_publication() +
-  theme(
-    panel.spacing    = unit(0.1, "lines"),
-    panel.grid.major.x = element_line(colour = "gray80"),
-    strip.background = element_blank(),
-    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
-    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
-    strip.placement = "outside",
-    legend.position  = "none"
-  ) +
-
-  geom_hline(aes(yintercept = threshold), colour="gray80", linetype="dashed", size=0.5) +
-  geom_hline(aes(yintercept = 0), colour="gray20", size=0.5) +
-  geom_hline(aes(yintercept = -threshold), colour="gray80", linetype="dashed", size=0.5) +
-
-  geom_line (aes(colour=factor(decade))) +
-  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
-
-  # coord_cartesian(xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-  coord_cartesian(ylim=c(-1, 1),
-                  xlim=c(1980, 2018)) +  # set the limits in cartesian space prevents outlier lines to be excluded
-
-  # ggMarginal(data=d.assessments, type="histogram")
-  scale_fill_manual(values = myBlues) +
-
-  scale_y_continuous(breaks = pretty_breaks(), labels=percent_format()) +
-  scale_x_continuous(breaks = pretty_breaks()) +
-
-  labs(x=" ", y="", title=mytitle) +
-  facet_grid(threshold ~ numberofyears)
-
-# filter(d.assessments, benchmark == 1) %>% View()
-
-```
-
-*`r fig_nums("metricsinpeels")`*
+*Figure 5: Scale-jumps for nine species derived from historical assessment records.
+Results from pairwise comparisons over periods of 5 years and a threshold of 10%.
+Large dots refer to years when the scale difference is larger than the threshold.
+Colours indicate the decades.
+Black triangles indicate benchmark assessment years.
+The bottom part of the panels shows the assessment methods being deployed, the upper part of the panels the stock definitions.*
 
@@
 
-```{r echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}
+```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}
 
@@
 
-*`r fig_nums("stockdefchange")`*
-
-<!-- ################################################################################## -->
-<!-- COmpare the number of scale jumps (as ratio) over time                             -->
-<!-- ################################################################################## -->
-
-```{r, eval=FALSE, fig.align="center", fig.asp=0.5, message=FALSE, warning=FALSE, include=FALSE}
-
-# NOT SHOWN, NOT USED
-
-d.jumps.byyear %>% 
-  filter(assessmentyear >= 1982) %>%
-
-  ggplot(aes(x=assessmentyear, y=ratio)) +
+*Figure 6: Comparisons of stock size (SSB) in assessments where stock definitions have changed. Blue-green is old stock definition, red is new stock definition*
+
+```{r, eval=TRUE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}
+
+# plot number of scale jumps (relative) over time
+
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+myvar         <- "stocksize"
+mytitle       <- paste0(myvar, ", removed years: ", myremoveyears)
+
+t <-
+  res$d.pcm %>% 
+  filter(variable        == myvar,
+         numberofyears   == mynyears,
+         removelastyears == myremoveyears,
+         threshold       == mythreshold) %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, 
+           removelastyears, numberofyears, threshold) %>%
+  filter(!is.na(scale_jump)) %>% 
+  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
+  group_by(variable, decade, assessmentyear, scale_jump) %>% 
+  summarize(n = n()) %>% 
+  group_by(variable, decade, assessmentyear) %>% 
+  mutate(prop = n / sum(n), 
+         nassess = sum(n))
+
+# scaling of the two axes
+ylim.prim <- c(0, 1)     # proportions
+ylim.sec  <- c(0, 9)     # number of assessments
+
+# calculations for two axes;
+# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
+b <- diff(ylim.prim)/diff(ylim.sec)
+a <- b*(ylim.prim[1] - ylim.sec[1])
+
+t %>%   
+  filter(scale_jump == TRUE) %>%
+
+  ggplot(aes(x=assessmentyear, y=prop)) +
   theme_publication() +
@@
     strip.background = element_blank(),
-    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
     strip.placement = "outside",
-    legend.title    = element_blank(),
     legend.position  = "none"
   ) +
+
   geom_bar(stat="identity") +
-  labs(x="", y="ratio")
+  geom_line(aes(y = a + nassess * b), color = "red", size=1) +
+  scale_y_continuous("Proportion scale jumps", 
+                     sec.axis = sec_axis(~ (. - a)/b, 
+                                         name = "N Assessments", 
+                                         breaks=pretty_breaks(n=4))) +
+  labs(x="")
 
@@
 
-*`r fig_nums("jumpratio")`*
+*Figure 7: Number of scale jumps in stocksize for the nine stocks considered, using 5 year means and a threshold of 10%.
+The red line indicates the number of assessments included in the comparison (right hand y-axis)*
+
+```{r echo=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE}
+
+# plot number of scale jumps by stock and decade
+
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+myvar         <- "stocksize"
+mytitle       <- paste0(myvar, ", removed years: ", myremoveyears)
+
+t <-
+  res$d.pcm %>% 
+  filter(variable        == myvar,
+         numberofyears   == mynyears,
+         removelastyears == myremoveyears,
+         threshold       == mythreshold) %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, 
+           removelastyears, numberofyears, threshold) %>%
+  filter(!is.na(scale_jump)) %>% 
+  mutate(decade = (10*floor(assessmentyear/10)) ) %>% 
+  group_by(stock, variable, decade, scale_jump) %>% 
+  summarize(n = n()) %>% 
+  group_by(stock, variable, decade) %>% 
+  mutate(prop = n / sum(n), 
+         nassess = sum(n))
+
+t2 <-
+  t %>% 
+  ungroup() %>% 
+  distinct(stock, variable, decade, .keep_all = TRUE) %>% 
+  dplyr::select(-scale_jump, -n, -prop)
+
+# scaling of the two axes
+ylim.prim <- c(0, 1)     # proportions
+ylim.sec  <- c(0, 10)     # number of assessments
+
+# calculations for two axes;
+# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
+b <- diff(ylim.prim)/diff(ylim.sec)
+a <- b*(ylim.prim[1] - ylim.sec[1])
+
+t %>%   
+  filter(scale_jump == TRUE) %>%
+
+  ggplot(aes(x=decade, y=prop)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_bar(stat="identity") +
+  geom_line(data=t2, aes(y = a + nassess * b), color = "red", size=1) +
+  scale_y_continuous("Proportion scale jumps", 
+                     sec.axis = sec_axis(~ (. - a)/b, 
+                                         name = "N Assessments", 
+                                         breaks=pretty_breaks(n=4))) +
+  labs(x="") +
+  facet_wrap(~stock)
+
+```
+
+*Figure 8: Proportion of scale jumps in stocksize by stock and decade, using 5 year means and a threshold of 10%.
+The red line indicates the number of assessments included in the comparison (right hand y-axis)*
+
+```{r echo=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE}
+
+# plot number of scale jumps by stock and decade
+
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+myvar         <- "stocksize"
+mytitle       <- paste0(myvar, ", removed years: ", myremoveyears)
+
+t <-
+  res$d.pcm %>% 
+  filter(variable        == myvar,
+         numberofyears   == mynyears,
+         removelastyears == myremoveyears,
+         threshold       == mythreshold) %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, 
+           removelastyears, numberofyears, threshold) %>%
+  filter(!is.na(scale_jump)) %>%
+
+  left_join(d.assessments, by=c("stock", "assessmentyear")) %>% 
+  filter(!is.na(assessmentmodel)) %>%
+
+  mutate(decade = (10*floor(assessmentyear/10)) ) %>% 
+  group_by(assessmentmodel, variable, scale_jump) %>% 
+  summarize(n = n()) %>% 
+  spread(key=scale_jump, value=n) %>% 
+  setNames(., gsub("TRUE","jump", names(.))) %>% 
+  setNames(., gsub("FALSE","nojump", names(.))) %>% 
+  mutate(jump    = replace_na(jump, 0),
+         nojump  = replace_na(nojump, 0),
+         nassess = jump + nojump, 
+         prop    = jump / nassess) %>% 
+  arrange(-prop) %>% 
+  ungroup() %>% 
+  mutate(rank = dense_rank(desc(prop))) %>% 
+  filter(nassess > 3)
+
+# scaling of the two axes
+ylim.prim <- c(0, 100)   # number of assessments  
+ylim.sec  <- c(0, 1)     # proportions
+
+# calculations for two axes;
+# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
+b <- diff(ylim.prim)/diff(ylim.sec)
+a <- b*(ylim.prim[1] - ylim.sec[1])
+
+t %>%   
+  # filter(scale_jump == TRUE) %>%
+
+  ggplot(aes(x=assessmentmodel, y=nassess)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_bar(aes(x = reorder(assessmentmodel, -rank), 
+                y = a + prop * b ), stat="identity") +
+  geom_line(aes(y = nassess, group=1), color = "red", size=0.8) +
+  geom_point(aes(y = nassess), color = "red", size=1.2) +
+  scale_y_continuous("N Assessments", 
+                     sec.axis = sec_axis(~ (. - a)/b, 
+                                         name = "Proportion scale jumps", 
+                                         breaks=pretty_breaks(n=4))) +
+  labs(x="") +
+  coord_flip()
+
+```
+
+*Figure 9: Proportion of scale jumps in stocksize by method, using 5 year means and a threshold of 10% (bars, top axis.
+The red line indicates the number of assessments included in the comparison (bottom axis)*
 

### 2019-08-15 R/iAssess figure historic retro by stock.r

@@ -72,7 +72,7 @@ d <-
   filter(year             >=   1980) %>% 
   # filter(year             >  2000) %>%  
   # filter(year             <= assessmentyear) %>% 
-  filter(assessmentyear   >  2008) %>% 
+  filter(assessmentyear   >  2000) %>% 
   # filter(assessmentyear   >  2014) %>% 
   
   # filter(purpose %in% c("advice", "update")) %>% 
@@ -194,6 +194,7 @@ p1 <-
   facet_grid(stockkeylabelold ~ ., scales="free_y")
   # facet_grid(stockkeylabelold ~ purpose, scales="free_y")
 
+p1
 
 # plot f
 p2 <-

### 2019-08-15 R/iDatabases generator.r

@@ -39,17 +39,7 @@
 # 23/03/2019 adapated after change to download data procedure
 # 06/05/2019 dplyr update. Changed "funs" to "list" in mutate_at functions 
 # 15/05/2019 removed double assessments from SAG
-
-# Warning messages:
-#   1: funs() is soft deprecated as of dplyr 0.8.0
-# please use list() instead
-# 
-# # Before:
-# funs(name = f(.)
-#      
-#      # After: 
-#      list(name = ~f(.))
-
+# 15/08/2019 removed doubles arising from iAdvice and SAG coupling on purpose
 # -----------------------------------------------------------------------------------------------
 
 # rm(list=ls())

### 2019-08-15 R/pdf analysis of advice.r

@@ -11,10 +11,10 @@ for (y in 1981:2018) {
   assign(paste("results",y,sep=""), {
     
     # print(y)
-    print(file.path("F:/ices/advice",y))
+    print(file.path("E:/ices/advice",y))
     
-    keyword_directory(file.path("F:/ices/advice",y),
-                      keyword = c('robust'),
+    keyword_directory(file.path("E:/ices/advice",y),
+                      keyword = c('rebuilding plan'),
                       file_pattern   = ".pdf",
                       ignore_case    = TRUE, recursive = TRUE, 
                       surround_lines = 1, full_names     = TRUE) %>% 

### 2019-09-05 R/ICES advice and assessment summaries WGWIDE 2019.Rmd
Updates after WGWIDE 2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 03/09/2019 updated after WGWIDE
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

fay <- 2011
fy  <- 2000
lay <- 2019
ly  <- 2019

stock="hom-nsea"
assessyear=lay
assessyear=2017

firstyear=fy
lastyear=ly
plot = c("stocksize", "fishingpressure", "recruitment")
my.purpose="advice"

# Function to plot single assessment
plot_assess <- function(stock,assessyear, firstyear, lastyear, 
                        my.purpose = "advice", plot=c("stocksize","fishingpressure")) {

  d <-
    iAssess %>% 
    ungroup() %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    filter(year             %in% firstyear:lastyear) %>% 
    filter(purpose %in% my.purpose) %>%

    #set type
    mutate(
      type = ifelse(purpose == "advice"      , "last"    , NA),
      type = ifelse(grepl("bench",purpose)   ,  "benchmark", type),
      type = ifelse(grepl("replaced",purpose), "replaced"  , type),
      type = ifelse(is.na(type), "other", type)) %>%

    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, type, 
                  assessmentyear, year,
                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
                  stocksize, lowstocksize, highstocksize, stocksizeunits,
                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
                  landings, discards, catches, catcheslandingsunits) %>%

    # convert to long data
    gather(key=variable, value=data, c(recruitment, lowrecruitment, highrecruitment, 
                  stocksize, lowstocksize, highstocksize,
                  fishingpressure, lowfishingpressure, highfishingpressure, 
                  landings, discards, catches)) %>%

    # set appropriate units and remove old units columns
    mutate(
      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%

    # separate estimate, low and high
    mutate(
      esttype = ifelse(grepl("low", variable), "low", NA),
      esttype = ifelse(grepl("high", variable), "high", esttype),
      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
    mutate(
      variable = gsub("low|high","", variable) ) %>%

    # select variables to plot
    filter(variable %in% plot) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # spread to make separate columns for est, low and high
    spread(key=esttype, value=data)

  # Reference points
  rp <-
    iAdvice %>% 
    ungroup() %>% 
    filter(purpose %in% my.purpose) %>% 
    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
    gather(key=refpoint, value=data, blim:fmsy) %>%

    # generate variables
    mutate(
      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # generate refpoint types
    mutate(
      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
    dplyr::select(-refpoint)

  # Titles
  tt <-
    d %>% 
    distinct(variable, unit, stockkeylabel, stockdescription) %>%
    filter(!is.na(unit))

  # d %>% View()

  # plot 
  d %>%

    ggplot(aes(x=year,y=est)) +

    theme_publication() +

    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10, hjust=0), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) +

    geom_line(aes(colour = type, size=type, linetype=type) ) +

    # geom_ribbon(aes(ymin=low, ymax=high), alpha=0.2, fill="red") +
    # geom_line(colour="red" ) +

    # plot limits
    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
               colour="red", linetype="solid", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    # plot msy
    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (
      values=c(last ="red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_fill_manual    (
      values=c(last = "red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_linetype_manual(
      values=c(last="solid",last_bench="solid",advice="solid",
               benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (
      values=c(last= 1.2,last_bench=1.2,advice=0.8,
               benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL, 
         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
    facet_wrap(~variable, scales="free_y")

  } # End of plot_assess function

# s   <- "mac-nea"
# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
#             plot = c("stocksize", "fishingpressure", "recruitment"), 
#             my.purpose="advice")
# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
#             plot = c("stocksize", "fishingpressure", "recruitment"), 
#             my.purpose="bench")

stock="whb-comb"
firstassessyear=fay; lastassessyear=lay; firstyear=fy;lastyear=ly;
plot = c("stocksize", "fishingpressure")
include.benchmark=TRUE; include.replaced=TRUE
plot.uncertainty=FALSE

# Function to plot historical assessments
plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, 
                         include.benchmark = TRUE, include.replaced=TRUE, 
                         plot=c("stocksize","fishingpressure"),
                         plot.uncertainty=FALSE) {

  d <-
    iAssess %>% 
    ungroup() %>%

    {if (include.benchmark == TRUE & include.replaced == TRUE) {
      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench", "replaced"))
    } else if (include.benchmark == TRUE & include.replaced == FALSE) {
      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
    } else if (include.benchmark == FALSE & include.replaced == TRUE) {
      filter(., purpose %in% c("advice","replaced"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   %in% c(firstassessyear:lastassessyear)) %>% 
    filter(year             %in% firstyear:lastyear) %>%

    group_by(assessmentyear) %>%

    #set type
    mutate(
      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"      , "last"      , NA),
      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
                    n_distinct(purpose)==1, "last_bench", type),
      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
                      n_distinct(purpose) > 1, "bench", type),
      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"      , "advice"    , type),
      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose)   ,  "benchmark", type),
      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose), "replaced"  , type),
      type = ifelse(is.na(type), "other", type)) %>%

    ungroup() %>%

    # set tyear and group
    mutate(tyear = substr(as.character(assessmentyear), 3,4)) %>% 
    mutate(group = paste0(tyear, type)) %>% 
    mutate(group = substr(group, 1,3)) %>% 
    mutate(group = gsub("a|l","", group)) %>%

    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, 
                  assessmentyear, year, tyear, group, type, purpose,
                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
                  stocksize, lowstocksize, highstocksize, stocksizeunits,
                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
                  landings, discards, catches, catcheslandingsunits) %>%

    # convert to long data
    gather(key=variable, value=data, 
           c(recruitment, lowrecruitment, highrecruitment, 
             stocksize, lowstocksize, highstocksize,
             fishingpressure, lowfishingpressure, highfishingpressure, 
             landings, discards, catches)) %>%

    # set appropriate units and remove old units columns
    mutate(
      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%

    # separate estimate, low and high
    mutate(
      esttype = ifelse(grepl("low", variable), "low", NA),
      esttype = ifelse(grepl("high", variable), "high", esttype),
      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
    mutate(
      variable = gsub("low|high","", variable) ) %>%

    # select variables to plot
    filter(variable %in% plot) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # spread to make separate columns for est, low and high
    spread(key=esttype, value=data) %>%

    # remove NA
    filter(!is.na(est))

  # Check
  # d %>% filter(assessmentyear == 2019) %>% View()

  # Reference points
  rp <-
    iAdvice %>% 
    ungroup() %>% 
    filter(purpose %in% my.purpose) %>% 
    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
    gather(key=refpoint, value=data, blim:fmsy) %>%

    # generate variables
    mutate(
      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # generate refpoint types
    mutate(
      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
    dplyr::select(-refpoint)

  # Titles
  tt <-
    d %>% 
    distinct(variable, unit, stockkeylabel, stockdescription) %>%
    filter(!is.na(unit))

  # plot
  d %>%

    ggplot(aes(year,est, group=group)) +

    theme_publication() +

    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10, hjust=0), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    {if (plot.uncertainty)     
      geom_ribbon(data=filter(d, type=="last"),
          aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) } +

    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = group, colour = type, group=group), 
            method = list(dl.combine("last.points"), cex = 0.8, inherit.aes=FALSE)) +

    # plot limits
    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
               colour="red", linetype="solid", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    # plot msy
    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (
      values=c(last ="red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_fill_manual    (
      values=c(last = "red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_linetype_manual(
      values=c(last="solid",last_bench="solid",advice="solid",
               benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (
      values=c(last= 1.2,last_bench=1.2,advice=0.8,
               benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL, 
         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
    facet_wrap(~variable, scales="free_y")

  # return(d)

  } # End of plot_history function

# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE

# Function to plot advice overviews
table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {

  d <-
    iAdvice %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%

    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%

    {if (include.replaced == TRUE) {
      filter(., purpose %in% c("advice","replaced"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    mutate(advicebasis= substr(advicebasis,1,25),

           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),

           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"),

           catch      = ifelse(!is.na(catches), catches, NA),
           catchtype  = ifelse(!is.na(catches), "catches", ""),

           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),

           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),

           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           cat_c       = 100 * (catch / lag(catch, n=1) -1)
           ) %>%

    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
           adv_c, tac_c, cat_c,
           advicetype, tactype, catchtype)

  d %>% 
  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
  arrange(tacyear, desc(purpose)) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 200, 
               justify      = "right",
               missing      =" ",
               big.mark     = '', 
               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)

```

**Summary of ICES 2019 WGWIDE assessments and advice**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

In this document an overview is presented of the ICES assessment and advice that resulted from the ICES Working Group on Widely Distributed Stocks (WGWIDE 2019) which deals with mackerel, blue whiting, horse mackerel and Atlanto-scandian herring. The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

**Northeast Atlantic Mackerel**

The ICES advice on Northeast Atlantic mackerel that was released in October 2018 has been received with a lot of scepticism and unbelief. According to that advice, the stock had been declining since 2011 and was estimated to be just below the MSY Btrigger. At the time, it was already found that the model results were very sensitive to the inclusion of the RFID tags. Leaving out only the RFID tags, gave a much higher stock estimate. To deal with this assessment problem, an inter-benchmark was set up which is lead to a new assessment configuration in March 2019 and an updated advice for 2019.

The new assessment that was carried out in WGWIDE 2019 has not created a different perception on stock trends of mackerel. The preliminary results of the triennial mackerel egg survey indicated a low SSB for mackerel. However, the other information sources indicate an increase in the recruitment and biomass. The egg survey carried a rather limited weight in the final model and therefore the assessment estimates a high stock in 2020 and fishing mortality only just above Fmsy.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "mac-nea"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 1a Northeast Atlantic mackerel. Summary of most recent assessment_

#####page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, 
             plot = c("stocksize", "fishingpressure")) 
```

_Figure 1b Northeast Atlantic mackerel. History of assessments. Red is most recent assessment, black is earlier assessments, blue is benchmarks and green is replaced assessments. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/mac.27.nea.pdf)

```

_Table 1 Northeast Atlantic mackerel. History of advice, tac and catch._

#####page break

*Blue whiting*

The blue whiting survey in 2019 went well and showed the highest biomass of the time series. Despite the high catches, the stock is still at a very high level. The stock size of blue whiting is estimated to be around 5 million tonnes. The stock size of blue whiting is estimated higher than last year, due to a high survey result for the strong 2014 year class. However, the stock is currently decreasing due to low recruitments coming in. Catches have been well above 1 Million tonnes over the last few years and above the scientific advice. Fishing mortality is estimated to be above Fmsy.

The advice for 2020 is 1.16 million tonnes, which is at the same levels as the advice for 2019. A decline in catching opportunities in the future years can be expected.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 2a Blue whiting. Summary of most recent assessment_

#####page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 2b Blue whiting. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/whb.27.1-91214.pdf)

```

_Table 2 Blue whiting. History of advice, tac and catch._

##### Page break

*Western horse mackerel*

The major discussion about Western horse mackerel was focussed on the reference points. An Interbenchmark protocol has been followed to re-estimate the reference points that had been set during the previous benchmark in 2017. At that time the Blim had been set at Bloss which was the last point of the time-series. Due to the retrospective revisions of the assessments since the benchmark, it became clear that that choice of Blim was inappropriate. The new reference points that have been proposed revert back to the earlier method of setting reference points for this stock: the Bpa / MSY Btrigger was selected as the lowest estimated stocks size in the previous low period of the stock (2003) during which no impaired recruitment was observed. The Blim and F reference points where then derived from that. The following reference points have been proposed: Blim = 834480 t (was 661917 t), Bpa = 1168272 t (was 911587 t), Fmsy = 0.074 (was 0.108).

The stock assessment of Western horse mackerel was reasonably comparable to last year's assessment. The retrospective revision is less than during the previous years. The stock is estimated just above the new Blim and fishing mortality is close to Fmsy. The recent recruitments appear to have been somewhat higher which leads to an modest increase in stock size.

An alternative assessment approach for Western horse mackerel has been presented in WGWIDE. This is using the SAM assessment model and largely the same data as in the standard assessment model (Stock Synthesis). The trends in the stock and fishing mortality are largely comparable although the absolute levels and uncertainties are different (higher F, lower stock and higher uncertainty in the SAM assessment).

The advice is based on the new reference points. Because the stock is between MSY Btrigger and Blim, a lower fishing mortality than Fmsy is used as the basis of the advice, which results in a substantially lower advice compared to last year. Given the indications we get from the fisheries, this could be a more realistic advice than the advice that was presented last year.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 3a Western horse mackerel. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 3b Western horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hom.27.2a4a5b6a7a-ce-k8.pdf)

```

_Table 3 Western horse mackerel. History of advice, tac and catch._

##### page break

*North Sea horse mackerel*

The assessment and advice of North Sea is carried out every second year. A new stock assessor was responsible for the assessment (as the previous stock assessor changed jobs), and she found a number of errors in the codes that had been used in 2017. These errors were now corrected and lead to a much lower estimate of stock size in 2016.

This is a so-called category 3 advice, which means that it only uses the relative trend in a (combined) survey to derive the advice. The trend is based on the abundance of the North Sea IBTS survey (quarter 3) and French Channel CGFS survey (quarter 4). The stock is considered to be at a low level even though recruitment appears to have been higher in recent year. A length-based proxy indicator suggests that fishing mortality is probably (just) above Fmsy.

The advice is lower than the advice presented in 2017. Note that the advice is a catch advice. The advice for the landed part is 13305 t.The advice is applicable for 2020 and 2021.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 4a North Sea horse mackerel. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 4b North Sea horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Table 4 North Sea horse mackerel. History of advice, tac and catch._

##### page break

*Atlanto-scandian herring*

The stock assessment is consistent with last year's assessment. The stock appears to be slowly declining since 2009. Fishing mortality is estimated to be below Fmsy. There are signs of a reasonable sized incoming year class 2016, but this year class is not showing up in the assessment yet, as the year class is still only seen in the Barents Sea. The year class would be expected to recruit to the fishery in 2020.

The advice is lower than last year due to the lower stock size.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "her-noss"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 5a Atlanto-scandian herring. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 5b Atlanto-scandian herring. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.1-24a514a.pdf)

```

_Table 5 Atlanto-scandian herring. History of advice, tac and catch._

*More information*

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.

### 2019-09-05 R/DownloadDataFromSAG.r

@@ -42,7 +42,7 @@ today <- format(Sys.time(), '%Y%m%d')
 # which year (set year = 0 for all years)
 # =====================================================================================
 
-myyear <- 2018
+myyear <- 2019
 
 # =====================================================================================
 # generate list of assessmentkeys

### 2019-09-05 R/iAssess figure historic retro by stock.r

@@ -61,9 +61,11 @@ d <-
   # filter(grepl("her-67bc", stockkeylabelold) ) %>% 
   # filter(grepl("her-4", stockkeylabelold) ) %>%
   # filter(grepl("her-nirs", stockkeylabelold) ) %>% 
+  # filter(stockkeylabelold == "hom-west") %>%
   # filter(grepl("hom-west", stockkeylabelold) ) %>%
-  filter(grepl("whb-comb", stockkeylabelold) ) %>%
-  
+  # filter(grepl("whb-comb", stockkeylabelold) ) %>%
+  filter(grepl("mac-nea", stockkeylabelold) ) %>%
+
   # filter(grepl("mac-nea|hom-west|whb-comb|her-noss", stockkeylabelold) ) %>% 
   # filter(grepl("whb", stockkeylabelold) ) %>% 
   # filter(grepl("noss", stockkeylabelold) ) %>% 
@@ -82,7 +84,8 @@ d <-
          landings, catches, 
          recruitment, lowrecruitment, highrecruitment, 
          fishingpressure, lowfishingpressure, highfishingpressure, 
-         stocksize, lowstocksize, highstocksize) %>%
+         stocksize, lowstocksize, highstocksize,
+         stocksizeunits, fishingpressureunits, catcheslandingsunits, unitofrecruitment) %>%
   
   mutate(purpose = ifelse(assessmentyear == max(assessmentyear) &
                                   purpose %in% c("advice","update"),
@@ -104,6 +107,7 @@ d <-
 
   data.frame()
 
+# iAssess %>% filter(grepl("hom-west", stockkeylabelold), assessmentyear == 2003) %>% View()
 
 # d %>% 
 #   filter(assessmentyear >= 2018,
@@ -122,7 +126,8 @@ last <-
   select(stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew, assessmentyear, year, 
          laststocksize = stocksize, lowstocksize, highstocksize, 
          lastfishingpressure=fishingpressure, lowfishingpressure, highfishingpressure, 
-         lastrecruitment = recruitment, lowrecruitment, highrecruitment, purpose)
+         lastrecruitment = recruitment, lowrecruitment, highrecruitment, purpose,
+         stocksizeunits, fishingpressureunits, unitofrecruitment, catcheslandingsunits)
 
 withdrawn <-
   d %>% 
@@ -131,9 +136,13 @@ withdrawn <-
[…]

### 2019-09-05 R/iDatabases generator.r

@@ -40,6 +40,8 @@
 # 06/05/2019 dplyr update. Changed "funs" to "list" in mutate_at functions 
 # 15/05/2019 removed double assessments from SAG
 # 15/08/2019 removed doubles arising from iAdvice and SAG coupling on purpose
+# 04/09/2019 added manual allocation of hom-nsea 2019 assessment
+# 04/09/2019 fixed strange error: mutate with multiple ifelse gave logical; now each line a separate mutate with ifelse works well
 # -----------------------------------------------------------------------------------------------
 
 # rm(list=ls())
@@ -215,7 +217,7 @@ sd <-
 # -----------------------------------------------------------------------------------------
 # load SAG full data (see: DownloadDataFromSAG.r)
 # -----------------------------------------------------------------------------------------
-# sag %>% filter(assessmentyear == 2019, stockkeylabel=="her.27.6a7bc") %>% View()
+# sag %>% filter(assessmentyear == 2019, stockkeylabelold=="hom-nsea") %>% View()
 
 sag <- 
   get(load(file=paste(dropboxdir, "/rdata/iSAGstock.RData",sep=""))) %>% 
@@ -224,7 +226,13 @@ sag <-
   # left_join(iRename[,c("stockkeylabel","stockkey")], by="stockkeylabel") %>%
   # left_join(iStockkey, by="stockkey") %>% 
   
-  # remove custom fields for now
+  # sag %>% filter(stockkeylabelold == "hom-nsea" & assessmentyear == 2019) %>% View()
+  
+  # !!!! manual allocation of hom-nsea assessment (from custom to stocksize)
+  # mutate() %>% 
+  #   stocksize = ifelse(stockkeylabelold == "hom-nsea" & assessmentyear == 2019, )
+  
+# remove custom fields for now
   dplyr::select(-starts_with("custom")) %>% 
   
   # remove assessmentkey (prevents appropriate merging)
@@ -296,8 +304,10 @@ only_in_sag <-
   anti_join(sag_unique, qcsexcel_unique)
 
 # incommon %>% filter(stockkey == 169049) %>% View()
+# incommon %>% filter(stockkeylabelold == "hom-nsea) %>% View()
 # only_in_qcsexcel %>% filter(stockkey == 169049) %>% View()
 # only_in_sag %>% filter(stockkey == 169049) %>% View()
+# only_in_sag %>% filter(stockkeylabelold == "hom-nsea") %>% View()
 
[…]

### 2019-09-05 R/plot stock and TAC area.r

@@ -119,7 +119,7 @@ stocks_toplot <-
   iAdvice %>% 
   
   filter(grepl("her", stockkeylabel),
-         !stockkeylabel %in% c("her.27.1-24a514a","her.27.5a"),
+         # !stockkeylabel %in% c("her.27.1-24a514a","her.27.5a"),
          assessmentyear == 2018) %>%
   # filter(stockkeylabelold %in% c("her-67bc", "her-vian","her-irlw"),
   #        assessmentyear == 2018) %>%
@@ -169,10 +169,13 @@ stocks_toplot %>%
   # geom_sf(aes(fill = factor(fishstock)), alpha=0.6, colour="black") +
   
   # coord_sf(xlim = c(-20,25), ylim = c(38,70)) +
-  geom_sf_label(aes(label = stockkeylabel), alpha=0.9) +
+  # geom_sf_label(aes(label = stockkeylabel), alpha=0.9) +
   ggmisc::scale_fill_crayola() +
   facet_wrap(~stockkeylabel)
 # facet_grid(variable~assessmentyear)
+
+# save to shape file
+st_write(stocks_toplot, dsn = "herring_stocks.shp", layer = "herring.shp", driver = "ESRI Shapefile")
 
 
 # ----------------------------------------------------------------------------------------------------------

### 2022-01-20 R/Analyse zero catch advice.Rmd
UPdates several files

--- 2019-06-10 R/Analyse changes in pelagic stocks.Rmd
@@
 # -----------------------------------------------------------------------------------------------
-# Analyse changes in pelagic stocks.Rmd
+# Analyse zero catch advice.Rmd
 #
-# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
-# 10/01/2018 made small changes to the table
-# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
-# 06/06/2019 changed the data frame d; now also has purpose
+# 26/02/2020 adapted from Analyse changes in pelagic stocks
 # -----------------------------------------------------------------------------------------------
@@
 
-  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss",
-               stockkeylabelold)) %>%
+  # filter(grepl("mac-67|mac-nea|whb-c|whb-n|her-noss|her-47d3",
+  #              stockkeylabelold)) %>%
+  # filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss",
+  #              stockkeylabelold)) %>%
   # filter(grepl("mac-67|mac-nea", 
@@
 
-  filter(assessmentyear >= 2008) %>% 
+  filter(assessmentyear >= 2000) %>% 
   ungroup()
@@
   filter(assessmentyear %in% unique(adv$assessmentyear)) %>% 
-  filter(year <  assessmentyear) %>% 
+  filter(year <=  assessmentyear) %>% 
   filter(year >= 2000) %>%
@@
 
-**Analysis of advice changes in a number of key pelagic stocks**
+**Analysis of zero catch advice**
 
@@
 The document outlines changes in reference points, changes in advice and a comparison between stock size assumed as the basis for the forecast and the most recent estimates of stock size from the 2018 assessments.
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+adv %>% 
+  filter(purpose == "advice") %>% 
+  filter(advisedlandingsmax == 0 | advisedcatchmax == 0) %>%  
+  filter(tacyear <= 2018) %>% 
+  filter(!is.na(tal)) %>% 
+  dplyr::select(speciesfaocode, stockkeylabelold, assessmentyear, tacyear, advisedlandingsmax, tal, landings, catches) %>%
+  ungroup() %>% 
+  mutate(catch     = ifelse(!is.na(catches), catches, landings)) %>% 
+  mutate(zerotac   = ifelse(tal   == 0, 1, 0)) %>% 
+  mutate(zerocatch = ifelse(catch == 0, 1, 0)) %>% 
+  # View()
+  group_by(tacyear) %>% 
+  summarize(nzeroadvice = n(),
+            nzerotac    = sum(zerotac, na.rm=TRUE),
+            nzerocatch  = sum(zerocatch, na.rm=TRUE)) %>% 
+  gather(key="variable", value="data", nzeroadvice:nzerocatch) %>%
+
+  ggplot(aes(x=tacyear)) +
+  theme_publication() +
+  theme(legend.title = element_blank()) +
+
+  geom_line(aes(y=data, group=variable, colour=factor(variable))) +
+  geom_point(aes(y=data, group=variable, colour=factor(variable))) +
+
+  scale_x_continuous(breaks=pretty_breaks()) +
+  expand_limits(y=0) + 
+  labs(y="N stocks") 
+#  facet_wrap(~variable)
+
+```
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+adv %>% 
+  filter(purpose == "advice") %>% 
+  filter(advisedlandingsmax == 0 | advisedcatchmax == 0) %>%  
+  filter(tacyear <= 2018) %>% 
+  filter(!is.na(tal)) %>% 
+  dplyr::select(speciesfaocode, stockkeylabelold, stockkeylabelnew, assessmentyear, tacyear, advisedlandingsmax, tal, landings, catches) %>%
+  ungroup() %>% 
+  mutate(catch     = ifelse(!is.na(catches), catches, landings)) %>% 
+  mutate(zerotac   = ifelse(tal   == 0, 1, 0)) %>% 
+  mutate(zerocatch = ifelse(catch == 0, 1, 0)) %>%
+
+  write.csv(file="zerocatch.csv", row.names = FALSE)
+
+```
 
@@
 
-ass_5year %>%
+my.stock <- c("whb-comb", "her-noss", "her-47d3", "mac-nea")
+my.year  <- 2010:2019
+
+# ass %>% 
+ass_10year %>%
+
+  filter(stockkeylabelold %in%  my.stock) %>% 
+  filter(assessmentyear %in%  my.year) %>%
 
@@
   geom_line(aes(size=purpose)) +
-  geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
+  geom_line(data=subset(ass_final, 
+                        stockkeylabelold %in% my.stock & assessmentyear %in% my.year), 
+            aes(colour=purpose ), size=1.5) +
   # geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +
@@
   facet_wrap(~stockkeylabelold, scales="free_y")
+  # facet_wrap(~assessmentyear, ncol=4)
 

### 2022-01-20 R/Analyse reference points.r

@@ -14,7 +14,7 @@ library(cowplot)       # multiplots
 library(directlabels)  # for printing labels at end of geom lines
 
 # Load utils code
-source("../mptools/r/my_utils.r")
+source("../mptools/r/my utils.r")
 
 # Set working directory to dropbox folder
 advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
@@ -93,6 +93,9 @@ iAdvice %>%
 # -----------------------------------------------------------------------------------------
 
 mystock <- "hom-west"
+mystock <- "her-47d3"
+mystock <- "her-3a22"
+mystock <- "ple-nsea"
 
 iAdvice %>% 
   filter(stockkeylabelold == mystock) %>% 

### 2022-01-20 R/DownloadDataFromSAG.r

@@ -22,7 +22,7 @@ library(stringr)       # for string manipulation
 options(icesSAG.use_token = TRUE)
 
 # Load utils code
-source("../mptools/r/my_utils.r")
+source("../mptools/r/my utils.r")
 
 # Set dropbox folder
 dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
@@ -42,14 +42,13 @@ today <- format(Sys.time(), '%Y%m%d')
 # which year (set year = 0 for all years)
 # =====================================================================================
 
-myyear <- 2019
+myyear <- 2020
+mystocks <- c("hom.27.2a4a5b6a7a-ce-k8","mac.27.nea","whb.27.1-91214","her.27.1-24a514a")
 
-# =====================================================================================
-# generate list of assessmentkeys
-# =====================================================================================
+# Load assessment data
+assessmentkeys <- sort(findAssessmentKey(mystocks, year=myyear))
 
-assessmentkeys <- sort(findAssessmentKey(year=myyear))
-
+# assessmentkeys <- sort(findAssessmentKey(year=myyear))
 # assessmentkeys <- findAssessmentKey(year=1983, full=TRUE)
 # assessmentkeys <- sort(findAssessmentKey())
 # assessmentkeys[assessmentkeys == 10362]
@@ -82,7 +81,7 @@ for (i in 1:length(t1)) {
     
     tmp <-
       t1[[i]] %>% 
-      mutate_all(funs("as.character")) %>% 
+      mutate_all(list("as.character")) %>% 
       lowcase() 
     
     t2  <- bind_rows(t2, tmp)

### 2022-01-20 R/ICES advice and assessment summaries WGWIDE 2020.Rmd
Assessment summaries

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 03/09/2019 updated after WGWIDE
# 30/03/2020 updated after HAWG; UNFINISHED. JUST TEMPORARY SOLUTIONS. 
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)
library(ggrepel)

# Load utils code
source("../../mptools/r/my utils.r")
source("iAssess functions.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
iAssess <- get(load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))) %>%

  # fix for 67bc in 2020
  mutate(purpose = ifelse(stockkeylabelold=="her-67bc" & assessmentyear==2020, "advice", purpose)) %>%

  # fix for North Sea sprat (ignore the realignment)
  mutate(stockkeylabelold = ifelse(stockkeylabel == "spr.27.3a4","spr-nsea", stockkeylabelold))

iAdvice <- get(load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")))

fay <- 2017
fy  <- 2000
lay <- 2020
ly  <- 2020

# iAssess %>% filter(stockkeylabelold=="her-67bc") %>% ungroup() %>% distinct(assessmentyear, purpose)
# iAssess %>% filter(stockkeylabelold=="spr-nsea") %>% ungroup() %>% distinct(assessmentyear, purpose)
# iAssess %>% filter(substr(stockkeylabelold,1,3)=="spr") %>% ungroup() %>% distinct(stockkeylabelold, stockkeylabel, assessmentyear, purpose) %>% View()

```

**Summary of ICES 2020 WGWIDE assessments and advice**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**EMBARGO**

**THIS DOCUMENT MAY NOT BE SHARED BEFORE THE ICES ADVICE HAS BEEN RELEASED (30 September 2020)**

In this document an overview is presented of the ICES assessment and advice that resulted from the ICES Working Group on Widely Distributed Stocks (WGWIDE 2020) which deals with mackerel, blue whiting, horse mackerel and Atlanto-scandian herring. The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

**Northeast Atlantic Mackerel**

The new assessment that was carried out in WGWIDE 2020 is consistent with the previous assessment.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "mac-nea"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 1a Northeast Atlantic mackerel. Summary of most recent assessment_

In previous years, there have been numerous revisions to the assessment of mackerel, leading to benchmarks, interbenchmarks and replace assessments.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "mac-nea"

plot_history(stock=s,
             firstassessyear=fay,
             lastassessyear=lay, 
             firstyear=fy, 
             lastyear=ly, 
             include.benchmark = TRUE,
             include.replaced = TRUE,
             plot = c("stocksize", "fishingpressure"))

```

_Figure 1b Northeast Atlantic mackerel. History of assessments. Red is most recent assessment, black is earlier assessments, blue is benchmarks and green is replaced assessments. _

##### page break

An overview of advice (red), TAC (or sum of unilateral quota, in blue) and observed catch (green) for NEA mackerel is shown below, both in absolute quantities (tonnes) and in deviations from the (final) ICES advice. Note that the advice may have been changed when an assessment has been replaced after a benchmark was carried out. Deviations from the ICES advice in the order of 20-60% can be observed from 2008 onwards.

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Figure 1c Northeast Atlantic mackerel. Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch. Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice. _

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

# [mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/mac.27.nea.pdf)

```

_Table 1.a Northeast Atlantic mackerel. History of advice, tac and catch._

#####page break

*Blue whiting*

Despite the lack of the 2020 survey observation of the blue whiting survey, the assessment in 2020 appeared to perform well and was very consistent with the previous year. Despite the high catches, the stock is still at a very high level although it is coming down. Low recruitments are coming in for this stock, so further declines in stock size and catches may be expected for the next years.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 2a Blue whiting. Summary of most recent assessment_

#####page break

The consistency of the assessment compared to the previous years appears to be reasonable.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "whb-comb"

plot_history(stock=s,
             firstassessyear=fay,
             lastassessyear=lay, 
             firstyear=fy, 
             lastyear=ly, 
             include.benchmark = TRUE, 
             include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=TRUE)

```

_Figure 2b Blue whiting. History of assessments. Red is most recent assessment and black is earlier assessments. _

##### page break

An overview of advice (red), TAC (or sum of unilateral quota, in blue) and observed catch (green) for blue whiting is shown below, both in absolute quantities (tonnes) and in deviations from the (final) ICES advice. Note that the advice may have been changed when an assessment has been replaced after a benchmark was carried out. Deviations from the ICES advice in the order of 20-60% can be observed from 2014 onwards. Much higher deviations were observed in the beginning of the 2000s.

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Figure 1c Blue whiting. Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch. Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

# [whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/whb.27.1-91214.pdf)

```

_Table 2 Blue whiting. History of advice, tac and catch._

##### Page break

*Western horse mackerel*

The stock assessment of Western horse mackerel was reasonably with last year's assessment. The retrospective revision is less than during the previous years. The stock is again estimated just above the new Blim and fishing mortality is above Fmsy. The recent recruitments appear to have been somewhat higher which leads to an modest increase in stock size.

An alternative assessment approach for Western horse mackerel has also been presented in WGWIDE. This is using the SAM assessment model and largely the same data as in the standard assessment model (Stock Synthesis). The trends in the stock and fishing mortality are largely comparable although the absolute levels and uncertainties are different (higher F, lower stock and higher uncertainty in the SAM assessment).

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

plot_assess(stock=s,
            assessyear=lay,
            firstyear=fy,
            lastyear=ly,
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 3a Western horse mackerel. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "hom-west"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly,
             include.benchmark = TRUE, include.replaced = FALSE,
             plot = c("stocksize", "fishingpressure"),
             plot.uncertainty=FALSE)

```

_Figure 3b Western horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

##### page break

An overview of advice (red), TAC and observed catch (green) for western horse mackerel is shown below, both in absolute quantities (tonnes) and in deviations from the (final) ICES advice. Note that the advice may have been changed when an assessment has been replaced after a benchmark was carried out. Deviations from the ICES advice have been relatively small and not in a specific direction.

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "hom-west"

plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
# plot_advice(stock="hom-nsea",firstadviceyear=1995,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Figure 1c western horse mackerel. Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch. Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice. _

The advice is based on the new reference points agreed during 2019. Because the stock is between MSY Btrigger and Blim, a lower fishing mortality than Fmsy is used as the basis of the advice, which results in a somewhat lower advice compared to last year.

The rebuilding plan proposed by the PELAC was presented to WGWIDE, although the formal request to evaluate the rebuilding plan analysis has not been submitted to ICES yet. The preferred HCR option in the PELAC rebuilding plan is included in the forecast table and suggests a yield of around 50 000 tonnes.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

# [hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hom.27.2a4a5b6a7a-ce-k8.pdf)

```

_Table 3 Western horse mackerel. History of advice, tac and catch._

##### page break

*Atlanto-scandian herring*

The stock assessment is consistent with last year's assessment. The stock appears to be slowly declining since 2009. Fishing mortality is estimated to be below Fmsy. There are signs of a reasonable sized incoming year class 2016, but this year class is only just starting to show up in the assessment and the forecast.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "her-noss"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 4a Atlanto-scandian herring. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "her-noss"

plot_history(stock=s,
             firstassessyear=fay,
             lastassessyear=lay, 
             firstyear=fy, 
             lastyear=ly, 
             include.benchmark = TRUE, 
             include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=TRUE)

```

_Figure 4b Atlanto-scandian herring. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

##### page break

An overview of advice (red), TAC (or sum of unilateral quota, in blue) and observed catch (green) for Atlanto-scandian herring  is shown below, both in absolute quantities (tonnes) and in deviations from the (final) ICES advice. Note that the advice may have been changed when an assessment has been replaced after a benchmark was carried out. Deviations from the ICES advice in the order of 20-40% can be observed from 2013 onwards.

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# s   <- "her-noss"

plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Figure 4c Atlanto-scandian herring. Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch. Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice. _

The advice is higher than last year due to the incoming 2016 year class.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)

# [her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.1-24a514a.pdf)

```

_Table 4 Atlanto-scandian herring. History of advice, tac and catch._

*More information*

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.

### 2022-01-20 R/iDatabases generator.r
Database generator and updates

@@ -42,6 +42,7 @@
 # 15/08/2019 removed doubles arising from iAdvice and SAG coupling on purpose
 # 04/09/2019 added manual allocation of hom-nsea 2019 assessment
 # 04/09/2019 fixed strange error: mutate with multiple ifelse gave logical; now each line a separate mutate with ifelse works well
+# 09/06/2020 updated version of iadvice tables (more details on stf)
 # -----------------------------------------------------------------------------------------------
 
 # rm(list=ls())
@@ -55,7 +56,7 @@ library(cowplot)       # multiplots
 library(directlabels)  # for printing labels at end of geom lines
 
 # Load utils code
-source("../mptools/r/my_utils.r")
+source("../mptools/r/my utils.r")
 
 # Set working directory to dropbox folder
 dropboxdir  <- paste(get_dropbox(), "/iAdvice", sep="")
@@ -78,7 +79,8 @@ iAdvice <-
               "advisedlandingsmax", "advisedcatchmax",
               "tal", "tac",
               "officiallandings", "landings", "ibc", "discards", "catches",
-              "fsqay", "ssbay", "fadvmax",
+              "fsqimy", "ssbimy", "fadvmax",
+              "catchrealized","catch1fcy","catch2fcy","f1fcy","f2fcy","ssb1fcy","ssb2fcy","fset","ssbset",
               "fmax", "f01", "fmed", "f35spr", "flim", "fpa", "fmsy",
               "blim", "bpa", "msybtrigger",
               "m1", "m5"),
@@ -468,31 +470,37 @@ save(iAssess, file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
 # filter(iAssess, purpose=="historical") %>% View()
 
 # create overviews
-t1 <-
-  sag_unique %>% 
-  ungroup() %>% 
-  mutate(purpose = toupper(substr(purpose,1,1))) %>% 
-  group_by(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear) %>% 
-  arrange(stockkey, stockkeylabel, stockkeylabelnew, stockkeylabelold, assessmentyear, purpose) %>% 
-  summarize(purpose= paste(purpose, collapse="")) 
+# t1 <-
+#   sag_unique %>% 
[…]

### 2022-01-20 R/pdf analysis of advice.r

@@ -2,121 +2,119 @@
 
 # install.packages("devtools")
 # devtools::install_github('lebebr01/pdfsearch')
-
-library(pdfsearch)
+library(pdftools)  # generic pdf reader
+library(pdfsearch) # wrapper around pdftools for searching
 library(tidyverse)
 
-# loop over years
-for (y in 1981:2018) {
-  assign(paste("results",y,sep=""), {
-    
-    # print(y)
-    print(file.path("E:/ices/advice",y))
-    
-    keyword_directory(file.path("E:/ices/advice",y),
-                      keyword = c('rebuilding plan'),
-                      file_pattern   = ".pdf",
-                      ignore_case    = TRUE, recursive = TRUE, 
-                      surround_lines = 1, full_names     = TRUE) %>% 
-      # keyword_directory(file.path("E:/PA",y),
-      #                 keyword = c('precautionary approach', 
-      #                             'biological reference point',
-      #                             "minimum biologically acceptable", "mbal", 
-      #                             "safe biological limit",
-      #                             "maximum sustainable yield", "msy",
-      #                             "btrigger", "bpa", "blim", "fpa", "fmsy"),
-      #                 # file_pattern   = ".pdf"
-      #                 ignore_case    = TRUE, recursive = FALSE, surround_lines = 1, full_names     = TRUE) %>% 
-      mutate(year = y) 
-    
-  })
-}
+# Load utils code
+source("../mptools/r/my_utils.r")
+
+# end of text pages 
+# y <- 2010
+# f <- files[92]
+
+pages <- NULL
+for (y in 1997:2019) {
+# for (y in 2010:2010) {
+  files <- list.files(path = file.path("D:/temp/advice",y),
+                      pattern = ".pdf", 
+                      recursive = FALSE,
+                      full.names = TRUE)  
   
-for (y in 1981:2018) {
-  if (y==1981) {
-    results <- get(eval(paste("results",y,sep="")))
-  } else {
-    results <- bind_rows(results, get(eval(paste("results",y,sep=""))))
-  }  
-}
+  for(f in files) {
+    gc()
+    print(f)
[…]

### 2022-01-20 R/plot stock and TAC area.r

@@ -22,7 +22,7 @@ library(maps)
 library(ggplot2)
 library(viridis)
 
-source("../mptools/r/my_utils.r")
+source("../mptools/r/my utils.r")
 # load("../prf/rdata/world.df.RData")
 
 # source("r/iDatabases generator.r")
@@ -35,6 +35,21 @@ dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
 # Load dataset
 load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+
+# ----------------------------------------------------------------------------------------------------------
+# count changes in areas
+# ----------------------------------------------------------------------------------------------------------
+
+iAdvice %>% 
+  filter(assessmentyear >= 1980) %>% 
+  filter(speciesfaocode=="cod") %>% 
+  ungroup() %>% 
+  distinct(stockkey, stockkeylabelold, assessmentyear, speciesfaocode, speciescommonname, stockarea) %>% 
+  # group_by(assessmentyear, speciesfaocode, speciescommonname) %>% 
+  # summarize(n=n()) %>% 
+  View()
+
+
 
 # ----------------------------------------------------------------------------------------------------------
 # load spatial objects in sf format
@@ -118,17 +133,19 @@ fao.eez.sf %>%
 stocks_toplot <- 
   iAdvice %>% 
   
-  filter(grepl("her", stockkeylabel),
-         # !stockkeylabel %in% c("her.27.1-24a514a","her.27.5a"),
-         assessmentyear == 2018) %>%
-  # filter(stockkeylabelold %in% c("her-67bc", "her-vian","her-irlw"),
-  #        assessmentyear == 2018) %>%
-  distinct(stockkeylabelold, stockkeylabel, stockarea) %>% 
+  filter(grepl("her", stockkeylabel)) %>% 
+  filter(assessmentyear >= 1980) %>%
+  filter(stockkeylabelold %in% c("her-vian","her-irlw", "her-irls", "her-nirs","her-67bc", 
+                                 "her-47d3","her-47d", "her-4a", "her-4b","her-4c7d","her-4ab")) %>% 
+  # filter(stockkeylabelold %in% c("her-vian","her-irlw", "her-irls", "her-nirs")) %>% 
[…]

### 2022-01-20 R/Presentation for pelagic forum 2019.Rmd
Pelagic forum 2019

---
output:
  powerpoint_presentation:
    reference_doc: ../powerpoint_template.pptx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# ---
# output:
#   powerpoint_presentation:
#     reference_doc: ../powerpoint_template.pptx
# ---

# -----------------------------------------------------------------------------------------------
# Presentation for pelagic forum 2019.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 29/09/2019 adapted from stock summary code
# 03/10/2019 converted to powerpoint presentation
# -----------------------------------------------------------------------------------------------

if (packageVersion("rmarkdown") <= "1.9") {
  install.packages("rmarkdown")  # update rmarkdown from CRAN
}

rmarkdown::pandoc_version()

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

library(tidyverse)
library(cowplot)
library(directlabels)  # for printing labels at end of geom lines
library(scales)
library(stringr)
library(readxl)
library(gridExtra)                   # for plotting table as image (bio data)
library(RColorBrewer)

# Load utils code
source("../../mptools/r/my_utils.r")
source("iPlot_utils.r")

# Set dropbox folder
dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# load the data
iSpecies <-
  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
  ungroup() %>% 
  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild)

iAssess <-
  get(load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))) %>% 
  left_join(iSpecies, by="speciesfaocode")

load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))

# add intermediate and forecast data
forecasts <-
  read_excel(path=paste(dropboxdir,"/excel/Pelagic forecasts 2019.xlsx", sep=""), col_names = TRUE) %>%
  left_join(iStockkey, by="stockkeylabelold") %>%
  mutate(stockkeylabel = stockkeylabelnew) %>%
  left_join(iRename,   by=c("stockkey", "stockkeylabel")) %>%
  left_join(iSpecies,  by="speciesfaocode") %>%
  mutate(catcheslandingsunits = "tonnes")

# t <-
#   icesSAGfull %>%
#   filter(year < assessmentyear) %>% 
#   bind_rows(forecasts)

fay <- 2011
fy  <- 1990
lay <- 2019
ly  <- 2019

```

# In the morning

## Catches by fisheries guild

```{r, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

# catches by fisheriesguild

d <-
  iAssess %>% 
  ungroup() %>% 
  # filter(assessmentyear      == lay | (assessmentyear == 2017 & stockkeylabelold == "cap-icel") ) %>% 
  filter(assessmentyear      == lay  ) %>% 
  # filter(catcheslandingsunits == "tonnes") %>% 
  filter(year >= fy, year < ly) %>% 
  filter(purpose %in% c("advice")) %>%

  mutate(catches = ifelse(is.na(catches) | catches == 0, landings, catches))  %>% 
  mutate(fisheriesguild = ifelse(fisheriesguild %in% c("pelagic","demersal"), fisheriesguild, "other"))  %>% 
  mutate(fisheriesguild = factor(fisheriesguild, levels=c("pelagic","demersal","other"))) %>%

  group_by(fisheriesguild, speciesfaocode, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>%

  group_by(fisheriesguild, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>% 
  group_by(year) %>% 
  mutate(perc = round(100 * catches/ sum(catches), digits=2)) %>% 
  arrange(catches) %>% 
  mutate(fisheriesguild = factor(fisheriesguild, fisheriesguild))

# d %>% filter(year == 2018) %>% arrange(-catches) %>% View()

d2 <-
  d %>% 
  group_by(fisheriesguild) %>% 
  summarize(perc = mean(perc),
            catches=mean(catches),
            year = mean(year)) %>% 
  ungroup() %>% 
  mutate(cumcatches=cumsum(catches))

# plot stacked catches: pelagic, demersal and other
p1 <-
  d %>%

  ggplot(aes(year,catches)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=(fisheriesguild)), stat="identity", position="stack", colour="black") +
  # geom_text(data=subset(d, year==ly-1), aes(label=round(perc,0), group=fisheriesguild), 
  #           position = position_stack(vjust = 0.5), colour="white") +
  geom_text(data=subset(d, year==ly-1), 
            aes(label=fisheriesguild, group=fisheriesguild, x=ly), 
            position = position_stack(vjust = 0.5), colour="black", hjust=0) +
  # geom_dl(aes(label= perc),
  #         position = position_stack(vjust = 0.5),
  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
  #                         fontface = 'bold',cex=.9, hjust = 0)) +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="tonnes",title="catches") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly+3))

# d %>% filter(year == 2018) %>% View()

# plot of filled relative catches (how to add the direct labels?)
p2 <-
  d %>% 
  ggplot(aes(year,perc)) +
  theme_publication() +
  theme(legend.position="none") +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(plot.background = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  geom_bar(aes(fill=fisheriesguild), stat="identity", position="stack", colour="black") +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="",title="") +
  # scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly))

vp <- grid::viewport(width = 0.3, height = 0.3, 
                     x = 0.15, y = 0.12, 
                     just = c("left", "bottom"))

 print(p1)
     theme_set(theme_bw(base_size = 8))
     # theme_white()
     print(p2, vp = vp)
     theme_set(theme_bw())

  # cowplot::plot_grid(p1, p2,
  #                ncol=2, scale=c(1.0, 0.25), align="hv",
  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))

```

## More information

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.

### 2022-12-17 Rmd/Analyse changes in mackerel.Rmd
Restructured into Rmd folder

--- 2019-05-16 R/Analyse changes in mackerel.Rmd
@@
   geom_line(data=ass_final, aes(colour=purpose ), size=1.5) +
-  geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +
+  # geom_dl(aes(label  = ayear, group=purpose), method = list(dl.combine("last.points"), cex = 0.8)) +
 
-  geom_segment(aes(x=2008, xend=2012, y=2000000, yend=2000000), colour="red", size=2) +
-  geom_segment(aes(x=2013, xend=2019, y=2000000, yend=2000000), colour="pink", size=2) +
-  geom_dl(aes(x=2008, y=1800000, label="ica"), colour="red", method = list(dl.combine("last.points") ) )  +
-  geom_dl(aes(x=2013, y=1800000, label="sam"), colour="pink", method = list(dl.combine("last.points") )) +
+  # geom_segment(aes(x=2008, xend=2012, y=2000000, yend=2000000), colour="red", size=2) +
+  # geom_segment(aes(x=2013, xend=2019, y=2000000, yend=2000000), colour="pink", size=2) +
+  # geom_dl(aes(x=2008, y=1800000, label="ica"), colour="red", method = list(dl.combine("last.points") ) )  +
+  # geom_dl(aes(x=2013, y=1800000, label="sam"), colour="pink", method = list(dl.combine("last.points") )) +
 

### 2022-12-17 Rmd/Analyse changes in pelagic stocks.Rmd
Restructured into Rmd folder

--- 2019-06-10 R/Analyse changes in pelagic stocks.Rmd
@@
 # 06/06/2019 changed the data frame d; now also has purpose
+# 03/09/2019 updated after WGWIDE 2019
+# 02/10/2019 updated for pelagic forum
 # -----------------------------------------------------------------------------------------------
@@
 
-  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss",
+  filter(grepl("mac-67|mac-nea|whb-c|whb-n|her-noss|her-47d3",
                stockkeylabelold)) %>%
+  # filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss",
+  #              stockkeylabelold)) %>%
   # filter(grepl("mac-67|mac-nea", 

### 2022-12-17 Rmd/ICES advice and assessment summaries HAWG 2020.Rmd
Restructured into Rmd folder

--- 2019-09-05 R/ICES advice and assessment summaries WGWIDE 2019.Rmd
@@
 # 03/09/2019 updated after WGWIDE
+# 30/03/2020 updated after HAWG; UNFINISHED. JUST TEMPORARY SOLUTIONS. 
 # -----------------------------------------------------------------------------------------------
@@
 library(lubridate)
+library(ggrepel)
 
 # Load utils code
-source("../../mptools/r/my_utils.r")
+source("../../mptools/r/my utils.r")
+source("iAssess functions.r")
 
@@
 # Load dataset
-load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
-load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+iAssess <- get(load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))) %>%
+
+  # fix for 67bc in 2020
+  mutate(purpose = ifelse(stockkeylabelold=="her-67bc" & assessmentyear==2020, "advice", purpose)) %>%
+
+  # fix for North Sea sprat (ignore the realignment)
+  mutate(stockkeylabelold = ifelse(stockkeylabel == "spr.27.3a4","spr-nsea", stockkeylabelold))
+
+iAdvice <- get(load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")))
 
@@
 fy  <- 2000
-lay <- 2019
-ly  <- 2019
+lay <- 2020
+ly  <- 2020
 
-stock="hom-nsea"
-assessyear=lay
-assessyear=2017
-
-firstyear=fy
-lastyear=ly
-plot = c("stocksize", "fishingpressure", "recruitment")
-my.purpose="advice"
-
-# Function to plot single assessment
-plot_assess <- function(stock,assessyear, firstyear, lastyear, 
-                        my.purpose = "advice", plot=c("stocksize","fishingpressure")) {
-
-  d <-
-    iAssess %>% 
-    ungroup() %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
-    filter(assessmentyear   == assessyear) %>% 
-    filter(year             %in% firstyear:lastyear) %>% 
-    filter(purpose %in% my.purpose) %>%
-
-    #set type
-    mutate(
-      type = ifelse(purpose == "advice"      , "last"    , NA),
-      type = ifelse(grepl("bench",purpose)   ,  "benchmark", type),
-      type = ifelse(grepl("replaced",purpose), "replaced"  , type),
-      type = ifelse(is.na(type), "other", type)) %>%
-
-    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>% 
-    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, type, 
-                  assessmentyear, year,
-                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
-                  stocksize, lowstocksize, highstocksize, stocksizeunits,
-                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
-                  landings, discards, catches, catcheslandingsunits) %>%
-
-    # convert to long data
-    gather(key=variable, value=data, c(recruitment, lowrecruitment, highrecruitment, 
-                  stocksize, lowstocksize, highstocksize,
-                  fishingpressure, lowfishingpressure, highfishingpressure, 
-                  landings, discards, catches)) %>%
-
-    # set appropriate units and remove old units columns
-    mutate(
-      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
-      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
-      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
-      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
-    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%
-
-    # separate estimate, low and high
-    mutate(
-      esttype = ifelse(grepl("low", variable), "low", NA),
-      esttype = ifelse(grepl("high", variable), "high", esttype),
-      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
-    mutate(
-      variable = gsub("low|high","", variable) ) %>%
-
-    # select variables to plot
-    filter(variable %in% plot) %>% 
-    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%
-
-    # spread to make separate columns for est, low and high
-    spread(key=esttype, value=data)
-
-  # Reference points
-  rp <-
-    iAdvice %>% 
-    ungroup() %>% 
-    filter(purpose %in% my.purpose) %>% 
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
-    filter(assessmentyear   == assessyear) %>% 
-    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
-              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
-    gather(key=refpoint, value=data, blim:fmsy) %>%
-
-    # generate variables
-    mutate(
-      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
-      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
-    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%
-
-    # generate refpoint types
-    mutate(
-      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
-      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
-      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
-    dplyr::select(-refpoint)
-
-  # Titles
-  tt <-
-    d %>% 
-    distinct(variable, unit, stockkeylabel, stockdescription) %>%
-    filter(!is.na(unit))
-
-  # d %>% View()
-
-  # plot 
-  d %>%
-
-    ggplot(aes(x=year,y=est)) +
-
-    theme_publication() +
-
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10, hjust=0), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) +
-
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    # geom_ribbon(aes(ymin=low, ymax=high), alpha=0.2, fill="red") +
-    # geom_line(colour="red" ) +
-
-    # plot limits
-    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
-               colour="red", linetype="solid", inherit.aes=FALSE) +
-    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
-               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +
-
-    # plot msy
-    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
-               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (
-      values=c(last ="red",last_bench="blue",advice="black",
-               benchmark="blue",replaced="green",other="gray80")) +
-    scale_fill_manual    (
-      values=c(last = "red",last_bench="blue",advice="black",
-               benchmark="blue",replaced="green",other="gray80")) +
-    scale_linetype_manual(
-      values=c(last="solid",last_bench="solid",advice="solid",
-               benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (
-      values=c(last= 1.2,last_bench=1.2,advice=0.8,
-               benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL, 
-         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
-    facet_wrap(~variable, scales="free_y")
-
-  } # End of plot_assess function
-
-# s   <- "mac-nea"
-# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
-#             plot = c("stocksize", "fishingpressure", "recruitment"), 
-#             my.purpose="advice")
-# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
-#             plot = c("stocksize", "fishingpressure", "recruitment"), 
-#             my.purpose="bench")
-
-stock="whb-comb"
-firstassessyear=fay; lastassessyear=lay; firstyear=fy;lastyear=ly;
-plot = c("stocksize", "fishingpressure")
-include.benchmark=TRUE; include.replaced=TRUE
-plot.uncertainty=FALSE
-
-# Function to plot historical assessments
-plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, 
-                         include.benchmark = TRUE, include.replaced=TRUE, 
-                         plot=c("stocksize","fishingpressure"),
-                         plot.uncertainty=FALSE) {
-
-  d <-
-    iAssess %>% 
-    ungroup() %>%
-
-    {if (include.benchmark == TRUE & include.replaced == TRUE) {
-      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench", "replaced"))
-    } else if (include.benchmark == TRUE & include.replaced == FALSE) {
-      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
-    } else if (include.benchmark == FALSE & include.replaced == TRUE) {
-      filter(., purpose %in% c("advice","replaced"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
-    filter(assessmentyear   %in% c(firstassessyear:lastassessyear)) %>% 
-    filter(year             %in% firstyear:lastyear) %>%
-
-    group_by(assessmentyear) %>%
-
-    #set type
-    mutate(
-      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"      , "last"      , NA),
-      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
-                    n_distinct(purpose)==1, "last_bench", type),
-      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
-                      n_distinct(purpose) > 1, "bench", type),
-      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"      , "advice"    , type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose)   ,  "benchmark", type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose), "replaced"  , type),
-      type = ifelse(is.na(type), "other", type)) %>%
-
-    ungroup() %>%
-
-    # set tyear and group
-    mutate(tyear = substr(as.character(assessmentyear), 3,4)) %>% 
-    mutate(group = paste0(tyear, type)) %>% 
-    mutate(group = substr(group, 1,3)) %>% 
-    mutate(group = gsub("a|l","", group)) %>%
-
-    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, 
-                  assessmentyear, year, tyear, group, type, purpose,
-                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
-                  stocksize, lowstocksize, highstocksize, stocksizeunits,
-                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
-                  landings, discards, catches, catcheslandingsunits) %>%
-
-    # convert to long data
-    gather(key=variable, value=data, 
-           c(recruitment, lowrecruitment, highrecruitment, 
-             stocksize, lowstocksize, highstocksize,
-             fishingpressure, lowfishingpressure, highfishingpressure, 
-             landings, discards, catches)) %>%
-
-    # set appropriate units and remove old units columns
-    mutate(
-      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
-      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
-      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
-      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
-    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%
-
-    # separate estimate, low and high
-    mutate(
-      esttype = ifelse(grepl("low", variable), "low", NA),
-      esttype = ifelse(grepl("high", variable), "high", esttype),
-      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
-    mutate(
-      variable = gsub("low|high","", variable) ) %>%
-
-    # select variables to plot
-    filter(variable %in% plot) %>% 
-    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%
-
-    # spread to make separate columns for est, low and high
-    spread(key=esttype, value=data) %>%
-
-    # remove NA
-    filter(!is.na(est))
-
-  # Check
-  # d %>% filter(assessmentyear == 2019) %>% View()
-
-  # Reference points
-  rp <-
-    iAdvice %>% 
-    ungroup() %>% 
-    filter(purpose %in% my.purpose) %>% 
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
-    filter(assessmentyear   == assessyear) %>% 
-    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
-              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
-    gather(key=refpoint, value=data, blim:fmsy) %>%
-
-    # generate variables
-    mutate(
-      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
-      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
-    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%
-
-    # generate refpoint types
-    mutate(
-      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
-      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
-      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
-    dplyr::select(-refpoint)
-
-  # Titles
-  tt <-
-    d %>% 
-    distinct(variable, unit, stockkeylabel, stockdescription) %>%
-    filter(!is.na(unit))
-
-  # plot
-  d %>%
-
-    ggplot(aes(year,est, group=group)) +
-
-    theme_publication() +
-
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10, hjust=0), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    {if (plot.uncertainty)     
-      geom_ribbon(data=filter(d, type=="last"),
-          aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) } +
-
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = group, colour = type, group=group), 
-            method = list(dl.combine("last.points"), cex = 0.8, inherit.aes=FALSE)) +
-
-    # plot limits
-    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
-               colour="red", linetype="solid", inherit.aes=FALSE) +
-    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
-               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +
-
-    # plot msy
-    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
-               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (
-      values=c(last ="red",last_bench="blue",advice="black",
-               benchmark="blue",replaced="green",other="gray80")) +
-    scale_fill_manual    (
-      values=c(last = "red",last_bench="blue",advice="black",
-               benchmark="blue",replaced="green",other="gray80")) +
-    scale_linetype_manual(
-      values=c(last="solid",last_bench="solid",advice="solid",
-               benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (
-      values=c(last= 1.2,last_bench=1.2,advice=0.8,
-               benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL, 
-         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
-    facet_wrap(~variable, scales="free_y")
-
-  # return(d)
-
-  } # End of plot_history function
-
-# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE
-
-# Function to plot advice overviews
-table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {
-
-  d <-
-    iAdvice %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
-
-    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%
-
-    {if (include.replaced == TRUE) {
-      filter(., purpose %in% c("advice","replaced"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    mutate(advicebasis= substr(advicebasis,1,25),
-
-           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),
-
-           tac        = ifelse(!is.na(tac), tac, tal),
-           tactype    = ifelse(!is.na(tac), "catch", "landings"),
-
-           catch      = ifelse(!is.na(catches), catches, NA),
-           catchtype  = ifelse(!is.na(catches), "catches", ""),
-
-           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
-           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),
-
-           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
-           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),
-
-           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
-           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
-           cat_c       = 100 * (catch / lag(catch, n=1) -1)
-           ) %>%
-
-    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
-           adv_c, tac_c, cat_c,
-           advicetype, tactype, catchtype)
-
-  d %>% 
-  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
-  arrange(tacyear, desc(purpose)) %>% 
-  pandoc.table(., 
-               style        = "simple",
-               split.tables = 200, 
-               justify      = "right",
-               missing      =" ",
-               big.mark     = '', 
-               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-}
-
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+# iAssess %>% filter(stockkeylabelold=="her-67bc") %>% ungroup() %>% distinct(assessmentyear, purpose)
+# iAssess %>% filter(stockkeylabelold=="spr-nsea") %>% ungroup() %>% distinct(assessmentyear, purpose)
+# iAssess %>% filter(substr(stockkeylabelold,1,3)=="spr") %>% ungroup() %>% distinct(stockkeylabelold, stockkeylabel, assessmentyear, purpose) %>% View()
 
@@
 
-**Summary of ICES 2019 WGWIDE assessments and advice**
+**Summary of ICES 2020 WGWIDE assessments and advice**
 
@@
 
-#####page break
-
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
+# s   <- "mac-nea"
+
+plot_history(stock=s,
+             firstassessyear=fay,
+             lastassessyear=lay, 
+             firstyear=fy, 
+             lastyear=ly, 
              include.benchmark = TRUE,
+             include.replaced = TRUE,
              plot = c("stocksize", "fishingpressure"))
+
 ```
@@
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+##### page break
 
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
+
+```
+
+_Figure 1c Northeast Atlantic mackerel.
+Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch.
+Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice.
+_
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
 
@@
 
-_Table 1 Northeast Atlantic mackerel. History of advice, tac and catch._
+_Table 1.a Northeast Atlantic mackerel. History of advice, tac and catch._
 
@@
 
-s   <- "whb-comb"
+# s   <- "whb-comb"
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
-             include.benchmark = TRUE, include.replaced = FALSE, 
+plot_history(stock=s,
+             firstassessyear=fay,
+             lastassessyear=lay, 
+             firstyear=fy, 
+             lastyear=ly, 
+             include.benchmark = TRUE, 
+             include.replaced = FALSE, 
              plot = c("stocksize", "fishingpressure"), 
-             plot.uncertainty=FALSE)
+             plot.uncertainty=TRUE)
 
@@
 
+##### page break
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
+
+```
+
+_Figure 1c Blue whiting.
+Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch.
+Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice.
+_
+
 ```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
 
-s   <- "whb-comb"
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)
+table_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
 
@@
 
-plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
+plot_assess(stock=s,
+            assessyear=lay,
+            firstyear=fy,
+            lastyear=ly,
             plot = c("stocksize", "fishingpressure", "recruitment"),
@@
 
-s   <- "hom-west"
+# s   <- "hom-west"
 
@@
 _Figure 3b Western horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _
+
+##### page break
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+# s   <- "hom-west"
+
+plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
+
+```
+
+_Figure 1c western horse mackerel.
+Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch.
+Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice.
+_
 
@@
 _Table 3 Western horse mackerel. History of advice, tac and catch._
-
-##### page break
-
-*North Sea horse mackerel*
-
-The assessment and advice of North Sea is carried out every second year.
-A new stock assessor was responsible for the assessment (as the previous stock assessor changed jobs), and she found a number of errors in the codes that had been used in 2017.
-These errors were now corrected and lead to a much lower estimate of stock size in 2016.
-
-This is a so-called category 3 advice, which means that it only uses the relative trend in a (combined) survey to derive the advice.
-The trend is based on the abundance of the North Sea IBTS survey (quarter 3) and French Channel CGFS survey (quarter 4).
-The stock is considered to be at a low level even though recruitment appears to have been higher in recent year.
-A length-based proxy indicator suggests that fishing mortality is probably (just) above Fmsy.
-
-The advice is lower than the advice presented in 2017. Note that the advice is a catch advice. The advice for the landed part is 13305 t.The advice is applicable for 2020 and 2021.
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "hom-nsea"
-
-plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
-            plot = c("stocksize", "fishingpressure", "recruitment"),
-            my.purpose="advice")
-
-```
-
-_Figure 4a North Sea horse mackerel. Summary of most recent assessment_
-
-##### page break
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "hom-nsea"
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
-             include.benchmark = TRUE, include.replaced = FALSE, 
-             plot = c("stocksize", "fishingpressure"), 
-             plot.uncertainty=FALSE)
-
-```
-
-_Figure 4b North Sea horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "hom-nsea"
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)
-
-```
-
-_Table 4 North Sea horse mackerel. History of advice, tac and catch._
 
@@
 
-_Figure 5a Atlanto-scandian herring. Summary of most recent assessment_
+_Figure 4a Atlanto-scandian herring. Summary of most recent assessment_
 
@@
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
-             include.benchmark = TRUE, include.replaced = FALSE, 
+# s   <- "her-noss"
+
+plot_history(stock=s,
+             firstassessyear=fay,
+             lastassessyear=lay, 
+             firstyear=fy, 
+             lastyear=ly, 
+             include.benchmark = TRUE, 
+             include.replaced = FALSE, 
              plot = c("stocksize", "fishingpressure"), 
-             plot.uncertainty=FALSE)
+             plot.uncertainty=TRUE)
 
@@
 
-_Figure 5b Atlanto-scandian herring. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _
+_Figure 4b Atlanto-scandian herring. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _
+
+##### page break
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+# s   <- "her-noss"
+
+plot_advice(stock=s,firstadviceyear=2001,lastadviceyear=lay+1, include.replaced = FALSE)
+
+```
+
+_Figure 4c Atlanto-scandian herring.
+Top: comparison of (max) scientific advice, TAC (or sum of unilateral quota) and Total Catch.
+Bottom: percentage deviation from ICES advice, CoA is Catch over Advice, ToA is TAC over Advice.
+_
 
@@
 
-_Table 5 Atlanto-scandian herring. History of advice, tac and catch._
+_Table 4 Atlanto-scandian herring. History of advice, tac and catch._
 

### 2022-12-17 Rmd/ICES advice and assessment summaries WGWIDE 2019.Rmd
Restructured into Rmd folder

--- 2019-09-05 R/ICES advice and assessment summaries WGWIDE 2019.Rmd
@@
       variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
+    filter(variable %in% plot) %>% 
     mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

### 2022-12-17 Rmd/ICES advice and assessment summaries.Rmd
Restructured into Rmd folder

--- 2019-06-10 R/ICES advice and assessment summaries.Rmd
@@
 
-# s   <- "hke-nrtn"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
-In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
-# 
-# [hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)
+s   <- "hke-nrtn"
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
 
 ```
+
+Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
+In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
+
+[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)
 
@@
 
-# s   <- "cod-347d"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
+s   <- "cod-347d"
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+
+```
+
+The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
 According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey.
@@
 This also leads to a substantially lower advice for 2019 compared to the advice for 2018.
-# 
-# [cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)
 
-```
+[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)
 
@@
 
-# s   <- "ple-nsea"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
+s   <- "ple-nsea"
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+
+```
+
+Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
 An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016).
@@
 The advice is based on the MSY approach.
-# 
-# [ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)
 
-```
+[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)
 
@@
 
-# s   <- "sol-nsea"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
-# 
-# [sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)
+s   <- "sol-nsea"
+
+cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+
+plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+
+table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
 
 ```
+
+Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea.
+
+[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)
 

### 2022-12-17 Rmd/ICES assessments averaged.Rmd
Restructured into Rmd folder

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES assessments averaged.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)

# Load utils code
source("../../prf/r/my utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

```

**Summary of ICES 2019 assessments and advice**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  # filter(grepl("her", stockkeylabelnew)) %>% 
  # distinct(stockkeylabelnew) %>% write.csv(file="C:/TEMP/stocks.csv", row.names = FALSE)

  group_by(year) %>% 
  summarise(
    stocksize       = sum(stocksize, na.rm=TRUE), 
    fishingpressure = mean(fishingpressure, na.rm=TRUE), 
    catches           = sum(catches, na.rm=TRUE)
  ) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  # filter(year >= 2020) %>% View()

  ggplot(aes(x=year, y=data)) +
  theme_publication() +
  geom_line(aes()) +
  expand_limits(y=0) +
  facet_wrap(~variable, scale="free_y", ncol=1)

```

\newpage

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  mutate(
    speciestype = case_when(
      grepl("cod|had|whg|sai|pol", stockkeylabelnew) ~ "roundfish",
      grepl("her|mac|whb|hom|anc|pil", stockkeylabelnew) ~ "pelagic",
      grepl("ple|sol|meg|tur|bll", stockkeylabelnew) ~ "flatfish",
      TRUE ~ as.character(NA)
    )
  ) %>% 
  drop_na(speciestype) %>%

  group_by(speciestype, year) %>% 
  summarise(
    stocksize       = sum(stocksize, na.rm=TRUE), 
    fishingpressure = mean(fishingpressure, na.rm=TRUE), 
    catches           = sum(catches, na.rm=TRUE)
  ) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  # filter(year >= 2020) %>% View()

  ggplot(aes(x=year, y=data)) +
  theme_publication() +
  geom_line(aes()) +
  expand_limits(y=0) +
  facet_wrap(~speciestype+variable, scale="free_y", ncol=3)

```

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

t <-
  iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>% 
  filter(purpose=="advice") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  mutate(
    speciestype = case_when(
      grepl("cod|had|whg|sai|pol", stockkeylabelnew) ~ "roundfish",
      grepl("her|mac|whb|hom|anc|pil", stockkeylabelnew) ~ "pelagic",
      grepl("ple|sol|meg|tur|bll", stockkeylabelnew) ~ "flatfish",
      TRUE ~ as.character(NA)
    )
  ) %>% 
  drop_na(speciestype)

t %>% 
  dplyr::select(speciestype, stockkeylabelnew, year, stocksize, catches) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  ggplot(aes(x=year, y=data, group=stockkeylabelnew)) +
  theme_publication() +
  geom_bar(aes(fill=stockkeylabelnew), stat="identity") +
  facet_wrap(~speciestype+variable, scales="free_y", ncol=2)

```

### 2022-12-17 Rmd/Metadata overview.Rmd
Restructured into Rmd folder

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Metadata overview.Rmd
#
# 14/11/2022 first coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
# library(ggExtra)   # for specific type of plots e.g. ggMarginal
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)

# captioner library
library(captioner) # deal with captions
tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")

# Load utils code
source("../../prf/r/my utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# Create subset of series (stocks) to be used for analysis
d1 <-
  iAdvice %>%
  filter(adviceonstock == TRUE) %>% 
  group_by(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold) %>% 
  summarise(
    nyears          = n_distinct(tacyear),
    nadvicelandings = sum(!is.na(advisedlandingsmax)),
    nadvicecatch    = sum(!is.na(advisedcatchmax)),
    nblim           = sum(!is.na(blim)),
    nfset           = sum(!is.na(fset)),
    nssbset         = sum(!is.na(ssbset)),
  ) %>%

  filter(nssbset > 0)

  # filter stocks
d2 <-
  iAdvice %>%
  filter(stockkey %in% d1$stockkey) %>% 
  filter(adviceonstock == TRUE) %>% 
  group_by(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold, tacyear) %>% 
  summarise(
    nblim           = sum(!is.na(blim)),
    nfmsy           = sum(!is.na(fmsy)),
    nadvicelandings = sum(!is.na(advisedlandingsmax)),
    nadvicecatch    = sum(!is.na(advisedcatchmax)),
    nfset           = sum(!is.na(fset)),
    nssbset         = sum(!is.na(ssbset)),
    ntal            = sum(!is.na(tal)),
    ntac            = sum(!is.na(tac)),
    nlandings       = sum(!is.na(landings)),
    ncatch          = sum(!is.na(catches))
  )

d2 %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", nblim:ncatch) %>% 
  filter(data > 0) %>% 
  mutate(
    variable = factor(variable,
                      levels =c("nblim","nfmsy","nadvicelandings","nadvicecatch", "nfset", "nssbset",
                                "ntal","ntac","nlandings","ncatch"))
  ) %>%

  ggplot(aes(x=tacyear, y=variable)) +
  theme_publication() +
  geom_point() +
  facet_wrap(~stockkeylabelold)

```

<!-- Plot assessment and reconstructions      ------------------------------------------------------ -->

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# Create subset of series (stocks) to be used for analysis
a <-
  iAssess %>%
  filter(stockkey %in% d1$stockkey) %>% 
  group_by(stockkey) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  mutate(catch = ifelse(is.na(catches), landings, catches)) %>% 
  dplyr::select(
    stockkey, stockkeylabelnew, stockkeylabelold, purpose, assessmentyear, year, 
    lowstocksize, stocksize, highstocksize,
    lowfishingpressure, fishingpressure, highfishingpressure,
    catch
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", lowstocksize:catch) %>% 
  mutate(
    metric = case_when(
      grepl("low", variable)  ~ "low",
      grepl("high", variable) ~ "high", 
      TRUE                    ~ "est"
    )
  ) %>% 
  mutate(variable = gsub("low|high", "", variable)) %>% 
  pivot_wider(names_from = metric, values_from = data)

d <-
  iAdvice %>%
  filter(stockkey %in% d1$stockkey) %>% 
  filter(adviceonstock == TRUE) %>% 
  filter(!is.na(fset) | !is.na(ssbset)) %>% 
  dplyr::select(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold, year=tacyear, 
                fishingpressure=fset, stocksize=ssbset, catch=catches) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "est", fishingpressure:catch)

for(myvar in c("stocksize", "fishingpressure", "catch")) {
  p <-
    ggplot() +
    theme_publication() +
    geom_point(data=filter(a, variable == myvar, year >= 2000), 
               aes(x=year, y=est)) +
    geom_line(data=filter(a, variable == myvar, year >= 2000), 
               aes(x=year, y=est)) +

    geom_point(data=filter(d, variable == myvar, year >= 2000), 
               aes(x=year, y=est),
               colour="red") +
    expand_limits(y=0) +
    labs(title=myvar) +
    facet_wrap(~stockkeylabelold, scales="free_y")

  print(p)
}

```

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

c <-
  a %>% 
  mutate(source="assessed") %>% 
  dplyr::select(-purpose, -assessmentyear) %>% 
  bind_rows(
    d %>% 
      mutate(source="predicted")
  ) %>% 
  dplyr::select(-low, -high, -speciescommonname) %>% 
  tidyr::pivot_wider(names_from = source, values_from = est) %>% 
  drop_na(predicted)

myvar <- "stocksize"

for(myvar in c("stocksize", "fishingpressure", "catch")) {
  p <-
    c %>% 
    filter(variable == myvar) %>%

    ggplot(aes(x=predicted, y=assessed)) +
    theme_publication() +
    geom_point() +
    geom_abline(slope=1, intercept=0) +
    expand_limits(x=0, y=0) +
    labs(title=myvar) +
    facet_wrap(~stockkeylabelold, scales="free")

  print(p)
}

```

### 2022-12-17 Rmd/Plot for pelagic forum 2019.Rmd
Restructured into Rmd folder

--- 2019-06-10 R/ICES advice and assessment summaries.Rmd
@@
 # -----------------------------------------------------------------------------------------------
-# ICES advice and assessment summaries.Rmd
+# Plots for pelagic forum 2019.Rmd
 #
@@
 # 01/06/2019 updated for herring advice round
+# 29/09/2019 adapted from stock summary code
 # -----------------------------------------------------------------------------------------------
@@
 
-# Libraries
-library(rmarkdown) # creating markdown documents
-library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
-library(reshape2)  # reshaping data; e.g. dcast
-library(ggthemes)  # for themes
-library(pander)    # for print tables
-library(readxl)    # read excel files
-library(cowplot)   # multiplots
-library(directlabels)  # label graphs
-library(RColorBrewer) # colours
-library(lubridate)
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+library(stringr)
+library(readxl)
+library(gridExtra)                   # for plotting table as image (bio data)
+library(RColorBrewer)
 
@@
 source("../../mptools/r/my_utils.r")
+source("iPlot_utils.r")
 
+# Set dropbox folder
 dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
 
-# Load dataset
-load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
-load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+# load the data
+iSpecies <-
+  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
+  ungroup() %>% 
+  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild)
+
+iAssess <-
+  get(load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))) %>% 
+  left_join(iSpecies, by="speciesfaocode")
+
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))
+
+# add intermediate and forecast data
+forecasts <-
+  read_excel(path=paste(dropboxdir,"/excel/Pelagic forecasts 2019.xlsx", sep=""), col_names = TRUE) %>%
+  left_join(iStockkey, by="stockkeylabelold") %>%
+  mutate(stockkeylabel = stockkeylabelnew) %>%
+  left_join(iRename,   by=c("stockkey", "stockkeylabel")) %>%
+  left_join(iSpecies,  by="speciesfaocode") %>%
+  mutate(catcheslandingsunits = "tonnes")
+
+# t <-
+#   icesSAGfull %>%
+#   filter(year < assessmentyear) %>% 
+#   bind_rows(forecasts)
 
 fay <- 2011
-fy  <- 2000
+fy  <- 1990
 lay <- 2019
@@
 
-# stock=c("spr-nsea","spr.27.3a4"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark==TRUE
-# stock=c("mac-nea"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark=TRUE; plot=c("stocksize","fishingpressure","recruitment")
-
-# d %>% filter(stockkeylabelold == "mac-nea", assessmentyear == 2019) %>% View()
-# unique(iAssess$purpose)
-
-# Function to plot historical assessments
-plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, include.benchmark = FALSE, 
-                         plot=c("stocksize","fishingpressure")) {
-
-  d <-
-    iAssess %>% 
-    ungroup() %>%
-
-    {if (include.benchmark == TRUE) {
-      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
-
-    filter(assessmentyear   %in% c(firstassessyear:lastassessyear), 
-           year             %in% firstyear:lastyear) %>%
-
-    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>%
-
-    #set type
-    mutate(
-      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"       , "last"      , NA),
-      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose),  "last_bench", type),  
-      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"       , "advice"    , type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose),  "benchmark" , type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose) , "replaced"  , type),
-      type = ifelse(is.na(type), "other", type)) %>%
-
-    # set group
-    mutate(group = paste0(tyear, type)) %>%
-
-    select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
-           type, group, source,
-           lowrecruitment, recruitment, highrecruitment, 
-           lowfishingpressure, fishingpressure, highfishingpressure, 
-           lowstocksize, stocksize, highstocksize) %>%
-    data.frame()
-
-  # find out what is the last assessment
-  last <- filter(d, assessmentyear == lastassessyear) %>% distinct(type) %>% as.character()
-
-  # plot stocksize
-  stocksize <-
-    d %>% 
-    filter(!is.na(stocksize)) %>%
-
-    ggplot(aes(year,stocksize, group=group)) +
-
-    theme_publication() +
-
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
-               colour="red", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes = FALSE) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=msybtrigger), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, msybtrigger, label = "msy btrigger", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "stocksize")
-
-  # fishingpressure
-  fishingpressure <-
-    d %>% 
-    filter(!is.na(fishingpressure)) %>%
-
-    ggplot(aes(year,fishingpressure, group=group)) +
-
-    theme_publication() +
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowfishingpressure, ymax=highfishingpressure, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=fmsy), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "fishingpressure")
-
-  # Recruitment
-  recruitment <-
-    d %>% 
-    filter(!is.na(recruitment)) %>%
-
-    ggplot(aes(year,recruitment, group=group)) +
-
-    theme_publication() +
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "Recruitment")
-
-    if (length(plot) == 1) {
-      if ("stocksize" %in% plot ) print(stocksize)
-      if ("fishingpressure" %in% plot ) print(fishingpressure)
-      if ("recruitment" %in% plot ) print(recruitment)
-
-    } else if (length(plot) == 2) {
-
-      if ("stocksize" %in% plot & "fishingpressure" %in% plot) plot_grid(stocksize, fishingpressure,  ncol=2, align = 'hv') 
-      if ("stocksize" %in% plot & "recruitment" %in% plot) plot_grid(stocksize, recruitment,  ncol=2, align = 'hv') 
-      if ("recruitment" %in% plot & "fishingpressure" %in% plot) plot_grid(fishingpressure, recruitment,  ncol=2, align = 'hv')
-
-    } else if (length(plot) == 3) {
-        plot_grid(stocksize, fishingpressure,  recruitment, ncol=length(plot), align = 'hv') 
-    }
-
-  } # End of plot_history function
-
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"), include.benchmark = TRUE)
-
-# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("spr-nsea","spr.27.3a4"); include.replaced <- TRUE
-# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE
-
-# Function to plot advice overviews
-table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {
-
-  d <-
-    iAdvice %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
-
-    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%
-
-    {if (include.replaced == TRUE) {
-      filter(., purpose %in% c("advice","replaced"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    mutate(advicebasis= substr(advicebasis,1,25),
-
-           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),
-
-           tac        = ifelse(!is.na(tac), tac, tal),
-           tactype    = ifelse(!is.na(tac), "catch", "landings"),
-
-           catch      = ifelse(!is.na(catches), catches, NA),
-           catchtype  = ifelse(!is.na(catches), "catches", ""),
-
-           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
-           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),
-
-           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
-           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),
-
-           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
-           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
-           cat_c       = 100 * (catch / lag(catch, n=1) -1)
-           ) %>%
-
-    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
-           adv_c, tac_c, cat_c,
-           advicetype, tactype, catchtype)
-
-  d %>% 
-  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
-  arrange(tacyear, desc(purpose)) %>% 
-  pandoc.table(., 
-               style        = "simple",
-               split.tables = 200, 
-               justify      = "right",
-               missing      =" ",
-               big.mark     = '', 
-               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-}
-
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-
 ```
 
-**Summary of ICES 2019 assessments and advice**
+**Status of pelagic stocks in the Northeast Atlantic**
 
@@
 
-In this document an overview is presented of the ICES assessment and advice for the most recent number of years for the stocks that are relevant to the PFA and Dutch fisheries in general.
-
-<!--Section 2: material and methods ----------------------------------------------------------- -->
-
-# Material and methods
-
+In this document an overview is presented of the ICES assessment and advice for pelagic stocks in the Northeast Atlantic.
 The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.
 
-<!--Section 3: advice and assessment overviews ------------------------------------------------ -->
+<!--Section 2: general trends ------------------------------------------------ -->
 
-# Assessment and advice overviews
+# General trends
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+# catches by fisheriesguild
+
+d <-
+  iAssess %>% 
+  ungroup() %>% 
+  # filter(assessmentyear      == lay | (assessmentyear == 2017 & stockkeylabelold == "cap-icel") ) %>% 
+  filter(assessmentyear      == lay  ) %>% 
+  # filter(catcheslandingsunits == "tonnes") %>% 
+  filter(year >= fy, year < ly) %>% 
+  filter(purpose %in% c("advice")) %>%
+
+  mutate(catches = ifelse(is.na(catches) | catches == 0, landings, catches))  %>% 
+  mutate(fisheriesguild = ifelse(fisheriesguild %in% c("pelagic","demersal"), fisheriesguild, "other"))  %>% 
+  mutate(fisheriesguild = factor(fisheriesguild, levels=c("pelagic","demersal","other"))) %>%
+
+  group_by(fisheriesguild, speciesfaocode, year) %>% 
+  summarize(catches  = sum(catches, na.rm=TRUE),
+            landings = sum(landings, na.rm=TRUE)) %>%
+
+  group_by(fisheriesguild, year) %>% 
+  summarize(catches  = sum(catches, na.rm=TRUE),
+            landings = sum(landings, na.rm=TRUE)) %>% 
+  group_by(year) %>% 
+  mutate(perc = round(100 * catches/ sum(catches), digits=2)) %>% 
+  arrange(catches) %>% 
+  mutate(fisheriesguild = factor(fisheriesguild, fisheriesguild))
+
+# d %>% filter(year == 2018) %>% arrange(-catches) %>% View()
+
+d2 <-
+  d %>% 
+  group_by(fisheriesguild) %>% 
+  summarize(perc = mean(perc),
+            catches=mean(catches),
+            year = mean(year)) %>% 
+  ungroup() %>% 
+  mutate(cumcatches=cumsum(catches))
+
+# plot stacked catches: pelagic, demersal and other
+p1 <-
+  d %>%
+
+  ggplot(aes(year,catches)) +
+  theme_publication() +
+  theme(legend.position="none") +
+  geom_bar(aes(fill=(fisheriesguild)), stat="identity", position="stack", colour="black") +
+  # geom_text(data=subset(d, year==ly-1), aes(label=round(perc,0), group=fisheriesguild), 
+  #           position = position_stack(vjust = 0.5), colour="white") +
+  geom_text(data=subset(d, year==ly-1), 
+            aes(label=fisheriesguild, group=fisheriesguild, x=ly), 
+            position = position_stack(vjust = 0.5), colour="black", hjust=0) +
+  # geom_dl(aes(label= perc),
+  #         position = position_stack(vjust = 0.5),
+  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
+  #                         fontface = 'bold',cex=.9, hjust = 0)) +
+  scale_fill_brewer(palette = "Paired") +
+  labs(x="",y="tonnes",title="catches") +
+  scale_y_continuous(labels = comma) +
+  scale_x_continuous(limits=c(fy,ly+3))
+
+# d %>% filter(year == 2018) %>% View()
+
+# plot of filled relative catches (how to add the direct labels?)
+p2 <-
+  d %>% 
+  ggplot(aes(year,perc)) +
+  theme_publication() +
+  theme(legend.position="none") +
+  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
+  theme(plot.background = element_blank()) +
+  theme(axis.text = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  geom_bar(aes(fill=fisheriesguild), stat="identity", position="stack", colour="black") +
+  scale_fill_brewer(palette = "Paired") +
+  labs(x="",y="",title="") +
+  # scale_y_continuous(labels = comma) +
+  scale_x_continuous(limits=c(fy,ly))
+
+vp <- grid::viewport(width = 0.3, height = 0.3, 
+                     x = 0.15, y = 0.12, 
+                     just = c("left", "bottom"))
+
+ print(p1)
+     theme_set(theme_bw(base_size = 8))
+     # theme_white()
+     print(p2, vp = vp)
+     theme_set(theme_bw())
+
+  # cowplot::plot_grid(p1, p2,
+  #                ncol=2, scale=c(1.0, 0.25), align="hv",
+  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))
+
+```
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+# catches of pelagic species
+
+d <-
+  iAssess %>% 
+  ungroup() %>% 
+  filter(year >= fy, year < ly) %>% 
+  filter(purpose %in% c("advice")) %>% 
+  filter(fisheriesguild == "pelagic") %>%
+  filter(assessmentyear == lay ) %>%
+
+  mutate(speciesfaocode = ifelse(speciesfaocode %in% c("mac","whb","her", "hom", "pil","spr", "cap"), 
+                                 speciesfaocode, "other")) %>%
+
+  # filter(speciesfaocode=="ane") %>% View()
+  # filter((speciesfaocode == "ane.27.8" & assessmentyear == lay-1) | 
+  #        (speciesfaocode != "ane.27.8") & assessmentyear == lay ) %>%
+
+  mutate(
+    catches = ifelse(!is.na(catches), catches, landings) ) %>% 
+  dplyr::select(-landings) %>% 
+  group_by(speciesfaocode, year) %>% 
+  summarize(
+    catches  = sum(catches, na.rm=TRUE)) %>% 
+  group_by(speciesfaocode) %>% 
+  mutate(
+    totalcatches = sum(catches, na.rm=TRUE) ) %>%
+  filter(totalcatches > 0) %>% 
+  group_by(year) %>% 
+  mutate(
+    perc = round(100 * catches/ sum(catches), digits=2) ) %>% 
+  arrange(catches) %>% 
+  mutate(speciesfaocode = factor(speciesfaocode, speciesfaocode))
+
+# d %>% filter(speciesfaocode=="ane.27.8") %>%  View()
+# d %>% filter(year == 2018) %>% View()
+
+d2 <-
+  d %>% 
+  group_by(speciesfaocode) %>% 
+  summarize(perc = mean(perc),
+            catches=mean(catches),
+            year = mean(year)) %>% 
+  ungroup() %>% 
+  mutate(cumcatches=cumsum(catches))
+
+# plot stacked catches: pelagic, demersal and other
+p1 <-
+  d %>%
+
+  ggplot(aes(year,catches)) +
+  theme_publication() +
+  theme(legend.position="none") +
+  geom_bar(aes(fill=(speciesfaocode)), stat="identity", position="stack", colour="black") +
+  geom_text(data=subset(d, year==ly-1), 
+            aes(x = year+0.7,label=speciesfaocode, group=speciesfaocode, colour=speciesfaocode ), 
+            position = position_stack(vjust = 0.5), hjust=0) +
+  # geom_dl(aes(label= perc),
+  #         position = position_stack(vjust = 0.5),
+  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
+  #                         fontface = 'bold',cex=.9, hjust = 0)) +
+  scale_fill_brewer(palette = "Paired") +
+  scale_colour_brewer(palette = "Paired") +
+  labs(x="",y="tonnes",title="catches by species") +
+  scale_y_continuous(labels = comma) +
+  scale_x_continuous(limits=c(fy,ly+1))
+
+d %>% filter(year==2018) %>% ungroup() %>% summarize(catches=sum(catches))
+
+# plot of filled relative catches (how to add the direct labels?)
+p2 <-
+  d %>% 
+  ggplot(aes(year,perc)) +
+  theme_publication() +
+  theme(legend.position="none") +
+  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
+  theme(plot.background = element_blank()) +
+  theme(axis.text = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  geom_bar(aes(fill=speciesfaocode), stat="identity", position="stack", colour="black") +
+  scale_fill_brewer(palette = "Paired") +
+  labs(x="",y="",title="") +
+  # scale_y_continuous(labels = comma) +
+  scale_x_continuous(limits=c(fy,ly))
+
+vp <- grid::viewport(width = 0.3, height = 0.3, 
+                     x = 0.15, y = 0.12, 
+                     just = c("left", "bottom"))
+
+ print(p1)
+     theme_set(theme_bw(base_size = 8))
+     # theme_white()
+     print(p2, vp = vp)
+     theme_set(theme_bw())
+
+  # cowplot::plot_grid(p1, p2,
+  #                ncol=2, scale=c(1.0, 0.25), align="hv",
+  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))
+
+```
+
+```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}
+
+# catches and advice for key pelagic species
+d <-
+  iAssess %>%
+  filter(year < ly) %>% 
+  bind_rows(forecasts) %>% 
+  filter(stockkeylabelold %in% c("her-47d3","her-noss","mac-nea","whb-comb")) %>% 
+  filter(assessmentyear == lay ) %>%
+  filter(year >= fy, year <= ly+1) %>% 
+  filter(purpose %in% c("advice")) %>%
+
+  mutate(
+    catches = ifelse(!is.na(catches), catches, landings) ) %>% 
+  dplyr::select(-landings) %>% 
+  group_by(stockkeylabelold, year) %>% 
+  summarize(
+    catches  = sum(catches, na.rm=TRUE)) %>% 
+  group_by(stockkeylabelold) %>% 
+  mutate(
+    totalcatches = sum(catches, na.rm=TRUE) ) %>%
+  filter(totalcatches > 0) %>% 
+  group_by(year) %>% 
+  mutate(
+    perc = round(100 * catches/ sum(catches), digits=2) ) %>% 
+  arrange(catches) %>% 
+  mutate(stockkeylabel = factor(stockkeylabelold, stockkeylabelold)) %>% 
+  arrange(stockkeylabel, year)
+
+# plot stacked catches: pelagic, demersal and other
+d %>% 
+  ggplot(aes(year,catches)) +
+  theme_publication() +
+  theme(legend.position="none") +
+  geom_bar(aes(fill=(stockkeylabelold)), stat="identity", position="stack", colour="black") +
+  geom_text(data=subset(d, year==ly+1), 
+            aes(x = year+0.7,label=stockkeylabelold, group=stockkeylabelold, colour=stockkeylabelold ), 
+            position = position_stack(vjust = 0.5), hjust=0) +
+  scale_fill_brewer(palette = "Paired") +
+  scale_colour_brewer(palette = "Paired") +
+  labs(x="",y="tonnes",title="catches by key pelagic species") +
+  scale_y_continuous(labels = comma) +
+  scale_x_continuous(limits=c(fy,ly+5))
+
+# d %>% 
+#   arrange(stockkeylabelold, year) %>% 
+#   filter(year >= 2019) %>% 
+#   group_by(year) %>% 
+#   summarize(catches=sum(catches)) %>% 
+#   View()
+
+```
+
+<!--Section 3: overviews by stock ------------------------------------------------ -->
+
+# Overviews by stock
 

### 2022-12-17 Rmd/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Restructured into Rmd folder

--- 2019-08-15 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
@@
 # -----------------------------------------------------------------------------------------------
-# Analyse changes in pelagic stocks.Rmd
+# Paper 02 Scale-jump analysis of historic retrospective patterns.Rmd
 #
@@
 myvar         <- "stocksize"
-mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
+# mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)
 
@@
 
-*Figure 6: Comparisons of stock size (SSB) in assessments where stock definitions have changed. Blue-green is old stock definition, red is new stock definition*
+*Figure 7: Comparisons of stock size (SSB) in assessments where stock definitions have changed. Blue-green is old stock definition, red is new stock definition*
 
 ```{r, eval=TRUE, fig.align="center", fig.asp=0.7, message=FALSE, warning=FALSE, include=FALSE}
+
+# NOT SHOWN
 
@@
 
+# *Figure 7: Number of scale jumps in stocksize for the nine stocks considered, using 5 year means and a threshold of 10%.
+The red line indicates the number of assessments included in the comparison (right hand y-axis)*
+
 ```
 
-*Figure 7: Number of scale jumps in stocksize for the nine stocks considered, using 5 year means and a threshold of 10%.
-The red line indicates the number of assessments included in the comparison (right hand y-axis)*
-
-```{r echo=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE}
+```{r eval=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE, include=FALSE}
+
+# NOT SHOWN
 
@@
 
-*Figure 8: Proportion of scale jumps in stocksize by stock and decade, using 5 year means and a threshold of 10%.
-The red line indicates the number of assessments included in the comparison (right hand y-axis)*
+```{r echo=FALSE, fig.align="center", fig.asp=0.4, message=FALSE, warning=FALSE}
+
+# plot number of scale jumps by decade
+
+mynyears      <- 5
+myremoveyears <- 0
+mythreshold   <- 0.1
+myvar         <- "stocksize"
+mytitle       <- paste0(myvar, ", removed years: ", myremoveyears)
+
+t <-
+  res$d.pcm %>% 
+  filter(variable        == myvar,
+         numberofyears   == mynyears,
+         removelastyears == myremoveyears,
+         threshold       == mythreshold) %>% 
+  distinct(stock, variable, assessmentyear, phi, scale_jump, 
+           removelastyears, numberofyears, threshold) %>%
+  filter(!is.na(scale_jump)) %>% 
+  mutate(decade = (10*floor(assessmentyear/10)) ) %>% 
+  # mutate(decade = paste0(decade, "s")) %>% 
+  group_by(variable, decade, scale_jump) %>% 
+  summarize(n = n()) %>% 
+  group_by(variable, decade) %>% 
+  mutate(prop = n / sum(n), 
+         nassess = sum(n)) %>% 
+  filter(scale_jump == TRUE)
+
+p1 <-
+  ggplot(t, aes(x=reorder(decade, -decade), y=prop)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    # plot.background   = element_rect(fill="red"),
+    plot.margin        = unit(c(0,0,0,0.0),"cm"),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_bar(stat="identity") +
+  scale_y_continuous("Proportion scale jumps", breaks=pretty_breaks(n=4)) +
+  labs(x="") +
+  coord_flip()
+
+p2 <-
+  t %>% 
+  ggplot(aes(x=reorder(decade, -decade), y=nassess)) +
+  theme_publication() +
+  theme(
+    panel.spacing      = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background   = element_blank(),
+    strip.text.x       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    axis.text.y = element_blank(), 
+    axis.ticks.y=element_blank(),
+    plot.margin        = unit(c(0,0,0,0.0),"cm"),
+    # plot.background   = element_rect(fill="blue"),
+    strip.placement    = "outside",
+    legend.position    = "none"
+  ) +
+
+  geom_path(colour="red") +
+  geom_point(colour="red") +
+  expand_limits(y=0) +
+  labs(x="") +
+  coord_flip()
+
+cowplot::plot_grid(p1, 
+                   p2 , 
+                   ncol=2,
+                   scale=0.99, align="hv", 
+                   rel_widths = c(1.2,0.6))
+
+# scaling of the two axes
+# ylim.prim <- c(0, 0.5)     # proportions
+# ylim.sec  <- c(0, 90)     # number of assessments
+
+# calculations for two axes;
+# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
+# b <- diff(ylim.prim)/diff(ylim.sec)
+# a <- b*(ylim.prim[1] - ylim.sec[1])
+
+# t %>%   
+#   ggplot(aes(x=reorder(decade, -decade), y=prop)) +
+#   theme_publication() +
+#   theme(
+#     panel.spacing    = unit(0.1, "lines"),
+#     panel.grid.major.x = element_line(colour = "gray80"),
+#     strip.background = element_blank(),
+#     strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+#     strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+#                                     angle = 180),
+#     strip.placement = "outside",
+#     legend.position  = "none"
+#   ) +
+#   
+#   geom_bar(stat="identity") +
+#   geom_line(data=t2, aes(x= reorder(decade, -decade), y = a + nassess * b), color = "red", size=1) +
+#   scale_y_continuous("Proportion scale jumps", 
+#                      sec.axis = sec_axis(~ (. - a)/b, 
+#                                          name = "N Assessments", 
+#                                          breaks=pretty_breaks(n=4))) +
+#   labs(x="") +
+#   coord_flip()
+#
+
+```
+
+*Figure 6: Proportion of scale jumps in stocksize by decade, using 5 year means and a threshold of 10%. The red line indicates the number of assessments included in the comparison (right hand y-axis)*
 
@@
 
+p1 <-
+  t %>%
+
+  ggplot(aes(x=reorder(assessmentmodel, -rank), y=prop)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    # axis.text.y = element_blank(), 
+    # axis.ticks.y=element_blank(),
+    plot.margin        = unit(c(0,0,0,0.0),"cm"),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_bar(stat="identity") +
+  labs(x="") +
+  scale_y_continuous("Proportion scale jumps", breaks=pretty_breaks(n=4)) +
+  coord_flip()
+
+p2 <-
+  t %>%
+
+  ggplot(aes(x=reorder(assessmentmodel, -rank), y=nassess)) +
+  theme_publication() +
+  theme(
+    panel.spacing    = unit(0.1, "lines"),
+    panel.grid.major.x = element_line(colour = "gray80"),
+    strip.background = element_blank(),
+    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
+    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
+                                    angle = 180),
+    axis.text.y = element_blank(), 
+    axis.ticks.y=element_blank(),
+    plot.margin        = unit(c(0,0,0,0.0),"cm"),
+    strip.placement = "outside",
+    legend.position  = "none"
+  ) +
+
+  geom_point(colour="red") +
+  geom_line(colour="red") +
+  labs(x="") +
+  scale_y_continuous("N assessments", breaks=pretty_breaks(n=4)) +
+  coord_flip()
+
+cowplot::plot_grid(p1, 
+                   p2 , 
+                   ncol=2,
+                   scale=0.99, align="hv", 
+                   rel_widths = c(1.2,0.6))
+
 # scaling of the two axes
-ylim.prim <- c(0, 100)   # number of assessments  
-ylim.sec  <- c(0, 1)     # proportions
+# ylim.prim <- c(0, 100)   # number of assessments  
+# ylim.sec  <- c(0, 1)     # proportions
 
@@
 # https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
-b <- diff(ylim.prim)/diff(ylim.sec)
-a <- b*(ylim.prim[1] - ylim.sec[1])
+# b <- diff(ylim.prim)/diff(ylim.sec)
+# a <- b*(ylim.prim[1] - ylim.sec[1])
 

### 2022-12-17 R/DownloadPDFfromICES.r

@@ -17,23 +17,22 @@ t <-
   readLines(fn) %>% 
   as.data.frame() %>% 
   setNames("string") %>% 
-  filter(grepl("T1 - |L4 - |ER -", string)) %>%
+  filter(grepl("T1 - |L4 - |KW - ", string)) %>%
   separate(string, into=c("code","data"), sep=" - ") %>% 
-  mutate(id = ifelse(row_number()==1,1,NA)) %>% 
+
+  # filter(KW = her.27.etc.)
   
-  # How to deal with records with one T1 and multiple L4?
-  # Can I set an index to each record, e.g. based on ER -?
-  
-  mutate(id = ifelse(is.na(data),lag(id)+1,lag(id))) %>% 
   bind_cols(id = sort(rep(seq(1,230,1),2))) %>% 
   tidyr::pivot_wider(names_from = code, values_from = data) %>% 
+  drop_na(T1) %>% 
   mutate(string = gsub("L4 - ","", string))
 
-class(t)
-glimpse(t)
+i <- 1
+for (i in 1:nrow(t)) {
+  print(paste(t$id[[i]],t$T1[[i]]))
+  httr::GET(t$L4[[i]], write_disk(paste0("C:/TEMP/",t$T1[[i]],".pdf"), overwrite=TRUE))
+}
 
-download.file(t[1], "C:/TEMP/test.pdf")
-httr::GET(t[[1,1]], write_disk("C:/TEMP/test.pdf", overwrite=TRUE))
 httr::GET(t[[2,1]], write_disk("C:/TEMP/test2.pdf", overwrite=TRUE))
 
 # assessmentkeys <- sort(findAssessmentKey(year=myyear))

### 2022-12-17 Rmd/Paper 03 short term forecast analysis.Rmd
First code for Paper 03 on forecasts

--- 2022-12-17 Rmd/ICES advice and assessment summaries.Rmd
@@
 # -----------------------------------------------------------------------------------------------
-# ICES advice and assessment summaries.Rmd
+# Paper 03 short term forecast analysis.Rmd
 #
-# 06/07/2018 first coding
-# 14/12/2018 updated for the new database structure
-# 01/06/2019 updated for herring advice round
+# 17/12/2022 first coding
 # -----------------------------------------------------------------------------------------------
@@
 # Load utils code
-source("../../mptools/r/my_utils.r")
+source("../../prf/r/my utils.r")
 
@@
 
-fay <- 2011
-fy  <- 2000
-lay <- 2019
-ly  <- 2019
+fay <- 1990
+fy  <- 1990
+lay <- 2022
+ly  <- 2021
 
@@
 
-**Summary of ICES 2019 assessments and advice**
+**Paper 03 Short term forecast analysis**
 
@@
 
-In this document an overview is presented of the ICES assessment and advice for the most recent number of years for the stocks that are relevant to the PFA and Dutch fisheries in general.
-
 <!--Section 2: material and methods ----------------------------------------------------------- -->
@@
 
-The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.
-
 <!--Section 3: advice and assessment overviews ------------------------------------------------ -->
@@
 # Assessment and advice overviews
-
-## North Sea herring
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "her-47d3"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               
-             plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-
-```
-
-The assessment of North Sea herring was benchmarked in February 2018.
-There have been some revisions in the model and in the data that is being used for North Sea herring.
-Most important it the way of handling natural mortality estimates.
-This is one of the few assessments where natural mortality is derived from the multispecies model of the North Sea.
-This multispecies model is rerun every three years leading  to new estimates of natural mortality for the whole time-series.
-Because of changes in the multispecies model and in the underlying data for predator species (e.g.
-hake, cod, saithe), the natural mortality estimates of herring have been revised upwards and downwards several times.
-This then also leads to rescaling of the estimates of stock size and fishing mortality.
-During the benchmark, a new method to standardize natural mortality estimates has been introduced which would hopefully circumvent re-scaling in the future?).
-
-Last year, the combination of low recruitment, fishing harder than recommended and a lower Fmsy, lead to a substantially lower advice for 2019 (291 000 tonnes).
-This year, the assessments estimates - overall - a somewhat higher stock size compared to the estimate in 2018.
-This is due to the 2018 herring acoustic survey, although the exact reason is not well understood.
-The advice for 2020 is higher than the advice that was given for 2019, despite that recruitment remains low and that the overall stock biomass is declining.
-The reason the advice is higher is because the catch forecast takes into account al the ages and weights of fish the make up the fishery.
-This is called the selectivity pattern.
-The fishery is heavily dependent on the abundance of the last strong year class (2013), which will be 7 year olds next year.
-The TAC advice uses only F on ages 2-6 years old, and this, together with the abundance and higher weight of 7 year olds is what is leading to a higher TAC for 2020.
-
-[her.27.3a47d.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.3a47d.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Atlanto-scandian herring
-
-# s   <- "her-noss"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# A new long-term management strategy for Atlanto-scandian herring has been evaluated and agreed.
-Fmsy and the fishing mortality target are now higher than the previous values being used.
-The advice for 2019 is based on following the new management plan, with a TAC of around 590 000 tonnes (compared to an advice of 384 000 tonnes that was advised for 2018).
-# 
-# The stock assessment is consistent with last year's assessment.
-The stock appears to be slowly declining.
-Fishing mortality has risen sharply in the last year (due to the erroneous advice in 2016).
-There are signs of a reasonable sized incoming year class.
-The reason that the advice is substantially higher than last year is due to the new management plan that has a higher target fishing mortality.
-Last year, the advice was based on a target fishing mortality of 0.086, this year it is based on 0.14.
-# 
-# [her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.1-24a514a.pdf)
-
-```
-
-## Northeast Atlantic Mackerel
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "mac-nea"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE, 
-             plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-
-# [mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/mac.27.nea.pdf)
-
-```
-
-The ICES advice on Northeast Atlantic mackerel that was released in October 2018 has been received with a lot of scepticism and unbelief.
-According to that advice, the stock had been declining since 2011 and was estimated to be just below the MSY Btrigger.
-At the time, it was already found that the model results were very sensitive to the inclusion of the RFID tags.
-Leaving out only the RFID tags, gave a much higher stock estimate.
-
-To deal with this assessment problem, an inter-benchmark has been set up which is lead to a new assessment configuration in March 2019 and an updated advice for 2019 in May 2019.
-The new assessment configuration uses a more limited tagging dataset (restricted in ages, years and years out at sea) and a different way of estimating recruitment in the model.
-This lead to a substantially higher stock estimate and a substantial revision of the advice for 2019.
-
-[mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/special_requests/no.2019.09.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Western horse mackerel
-
-# s   <- "hom-west"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# The stock assessment of Western horse mackerel is comparable to last year but there has been another revision upwards of the stock.
-This is especially relevant in conjunction with the biomass reference points that have been set after the benchmark in 2017 and where the biomass in 2015 (the last year in the time series) was selected as Blim.
-Because of the upward revisions in the assessments of last year and this year, the stock is now well above the absolute Blim and in fact close to MSY Btrigger even though the stock is at the lowest point in the time-series!
-Fishing mortality is estimated to be well below the Fmsy.
-The advice for 2019 is 145 000 tonnes, which is 24% higher than the advice for 2018.
-The increase in the advice is because the stock size has been revised upwards and because the recent catches have been taken with a fishing mortality which was lower than Fmsy.
-The advice is based on the application of Fmsy.
-# 
-# The assessment of western horse mackerel is very uncertain and there has been frequent re-scaling of the assessment during every new assessment.
-This re-scaling is still happening.
-For western horse mackerel this results in the overall stock to be estimated higher than the year before.
-That is why the stock is now estimated to be close to MSY Btrigger while the stock is at the same time around the lowest point in the time series.
-The resulting advice is a direct application of the ICES advice rule to an assessment that could very well be too optimistic.
-#  
-# [hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hom.27.2a4a5b6a7a-ce-k8.pdf)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Blue whiting
-
-# s   <- "whb-comb"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# The stock size of blue whiting was high in 2018 but is predicted to be substantially lower in 2019 due to low recruitments coming in.
-The stock size of blue whiting appears to have been overestimated.
-Catches have been well above 1 Million tonnes over the last few years and above the scientific advice.
-Fishing mortality is estimated to be above Fmsy.
-The advice for 2019 is 1 144 000 tonnes, which is 18% lower than the advice for 2018.
-# 
-# The blue whiting survey went well in 2018 and showed the highest biomass of the time series.
-Despite the high catches, the stock is still at a very high level although a decline does appear to have started.
-The most recent year classes of blue whiting are very weak so that decline in catching opportunities in the further years is foreseen.
-# 
-# [whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/whb.27.1-91214.pdf)
-
-```
-
-## Western Baltic Spring Spawning herring
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-s   <- "her-3a22"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-
-```
-
-Assessment was benchmarked in 2018 when reference points were also re-estimated.
-Redefinition of Blim from 90 kT to 120 kT which lead to stock having been below Blim for a number of years already.
-No possibility of rebuilding stock to Blim within one year (which is the unwritten new ICES rule), and therefore a zero catch advice.
-
-The new assessment has estimated the stock to be lower compared to the assessment of last year. This lead to a repeat of the zeor advice.
-
-[her.27.20-24.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.20-24.pdf)
-
-## Baltic Sea herring
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "her-2532-gor"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               
-             plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-
-```
-
-Assessment largely consistent with last year's assessment. The 2014 year class is estimated to be strong.
-
-[her.27.25-2932.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.25-2932.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Herring in 6a and 7bc
-
-# s   <- "her-67bc"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
-# 
-# The advice for a zero TAC is the same as in the previous years.
-Unfortunately ICES has not presented advice on the proposed rebuilding plan that has been developed by the PELAC and that has been submitted for review to ICES and that had been evaluated by the herring working group in March 2018.
-In previous years the monitoring TAC for herring in 6a and 7bc had been suggested by ICES as a means to maintain the knowledge base for this herring stock.
-The monitoring TAC is still mentioned under issues relevant for the advice.
-# 
-# [her.27.6a7bc.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.6a7bc.pdf)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Herring in the Irish Sea
-
-# s   <- "her-nirs"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Stock was benchmarked in 2017. SSB is well above MSY Btrigger and fishing mortality well below Fmsy. Potential mixing of herring in the Irish Sea and the Celtic Sea. 
-# 
-# [her.27.nirs.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.nirs.pdf)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Herring in the Celtic Sea
-
-# s   <- "her-irls"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Stock went through interbenchmark in 2018 however perception of the stock trends is very similar to last year.
-Evaluations conducted in 2018 showed that the agreed management strategy is no longer precautionary.
-Therefore the advice is according to the MSY approach.
-# 
-# [her.27.irls.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.irls.pdf)
-
-```
-
-## Sprat in the North Sea
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- c("spr-nsea", "spr.27.3a4")
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-
-```
-
-Assessment was benchmarked in 2019, leading to a combination of sprat in 4 and sprat in 3a into a single stock.
-Some changes to the model al well, leading to a different perception of the history of the stock and lower stocksize in recent years.
-
-[spr.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.4.pdf)
-
-## Sprat in the Baltic Sea
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-s   <- "spr-2232"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-
-```
-
-Advice according to the Baltic Multiannual Plan (MAP).
-
-[spr.27.22-32.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.22-32.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Greater argentine in 5b and 6a
-
-# s   <- "arg-5b6a"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# Latest advice was from 2017, applicable for two years. Next advice foreseen in 2019. 
-# 
-# [aru.27.5b6a.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2017/2017/aru.27.5b6a.pdf)
-
-```
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Northern Hake
-
-s   <- "hke-nrtn"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-
-```
-
-Stock size is well above MSY Btrigger and fishing mortality is below Fmsy.
-In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.
-
-[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## North Sea cod
-
-s   <- "cod-347d"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-
-```
-
-The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years.
-According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey.
-The 2017 SSB is 32% lower compared to the estimate that was presented last year.
-This also leads to a substantially lower advice for 2019 compared to the advice for 2018.
-
-[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)
-
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## North Sea plaice
-
-s   <- "ple-nsea"
-
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
-
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-
-```
-
-Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy.
-An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016).
-This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock.
-The advice is based on the MSY approach.
-
-[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)
 
@@
 
+d1 <-
+  iAdvice %>% 
+  filter(stockkeylabelold == s) %>% 
+  filter(adviceonstock) %>% 
+  filter(tacyear >= fy, tacyear <= ly) %>% 
+  dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear, year=tacyear, f=fset, ssb=ssbset) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+d2 <-
+  iAssess %>% 
+  filter(stockkeylabelold==s) %>% 
+  # filter(assessmentyear == max(assessmentyear)) %>% 
+  filter(year >= fy, year <= ly) %>% 
+  dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear, year, ssb=stocksize, f=fishingpressure) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+d2 %>% 
+  ggplot(aes(x=year, y=data)) +
+  theme_bw() +
+  geom_line(aes(colour=as.character(assessmentyear), group=assessmentyear)) +
+  geom_point(data=d1, aes(colour=as.character(assessmentyear))) +
+  labs(title=s) +
+  expand_limits(y=0) +
+  facet_wrap(~variable, scales="free_y")
+
+d2 %>% 
+  filter(variable=="f") %>%
+
+  ggplot(aes(x=year, y=data)) +
+  theme_bw() +
+  theme(legend.position="none") +
+  geom_line(aes(colour=as.character(assessmentyear), group=assessmentyear)) +
+  geom_point(data=filter(d1,
+                         variable=="f"), 
+             aes(colour=as.character(assessmentyear))) +
+  labs(title=s) +
+  expand_limits(y=0) +
+  facet_wrap(~assessmentyear, scales="free_y")
+
 ```
@@
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
-
-## Atlantic Seabass
-
-# s   <- "bss-47"
-# 
-# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
-# 
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
-# 
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
-# 
-# ICES has reissued the advice for 2018 and issued an advice for 2019.
-ICES was requested by the EU to provide updated advice on fishing opportunities for 2018 because a benchmark of the assessment had been conducted in early 2018.
-The 2018 benchmark has not changed the perception of the current SSB (B below Blim).
-However, given the current low fishing mortality and the above-average recruitment in 2013 and 2014, SSB in 2019 and 2020 is predicted to be above Blim, allowing for some catches in 2018 and 2019.
-# 
-# [bss.27.4bc7ad-h.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/bss.27.4bc7ad-h.pdf)
-
-```
-
 # More information

### 2023-04-20 Rmd/Metadata overview.Rmd

@@
 
-  filter(nssbset > 0)
+  filter(nssbset > 0 | nfset > 0 )
+
+# writexl::write_xlsx(d1, "meta.xlsx")
 
@@
 d2 %>% 
+  mutate(
+    stockkeylabelold = case_when(
+      stockkeylabelold == "cod-nsea" ~ "cod-347d",
+      stockkeylabelold == "ple-nsea2" ~ "ple-nsea",
+      TRUE                            ~ stockkeylabelold
+    ) 
+  ) %>% 
   tidyr::pivot_longer(names_to = "variable", values_to = "data", nblim:ncatch) %>% 

### 2023-04-20 Rmd/Paper 03 short term forecast analysis.Rmd

@@
 # Load dataset
-load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
-load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+iAssess <-
+  loadRData(file=paste(dropboxdir, "/rdata/iAssess.RData",sep="")) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", "ple-nsea", stockkeylabelold))
+
+iAdvice <-
+  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", "ple-nsea", stockkeylabelold))
 
@@
 
-s   <- "sol-nsea"
+s   <- "ple-nsea"
 

### 2023-04-24 Rmd/Paper 03 short term forecast analysis.Rmd
short term forecast paper

--- 2022-12-17 Rmd/Paper 03 short term forecast analysis.Rmd
@@
 library(lubridate)
+library(ggrepel)
+library(patchwork)
 
@@
 source("../../prf/r/my utils.r")
+source("../r/iAssess functions.r")
 
@@
 
+mystocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west", 
+              "ple-nsea", "cod-nsea", "sol-nsea","had-nsea")
+
+demstocks <- c("ple-nsea", "cod-nsea", "sol-nsea","had-nsea")
+pelstocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west")
+
 # Load dataset
-load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
-load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))
+iAssess <-
+  loadRData(file=paste(dropboxdir, "/rdata/iAssess.RData",sep="")) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
+                                   "ple-nsea", 
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
+                                   "had-nsea",
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
+                                   "cod-nsea",
+                                   stockkeylabelold)) %>% 
+  filter(stockkeylabelold %in% mystocks)
+
+# iAdvice %>% ungroup() %>% arrange(stockkeylabelold) %>% distinct(stockkeylabelold) %>% View()
+
+iAdvice <-
+  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
+                                   "ple-nsea", 
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
+                                   "had-nsea",
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
+                                   "cod-nsea",
+                                   stockkeylabelold)) %>% 
+  filter(stockkeylabelold %in% mystocks)
+
+# iAdvice %>% ungroup() %>% distinct(stockkeylabelold) %>% filter(grepl("had-",stockkeylabelold)) %>% View()
 
@@
 
-# stock=c("spr-nsea","spr.27.3a4"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark==TRUE
-# stock=c("mac-nea"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark=TRUE; plot=c("stocksize","fishingpressure","recruitment")
-
-# d %>% filter(stockkeylabelold == "mac-nea", assessmentyear == 2019) %>% View()
-# unique(iAssess$purpose)
-
-# Function to plot historical assessments
-plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, include.benchmark = FALSE, 
-                         plot=c("stocksize","fishingpressure")) {
-
-  d <-
-    iAssess %>% 
-    ungroup() %>%
-
-    {if (include.benchmark == TRUE) {
-      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
-
-    filter(assessmentyear   %in% c(firstassessyear:lastassessyear), 
-           year             %in% firstyear:lastyear) %>%
-
-    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>%
-
-    #set type
-    mutate(
-      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"       , "last"      , NA),
-      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose),  "last_bench", type),  
-      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"       , "advice"    , type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose),  "benchmark" , type),
-      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose) , "replaced"  , type),
-      type = ifelse(is.na(type), "other", type)) %>%
-
-    # set group
-    mutate(group = paste0(tyear, type)) %>%
-
-    select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
-           type, group, source,
-           lowrecruitment, recruitment, highrecruitment, 
-           lowfishingpressure, fishingpressure, highfishingpressure, 
-           lowstocksize, stocksize, highstocksize) %>%
-    data.frame()
-
-  # find out what is the last assessment
-  last <- filter(d, assessmentyear == lastassessyear) %>% distinct(type) %>% as.character()
-
-  # plot stocksize
-  stocksize <-
-    d %>% 
-    filter(!is.na(stocksize)) %>%
-
-    ggplot(aes(year,stocksize, group=group)) +
-
-    theme_publication() +
-
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
-               colour="red", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes = FALSE) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=msybtrigger), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, msybtrigger, label = "msy btrigger", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "stocksize")
-
-  # fishingpressure
-  fishingpressure <-
-    d %>% 
-    filter(!is.na(fishingpressure)) %>%
-
-    ggplot(aes(year,fishingpressure, group=group)) +
-
-    theme_publication() +
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowfishingpressure, ymax=highfishingpressure, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes(x=assessmentyear, yintercept=fmsy), 
-               colour="red", linetype="dashed", inherit.aes=FALSE) +
-    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
-               aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
-               colour="red", inherit.aes=FALSE) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "fishingpressure")
-
-  # Recruitment
-  recruitment <-
-    d %>% 
-    filter(!is.na(recruitment)) %>%
-
-    ggplot(aes(year,recruitment, group=group)) +
-
-    theme_publication() +
-    theme(legend.title=element_blank(),
-          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
-          axis.text.y = element_text(size=9),
-          plot.title  = element_text(size=10), 
-          plot.margin = margin(2, 2, 2, 2, "mm"),
-          # strip.background = element_blank(),
-          legend.position = "null") +
-
-    geom_ribbon(data=filter(d, type==last), 
-                aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
-                alpha=0.2) +
-    geom_line(aes(colour = type, size=type, linetype=type) ) +
-
-    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +
-
-    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
-    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
-    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +
-
-    expand_limits(y = 0) +
-    xlim(firstyear,lastyear) +
-    labs(x = NULL, y = NULL , title = "Recruitment")
-
-    if (length(plot) == 1) {
-      if ("stocksize" %in% plot ) print(stocksize)
-      if ("fishingpressure" %in% plot ) print(fishingpressure)
-      if ("recruitment" %in% plot ) print(recruitment)
-
-    } else if (length(plot) == 2) {
-
-      if ("stocksize" %in% plot & "fishingpressure" %in% plot) plot_grid(stocksize, fishingpressure,  ncol=2, align = 'hv') 
-      if ("stocksize" %in% plot & "recruitment" %in% plot) plot_grid(stocksize, recruitment,  ncol=2, align = 'hv') 
-      if ("recruitment" %in% plot & "fishingpressure" %in% plot) plot_grid(fishingpressure, recruitment,  ncol=2, align = 'hv')
-
-    } else if (length(plot) == 3) {
-        plot_grid(stocksize, fishingpressure,  recruitment, ncol=length(plot), align = 'hv') 
-    }
-
-  } # End of plot_history function
-
-# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"), include.benchmark = TRUE)
-
-# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("spr-nsea","spr.27.3a4"); include.replaced <- TRUE
-# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE
-
-# Function to plot advice overviews
-table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {
-
-  d <-
-    iAdvice %>%
-
-    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%
-
-    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%
-
-    {if (include.replaced == TRUE) {
-      filter(., purpose %in% c("advice","replaced"))
-    } else {
-      filter(., purpose %in% c("advice")) 
-    }} %>%
-
-    mutate(advicebasis= substr(advicebasis,1,25),
-
-           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
-           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),
-
-           tac        = ifelse(!is.na(tac), tac, tal),
-           tactype    = ifelse(!is.na(tac), "catch", "landings"),
-
-           catch      = ifelse(!is.na(catches), catches, NA),
-           catchtype  = ifelse(!is.na(catches), "catches", ""),
-
-           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
-           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),
-
-           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
-           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),
-
-           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
-           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
-           cat_c       = 100 * (catch / lag(catch, n=1) -1)
-           ) %>%
-
-    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
-           adv_c, tac_c, cat_c,
-           advicetype, tactype, catchtype)
-
-  d %>% 
-  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
-  arrange(tacyear, desc(purpose)) %>% 
-  pandoc.table(., 
-               style        = "simple",
-               split.tables = 200, 
-               justify      = "right",
-               missing      =" ",
-               big.mark     = '', 
-               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
-}
-
-# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)
+myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,6,8,10)]
+names(myColors)         <- c("1990", "2000", "2010", "2020", "2022")
 
@@
 
-```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}
+**Demersal stocks**
 
-## North Sea sole
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
 
-s   <- "sol-nsea"
+## Demersal
 
-cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
+# advice
+d1 <-
+  iAdvice %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  filter(adviceonstock) %>%
+  filter(tacyear >= fy, tacyear <= ly) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, 
+                f=fset, ssb=ssbset) %>%
+  mutate(decade = as.character(10 * floor(year/10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
+                               width=2, pad="0")) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
 
-plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"))
+# assess
+d2 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  filter(purpose=="advice") %>% 
+  # filter(assessmentyear == max(assessmentyear)) %>%
+  filter(year >= fy, year <= ly) %>%
+  dplyr::select(stockkeylabelold, assessmentyear, year, 
+                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
+  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
 
-table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
+i <- 1
+for (i in 1:length(demstocks)) {
+
+  assign(paste0("p",i),
+         d2 %>% 
+           filter(stockkeylabelold == demstocks[[i]]) %>%
+
+           ggplot(aes(x=year, y=data, group=assessmentyear)) +
+           theme_publication() +
+           geom_line(aes(colour=decade)) +
+           geom_point(data=filter(d1, stockkeylabelold == demstocks[[i]]), 
+                      aes(x=year, y=data, colour=decade),
+                      inherit.aes=FALSE) +
+           ggrepel::geom_text_repel(
+                      data=filter(d1, stockkeylabelold == demstocks[[i]]),
+                      aes(x=year, y=data, colour=decade, label=yr),
+                      inherit.aes=FALSE) +
+           expand_limits(y=0) +
+           scale_colour_manual(values=myColors) +
+           labs(x="",y="", title=demstocks[[i]]) +
+           facet_wrap(~variable, scales="free_y")
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
+
+```
+
+\newpage
+
+**Pelagic stocks**
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+## Pelagic
+
+# advice
+d1 <-
+  iAdvice %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(adviceonstock) %>%
+  filter(tacyear >= fy, tacyear <= ly) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, 
+                f=fset, ssb=ssbset) %>%
+  mutate(decade = as.character(10 * floor(year/10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
+                               width=2, pad="0")) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
+
+# assess
+d2 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(purpose=="advice") %>% 
+  # filter(assessmentyear == max(assessmentyear)) %>%
+  filter(year >= fy, year <= ly) %>%
+  dplyr::select(stockkeylabelold, assessmentyear, year, 
+                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
+  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
+
+i <- 1
+for (i in 1:length(pelstocks)) {
+
+  assign(paste0("p",i),
+         d2 %>% 
+           filter(stockkeylabelold == pelstocks[[i]]) %>%
+
+           ggplot(aes(x=year, y=data, group=assessmentyear)) +
+           theme_publication() +
+           geom_line(aes(colour=decade)) +
+           geom_point(data=filter(d1, stockkeylabelold == pelstocks[[i]]), 
+                      aes(x=year, y=data, colour=decade),
+                      inherit.aes=FALSE) +
+           ggrepel::geom_text_repel(
+                      data=filter(d1, stockkeylabelold == pelstocks[[i]]),
+                      aes(x=year, y=data, colour=decade, label=yr),
+                      inherit.aes=FALSE) +
+           scale_colour_manual(values=myColors) +
+           expand_limits(y=0) +
+           labs(x="",y="", title=demstocks[[i]]) +
+           facet_wrap(~variable, scales="free_y")
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
+
+```
+
+\newpage
+
+**Demersal F**
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot of F against Fset and SSB against SSBset for multiple assessments
+
+mytype    <- "Demersal"
+myvar     <- "f"
 
@@
   iAdvice %>%
-  filter(stockkeylabelold == s) %>% 
+  filter(stockkeylabelold %in% demstocks) %>%
   filter(adviceonstock) %>%
-  filter(tacyear >= fy, tacyear <= ly) %>% 
-  dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear, year=tacyear, f=fset, ssb=ssbset) %>% 
+  filter(tacyear >= fay, tacyear <= lay) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
@@
   iAssess %>%
-  filter(stockkeylabelold==s) %>% 
-  # filter(assessmentyear == max(assessmentyear)) %>% 
-  filter(year >= fy, year <= ly) %>% 
-  dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear, year, ssb=stocksize, f=fishingpressure) %>% 
+  filter(stockkeylabelold %in% demstocks) %>%
+  filter(purpose=="advice") %>% 
+  filter(assessmentyear %in% c(2000, 2010, 2020, 2022)) %>% 
+  filter(year >= fay, year <= lay) %>%
+  mutate(source=paste0("assess", assessmentyear)) %>% 
+  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
 
+t <-
   d2 %>% 
-  ggplot(aes(x=year, y=data)) +
-  theme_bw() +
-  geom_line(aes(colour=as.character(assessmentyear), group=assessmentyear)) +
-  geom_point(data=d1, aes(colour=as.character(assessmentyear))) +
-  labs(title=s) +
-  expand_limits(y=0) +
-  facet_wrap(~variable, scales="free_y")
+  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
+  rename(
+    assess=data.x,
+    forecast=data.y
+  ) %>% 
+  drop_na(forecast) %>% 
+  mutate(decade = as.character(10 * floor(year / 10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))
 
-d2 %>% 
-  filter(variable=="f") %>%
+# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
+#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
+#   as.numeric()
 
-  ggplot(aes(x=year, y=data)) +
-  theme_bw() +
-  theme(legend.position="none") +
-  geom_line(aes(colour=as.character(assessmentyear), group=assessmentyear)) +
-  geom_point(data=filter(d1,
-                         variable=="f"), 
-             aes(colour=as.character(assessmentyear))) +
-  labs(title=s) +
-  expand_limits(y=0) +
-  facet_wrap(~assessmentyear, scales="free_y")
+i <- 1
+for (i in 1:length(demstocks)) {
+
+  max  <- t %>% 
+    filter(variable==myvar) %>% 
+    filter(stockkeylabelold == demstocks[[i]]) %>% 
+    ungroup() %>%
+    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
+    as.numeric()
+
+  assign(paste0("p",i),
+         t %>% 
+           filter(variable == myvar) %>% 
+           filter(stockkeylabelold == demstocks[[i]]) %>%
+
+           ggplot(aes(x=assess, y=forecast)) +
+           theme_publication() +
+           geom_point(aes(colour=decade)) +
+           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
+           scale_colour_manual(values=myColors) +
+           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
+           scale_x_continuous(limits=c(0,max)) +
+           scale_y_continuous(limits=c(0,max)) +
+           facet_grid(stockkeylabelold~source)
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
 
@@
 
-Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea.
+\newpage
 
-[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)
+**Demersal SSB**
 
-# More information
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
 
-Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.
+# plot of F against Fset and SSB against SSBset for multiple assessments
+
+mytype    <- "Demersal"
+myvar     <- "ssb"
+
+d1 <-
+  iAdvice %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  filter(adviceonstock) %>%
+  filter(tacyear >= fay, tacyear <= lay) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+d2 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  filter(purpose=="advice") %>% 
+  filter(assessmentyear %in% c(2000, 2010, 2020, 2022)) %>% 
+  filter(year >= fay, year <= lay) %>%
+  mutate(source=paste0("assess", assessmentyear)) %>% 
+  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+t <-
+  d2 %>% 
+  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
+  rename(
+    assess=data.x,
+    forecast=data.y
+  ) %>% 
+  drop_na(forecast) %>% 
+  mutate(decade = as.character(10 * floor(year / 10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))
+
+# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
+#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
+#   as.numeric()
+
+i <- 1
+for (i in 1:length(demstocks)) {
+
+  max  <- t %>% 
+    filter(variable==myvar) %>% 
+    filter(stockkeylabelold == demstocks[[i]]) %>% 
+    ungroup() %>%
+    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
+    as.numeric()
+
+  assign(paste0("p",i),
+         t %>% 
+           filter(variable == myvar) %>% 
+           filter(stockkeylabelold == demstocks[[i]]) %>%
+
+           ggplot(aes(x=assess, y=forecast)) +
+           theme_publication() +
+           geom_point(aes(colour=decade)) +
+           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
+           scale_colour_manual(values=myColors) +
+           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
+           scale_x_continuous(limits=c(0,max)) +
+           scale_y_continuous(limits=c(0,max)) +
+           facet_grid(stockkeylabelold~source)
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
+
+```
+
+\newpage
+
+**Pelagic F**
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot of F against Fset and SSB against SSBset for multiple assessments
+
+mytype    <- "Pelagic"
+myvar     <- "f"
+
+d1 <-
+  iAdvice %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(adviceonstock) %>%
+  filter(tacyear >= fay, tacyear <= lay) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+d2 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(purpose=="advice") %>% 
+  filter(assessmentyear %in% c(2000, 2010, 2020, 2022)) %>% 
+  filter(year >= fay, year <= lay) %>%
+  mutate(source=paste0("assess", assessmentyear)) %>% 
+  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+t <-
+  d2 %>% 
+  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
+  rename(
+    assess=data.x,
+    forecast=data.y
+  ) %>% 
+  drop_na(forecast) %>% 
+  mutate(decade = as.character(10 * floor(year / 10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))
+
+# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
+#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
+#   as.numeric()
+
+i <- 1
+for (i in 1:length(pelstocks)) {
+
+  max  <- t %>% 
+    filter(variable==myvar) %>% 
+    filter(stockkeylabelold == pelstocks[[i]]) %>% 
+    ungroup() %>%
+    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
+    as.numeric()
+
+  assign(paste0("p",i),
+         t %>% 
+           filter(variable == myvar) %>% 
+           filter(stockkeylabelold == pelstocks[[i]]) %>%
+
+           ggplot(aes(x=assess, y=forecast)) +
+           theme_publication() +
+           geom_point(aes(colour=decade)) +
+           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
+           scale_colour_manual(values=myColors) +
+           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
+           scale_x_continuous(limits=c(0,max)) +
+           scale_y_continuous(limits=c(0,max)) +
+           facet_grid(stockkeylabelold~source)
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
+
+```
+
+\newpage
+
+**Pelagic SSB**
+
+```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+
+# plot of F against Fset and SSB against SSBset for multiple assessments
+
+mytype    <- "Pelagic"
+myvar     <- "ssb"
+
+d1 <-
+  iAdvice %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(adviceonstock) %>%
+  filter(tacyear >= fay, tacyear <= lay) %>%
+  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+d2 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% pelstocks) %>%
+  filter(purpose=="advice") %>% 
+  filter(assessmentyear %in% c(2000, 2010, 2020, 2022)) %>% 
+  filter(year >= fay, year <= lay) %>%
+  mutate(source=paste0("assess", assessmentyear)) %>% 
+  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+t <-
+  d2 %>% 
+  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
+  rename(
+    assess=data.x,
+    forecast=data.y
+  ) %>% 
+  drop_na(forecast) %>% 
+  mutate(decade = as.character(10 * floor(year / 10))) %>% 
+  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))
+
+# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
+#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
+#   as.numeric()
+
+i <- 1
+for (i in 1:length(pelstocks)) {
+
+  max  <- t %>% 
+    filter(variable==myvar) %>% 
+    filter(stockkeylabelold == pelstocks[[i]]) %>% 
+    ungroup() %>%
+    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
+    as.numeric()
+
+  assign(paste0("p",i),
+         t %>% 
+           filter(variable == myvar) %>% 
+           filter(stockkeylabelold == pelstocks[[i]]) %>%
+
+           ggplot(aes(x=assess, y=forecast)) +
+           theme_publication() +
+           geom_point(aes(colour=decade)) +
+           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
+           scale_colour_manual(values=myColors) +
+           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
+           scale_x_continuous(limits=c(0,max)) +
+           scale_y_continuous(limits=c(0,max)) +
+           facet_grid(stockkeylabelold~source)
+         )
+}
+
+p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom')
+
+```

### 2023-07-19 Rmd/Paper 03 short term forecast analysis.Rmd
Several updates

@@
 mystocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west", 
-              "ple-nsea", "cod-nsea", "sol-nsea","had-nsea")
+              "ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")
 
-demstocks <- c("ple-nsea", "cod-nsea", "sol-nsea","had-nsea")
+demstocks <- c("ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")
 pelstocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west")
@@
 
+assessmentyears         <- c(2015, 2020, 2022)
+
 myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,6,8,10)]
 names(myColors)         <- c("1990", "2000", "2010", "2020", "2022")
+
+# iAdvice %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
+# iAssess %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
+
+# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()
 
@@
 
+```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
+
+iAdvice %>% 
+  group_by(stockkeylabelold, tacyear) %>% 
+  summarize(
+    nF = sum(!is.na(fset)),
+    nSSB = sum(!is.na(ssbset))
+  ) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", nF:nSSB) %>% 
+  # mutate(data = ifelse(data==0, NA, data)) %>% 
+  filter(data > 0) %>% 
+  ggplot(aes(x=tacyear, y=variable)) +
+  theme_publication() +
+  geom_point(aes(colour=variable)) +
+  facet_wrap(~stockkeylabelold)
+
+```
+
 **Demersal stocks**
 
-```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
 
@@
 
-p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
+p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
   theme(legend.position='bottom')
@@
            expand_limits(y=0) +
-           labs(x="",y="", title=demstocks[[i]]) +
+           labs(x="",y="", title=pelstocks[[i]]) +
            facet_wrap(~variable, scales="free_y")
@@
 
-```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
 
@@
   filter(purpose=="advice") %>% 
-  filter(assessmentyear %in% c(2000, 2010, 2020, 2022)) %>% 
+  filter(assessmentyear %in% assessmentyears) %>% 
   filter(year >= fay, year <= lay) %>%
@@
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()
 
@@
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)
+
+# d2 %>% filter(stockkeylabelold=="hke-nrtn", source=="assess1990") %>% View()
 

### 2023-07-19 R/Analyse meta data.r

@@ -1,0 +1,98 @@
+# -----------------------------------------------------------------------------------------------
+# analyse meta data
+#
+# 15/11/2018 first coding
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
+library(reshape2)      # reshaping data; e.g. dcast
+library(pander)        # for print tables
+library(readxl)        # read excel files
+library(lubridate)     # dates
+library(cowplot)       # multiplots
+
+# Load utils code
+source("../prf/r/my utils.r")
+
+# Set working directory to dropbox folder
+advicedir  <- paste(get_dropbox(), "/iAdvice", sep="")
+
+# load(file=paste(advicedir, "/rdata/iStock.RData",sep=""))
+# load(file=paste(advicedir, "/rdata/qcsexcel.RData",sep=""))
+load(file=paste(advicedir, "/rdata/iAdvice.RData",sep=""))
+
+# -----------------------------------------------------------------------------------------
+# number of stockassessments per stock
+# -----------------------------------------------------------------------------------------
+
+p <-
+  iAdvice %>% 
+  filter(!grepl("benchmark|unofficial", purpose) ) %>% 
+  filter(!grepl("elasmobranch", speciestype) ) %>% 
+  filter(!grepl("grenadier|forkbeard|blue ling|roughy|blackspot|ling|skates", speciescommonname) ) %>% 
+  mutate(speciescommonname = ifelse(grepl("megrim", speciescommonname), "megrims",speciescommonname)) %>% 
+  mutate(speciescommonname = ifelse(grepl("anglerfish", speciescommonname), "anglerfishes",speciescommonname)) %>% 
+  mutate(speciescommonname = ifelse(grepl("redfish", speciescommonname), "redfishes",speciescommonname)) %>% 
+  mutate(assessmentmodel = substr(assessmentmodel2, 1, 40)) %>% 
+  group_by(speciescommonname, assessmentmodel) %>% 
+  summarize(n = n()) %>% 
+  ggplot(aes(y=assessmentmodel, x=speciescommonname)) +
+  theme_publication() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  geom_point(aes(size=n))
+
+png(filename=file.path("methods_species.png"),
[…]

### 2023-07-19 R/DownloadDataFromSAG.r

@@ -42,7 +42,7 @@ today <- format(Sys.time(), '%Y%m%d')
 # which year (set year = 0 for all years)
 # =====================================================================================
 
-myyear <- 2022
+myyear <- 2023
 # mystocks <- c("hom.27.2a4a5b6a7a-ce-k8","mac.27.nea","whb.27.1-91214","her.27.1-24a514a")
 
 # Load assessment data

### 2023-07-19 R/Plot of number of assessments per stock.R

@@ -18,7 +18,7 @@ library(lubridate)
 # library(ggstance)  # extension to ggplot; vertical extensions; only works for ggplot >= 3.0.0
 
 # Load utils code
-source("../mptools/r/my_utils.r")
+source("../prf/r/my utils.r")
 
 # Set working directory to dropbox folder
 dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

### 2023-07-19 R/iDatabases generator.r

@@ -175,6 +175,9 @@ iStockkey <-
 # write.csv(iRename, file=paste(dropboxdir, "/excel/iRename.csv",sep=""), row.names = FALSE)
 # write.csv(iStockkey, file=paste(dropboxdir, "/excel/iStockkey.csv",sep=""), row.names = FALSE)
 
+# load(file=paste(dropboxdir, "/rdata/iRename.RData",sep=""))
+# load(file=paste(dropboxdir, "/rdata/iStockkey.RData",sep=""))
+
 # -----------------------------------------------------------------------------------------
 # load Excel and QSC data
 # -----------------------------------------------------------------------------------------
@@ -230,16 +233,16 @@ save(qcsexcel, file=paste(dropboxdir, "/rdata/qcsexcel.RData",sep=""))
 # -----------------------------------------------------------------------------------------
 
 sd <-
-  sd <- icesSD::getSD() %>% 
-  # loadRData(file=paste(dropboxdir, "/rdata/icesSD 20181023.RData",sep="")) %>% 
-  lowcase() %>% 
-  drop_na(stockkey) %>% 
-  dplyr::select(-stockkey) %>% 
-  distinct() %>% 
-  # dplyr::select(-stockkey, -previousstockkey, -previousstockkeylabel, -stockdatabaseid) %>% 
+  sd <- icesSD::getSD() %>%
+  # loadRData(file=paste(dropboxdir, "/rdata/icesSD 20181023.RData",sep="")) %>%
+  lowcase() %>%
+  drop_na(stockkey) %>%
+  dplyr::select(-stockkey) %>%
+  distinct() %>%
+  # dplyr::select(-stockkey, -previousstockkey, -previousstockkeylabel, -stockdatabaseid) %>%
   left_join(iRename[,c("stockkeylabel","stockkey")], by="stockkeylabel") %>%
-  left_join(iStockkey, by="stockkey") %>% 
-  rename(assessmentyear = activeyear) 
+  left_join(iStockkey, by="stockkey") %>%
+  rename(assessmentyear = activeyear)
 
 
 # -----------------------------------------------------------------------------------------
@@ -538,6 +541,7 @@ save(iAssess, file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
 #   distinct(stockkeylabel) %>% 
 #   View()
 
-
[…]

### 2023-07-19 R/presentation for MSC NE Atlantic event 2023.r

@@ -1,0 +1,1391 @@
+# -----------------------------------------------------------------------------------------------
+# Presentation for pelagic forum
+#
+# 03/11/2017 first coding
+# 15/10/2018 updated for 2018 version of the talk
+# 28/09/2019 updated for 2019 version of the talk
+# 18/04/2023 using Officer package (while saving files as png and jpg)
+# -----------------------------------------------------------------------------------------------
+
+library(tidyverse)
+library(cowplot)
+library(directlabels)  # for printing labels at end of geom lines
+library(scales)
+library(stringr)
+library(readxl)
+library(officer)
+library(flextable)
+
+# Load utils code
+source("../prf/r/my utils.r")
+source("R/iAssess functions.r")
+
+# Set dropbox folder
+dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")
+
+figuresdir <- file.path(dropboxdir, "figures")
+dir.create(figuresdir, showWarnings = FALSE)
+
+figuresdir <- file.path(dropboxdir, "figures", "msc2023")
+dir.create(figuresdir, showWarnings = FALSE)
+# file.remove(dir(figuresdir, full.names = TRUE))
+
+# load the data
+iSpecies <-
+  loadRData(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep="")) %>% 
+  group_by(speciesfaocode, trophicguild, fisheriesguild) %>% 
+  filter(row_number()==1) %>% 
+  ungroup() %>% 
+  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild) 
+
+iAssess <-
+  loadRData(file=paste(dropboxdir,"/rdata/iAssess.RData", sep="")) %>% 
+  distinct() %>% 
+  left_join(iSpecies, by="speciesfaocode")
+
+load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iForecast.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
+load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))
+
+lay               <- 2022
+my.lastyear       <- 2022
+my.firstyear      <- 2005
+
+# top colours
+# colourCount             <- nrow(top)+1
+# getPalette              <- colorRampPalette(brewer.pal(12, "Paired"))
[…]

### 2023-08-15 Rmd/Paper 03 short term forecast analysis v2.Rmd
UPdates STF paper

--- 2023-07-19 Rmd/Paper 03 short term forecast analysis.Rmd
@@
 # 17/12/2022 first coding
+# August 2023 updated coding
 # -----------------------------------------------------------------------------------------------
@@
 
+fay <- 1980
+fy  <- 1980
+lay <- 2023
+ly  <- 2023
+
 # Load dataset
@@
                                    stockkeylabelold)) %>% 
-  filter(stockkeylabelold %in% mystocks)
+  filter(purpose == "advice") %>% 
+  filter(stockkeylabelold %in% mystocks) %>% 
+  filter(year >= assessmentyear -9, year < assessmentyear) %>%
 
-# iAdvice %>% ungroup() %>% arrange(stockkeylabelold) %>% distinct(stockkeylabelold) %>% View()
+  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
+  mutate(pentad = as.character(5 * floor(assessmentyear/5))) %>% 
+  arrange(stockkeylabelold, assessmentyear, year) %>% 
+  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, 
+              ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
+  group_by(stockkeylabelold, pentad) %>% 
+  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
+  ungroup()
+
+# iAssess %>% filter(stockkeylabelold == "cod-nsea", pentad=="2010") %>% View()
+
+# add intermediate year from iAdvice database
+iAssess2 <-
+  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
+                                   "ple-nsea", 
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
+                                   "had-nsea",
+                                   stockkeylabelold)) %>% 
+  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
+                                   "cod-nsea",
+                                   stockkeylabelold)) %>% 
+  filter(adviceonstock == TRUE & purpose=="advice") %>% 
+  filter(stockkeylabelold %in% mystocks) %>% 
+  filter(tacyear >= fy, tacyear <= ly) %>%
+  rename(ssb = ssbimy, f = fsqimy) %>% 
+  mutate(year = assessmentyear + 1) %>% 
+  mutate(pentad = as.character(5 * floor(year/5))) %>% 
+  mutate(decade = as.character(10 * floor(year/10))) %>% 
+  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, ssb, f)
+
+# iAssess %>% ungroup() %>% arrange(stockkeylabelold) %>% distinct(stockkeylabelold, decade, assessmentyear) %>% View()
+# iAssess %>% group_by(stockkeylabelold, decade, assessmentyear) %>% summarise(minyear = min(year), maxyear = max(year)) %>% View()
 
@@
                                    stockkeylabelold)) %>% 
-  filter(stockkeylabelold %in% mystocks)
+  filter(adviceonstock == TRUE & purpose == "advice") %>% 
+  filter(stockkeylabelold %in% mystocks) %>% 
+  filter(tacyear >= fy, tacyear <= ly) %>%
 
-# iAdvice %>% ungroup() %>% distinct(stockkeylabelold) %>% filter(grepl("had-",stockkeylabelold)) %>% View()
+  rename(year = tacyear) %>%
 
-fay <- 1990
-fy  <- 1990
-lay <- 2022
-ly  <- 2021
+  mutate(pentad = as.character(5 * floor(year/5))) %>% 
+  mutate(decade = as.character(10 * floor(year/10))) %>% 
+  arrange(stockkeylabelold, year) %>% 
+  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, 
+              ssb=ssbset, f=fset) %>%
+  ungroup()
 
-assessmentyears         <- c(2015, 2020, 2022)
+# iAdvice %>% ungroup() %>% distinct(stockkeylabelold) %>%  View()
+# iAdvice %>% distinct(pentad) %>% View()
+# iAdvice %>% filter(pentad==1975) %>% View()
 
-myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,6,8,10)]
-names(myColors)         <- c("1990", "2000", "2010", "2020", "2022")
+# year strings
+assessmentyears         <- iAdvice %>% distinct(pentad, year) %>% group_by(pentad) %>% filter(year == max(year)) %>% pull(year)
+pentadyears             <- iAdvice %>% distinct(pentad) %>% pull(pentad)
+
+# colours
+myColors                <- RColorBrewer::brewer.pal(12, "Paired")[1:length(pentadyears)]
+names(myColors)         <- pentadyears
+
+# scales::show_col(RColorBrewer::brewer.pal(12, "Paired"))
 
 # iAdvice %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
+
 # iAssess %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
-
 # iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()
@@
 
+Reconstructed F and SSB by stock, calculated from observed catches and implied F and SSB from short term forecasts in different years
+
 ```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
@@
 iAdvice %>% 
-  group_by(stockkeylabelold, tacyear) %>% 
+  group_by(stockkeylabelold, year) %>% 
   summarize(
-    nF = sum(!is.na(fset)),
-    nSSB = sum(!is.na(ssbset))
+    nSTF_F = sum(!is.na(f)),
+    nSTF_SSB = sum(!is.na(ssb))
   ) %>% 
-  tidyr::pivot_longer(names_to = "variable", values_to = "data", nF:nSSB) %>% 
-  # mutate(data = ifelse(data==0, NA, data)) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", nSTF_F:nSTF_SSB) %>% 
   filter(data > 0) %>%
-  ggplot(aes(x=tacyear, y=variable)) +
+
+  bind_rows(
+    iAssess %>% 
+      ungroup() %>% 
+      distinct(stockkeylabelold, assessmentyear) %>% 
+      rename(year = assessmentyear) %>% 
+      mutate(variable = "nAssess", data=1)
+  ) %>%
+
+  ggplot(aes(x=year, y=variable)) +
   theme_publication() +
+  theme(legend.position = "none") +
   geom_point(aes(colour=variable)) +
@@
 
+Historical retrospective assessments of demersal stocks (SSB)
+
 ```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}
@@
 ## Demersal
+my_var <- "f"
 
@@
   filter(stockkeylabelold %in% demstocks) %>%
-  filter(adviceonstock) %>%
-  filter(tacyear >= fy, tacyear <= ly) %>%
-  dplyr::select(stockkeylabelold, year=tacyear, 
-                f=fset, ssb=ssbset) %>%
-  mutate(decade = as.character(10 * floor(year/10))) %>% 
+  # filter(stockkeylabelold == "cod-nsea") %>% 
   mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
@@
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
-  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
+  mutate(variable = factor(variable, levels=c("ssb","f","rec"))) %>%
+
+  filter(variable == my_var)
 
@@
   filter(stockkeylabelold %in% demstocks) %>%
-  filter(purpose=="advice") %>% 
-  # filter(assessmentyear == max(assessmentyear)) %>%
-  filter(year >= fy, year <= ly) %>%
-  dplyr::select(stockkeylabelold, assessmentyear, year, 
-                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
-  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
+  # filter(stockkeylabelold == "cod-nsea") %>%
+
+  group_by(stockkeylabelold, pentad) %>% 
+  filter(assessmentyear == max(assessmentyear)) %>%
+
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
-  mutate(variable = factor(variable, levels=c("ssb","f","rec")))
+  mutate(variable = factor(variable, levels=c("ssb","f","rec"))) %>%
 
-i <- 1
-for (i in 1:length(demstocks)) {
+  filter(variable == my_var)
 
-  assign(paste0("p",i),
+# iAssess %>% filter(stockkeylabelold == "cod-nsea", pentad=="2010") %>% View()
+
+# IMY assumptions for ssb and F
+d3 <-
+  iAssess2 %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  # filter(stockkeylabelold == "cod-nsea") %>%
+
+  group_by(stockkeylabelold, pentad) %>% 
+  filter(assessmentyear == max(assessmentyear)) %>%
+
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f"))) %>%
+
+  filter(variable == my_var)
+
+# connecting last assessment point with IMY assumptions
+d4 <-
+  iAssess %>%
+  filter(stockkeylabelold %in% demstocks) %>%
+  # filter(stockkeylabelold == "cod-nsea") %>%
+
+  group_by(stockkeylabelold, pentad) %>% 
+  filter(assessmentyear == max(assessmentyear)) %>% 
+  filter(year == max(year)) %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
+  mutate(variable = factor(variable, levels=c("ssb","f"))) %>%
+
+  bind_rows(d3) %>%
+
+  filter(variable == my_var)
+
+int_breaks <- function(x, n = 5) {
+  l <- pretty(x, n)
+  l[abs(l %% 1) < .Machine$double.eps ^ 0.5] 
+}
+
 d2 %>% 
-           filter(stockkeylabelold == demstocks[[i]]) %>%
-
   ggplot(aes(x=year, y=data, group=assessmentyear)) +
   theme_publication() +
-           geom_line(aes(colour=decade)) +
-           geom_point(data=filter(d1, stockkeylabelold == demstocks[[i]]), 
-                      aes(x=year, y=data, colour=decade),
+  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
+  theme(plot.title = element_text(size=12)) +
+  theme(panel.spacing.x = unit(0.1, "mm"), panel.spacing.y = unit(0.1, "mm")) +
+  theme(legend.position = "none") +
+
+  geom_line(aes(colour=pentad), alpha=1) +
+  ggrepel::geom_text_repel(
+             data=d2 %>% group_by(stockkeylabelold, decade, pentad, assessmentyear) %>% filter(row_number()==1),
+             aes(x=year, y=data, colour=pentad, label=assessmentyear),
              inherit.aes=FALSE) +
+  geom_point(data=d3, 
+             aes(x=year, y=data, colour=pentad),
+             inherit.aes=FALSE, alpha=1, shape=1) +
+  geom_line(data=d4, 
+             aes(x=year, y=data, colour=pentad),
+             inherit.aes=FALSE, alpha=1, linetype="dotted") +
+  geom_point(data=d1,
+             aes(x=year, y=data, colour=pentad),
+             inherit.aes=FALSE, alpha=1) +
   ggrepel::geom_text_repel(
-                      data=filter(d1, stockkeylabelold == demstocks[[i]]),
-                      aes(x=year, y=data, colour=decade, label=yr),
+             data=d1,
+             aes(x=year, y=data, colour=pentad, label=yr),
              inherit.aes=FALSE) +
@@
   scale_colour_manual(values=myColors) +
-           labs(x="",y="", title=demstocks[[i]]) +
-           facet_wrap(~variable, scales="free_y")
-         )
-}
-
-p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
-  theme(legend.position='bottom')
+  scale_x_continuous(breaks = int_breaks) +
+  labs(x="",y="", title=my_var) +
+  # facet_wrap(~pentad, scales="free")
+  facet_grid(stockkeylabelold~pentad, scales="free")
 

### 2023-08-15 Rmd/Paper 03 short term forecast analysis.Rmd
UPdates STF paper

@@
                                    stockkeylabelold)) %>% 
-  filter(stockkeylabelold %in% mystocks)
+  filter(stockkeylabelold %in% mystocks) %>% 
+  mutate(decade = as.character(10 * floor(assessmentyear/10)))
 
@@
                                    stockkeylabelold)) %>% 
-  filter(stockkeylabelold %in% mystocks)
+  filter(stockkeylabelold %in% mystocks) %>% 
+  mutate(decade = as.character(10 * floor(tacyear/10)))
 
@@
 
-fay <- 1990
-fy  <- 1990
-lay <- 2022
-ly  <- 2021
+fay <- 1980
+fy  <- 1980
+lay <- 2021
+ly  <- 2020
 
-assessmentyears         <- c(2015, 2020, 2022)
+assessmentyears         <- c(2005, 2010, 2015, 2020)
 
-myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,6,8,10)]
-names(myColors)         <- c("1990", "2000", "2010", "2020", "2022")
+myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,5, 6,8,10)]
+names(myColors)         <- c("1990", "2000", "2005", "2010", "2015", "2020")
 
@@
   tidyr::pivot_longer(names_to = "variable", values_to = "data", nF:nSSB) %>% 
-  # mutate(data = ifelse(data==0, NA, data)) %>% 
+  rename(year = tacyear) %>% 
   filter(data > 0) %>%
-  ggplot(aes(x=tacyear, y=variable)) +
+
+  bind_rows(
+    iAssess %>% 
+      ungroup() %>% 
+      distinct(stockkeylabelold, assessmentyear) %>% 
+      rename(year = assessmentyear) %>% 
+      mutate(variable = "nAssess", data=1)
+  ) %>% 
+  ggplot(aes(x=year, y=variable)) +
   theme_publication() +
@@
 **Demersal stocks**
+
+Historical retrospective assessments of demersal stocks
 
@@
   filter(tacyear >= fy, tacyear <= ly) %>%
-  dplyr::select(stockkeylabelold, year=tacyear, 
+  dplyr::select(stockkeylabelold, decade, year=tacyear, 
                 f=fset, ssb=ssbset) %>%
-  mutate(decade = as.character(10 * floor(year/10))) %>% 
   mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
@@
   filter(year >= fy, year <= ly) %>%
-  dplyr::select(stockkeylabelold, assessmentyear, year, 
+  dplyr::select(stockkeylabelold, decade, assessmentyear, year, 
                 ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
-  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
   tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
@@
            theme_publication() +
+           theme(plot.margin = unit(c(0,0,0,0), "cm")) +
+           theme(plot.title = element_text(size=12)) +
+           theme(panel.spacing.x = unit(0.1, "mm"), panel.spacing.y = unit(0.1, "mm")) +
+
            geom_line(aes(colour=decade)) +
@@
 
-p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
-  theme(legend.position='bottom')
+p1 + p2 + p3 + p4 + p5 + 
+  plot_layout(ncol = 1, guides="collect") &
+  theme(legend.position='bottom') &
+  theme(plot.margin = unit(c(0,0,0,0), "cm"))
 
